### 2017-01-21 paper/AlignedIsotropicGW.tex
Paper skeleton, with plots.

% !TEX TS-program = pdflatexmk

\documentclass[modern]{aastex61}

\usepackage{acronym}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\chieff}{\chi_\mathrm{eff}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\diff}[2]{\frac{\dd #1}{\dd #2}}
\newcommand{\pali}{p_\mathrm{ali}}
\newcommand{\piso}{p_\mathrm{iso}}

\newcommand{\checkme}[1]{\textcolor{red}{#1}}

\newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
\newcommand{\OOneOddsIsoAligned}{\checkme{0.044}}
\newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
\newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0012}}

\begin{document}

\acrodef{BBH}{binary black hole}
\acrodef{BH}{black hole}
\acrodef{EM}{electromagnetic}
\acrodef{GW}{gravitational wave}

\title{Distinguishing Between Spin-Aligned and Isotropic Binary Black
  Hole Populations in Advanced LIGO}

\author[0000-0003-1540-8562]{Will M. Farr}

\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\author{M. Coleman Miller}

\affiliation{Department of Astronomy and Maryland Center for Theory
  and Computation, University of Maryland, College Park MD
  20742, United States}

\author[0000-0002-6254-1617]{Alberto Vecchio}

\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\author[0000-0002-6134-8946]{Ilya Mandel}

\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
  University of Birmingham, Birmingham, B15 2TT, United Kingdom}

\author{Friends}

\email{w.farr@bham.ac.uk,miller@astro.umd.edu,av@star.sr.bham.ac.uk,imandel@star.sr.bham.ac.uk}

\begin{abstract}
  The first direct detections of \acp{GW} from merging \acp{BBH} in
  the last year open a unique window into the formation environment of
  these systems.  One promising signature of the formation environment
  is the angular distribution of the \ac{BH} spins; systems formed
  through dynamical interacions among already-compact objects are
  expected to have isotropic spin orientations while binaries from
  pairs of stars born together are more likely to have spins aligned
  with the binary orbit as a consequence of their joint evolution
  toward a \ac{BBH} system.  By considering existing \ac{GW}
  measurements of $\chieff$, the best-measured combination of spin
  parameters, in the three likely binary black hole detections
  GW150914, LVT151012, and GW151226, we show that the data already
  exhibit a $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
  ratio) preference for an isotropic angular distribution amongst a
  suite of models for the spin distribution.  By considering the
  effect of an additional 10 detections (a possible outcome of the
  ongoing second observing run of Advanced LIGO) drawn from the
  various models in the suite we show that such an agumented data set
  would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
  isotropic and aligned models, and in most cases better than
  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).
\end{abstract}

\acresetall{}

\section{Introduction}

This is a reference dump at the moment.

\section{Model Selection Formalism}
\label{sec:formalism}

This is currently pasted from the iPython notebook:

Cole had a great idea, the thrust of which is the following: assume
that you have some distribution for the spin magnitudes of black holes
(and it is common to the more- and less-massive BH in any binary).
The distribution of $\chieff$ that is implied by this magnitude
distribution differs depending on whether you assume that the angular
distribution is isotropic or aligned.  In general, isotropic angular
distributions concentrate more of the $\chieff$ distribution near
zero, since that is where most of the angular phase space is.  Here we
derive the distributions for $\chieff$ under various reasonable, yet
straw-man assumptions.

First, $\chieff$ is defined by
\begin{equation}
\chieff = \frac{c}{GM} \left( \frac{s_1^z}{m_1} + \frac{s_2^z}{m_2}
\right) = \frac{1}{M} \left( m_1 a_1^z + m_2 a_2^z \right)
\end{equation}
where the $\hat{z}$ axis is aligned with the orbital angular momentum,
$s_{1,2}$ are the spin angular momentum of the components of the
binary, and $a_{1,2}$ are the spin parameters of the black holes in
the binary.

First we derive the distribution for $a^z$ when we know the
distribution for $a$ and the angular distribution is (a) isotropic and
(b) aligned.  First isotropic: Then $a^z$ is equal to the product of
$a$ with a uniform random variable on $[-1,1]$ ($\iota = \cos \theta$,
with $\theta$ the colatitude on a sphere).  Therefore
\begin{equation}
  p\left( a^z \right) = \int_{-1}^{1} \dd \iota \, \int_0^1 \dd a \, \delta\left( a^z - a \iota \right) \frac{p(a)}{2} = \int_{-1}^{1} \dd \iota \, \int_0^1 \dd a \, \frac{\delta\left( a - \frac{a^z}{\iota} \right)}{\left| \iota \right|} \frac{p(a)}{2}.
\end{equation}
Now, noting that the distribution of $a^z$ is an even function, assume
that $a^z > 0$; since $0 \leq a \leq 1$, we must have $\iota \geq a^z$
if the $\delta$-function is to be "activated," so
\begin{equation}
p\left( a^z \right) = \int_{a^z}^1 \dd \iota \, \frac{p\left( \frac{a^z}{\iota} \right)}{2 \iota}.
\end{equation}
The even assumption implies 
\begin{equation}
p\left( a^z \right) = \int_{\left| a^z \right|}^1 \dd \iota \,
\frac{p\left( \frac{\left|a^z\right|}{\iota} \right)}{2 \iota}
\end{equation}
When the angles are aligned, we have
\begin{equation}
p\left( a^z \right) = \begin{cases}
p_a\left(a^z \right) & a^z > 0\\
0 & \mathrm{otherwise}
\end{cases}.
\end{equation}
In other words, the distribution of $a^z$ follows $a$ when $a^z$ is
positive, and is zero otherwise.

For example, when the magnitude distribution is flat, $p(a) = 1$, the
isotropic angle assumption implies that
\begin{equation}
\piso\left( a^z \right) = -\frac{\log \left| a^z \right|}{2},
\end{equation}
while the aligned assumption implies that
\begin{equation}
\pali\left( a^z \right) = \begin{cases}
1 & a^z > 0 \\
0 & a^z < 0
\end{cases}
\end{equation}

If the mass ratio, $q$, (recall: $0 \leq q \leq 1$) is zero, this is
all we need; if not, then
\begin{equation}
\chieff = \frac{1}{1+q}\left(a_1^z + q a_2^z\right),
\end{equation}
and the distribution is given by a convolution of the distributions
for $a^z$:
\begin{equation}
p\left(\chieff \right) = \int_{-1}^1 \dd a_1^z \, \int_{-1}^1 \dd a_2^z \, \delta\left( \chieff - \frac{1}{(1+q)} \left[a_1^z + q a_2^z \right]\right) p\left( a_1^z\right) p\left(a_2^z\right),
\end{equation}
which becomes, after some manipulation,
\begin{equation}
p\left(\chieff \right) = \int_{-1}^1 \dd a_1^z \, \int_{-1}^1 \dd a_2^z \, \delta\left( a_1^z + q a_2^z - (1+q)\chieff \right) (1+q) p\left( a_1^z\right) p\left(a_2^z\right),
\end{equation}
and more manipulation:
\begin{equation}
p\left(\chieff \right) = \int_{\max\left(-1, \frac{(1+q)\chieff - 1}{q} \right)}^{\min\left(1, \frac{(1+q)\chieff+1}{q}\right)} \dd a_2^z \, (1+q) p\left(a_2^z\right) p\left((1+q)\chieff - q a_2^z \right).
\end{equation}

Here we will consider three models for the spin magnitude distribution
of the two \ac{BH} spins.  The ``flat'' model assumes that magnitudes
are drawn uniformly in $[0, 1]$:
\begin{equation}
  \label{eq:flat-def}
  p(a) = 1.
\end{equation}
This model is maximally uncertain about the spin magnitude.  The
increasing distribution is meant to be more consistent with the
existing \ac{EM} observations of \acp{BH} in X-ray binaries
\citep{Miller2015}
\begin{equation}
  \label{eq:increasing-def}
  p(a) = 2a.
\end{equation}
The ``decreasing'' distribution is meant to be more consistent with
the observed low-spins in the existing \ac{GW} detections
\citep{O1-BBH}, with
\begin{equation}
  \label{eq:decreasing-def}
  p(a) = 2(1-a).
\end{equation}

This is not going to be analytic, even if we specialise to $q = 1$
(equal mass); though a good check of the math is that it works out to
the above single-spin distribution when $q=0$.  We should just
simulate the distribution empirically.  The result appears in Figure
\ref{fig:chieff-distribution-models}

\begin{figure}
  \plotone{../plots/chi-eff-distributions}
  \caption{\label{fig:chieff-distribution-models} The models for the
    distribution of $\chieff$ considered in this paper.  In all models
    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$.  The
    ``flat'' model (blue curves) for the spin magnitudes is defined in
    Eq.\ \eqref{eq:flat-def}; the ``increasing'' model (green curves)
    for the spin magnitudes is defined in Eq.\
    \eqref{eq:increasing-def}; and the ``decreasing'' model (red
    curves) is defined in Eq.\ \eqref{eq:decreasing-def}.  Solid lines
    give the $\chieff$ distribution under the assumption that the
    orientations of the spins are isotropic; dashed lines give the
    distribution under the assumption that both objects' spins are
    aligned with the orbital angular momentum.}
\end{figure}

LIGO measures $\chieff$ better than any other spin parameter, but
still with significant uncertainty, so we need to properly incorporate
measurement uncertainty in our analysis; thus our analysis must be
\emph{hierarchical}.  In a hierarchical analysis, we assume that each
event has a true, but unknown, value of the effective spin, drawn from
the population distribution, which may have some parameters $\lambda$;
then the system is observed, represented by the likelihood function,
which results in a distribution for the true effective spin (and all
other paremeters describing the system) consistent with the data.
Combining, the *joint* posterior on each system's $\chieff^i$
parameters and the population parameters $\lambda$ implied by a set of
observations each with data $d^i$, is
\begin{equation}
  p\left( \left\{ \chieff^i \right\}, \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
\end{equation}

The components of this formula are
\begin{itemize}
\item The GW (marginal) likelihood, $p\left(d \mid \chieff\right)$.
  "Marginal" because we are (implicitly) *integrating* over all
  parameters of the signal but $\chieff$.  Note that it is the
  *likelihood*, not the *posterior* that matters for the hierarchical
  analysis; if we are given posterior distributions or posterior
  samples, we need to re-weight to "remove" the prior and obtain the
  likelihood.
\item The population distribution for $\chieff$,
  $p\left( \chieff \mid \lambda \right)$.  This function can be
  parameterised by population-level parameters, $\lambda$.  (In the
  cases discussed above, there are no parameters for the population.)
\item The prior on the population-level parameters, $p(\lambda)$.
\end{itemize}
If we don't care about the individual event $\chieff$ paremetrs, we
can integrate them out, obtaining
\begin{equation}
  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \int \dd \chieff^i \, p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
\end{equation}
If we are given posterior samples of $\chieff^{ij}$ ($i$ labels the
event, $j$ labels the particular posterior sample) drawn from an
analysis using a prior $p\left( \chieff \right)$, then we can
approximate the integral by an re-weighted average of the population
distribution over the samples (here $p\left( \chieff^{ij} \right)$ is
the prior used to produce the posterior samples):
\begin{equation}
  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \frac{1}{N_i} \sum_{j=1}^{N_i} \frac{p\left( \chieff^{ij} \mid \lambda \right)}{p\left( \chieff^{ij} \right)} \right] p\left(\lambda\right).
\end{equation}

In a fully-spinning `LALInference` analysis, our prior for $\chieff$
is given by the combination of a flat prior in the spin magnitudes
with an isotropic prior in angles.

\section{Results From O1}
\label{sec:O1}

We approximate the posteriors on $\chieff$ for the three \ac{GW}
events reported in \cite{O1-BBH} by Gaussians with the same median and
90\% range reported in that paper.  Our approximations are shown in
Figure \ref{fig:O1-posteriors}.

\begin{figure}
  \plotone{../plots/chi-eff-mock-posteriors}
  \caption{\label{fig:O1-posteriors} ``Mock'' posteriors from the O1
    observations.  We approximate the posteriors reported in
    \citet{O1-BBH} with Gaussians with the same median and 90\%
    range.}
\end{figure}

The hierarchical analysis described in Section \ref{sec:formalism}
results in the odds ratios shown in Figure \ref{fig:O1-odds}.  The
best-fitting isotropic model (``decreasing'') is favoured over the
best-fitting aligned model (``decreasing'') by an odds ratio of
$\OOneOddsIsoAligned$, which corresponds to $\OOneSigmaIsoAligned$.

\begin{figure}
  \plotone{../plots/O1-model-selection}
  \caption{Odds ratios among our six models using the approximations
    to the posteriors on $\chieff$ from the O1 observations shown in
    Figure \ref{fig:O1-posteriors}.  The flat, increasing, and
    decreasing spin magnitude distributions (``F'', ``I'', ``D'';
    Eqs.\ \eqref{eq:flat-def}, \eqref{eq:increasing-def},
    \eqref{eq:decreasing-def}) are paired with isotropic and aligned
    angular distributions (``I'', ``A'').  The most-favoured isotropic
    model has a decreasing distribution of spin magnitudes, as does
    the most-favoured aligned model.  The odds ratio between these
    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.}
  \label{fig:O1-odds}
\end{figure}

\section{Predictions for the Future}

See Figure \ref{fig:O2-predictions}.

\begin{figure}
  \plotone{../plots/six-way-O2-model-selection}
  \caption{\label{fig:O2-predictions} Distribution of odds ratios
    predicted with 10 additional observations above the three
    discussed in Section \ref{sec:O1}.  Each panel corresponds to
    additional observations drawn from one of the $\chieff$
    distribution models.  The model from which the additional
    observations are drawn is outlined in red.  The height of the blue
    bar gives the median odds ratio relative to the model from which
    the additional observations are drawn; the green line gives the
    68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
    separate draws from the model distribution.  The closest ratio
    between the most-favoured isotropic model and the most-favoured
    aligned model is $\OTwoOddsIsoAlignedMin$, corresponding to
    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
    distribution; most models result in more than $5\sigma$ preference
    for the correct angular distribution.  Because the three
    observations from Section \ref{sec:O1} are included in each data
    set the ``correct'' model is not necessarily preferred over the
    others, particularly when that model uses the ``increasing''
    magnitude distribution, which is strongly dis-favoured from the O1
    observations alone.}
\end{figure}

\section{Conclusion}

\bibliography{AlignedIsotropicGW}

\end{document}

### 2017-01-21 paper/AlignedIsotropicGW.tex
Dumped a couple of references in intro.

@@
 \newcommand{\diff}[2]{\frac{\dd #1}{\dd #2}}
+\newcommand{\fixme}[1]{\textcolor{red}{FIXME: #1}}
 \newcommand{\pali}{p_\mathrm{ali}}
@@
 This is a reference dump at the moment.
+
+We should be sure to discuss \citet{Rodriguez2016} here; they note the
+small $\chieff$ of GW150914, but only in the context of looking for a
+single system that is \emph{definitely} formed from dynamical
+interactions; obviously, the population provides a more stringent
+origin test.
+
+We find much higher resolving power than \citet{Vitale2015}, probably
+because they focus on measurements of spin \emph{angles}, using a
+mixture model comprising an aligned and isotropic component.
+
+\citet{Miller2015} summarises the current best \ac{EM} constraints on
+the black hole spin distribution; it is consistent with flat but could
+be increasing toward large spins (\fixme{Cole: can you confirm and
+  summarise better?}).
 

### 2017-01-23 paper/AlignedIsotropicGW.tex
Update AlignedIsotropicGW.tex

@@
 
-\citet{Miller2015} summarises the current best \ac{EM} constraints on
-the black hole spin distribution; it is consistent with flat but could
-be increasing toward large spins (\fixme{Cole: can you confirm and
-  summarise better?}).
+In order to perform our analysis we need to select models for the distribution of the dimensionless spin parameters of stellar-mass black holes.
+Data for such model selection is sparse.
+\citet{2015PhR...548....1M} give current estimates of the spin parameters for stellar-mass black holes, obtained using disk reflection and/or disk continuum methods.
+
+Most of the systems studied are low-mass X-ray binaries (i.e., the donor star has a mass $M_{\rm donor}<1~M_\odot$) rather than the high-mass X-ray binaries that are likely to be the progenitors of double black hole binaries.
+In addition, there are substantial systematic errors that can complicate either type of analysis (see \citealt{2015PhR...548....1M} for a discussion).
+Nonetheless, if we take the reported spin magnitudes as representative then we find that there is a preference for high spins; for example, 14 of the 19 systems with reported spins have spin parameters in excess of 0.5.
+The masses and spin parameters of stellar-mass black holes are unlikely to be altered significantly by accetion (low-mass donors simply do not have enough mass and high-mass donors have a very short phase in which they transfer mass; a variant of this long-standing argument was presented by \citealt{1999MNRAS.305..654K}).
+Thus the current spin parameters probably are close to their values upon core collapse.
+We emphasize, however, that the specific processes involved in the production of black hole binaries could alter the spin distribution of those holes: for example, close tidal interactions could spin up the core, or stripping of the envelope could reduce the available angular momentum.
+
+With this in mind, we do our analysis using three straw-man spin distributions for spin parameters from $a/M=0$ to 1: flat ($P(a/M)=1$); rising ($P(a/M)=2a/M$); and falling ($P(a/M)=2(1-a/M)$).
+These are not meant to represent any particular physical model, and indeed neither observations nor population synthesis codes can authoritatively suggest {\it any} particular spin distribution.
+This does, however, allow us to see how sensitive the $\chieff$ distribution is to spin alignment models given uncertainties about the spin magnitudes.
 

### 2017-02-01 paper/AlignedIsotropicGW.tex
Update AlignedIsotropicGW.tex
Added mixture modelling results to git and added words about them to the paper draft

@@
 \author[0000-0003-1540-8562]{Will M. Farr}
+
+\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
+  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+
+\author[0000-0002-6100-537X]{Simon Stevenson}
 
@@
 
-In order to perform our analysis we need to select models for the distribution of the dimensionless spin parameters of stellar-mass black holes.
-Data for such model selection is sparse.
-\citet{2015PhR...548....1M} give current estimates of the spin parameters for stellar-mass black holes, obtained using disk reflection and/or disk continuum methods.
+In order to perform our analysis we need to select models for the 
+distribution of the dimensionless spin parameters of stellar-mass black 
+holes.  Data for such model selection is sparse.  \citet{2015PhR...548....1M} 
+give current estimates of the spin parameters for stellar-mass black holes, 
+obtained using disk reflection and/or disk continuum methods.
 
-Most of the systems studied are low-mass X-ray binaries (i.e., the donor star has a mass $M_{\rm donor}<1~M_\odot$) rather than the high-mass X-ray binaries that are likely to be the progenitors of double black hole binaries.
-In addition, there are substantial systematic errors that can complicate either type of analysis (see \citealt{2015PhR...548....1M} for a discussion).
-Nonetheless, if we take the reported spin magnitudes as representative then we find that there is a preference for high spins; for example, 14 of the 19 systems with reported spins have spin parameters in excess of 0.5.
-The masses and spin parameters of stellar-mass black holes are unlikely to be altered significantly by accetion (low-mass donors simply do not have enough mass and high-mass donors have a very short phase in which they transfer mass; a variant of this long-standing argument was presented by \citealt{1999MNRAS.305..654K}).
-Thus the current spin parameters probably are close to their values upon core collapse.
-We emphasize, however, that the specific processes involved in the production of black hole binaries could alter the spin distribution of those holes: for example, close tidal interactions could spin up the core, or stripping of the envelope could reduce the available angular momentum.
+Most of the systems studied are low-mass X-ray binaries (i.e., the donor star 
+has a mass $M_{\rm donor}<1~M_\odot$) rather than the high-mass X-ray binaries 
+that are likely to be the progenitors of double black hole binaries.  In 
+addition, there are substantial systematic errors that can complicate either 
+type of analysis (see \citealt{2015PhR...548....1M} for a discussion). Nonetheless, 
+if we take the reported spin magnitudes as representative then we find that there 
+is a preference for high spins; for example, 14 of the 19 systems with reported 
+spins have spin parameters in excess of 0.5.  The masses and spin parameters of 
+stellar-mass black holes are unlikely to be altered significantly by accetion 
+(low-mass donors simply do not have enough mass and high-mass donors have a very 
+short phase in which they transfer mass; a variant of this long-standing argument 
+was presented by \citealt{1999MNRAS.305..654K}).  Thus the current spin parameters 
+probably are close to their values upon core collapse.  We emphasize, however, 
+that the specific processes involved in the production of black hole binaries 
+could alter the spin distribution of those holes: for example, close tidal 
+interactions could spin up the core, or stripping of the envelope could reduce the 
+available angular momentum.
 
-With this in mind, we do our analysis using three straw-man spin distributions for spin parameters from $a/M=0$ to 1: flat ($P(a/M)=1$); rising ($P(a/M)=2a/M$); and falling ($P(a/M)=2(1-a/M)$).
-These are not meant to represent any particular physical model, and indeed neither observations nor population synthesis codes can authoritatively suggest {\it any} particular spin distribution.
-This does, however, allow us to see how sensitive the $\chieff$ distribution is to spin alignment models given uncertainties about the spin magnitudes.
+With this in mind, we do our analysis using three straw-man spin distributions for 
+spin parameters from $a/M=0$ to 1: flat ($P(a/M)=1$); rising ($P(a/M)=2a/M$); and 
+falling ($P(a/M)=2(1-a/M)$).  These are not meant to represent any particular 
+physical model, and indeed neither observations nor population synthesis codes 
+can authoritatively suggest {\it any} particular spin distribution.  This does, 
+however, allow us to see how sensitive the $\chieff$ distribution is to spin alignment 
+models given uncertainties about the spin magnitudes.
 
@@
 
+%\begin{figure}
+%  \plotone{../plots/O1-model-selection}
+%  \caption{Odds ratios among our six models using the approximations
+%    to the posteriors on $\chieff$ from the O1 observations shown in
+%    Figure \ref{fig:O1-posteriors}.  The flat, increasing, and
+%    decreasing spin magnitude distributions (``F'', ``I'', ``D'';
+%    Eqs.\ \eqref{eq:flat-def}, \eqref{eq:increasing-def},
+%    \eqref{eq:decreasing-def}) are paired with isotropic and aligned
+%    angular distributions (``I'', ``A'').  The most-favoured isotropic
+%    model has a decreasing distribution of spin magnitudes, as does
+%    the most-favoured aligned model.  The odds ratio between these
+%    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.}
+%  \label{fig:O1-odds}
+%\end{figure}
+
 \begin{figure}
-  \plotone{../plots/O1-model-selection}
-  \caption{Odds ratios among our six models using the approximations
+  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models.png}
+  \caption{Odds ratios among our models using the approximations
     to the posteriors on $\chieff$ from the O1 observations shown in
     Figure \ref{fig:O1-posteriors}.  The flat, increasing, and
-    decreasing spin magnitude distributions (``F'', ``I'', ``D'';
+    decreasing spin magnitude distributions (``$\mathrm{F}$'', ``$\mathrm{I}$'', ``$\mathrm{D}$'';
     Eqs.\ \eqref{eq:flat-def}, \eqref{eq:increasing-def},
-    \eqref{eq:decreasing-def}) are paired with isotropic and aligned
-    angular distributions (``I'', ``A'').  The most-favoured isotropic
-    model has a decreasing distribution of spin magnitudes, as does
-    the most-favoured aligned model.  The odds ratio between these
+    \eqref{eq:decreasing-def}) are paired with isotropic, aligned
+    angular distributions (``$\mathrm{I}$'', ``$\mathrm{A}$''), as well as a mixture model of the two (``$\mathrm{M}$'').
+The most-favoured models have a decreasing distribution of spin magnitudes.
+The odds ratio between these
     models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.}
@@
 \end{figure}
+
+\section{Mixture model}
+
+BBHs can form as the end point of isolated binary evolution \cite{refs}, or dynamically in dense stellar environments \cite{refs}.
+The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.
+
+Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
+In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.
+
+In this section we assume that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).
+
+%
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{../plots/posterior_on_isotropic_fraction.png}
+\caption{\textbf{Fraction of the BBH population coming from an isotropic distribution} The dotted (blue) line shows the flat prior on the fraction of BBHs coming from an isotropic distribution $f_i$.
+The solid (red) line shows the posterior after O1, assuming that all BHs have their spin magnitude drawn from a uniform distribution.
+The dashed line assumes a `increasing' distribution $p(a) = 2a$ for BH spin magnitudes, whilst the dot-dash line assumes a `decreasing' distribution $p(a) = 2(1-a)$.
+We see that regardless of our assumption regarding BH spin magnitudes, there is a preference for a large fraction coming from an isotropic distribution.}
+\label{fig:mixture_fraction_posterior}
+\end{figure}
+%
+
+%-- could also use the cumulative posterior
+%%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.45\textwidth]{../plots/isotropic_fraction_cumulative_posterior.png}
+%\caption{\textbf{Fraction of the BBH population coming from an isotropic distribution} The dotted (blue) line shows the cumulative distribution for the flat prior on the fraction of BBHs coming from an isotropic distribution $f_i$.
+The solid (red) line shows the cumulative distribution for the posterior after O1, assuming that all BHs have their spin magnitude drawn from a uniform distribution.
+The dashed line assumes a `increasing' distribution $p(a) = 2a$ for BH spin magnitudes, whilst the dot-dash line assumes a `decreasing' distribution $p(a) = 2(1-a)$.
+We see that regardless of our assumption regarding BH spin magnitudes, there is a preference for a large fraction coming from an isotropic distribution.}
+%\label{fig:mixture_fraction_posterior}
+%\end{figure}
+%
+
+We fit a mixture model (labelled model 'M') where a fraction $f_i$ ($f_i=\lambda$ in the derivations above) of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
+We assume a flat prior on the fraction $f_i$.
+To test the robustness of our result, we vary the distribution we assume for BH spin magnitude distributions as with the aligned and isotropic models.
+We use the `uniform', `increasing' and `decreasing' distributions introduced in Section X, assuming all BHs have their spin magnitude drawn from the same distribution.
+We calculate the posterior given by EquationX plot the posterior distribution/cumulative posterior in Figure~\ref{fig:mixture_fraction_posterior}.
+We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the `decreasing', `uniform' and `increasing' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
+The lower 90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the prior of 0.1.
+Thus, irrespective of our ignorance of the BH spin magnitude distribution, we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
+
+If we assume that the subpopulation of BBHs with isotropically distributed spin orientations corresponds to a subpopulation of dynamically formed BBHs, our result suggests that the majority of merging BBHs observed by LIGO are formed dynamically, rather than through isolated binary evolution.
+
+The evidence ratios of these models to the isotropic distribution with uniform spin magnitudes are 0.85, 0.39 and 0.19. Thus we cannot rule out a mixture with the current data.
 

### 2017-02-28 paper/AlignedIsotropicGW.tex
Added Ilya's caveats (no good formatting at this point).

@@
 \newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
-\newcommand{\OOneOddsIsoAligned}{\checkme{0.044}}
+\newcommand{\OOneOddsIsoAligned}{\checkme{0.087}}
 \newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
 \newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0012}}
+
+\newcommand{\ilya}[1]{\textcolor{purple}{#1}}
 
@@
 
+\author[0000-0002-6134-8946]{Ilya Mandel}
+
+\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
+  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+
 \author[0000-0002-6254-1617]{Alberto Vecchio}
@@
 
-\author[0000-0002-6134-8946]{Ilya Mandel}
-
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
-  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\author{Friends}
-
-\email{w.farr@bham.ac.uk,miller@astro.umd.edu,av@star.sr.bham.ac.uk,imandel@star.sr.bham.ac.uk}
+\email{w.farr@bham.ac.uk,simons@star.sr.bham.ac.uk,miller@astro.umd.edu,imandel@star.sr.bham.ac.uk,av@star.sr.bham.ac.uk}
 
@@
   isotropic and aligned models, and in most cases better than
-  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).
+  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing data
+  strongly suggest a dynamical formation secnario for the observed
+  systems, and this will be confirmed (or overturned) confidently in
+  the near future by subsequent observations.
 \end{abstract}
@@
 models given uncertainties about the spin magnitudes.
+
+\subsection{Ilya's Caveats}
+
+We have analysed the observed spin distribution for BBH mergers detected by aLIGO during the O1 science run and compared it to a few simple models.
+Here, we discuss observational and modelling constrains on the predicted spin distribution.
+We separate these into a discussion of spin magnitudes and spin directions (spin-orbit misalignment angles) as in the preceding analysis; however, it is worth pointing out that, contrary to our simplified assumptions, the magnitude and direction distributions may well be coupled.
+
+Other than the LIGO observations of binary black holes discussed above, the best constraints on BH spin magnitudes and directions come from BH X-ray binaries(XRBs).
+Observations of spin magnitudes in BH XRBs have been claimed through three different techniques: continuum flux measurements, iron K-$\alpha$ line measurements, and quasi-periodic oscillations (QPOs) \citep[e.g.,][for a review]{MillerMiller:2015}.
+All measurements suffer from both statistical uncertainty and possibly very significant systematics.
+Both continuum flux and iron K-$\alpha$ line measurements indicate that some black holes may have very rapid spins, $\chi>0.9$; however, they also disagree on several spin measurements.
+Meanwhile, QPO measurements find much lower BH spins \ilya{[check numbers and add reference]}.
+
+Although BH spin-orbit alignment is assumed in continuum flux measurements \citep{MillerMiller:2015}, some BH XRBs, including GRO J1655-40 \ilya{and 4U 1543-475?} appear to indicate that the micro quasar jet, presumably aligned with the BH spin axis, is misaligned with the orbit.
+Moreover, initial stellar spins in massive binaries have been observed to be misaligned \ilya{[ref]}, though there are opportunities for realignment during binary evolution, as described below.
+
+Several models for binary black hole formation have been proposed \citep[e.g.,][]{GW150914:astro}, yielding potentially different spin distributions.
+
+The spin directions of binary black holes formed dynamically through interactions in dense stellar environments \ilya{[add refs.
+if not discussed elsewhere]} are expected to be isotropic given the absence of a preferred direction \citep[e.g.,][]{Rodriguez:2016spins} and the persistence of an isotropic distribution through post-Newtonian evolution \citep{Schnittman:2004}.
+
+Chemically homogeneous evolution is a possible BBH formation channel specific to rapidly spinning massive stars aligned by tidal effects \ilya{[add refs.
+if not discussed elsewhere]}.
+However, wind-driven mass loss can spin down even some of these stars to the point that values of $\chi_\textrm{eff}$ as low as $\sim 0.2$ are possible, though alignment should be preserved \citep{Marchant:2017}.
+
+The classical isolated binary evolution channel through the common-envelope (CE) phase \ilya{[add refs.
+if not discussed elsewhere]} yields the following typical formation channel for binary black holes: possible mass transfer from the primary to the secondary; the collapse of the primary into a BH; the formation and ejection of a CE as a consequence of dynamically unstable mass transfer from the secondary to the primary; possible further mass transfer from remaining secondary core to primary; collapse of the secondary into a BH; a merger induced by GW emission \citep[e.g.,][]{Belczynski:2016,Stevenson:2017}.
+Interaction before the collapse of the primary could realign the binary, but the primary’s spin in the wide (separation $\gtrsim 10$ AU) binary can easily be misaligned by even moderate \citep{Mandel:2015kicks} natal kicks.
+Realignment of the primary through subsequent mass accretion is unlikely given the small amount of accreted mass, except perhaps during hyper-critical accretion during CE \ilya{[add ref.
+to Linden?]}; realignment through dissipation in the possibly warped accretion disk \citep{BardeenPatterson} is possible, but calculations suggest it is unlikely for stellar-mass binaries \citep{KLOP}.
+
+The secondary should tidally align with the orbit before the onset of mass transfer \citep[e.g.,][]{Kalogera:2000,Gerosa:2013}; however, most of its angular momentum would be lost to the envelope, which will be ejected during the CE phase.
+The remnant core may or may not be efficiently re-aligned, depending on the separation between the secondary’s core and the BH immediately after the CE is ejected \citep{Kushnir:2017}.
+The secondary’s natal kick will be small relative to the orbital velocity of the close post-CE binary, so any further kick-induced misalignment will be moderate.
+
+The analysis above assumed the conservation of angular momentum during core collapse.
+However, we know that this may not always be the case.
+For example, the magnitude of the angular momentum cannot possibly be conserved during the collapse to a BH if the pre-collapse dimensionless spin exceeds $1$.
+The direction of the angular momentum may also change; for example, the misalignment of the secondary pulsar in the double pulsar \ilya{J0737} by $\sim 135^{\circ}$ relative to the orbit in the apparent absence of a kick indicated that, at least for slowly spinning objects, a spin tilt during a supernova may be possible.
+The dependence of the magnitude and direction of the spin on its pre-collapse values depends on the stellar structure (i.e., the level of angular momentum coupling between the collapsing and ejected material), the supernova explosion (if any), and the details of the post-collapse fallback process \ilya{[Check if Levin and Matzner are done]}.
 

### 2017-03-23 paper/AlignedIsotropicGW.tex
Massive commit, rough draft done, needs some polishing and sections added to appendix.

--- 2017-02-01 paper/AlignedIsotropicGW.tex
@@
 
-\documentclass[modern]{aastex61}
+\documentclass[modern,linenumbers]{aastex61}
 
@@
 \newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
-\newcommand{\OOneOddsIsoAligned}{\checkme{0.044}}
+\newcommand{\OOneOddsIsoAligned}{\checkme{0.087}}
 \newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
 \newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0012}}
+
+\newcommand{\ilya}[1]{\textcolor{purple}{#1}}
+\newcommand{\will}[1]{\textcolor{cyan}{#1}}
 
@@
 \acrodef{GW}{gravitational wave}
+\acrodef{PE}{parameter estimation}
 
 \title{Distinguishing Between Spin-Aligned and Isotropic Binary Black
-  Hole Populations in Advanced LIGO}
+  Hole Populations Using Gravitational Wave Observations}
 
@@
 
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
-  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
+  School of Physics and Astronomy, University of Birmingham,
+  Birmingham, B15 2TT, United Kingdom}
 
@@
 
+\author[0000-0002-6134-8946]{Ilya Mandel}
+
+\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
+  School of Physics and Astronomy, University of Birmingham,
+  Birmingham, B15 2TT, United Kingdom}
+
 \author[0000-0002-6254-1617]{Alberto Vecchio}
 
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
-  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
+\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
+  School of Physics and Astronomy, University of Birmingham,
+  Birmingham, B15 2TT, United Kingdom}
 
-\author[0000-0002-6134-8946]{Ilya Mandel}
-
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy,
-  University of Birmingham, Birmingham, B15 2TT, United Kingdom}
-
-\author{Friends}
-
-\email{w.farr@bham.ac.uk,miller@astro.umd.edu,av@star.sr.bham.ac.uk,imandel@star.sr.bham.ac.uk}
+\email{w.farr@bham.ac.uk,simons@star.sr.bham.ac.uk,miller@astro.umd.edu,imandel@star.sr.bham.ac.uk,av@star.sr.bham.ac.uk}
 
 \begin{abstract}
-  The first direct detections of \acp{GW} from merging \acp{BBH} in
-  the last year open a unique window into the formation environment of
-  these systems.  One promising signature of the formation environment
-  is the angular distribution of the \ac{BH} spins; systems formed
-  through dynamical interacions among already-compact objects are
-  expected to have isotropic spin orientations while binaries from
-  pairs of stars born together are more likely to have spins aligned
-  with the binary orbit as a consequence of their joint evolution
-  toward a \ac{BBH} system.  By considering existing \ac{GW}
-  measurements of $\chieff$, the best-measured combination of spin
-  parameters, in the three likely binary black hole detections
+  The first direct detections of \acp{GW} from merging \acp{BBH} open
+  a unique window into the formation environment of massive stars and
+  their compact remnants.  One promising signature of the formation
+  environment is the angular distribution of the \ac{BH} spins;
+  systems formed through dynamical interacions among already-compact
+  objects are expected to have isotropic spin orientations while
+  binaries formed from pairs of stars born together are more likely to
+  have spins aligned with the binary orbit as a consequence of their
+  joint evolution toward a \ac{BBH} system.  By considering existing
+  \ac{GW} measurements of $\chieff$, the best-measured combination of
+  spin parameters, in the three likely binary black hole detections
   GW150914, LVT151012, and GW151226, we show that the data already
@@
   suite of models for the spin distribution.  By considering the
-  effect of an additional 10 detections (a possible outcome of the
-  ongoing second observing run of Advanced LIGO) drawn from the
-  various models in the suite we show that such an agumented data set
-  would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
-  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
-  isotropic and aligned models, and in most cases better than
-  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).
+  effect of an additional 10 detections drawn from the various models
+  in the suite we show that such an agumented data set would enable at
+  least a $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
+  odds ratio) distinction between the isotropic and aligned models,
+  and in most cases better than $5\sigma$ ($2.9 \times 10^{-7}$ odds
+  ratio).  The existing data strongly suggest a dynamical formation
+  secnario for the observed systems, and this will be confirmed (or
+  overturned) confidently in the near future by subsequent
+  observations.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{\ac{GW} Spin Measurements and Model Selection}
+\label{sec:O1}
 
-This is a reference dump at the moment.
+Following the detection of a merging \ac{BBH} system, \ac{PE} tools
+\citep{2015PhRvD..91d2003V} compare model gravitational waveforms
+\citep{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
+against the observed data to obtain a posterior distribution on the
+parameter space that describes the compact binary source.  The spin
+parameter with the largest effect on waveforms, and a correspondingly
+tight constraint from the data, is a mass-weighted, aligned,
+``effective spin,''
+\begin{equation}
+  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
+  \right) \cdot \frac{\vec{L}}{\hat{L}} = \frac{1}{M} \left( m_1 a_1^z + m_2 a_2^z \right),
+\end{equation}
+where the $\hat{z}$ axis is aligned with the orbital angular momentum
+of the binary, $m_{1,2}$ are the masses of the more-massive (1) and
+less-massive (2) components, $M = m_1 + m_2$ is the total mass,
+$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
+holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
+and $a_{1,2}$ are the corresponding dimensionless spin parameters
+\citep{2016PhRvL.116x1102A}.
 
-We should be sure to discuss \citet{Rodriguez2016} here; they note the
-small $\chieff$ of GW150914, but only in the context of looking for a
-single system that is \emph{definitely} formed from dynamical
-interactions; obviously, the population provides a more stringent
-origin test.
+Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
+inferred on $\chieff$ for the three likely \ac{GW} detections
+GW150914, GW151226 and LVT151012 from \citet{O1-BBH}.  Because samples
+drawn from the posterior on $\chieff$ are not publicly released at
+this time, we have approximated the posterior as a Gaussian
+distribution with the same mean and 90\% credible interval as quoted
+in \citet{O1-BBH}.  There is essentially no posterior support for
+$\chieff \gtrsim 0.5$ as observed in the majority of the
+electromagnetically-detected stellar-mass black hole population.
 
-We find much higher resolving power than \citet{Vitale2015}, probably
-because they focus on measurements of spin \emph{angles}, using a
-mixture model comprising an aligned and isotropic component.
+\begin{figure}
+  \plotone{../plots/chi-eff-mock-posteriors}
+  \caption{\label{fig:O1-posteriors} Approximate posteriors on
+    $\chieff$ from the LIGO O1 observations in \citet{O1-BBH}.  We
+    approximate the posteriors reported in \citet{O1-BBH} using
+    Gaussians with the same median and 90\% credible interval as
+    reported in \citet{O1-BBH}.  It is notable that none of the
+    $\chieff$ posteriors extends to the high spins in
+    electromagnetically-detected stellar-mass black hole systems
+    \citep{2015PhR...548....1M}.}
+\end{figure}
 
-In order to perform our analysis we need to select models for the 
-distribution of the dimensionless spin parameters of stellar-mass black 
-holes.  Data for such model selection is sparse.  \citet{2015PhR...548....1M} 
-give current estimates of the spin parameters for stellar-mass black holes, 
-obtained using disk reflection and/or disk continuum methods.
+Small values of $\chieff$ can result from either intrinsically small
+spins or larger spins whose direction is mis-aligned with the orbital
+plane of the binary (i.e.\ spin vectors with with small
+$z$-components).  Mis-alignment is capable of producing
+\emph{negative} values of $\chieff$, however, while small, aligned
+spins will always have $\chieff \geq 0$.  To quantify the degree of
+support for these two alternate explanations of small $\chieff$ values
+in the merging \ac{BBH} population, we compared the Bayesian evidence
+for various simple models of the spin population using the \ac{GW}
+data set.
 
-Most of the systems studied are low-mass X-ray binaries (i.e., the donor star 
-has a mass $M_{\rm donor}<1~M_\odot$) rather than the high-mass X-ray binaries 
-that are likely to be the progenitors of double black hole binaries.  In 
-addition, there are substantial systematic errors that can complicate either 
-type of analysis (see \citealt{2015PhR...548....1M} for a discussion). Nonetheless, 
-if we take the reported spin magnitudes as representative then we find that there 
-is a preference for high spins; for example, 14 of the 19 systems with reported 
-spins have spin parameters in excess of 0.5.  The masses and spin parameters of 
-stellar-mass black holes are unlikely to be altered significantly by accetion 
-(low-mass donors simply do not have enough mass and high-mass donors have a very 
-short phase in which they transfer mass; a variant of this long-standing argument 
-was presented by \citealt{1999MNRAS.305..654K}).  Thus the current spin parameters 
-probably are close to their values upon core collapse.  We emphasize, however, 
-that the specific processes involved in the production of black hole binaries 
-could alter the spin distribution of those holes: for example, close tidal 
-interactions could spin up the core, or stripping of the envelope could reduce the 
-available angular momentum.
-
-With this in mind, we do our analysis using three straw-man spin distributions for 
-spin parameters from $a/M=0$ to 1: flat ($P(a/M)=1$); rising ($P(a/M)=2a/M$); and 
-falling ($P(a/M)=2(1-a/M)$).  These are not meant to represent any particular 
-physical model, and indeed neither observations nor population synthesis codes 
-can authoritatively suggest {\it any} particular spin distribution.  This does, 
-however, allow us to see how sensitive the $\chieff$ distribution is to spin alignment 
-models given uncertainties about the spin magnitudes.
-
-\section{Model Selection Formalism}
-\label{sec:formalism}
-
-This is currently pasted from the iPython notebook:
-
-Cole had a great idea, the thrust of which is the following: assume
-that you have some distribution for the spin magnitudes of black holes
-(and it is common to the more- and less-massive BH in any binary).
-The distribution of $\chieff$ that is implied by this magnitude
-distribution differs depending on whether you assume that the angular
-distribution is isotropic or aligned.  In general, isotropic angular
-distributions concentrate more of the $\chieff$ distribution near
-zero, since that is where most of the angular phase space is.  Here we
-derive the distributions for $\chieff$ under various reasonable, yet
-straw-man assumptions.
-
-First, $\chieff$ is defined by
+Each of our models for the merging \ac{BBH} spin population assumes
+that the merging black holes are of equal mass (this is marginally
+consistent with the three observations \citep{O1-BBH}, and the
+$\chieff$ distribution is not particularly sensitive to the mass ratio
+between the merging objects---see Section \ref{sec:mass-ratio}).  We
+assume that the population spin distribution factorises into a
+distribution for the spin magnitude, $a$, and a distribution for the
+spin angles.  Finally, we assume that the distribution of spins is
+common to each component in a merging binary.  Choosing one of three
+magnitude distributions
 \begin{equation}
-\chieff = \frac{c}{GM} \left( \frac{s_1^z}{m_1} + \frac{s_2^z}{m_2}
-\right) = \frac{1}{M} \left( m_1 a_1^z + m_2 a_2^z \right)
+  \label{eq:magnitude-dists}
+  p(a) = \begin{cases}
+    2\left(1-a\right) & \textnormal{``decreasing''} \\
+    1 & \textnormal{``flat''} \\
+    2\left(a-1\right) & \textnormal{``increasing''}
+  \end{cases},
 \end{equation}
-where the $\hat{z}$ axis is aligned with the orbital angular momentum,
-$s_{1,2}$ are the spin angular momentum of the components of the
-binary, and $a_{1,2}$ are the spin parameters of the black holes in
-the binary.
-
-First we derive the distribution for $a^z$ when we know the
-distribution for $a$ and the angular distribution is (a) isotropic and
-(b) aligned.  First isotropic: Then $a^z$ is equal to the product of
-$a$ with a uniform random variable on $[-1,1]$ ($\iota = \cos \theta$,
-with $\theta$ the colatitude on a sphere).  Therefore
-\begin{equation}
-  p\left( a^z \right) = \int_{-1}^{1} \dd \iota \, \int_0^1 \dd a \, \delta\left( a^z - a \iota \right) \frac{p(a)}{2} = \int_{-1}^{1} \dd \iota \, \int_0^1 \dd a \, \frac{\delta\left( a - \frac{a^z}{\iota} \right)}{\left| \iota \right|} \frac{p(a)}{2}.
-\end{equation}
-Now, noting that the distribution of $a^z$ is an even function, assume
-that $a^z > 0$; since $0 \leq a \leq 1$, we must have $\iota \geq a^z$
-if the $\delta$-function is to be "activated," so
-\begin{equation}
-p\left( a^z \right) = \int_{a^z}^1 \dd \iota \, \frac{p\left( \frac{a^z}{\iota} \right)}{2 \iota}.
-\end{equation}
-The even assumption implies 
-\begin{equation}
-p\left( a^z \right) = \int_{\left| a^z \right|}^1 \dd \iota \,
-\frac{p\left( \frac{\left|a^z\right|}{\iota} \right)}{2 \iota}
-\end{equation}
-When the angles are aligned, we have
-\begin{equation}
-p\left( a^z \right) = \begin{cases}
-p_a\left(a^z \right) & a^z > 0\\
-0 & \mathrm{otherwise}
-\end{cases}.
-\end{equation}
-In other words, the distribution of $a^z$ follows $a$ when $a^z$ is
-positive, and is zero otherwise.
-
-For example, when the magnitude distribution is flat, $p(a) = 1$, the
-isotropic angle assumption implies that
-\begin{equation}
-\piso\left( a^z \right) = -\frac{\log \left| a^z \right|}{2},
-\end{equation}
-while the aligned assumption implies that
-\begin{equation}
-\pali\left( a^z \right) = \begin{cases}
-1 & a^z > 0 \\
-0 & a^z < 0
-\end{cases}
-\end{equation}
-
-If the mass ratio, $q$, (recall: $0 \leq q \leq 1$) is zero, this is
-all we need; if not, then
-\begin{equation}
-\chieff = \frac{1}{1+q}\left(a_1^z + q a_2^z\right),
-\end{equation}
-and the distribution is given by a convolution of the distributions
-for $a^z$:
-\begin{equation}
-p\left(\chieff \right) = \int_{-1}^1 \dd a_1^z \, \int_{-1}^1 \dd a_2^z \, \delta\left( \chieff - \frac{1}{(1+q)} \left[a_1^z + q a_2^z \right]\right) p\left( a_1^z\right) p\left(a_2^z\right),
-\end{equation}
-which becomes, after some manipulation,
-\begin{equation}
-p\left(\chieff \right) = \int_{-1}^1 \dd a_1^z \, \int_{-1}^1 \dd a_2^z \, \delta\left( a_1^z + q a_2^z - (1+q)\chieff \right) (1+q) p\left( a_1^z\right) p\left(a_2^z\right),
-\end{equation}
-and more manipulation:
-\begin{equation}
-p\left(\chieff \right) = \int_{\max\left(-1, \frac{(1+q)\chieff - 1}{q} \right)}^{\min\left(1, \frac{(1+q)\chieff+1}{q}\right)} \dd a_2^z \, (1+q) p\left(a_2^z\right) p\left((1+q)\chieff - q a_2^z \right).
-\end{equation}
-
-Here we will consider three models for the spin magnitude distribution
-of the two \ac{BH} spins.  The ``flat'' model assumes that magnitudes
-are drawn uniformly in $[0, 1]$:
-\begin{equation}
-  \label{eq:flat-def}
-  p(a) = 1.
-\end{equation}
-This model is maximally uncertain about the spin magnitude.  The
-increasing distribution is meant to be more consistent with the
-existing \ac{EM} observations of \acp{BH} in X-ray binaries
-\citep{Miller2015}
-\begin{equation}
-  \label{eq:increasing-def}
-  p(a) = 2a.
-\end{equation}
-The ``decreasing'' distribution is meant to be more consistent with
-the observed low-spins in the existing \ac{GW} detections
-\citep{O1-BBH}, with
-\begin{equation}
-  \label{eq:decreasing-def}
-  p(a) = 2(1-a).
-\end{equation}
-
-This is not going to be analytic, even if we specialise to $q = 1$
-(equal mass); though a good check of the math is that it works out to
-the above single-spin distribution when $q=0$.  We should just
-simulate the distribution empirically.  The result appears in Figure
-\ref{fig:chieff-distribution-models}
+and pairing with either an isotropic angular distribution or a
+distribution that generates perfect alignment with the positive $z$
+axis yields six different models for the $\chieff$ distribution.
+These models are shown in Figure \ref{fig:chieff-distribution-models}.
 
@@
     distribution of $\chieff$ considered in this paper.  In all models
-    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$.  The
-    ``flat'' model (blue curves) for the spin magnitudes is defined in
-    Eq.\ \eqref{eq:flat-def}; the ``increasing'' model (green curves)
-    for the spin magnitudes is defined in Eq.\
-    \eqref{eq:increasing-def}; and the ``decreasing'' model (red
-    curves) is defined in Eq.\ \eqref{eq:decreasing-def}.  Solid lines
-    give the $\chieff$ distribution under the assumption that the
-    orientations of the spins are isotropic; dashed lines give the
-    distribution under the assumption that both objects' spins are
-    aligned with the orbital angular momentum.}
+    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$ and
+    that the same distribution of spin vectors obtains for each
+    component.  The ``flat'' (blue lines), ``increasing,'' (green
+    lines), and ``decreasing'' (red lines) magnitude distributions are
+    defined in Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the
+    $\chieff$ distribution under the assumption that the orientations
+    of the spins are isotropic; dashed lines give the distribution
+    under the assumption that both objects' spins are aligned with the
+    orbital angular momentum.  The isotropic distributions are readily
+    distinguished from the aligned by the production of negative
+    $\chieff$ values, while the distinction between the three models
+    for the spin magnitude distribution is less sharp.}
 \end{figure}
 
-LIGO measures $\chieff$ better than any other spin parameter, but
-still with significant uncertainty, so we need to properly incorporate
-measurement uncertainty in our analysis; thus our analysis must be
-\emph{hierarchical}.  In a hierarchical analysis, we assume that each
-event has a true, but unknown, value of the effective spin, drawn from
-the population distribution, which may have some parameters $\lambda$;
-then the system is observed, represented by the likelihood function,
-which results in a distribution for the true effective spin (and all
-other paremeters describing the system) consistent with the data.
-Combining, the *joint* posterior on each system's $\chieff^i$
-parameters and the population parameters $\lambda$ implied by a set of
-observations each with data $d^i$, is
-\begin{equation}
-  p\left( \left\{ \chieff^i \right\}, \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
-\end{equation}
-
-The components of this formula are
-\begin{itemize}
-\item The GW (marginal) likelihood, $p\left(d \mid \chieff\right)$.
-  "Marginal" because we are (implicitly) *integrating* over all
-  parameters of the signal but $\chieff$.  Note that it is the
-  *likelihood*, not the *posterior* that matters for the hierarchical
-  analysis; if we are given posterior distributions or posterior
-  samples, we need to re-weight to "remove" the prior and obtain the
-  likelihood.
-\item The population distribution for $\chieff$,
-  $p\left( \chieff \mid \lambda \right)$.  This function can be
-  parameterised by population-level parameters, $\lambda$.  (In the
-  cases discussed above, there are no parameters for the population.)
-\item The prior on the population-level parameters, $p(\lambda)$.
-\end{itemize}
-If we don't care about the individual event $\chieff$ paremetrs, we
-can integrate them out, obtaining
-\begin{equation}
-  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \int \dd \chieff^i \, p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
-\end{equation}
-If we are given posterior samples of $\chieff^{ij}$ ($i$ labels the
-event, $j$ labels the particular posterior sample) drawn from an
-analysis using a prior $p\left( \chieff \right)$, then we can
-approximate the integral by an re-weighted average of the population
-distribution over the samples (here $p\left( \chieff^{ij} \right)$ is
-the prior used to produce the posterior samples):
-\begin{equation}
-  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \frac{1}{N_i} \sum_{j=1}^{N_i} \frac{p\left( \chieff^{ij} \mid \lambda \right)}{p\left( \chieff^{ij} \right)} \right] p\left(\lambda\right).
-\end{equation}
-
-In a fully-spinning `LALInference` analysis, our prior for $\chieff$
-is given by the combination of a flat prior in the spin magnitudes
-with an isotropic prior in angles.
-
-\section{Results From O1}
-\label{sec:O1}
-
-We approximate the posteriors on $\chieff$ for the three \ac{GW}
-events reported in \cite{O1-BBH} by Gaussians with the same median and
-90\% range reported in that paper.  Our approximations are shown in
-Figure \ref{fig:O1-posteriors}.
-
-\begin{figure}
-  \plotone{../plots/chi-eff-mock-posteriors}
-  \caption{\label{fig:O1-posteriors} ``Mock'' posteriors from the O1
-    observations.  We approximate the posteriors reported in
-    \citet{O1-BBH} with Gaussians with the same median and 90\%
-    range.}
-\end{figure}
-
-The hierarchical analysis described in Section \ref{sec:formalism}
-results in the odds ratios shown in Figure \ref{fig:O1-odds}.  The
-best-fitting isotropic model (``decreasing'') is favoured over the
-best-fitting aligned model (``decreasing'') by an odds ratio of
-$\OOneOddsIsoAligned$, which corresponds to $\OOneSigmaIsoAligned$.
-
-%\begin{figure}
-%  \plotone{../plots/O1-model-selection}
-%  \caption{Odds ratios among our six models using the approximations
-%    to the posteriors on $\chieff$ from the O1 observations shown in
-%    Figure \ref{fig:O1-posteriors}.  The flat, increasing, and
-%    decreasing spin magnitude distributions (``F'', ``I'', ``D'';
-%    Eqs.\ \eqref{eq:flat-def}, \eqref{eq:increasing-def},
-%    \eqref{eq:decreasing-def}) are paired with isotropic and aligned
-%    angular distributions (``I'', ``A'').  The most-favoured isotropic
-%    model has a decreasing distribution of spin magnitudes, as does
-%    the most-favoured aligned model.  The odds ratio between these
-%    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.}
-%  \label{fig:O1-odds}
-%\end{figure}
+We fit hierarchical models of the three LIGO O1 observations using
+these six different, zero-parameter population distributions (see
+Section \ref{sec:hierarchical}).  We also fit three mixture models for
+the population, where the spin magnitude distribution is fixed but the
+angular distribution is a weighted sum of the isotropic and aligned
+distribution.  The evidence, or marginal likelihood, for each of the
+models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
+distributions, the mixture models' posterior on the mixing fraction
+peaked at 100\% isotropic, explaining why the zero-parameter,
+pure-isotropic models are preferred over the single-parameter mixture
+models for every magnitude distribution with this data set.  Not
+surprisingly, given the small $\chieff$ values in the three detected
+systems, the most-favoured model among those with an isotropic angular
+distribution has the ``decreasing'' magnitude distribution; the most
+favoured model among those with an aligned distribution also has the
+``decreasing'' magnitude distribution.  The odds ratio between the
+aligned and isotropic models is $\OOneOddsIsoAligned$, or
+$\OOneSigmaIsoAligned\sigma$; thus the data strongly favour isotropic
+spins among our suite of models.
 
@@
   \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models.png}
-  \caption{Odds ratios among our models using the approximations
-    to the posteriors on $\chieff$ from the O1 observations shown in
-    Figure \ref{fig:O1-posteriors}.  The flat, increasing, and
-    decreasing spin magnitude distributions (``$\mathrm{F}$'', ``$\mathrm{I}$'', ``$\mathrm{D}$'';
-    Eqs.\ \eqref{eq:flat-def}, \eqref{eq:increasing-def},
-    \eqref{eq:decreasing-def}) are paired with isotropic, aligned
-    angular distributions (``$\mathrm{I}$'', ``$\mathrm{A}$''), as well as a mixture model of the two (``$\mathrm{M}$'').
-The most-favoured models have a decreasing distribution of spin magnitudes.
-The odds ratio between these
-    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.}
+  \caption{Odds ratios among our models using the approximations to
+    the posteriors on $\chieff$ from the O1 observations shown in
+    Figure \ref{fig:O1-posteriors}.  The flat (``F''), increasing
+    (``I''), and decreasing (``D'') spin magnitude distributions (see
+    Eq.\ \eqref{eq:magnitude-dists}) are paired with isotropic,
+    aligned angular distributions (``$\mathrm{I}$'',
+    ``$\mathrm{A}$''), as well as a mixture model of the two
+    (``$\mathrm{M}$'').  The most-favoured models have a decreasing
+    distribution of spin magnitudes.  The odds ratio between these
+    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.
+    For all magnitude distributions the pure-isotropic models are
+    preferred over the mixture models; correspondingly, the posterior
+    on the mixture fraction peaks at 100\% isotropic.}
   \label{fig:O1-odds}
 \end{figure}
+
+\subsection{Future Spin Measurements}
+\label{subsec:future}
+
+Estimates of the rate of \ac{BBH} coalescences give a reasonable
+chance of 10 additional \ac{BBH} detections in the near future
+\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming an
+additional 10 detections drawn from each of our six zero-parameter
+models for the spin distribution, we find the odds ratios shown in
+Figure \ref{fig:O2-predictions}.  We find that most scenarios with an
+additional 10 detections allow to distinguish the correct angular
+distribution with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
+credibility; and in the worst case the distinction is
+$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
+
+\begin{figure}
+  \plotone{../plots/six-way-O2-model-selection}
+  \caption{\label{fig:O2-predictions} Distribution of odds ratios
+    predicted with 10 additional observations above the three
+    discussed in Section \ref{sec:O1}.  Each panel corresponds to
+    additional observations drawn from one of the $\chieff$
+    distribution models.  The model from which the additional
+    observations are drawn is outlined in red.  The height of the blue
+    bar gives the median odds ratio relative to the model from which
+    the additional observations are drawn; the green line gives the
+    68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
+    separate draws from the model distribution.  The closest ratio
+    between the most-favoured isotropic model and the most-favoured
+    aligned model is $\OTwoOddsIsoAlignedMin$, corresponding to
+    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
+    distribution; most models result in more than $5\sigma$ preference
+    for the correct angular distribution.  Because the three
+    observations from Section \ref{sec:O1} are included in each data
+    set the ``correct'' model is not necessarily preferred over the
+    others, particularly when that model uses the ``increasing''
+    magnitude distribution, which is strongly dis-favoured from the O1
+    observations alone.}
+\end{figure}
+
+\section{Discussion}
+
+Most of our resolving power for the spin angular distribution is a
+result of the fact that our ``aligned'' models cannot produce
+$\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
+spins are intrinsically very small, with $a \lesssim 0.1$ always, then
+it is no longer possible to resolve the negative spin with a small
+number of observations.  As noted below, however, spins observed in
+X-ray binaries are generally large, and most models of isolated binary
+evolution do not produce systems with spins uniformly bounded to small
+values.
+
+\citet{2017CQGra..34cLT01V} also studied the possibility of
+distinguishing aligned and isotropic angular distributions of \ac{BBH}
+spins, but concluded that several hundred sources would be required to
+adequately separate the models.  The ``aligned'' population in that
+study permitted spins nearly parallel or anti-parallel to the orbital
+axis, i.e.\ $\hat{S}\cdot \hat{L} \simeq \pm 1$, in contrast to our
+aligned models.  Anti-parallel spins are strongly disfavoured in
+models of isolated binary evolution.  \ilya{This would be a good place
+  for some review articles about pop.\ synth.}
+
+In order to perform our analysis we need to select models for the
+distribution of the dimensionless spin parameters of stellar-mass
+black holes.  Data for such model selection is sparse.
+\citet{2015PhR...548....1M} give current estimates of the spin
+parameters for stellar-mass black holes, obtained using disk
+reflection and/or disk continuum methods.
+
+Most of the systems studied are low-mass X-ray binaries (i.e., the
+donor star has a mass $M_{\rm donor}<1~M_\odot$) rather than the
+high-mass X-ray binaries that are likely to be the progenitors of
+double black hole binaries.  In addition, there are substantial
+systematic errors that can complicate either type of analysis (see
+\citealt{2015PhR...548....1M} for a discussion). Nonetheless, if we
+take the reported spin magnitudes as representative then we find that
+there is a preference for high spins; for example, 14 of the 19
+systems with reported spins have spin parameters in excess of 0.5.
+The masses and spin parameters of stellar-mass black holes are
+unlikely to be altered significantly by accretion (low-mass donors
+simply do not have enough mass and high-mass donors have a very short
+phase in which they transfer mass; a variant of this long-standing
+argument was presented by \citet{1999MNRAS.305..654K}).  Thus the
+current spin parameters probably are close to their values upon core
+collapse.  We emphasize, however, that the specific processes involved
+in the production of black hole binaries could alter the spin
+distribution of those holes: for example, close tidal interactions
+could spin up the core, or stripping of the envelope could reduce the
+available angular momentum.
+
+With this in mind, we performed our analysis using three straw-man
+spin distributions for spin parameters from $a/M=0$ to 1: flat
+($P(a/M)=1$); rising ($P(a/M)=2a/M$); and falling ($P(a/M)=2(1-a/M)$).
+These are not meant to represent any particular physical model, and
+indeed neither observations nor population synthesis codes can
+authoritatively suggest {\it any} particular spin distribution.  This
+does, however, allow us to see how sensitive the $\chieff$
+distribution is to spin alignment models given uncertainties about the
+spin magnitudes.
+
+Although BH spin-orbit alignment is assumed in continuum flux
+measurements \citep{2015PhR...548....1M}, some BH XRBs, including GRO
+J1655-40 \ilya{and 4U 1543-475?} appear to indicate that the micro
+quasar jet, presumably aligned with the BH spin axis, is misaligned
+with the orbit.  Moreover, initial stellar spins in massive binaries
+have been observed to be misaligned \ilya{[ref]}, though there are
+opportunities for realignment during binary evolution, as described
+below.
+
+There are a few modelling constrains on the predicted spin
+distribution.  We separate these into a discussion of spin magnitudes
+and spin directions (spin-orbit misalignment angles) as in the
+preceding analysis; however, it is worth pointing out that, contrary
+to our simplified assumptions, the magnitude and direction
+distributions may well be coupled.
+
+Several models for binary black hole formation have been proposed
+\citep[e.g.,][]{2016ApJ...818L..22A}, yielding potentially different spin
+distributions.
+
+The spin directions of binary black holes formed dynamically through
+interactions in dense stellar environments \ilya{[add refs. if not
+  discussed elsewhere]} are expected to be isotropic given the absence
+of a preferred direction \citep[e.g.,][]{2016ApJ...832L...2R} and the
+persistence of an isotropic distribution through post-Newtonian
+evolution \citep{2004PhRvD..70l4020S}.
+
+Chemically homogeneous evolution is a possible BBH formation channel
+specific to rapidly spinning massive stars aligned by tidal effects
+\ilya{[add refs. if not discussed elsewhere]}.  However, wind-driven
+mass loss can spin down even some of these stars to the point that
+values of $\chi_\textrm{eff}$ as low as $\sim 0.2$ are possible,
+though alignment should be preserved \citep{2016AnA...588A..50M}.
+
+The classical isolated binary evolution channel through the
+common-envelope (CE) phase \ilya{[add refs. if not discussed
+  elsewhere]} yields the following typical formation channel for
+binary black holes: possible mass transfer from the primary to the
+secondary; the collapse of the primary into a BH; the formation and
+ejection of a CE as a consequence of dynamically unstable mass
+transfer from the secondary to the primary; possible further mass
+transfer from remaining secondary core to primary; collapse of the
+secondary into a BH; a merger induced by GW emission
+\citep[e.g.,][]{2016Natur.534..512B,Stevenson:2017}.  Interaction
+before the collapse of the primary could realign the binary, but the
+primary’s spin in the wide (separation $\gtrsim 10$ AU) binary can
+easily be misaligned by even moderate \citep{2015MNRAS.450L..85M}
+natal kicks.  Realignment of the primary through subsequent mass
+accretion is unlikely given the small amount of accreted mass, except
+perhaps during hyper-critical accretion during CE \ilya{[add ref. to
+  Linden?]}; realignment through dissipation in the possibly warped
+accretion disk \citep{1975ApJ...195L..65B} is possible, but calculations
+suggest it is unlikely for stellar-mass binaries \citep{KLOP}.
+
+The secondary should tidally align with the orbit before the onset of
+mass transfer
+\citep[e.g.,][]{2000ApJ...541..319K,2013PhRvD..87j4028G}; however,
+most of its angular momentum would be lost to the envelope, which will
+be ejected during the CE phase.  The remnant core may or may not be
+efficiently re-aligned, depending on the separation between the
+secondary’s core and the BH immediately after the CE is ejected
+\citep{2016MNRAS.462..844K}.  The secondary’s natal kick will be small
+relative to the orbital velocity of the close post-CE binary, so any
+further kick-induced misalignment will be moderate
+\citep{2016ApJ...832L...2R}.
+
+The analysis above assumed the conservation of angular momentum during
+core collapse. However, we know that this may not always be the case.
+For example, the magnitude of the angular momentum cannot possibly be
+conserved during the collapse to a BH if the pre-collapse
+dimensionless spin exceeds $1$.  The direction of the angular momentum
+may also change; for example, the misalignment of the secondary pulsar
+in the double pulsar \ilya{J0737} by $\sim 135^{\circ}$ relative to
+the orbit in the apparent absence of a kick indicated that, at least
+for slowly spinning objects, a spin tilt during a supernova may be
+possible \citep{2011ApJ...742...81F}.  The dependence of the magnitude
+and direction of the spin on its pre-collapse values depends on the
+stellar structure (i.e., the level of angular momentum coupling
+between the collapsing and ejected material), the supernova explosion
+(if any), and the details of the post-collapse fallback process
+\ilya{[Check if Levin and Matzner are done]}.
+
+In short, the angular distribution of spins in \acp{BBH} formed
+through isolated binary evolution is uncertain; nevertheless, it is
+more probable that spins are aligned and non-zero after evolution
+through this channel than not.
+
+\appendix
 
@@
 
-\section{Predictions for the Future}
+\section{Hierarchical Modelling} 
+\label{sec:hierarchical}
 
-See Figure \ref{fig:O2-predictions}.
+LIGO measures $\chieff$ better than any other spin parameter, but
+still with significant uncertainty, so we need to properly incorporate
+measurement uncertainty in our analysis; thus our analysis must be
+\emph{hierarchical} \citep{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
+In a hierarchical analysis, we assume that each event has a true, but
+unknown, value of the effective spin, drawn from the population
+distribution, which may have some parameters $\lambda$; then the
+system is observed, represented by the likelihood function, which
+results in a distribution for the true effective spin (and all other
+paremeters describing the system) consistent with the data.
+Combining, the joint posterior on each system's $\chieff^i$ parameters
+and the population parameters $\lambda$ implied by a set of
+observations each with data $d^i$, is
+\begin{equation}
+  p\left( \left\{ \chieff^i \right\}, \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
+\end{equation}
 
-\begin{figure}
-  \plotone{../plots/six-way-O2-model-selection}
-  \caption{\label{fig:O2-predictions} Distribution of odds ratios
-    predicted with 10 additional observations above the three
-    discussed in Section \ref{sec:O1}.  Each panel corresponds to
-    additional observations drawn from one of the $\chieff$
-    distribution models.  The model from which the additional
-    observations are drawn is outlined in red.  The height of the blue
-    bar gives the median odds ratio relative to the model from which
-    the additional observations are drawn; the green line gives the
-    68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
-    separate draws from the model distribution.  The closest ratio
-    between the most-favoured isotropic model and the most-favoured
-    aligned model is $\OTwoOddsIsoAlignedMin$, corresponding to
-    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
-    distribution; most models result in more than $5\sigma$ preference
-    for the correct angular distribution.  Because the three
-    observations from Section \ref{sec:O1} are included in each data
-    set the ``correct'' model is not necessarily preferred over the
-    others, particularly when that model uses the ``increasing''
-    magnitude distribution, which is strongly dis-favoured from the O1
-    observations alone.}
-\end{figure}
+The components of this formula are
+\begin{itemize}
+\item The GW (marginal) likelihood, $p\left(d \mid \chieff\right)$.
+  ``Marginal'' because we are (implicitly) integrating over all
+  parameters of the signal but $\chieff$.  Note that it is the
+  likelihood, not the posterior that matters for the hierarchical
+  analysis; if we are given posterior distributions or posterior
+  samples, we need to re-weight to "remove" the prior and obtain the
+  likelihood.
+\item The population distribution for $\chieff$,
+  $p\left( \chieff \mid \lambda \right)$.  This function can be
+  parameterised by population-level parameters, $\lambda$.  (In the
+  cases discussed above, there are no parameters for the population.)
+\item The prior on the population-level parameters, $p(\lambda)$.
+\end{itemize}
+If we don't care about the individual event $\chieff$ paremetrs, we
+can integrate them out, obtaining
+\begin{equation}
+  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \int \dd \chieff^i \, p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
+\end{equation}
+If we are given posterior samples of $\chieff^{ij}$ ($i$ labels the
+event, $j$ labels the particular posterior sample) drawn from an
+analysis using a prior $p\left( \chieff \right)$, then we can
+approximate the integral by an re-weighted average of the population
+distribution over the samples (here $p\left( \chieff^{ij} \right)$ is
+the prior used to produce the posterior samples):
+\begin{equation}
+  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \frac{1}{N_i} \sum_{j=1}^{N_i} \frac{p\left( \chieff^{ij} \mid \lambda \right)}{p\left( \chieff^{ij} \right)} \right] p\left(\lambda\right).
+\end{equation}
 
-\section{Conclusion}
+\section{Mass Ratio}
+\label{sec:mass-ratio}
+
+\will{Some plots go here showing how the distribution changes as a
+  function of mass ratio.}
 

### 2017-03-23 paper/AlignedIsotropicGW.tex
Updated bibliography.

@@
 aligned models.  Anti-parallel spins are strongly disfavoured in
-models of isolated binary evolution.  \ilya{This would be a good place
-  for some review articles about pop.\ synth.}
+models of isolated binary evolution \citep[e.g.,]{Kalogera:2000}.
 
@@
 measurements \citep{2015PhR...548....1M}, some BH XRBs, including GRO
-J1655-40 \ilya{and 4U 1543-475?} appear to indicate that the micro
+J1655-40 \citep{Martin:2008}, 4U 1543-47 \citep{MorningstarMiller:2014} and particularly V4641 Sgr \citep{Orosz:2001,Martin:2008b} appear to indicate that the micro
 quasar jet, presumably aligned with the BH spin axis, is misaligned
 with the orbit.  Moreover, initial stellar spins in massive binaries
-have been observed to be misaligned \ilya{[ref]}, though there are
+have been observed to be misaligned \citep[e.g.,][]{Albrecht:2009}, though there are
 opportunities for realignment during binary evolution, as described
@@
 The spin directions of binary black holes formed dynamically through
-interactions in dense stellar environments \ilya{[add refs. if not
-  discussed elsewhere]} are expected to be isotropic given the absence
+interactions in dense stellar environments \citep{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016} 
+are expected to be isotropic given the absence
 of a preferred direction \citep[e.g.,][]{2016ApJ...832L...2R} and the
@@
 specific to rapidly spinning massive stars aligned by tidal effects
-\ilya{[add refs. if not discussed elsewhere]}.  However, wind-driven
+\citep{MandeldeMink:2016,Marchant:2016}.  However, wind-driven
 mass loss can spin down even some of these stars to the point that
 values of $\chi_\textrm{eff}$ as low as $\sim 0.2$ are possible,
-though alignment should be preserved \citep{2016AnA...588A..50M}.
+though alignment should be preserved \citep{Marchant:2016}.
 
-The classical isolated binary evolution channel through the
-common-envelope (CE) phase \ilya{[add refs. if not discussed
-  elsewhere]} yields the following typical formation channel for
+The classical isolated binary evolution channel \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,Dominik:2015} through the
+common-envelope (CE) phase \citep{Ivanova:2013} 
+yields the following typical formation channel for
 binary black holes: possible mass transfer from the primary to the
@@
 before the collapse of the primary could realign the binary, but the
-primary’s spin in the wide (separation $\gtrsim 10$ AU) binary can
-easily be misaligned by even moderate \citep{2015MNRAS.450L..85M}
+primary'۪s spin in the wide (separation $\gtrsim 10$ AU) binary can
+easily be misaligned by even moderate \citep{Mandel:2015kicks}
 natal kicks.  Realignment of the primary through subsequent mass
 accretion is unlikely given the small amount of accreted mass, except
-perhaps during hyper-critical accretion during CE \ilya{[add ref. to
-  Linden?]}; realignment through dissipation in the possibly warped
+perhaps during hyper-critical accretion during CE \citep{2005ApJ...632.1035O}, though even this accretion is likely to be small according to recent results \citep{MacLeodRamirezRuiz:2015}; realignment through dissipation in the possibly warped
 accretion disk \citep{1975ApJ...195L..65B} is possible, but calculations
@@
 efficiently re-aligned, depending on the separation between the
-secondary’s core and the BH immediately after the CE is ejected
-\citep{2016MNRAS.462..844K}.  The secondary’s natal kick will be small
+secondary's core and the BH immediately after the CE is ejected
+\citep{2016MNRAS.462..844K}.  The secondary'۪s natal kick will be small
 relative to the orbital velocity of the close post-CE binary, so any
@@
 may also change; for example, the misalignment of the secondary pulsar
-in the double pulsar \ilya{J0737} by $\sim 135^{\circ}$ relative to
+in the double pulsar J0737-3039 by $\sim 135^{\circ}$ relative to
 the orbit in the apparent absence of a kick indicated that, at least
@@
 between the collapsing and ejected material), the supernova explosion
-(if any), and the details of the post-collapse fallback process
-\ilya{[Check if Levin and Matzner are done]}.
+(if any), and the details of the post-collapse fallback process.
 
@@
 
-BBHs can form as the end point of isolated binary evolution \cite{refs}, or dynamically in dense stellar environments \cite{refs}.
+As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
 The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.

### 2017-03-27 paper/AlignedIsotropicGW.tex
Removed some un-printable characters and added arXiver comment line.

@@
 % !TEX TS-program = pdflatexmk
+
+% arXiver will snag the right plots so they show up on Twitter :)
+%@arxiver{chi-eff-distributions.pdf,Wills_evidence_ratio_figure_with_mixture_models.png,chi-eff-mock-posteriors.pdf,six-way-O2-model-selection.pdf}
 
@@
 before the collapse of the primary could realign the binary, but the
-primary'۪s spin in the wide (separation $\gtrsim 10$ AU) binary can
+primary's spin in the wide (separation $\gtrsim 10$ AU) binary can
 easily be misaligned by even moderate \citep{Mandel:2015kicks}
@@
 secondary's core and the BH immediately after the CE is ejected
-\citep{2016MNRAS.462..844K}.  The secondary'۪s natal kick will be small
+\citep{2016MNRAS.462..844K}.  The secondary's natal kick will be small
 relative to the orbital velocity of the close post-CE binary, so any

### 2017-03-27 paper/AlignedIsotropicGW.tex
Changed "increasing" to "high", "decreasing" to "low" to avoid letter-collision.
Also implemented distribution plot comparing q = 1 and q = 0.5 mass ratios.  No appreciable change.

@@
 
-\newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
-\newcommand{\OOneOddsIsoAligned}{\checkme{0.087}}
+\newcommand{\OOneSigmaIsoAligned}{\checkme{1.8}}
+\newcommand{\OOneOddsIsoAligned}{\checkme{0.08}}
 \newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
@@
   p(a) = \begin{cases}
-    2\left(1-a\right) & \textnormal{``decreasing''} \\
+    2\left(1-a\right) & \textnormal{``low''} \\
     1 & \textnormal{``flat''} \\
-    2\left(a-1\right) & \textnormal{``increasing''}
+    2\left(a-1\right) & \textnormal{``high''}
   \end{cases},
@@
     that the same distribution of spin vectors obtains for each
-    component.  The ``flat'' (blue lines), ``increasing,'' (green
-    lines), and ``decreasing'' (red lines) magnitude distributions are
+    component.  The ``flat'' (blue lines), ``high,'' (green
+    lines), and ``low'' (red lines) magnitude distributions are
     defined in Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the
@@
     the posteriors on $\chieff$ from the O1 observations shown in
-    Figure \ref{fig:O1-posteriors}.  The flat (``F''), increasing
-    (``I''), and decreasing (``D'') spin magnitude distributions (see
-    Eq.\ \eqref{eq:magnitude-dists}) are paired with isotropic,
-    aligned angular distributions (``$\mathrm{I}$'',
-    ``$\mathrm{A}$''), as well as a mixture model of the two
-    (``$\mathrm{M}$'').  The most-favoured models have a decreasing
-    distribution of spin magnitudes.  The odds ratio between these
-    models is $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.
-    For all magnitude distributions the pure-isotropic models are
-    preferred over the mixture models; correspondingly, the posterior
-    on the mixture fraction peaks at 100\% isotropic.}
+    Figure \ref{fig:O1-posteriors}.  The flat (``F''), high (``H''),
+    and low (``L'') spin magnitude distributions (see Eq.\
+    \eqref{eq:magnitude-dists}) are paired with isotropic, aligned
+    angular distributions (``$\mathrm{I}$'', ``$\mathrm{A}$''), as
+    well as a mixture model of the two (``$\mathrm{M}$'').  The
+    most-favoured models have a decreasing distribution of spin
+    magnitudes.  The odds ratio between these models is
+    $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.  For all
+    magnitude distributions the pure-isotropic models are preferred
+    over the mixture models; correspondingly, the posterior on the
+    mixture fraction peaks at 100\% isotropic.}
   \label{fig:O1-odds}
@@
     set the ``correct'' model is not necessarily preferred over the
-    others, particularly when that model uses the ``increasing''
-    magnitude distribution, which is strongly dis-favoured from the O1
+    others, particularly when that model uses the ``high'' magnitude
+    distribution, which is strongly dis-favoured from the O1
     observations alone.}

### 2017-03-27 paper/AlignedIsotropicGW.tex
Added text for mass ratio sensitivity.

@@
 $\chieff \gtrsim 0.5$ as observed in the majority of the
-electromagnetically-detected stellar-mass black hole population.
+electromagnetically-detected stellar-mass black hole population.  The
+hierarchical modelling is relatively insensitive to the precise
+details of the posterior distributions, unlike single-source analysis;
+notably our approximation does permit $\chieff = 0$ for GW151226 while
+the true posterior rules this out at high confidence
+\citep{2016PhRvL.116x1103A,O1-BBH}.
 
@@
 
-\will{Some plots go here showing how the distribution changes as a
-  function of mass ratio.}
+Figure \ref{fig:mass-ratio-sensitivity} shows the distributions of
+$\chieff$ that would obtain with a mass ratio $q = m_2/m_1 = 0.5$
+compared to the distributions with $q = 1$ used above.  The details of
+the distribution are sensitive to the mass ratio, but in our analysis
+we are primarily sensitive to the changing \emph{sign} of $\chieff$
+under the isotropic models.  This latter property is insensitive to
+mass ratio.  As an example, the distinction between the three
+different spin amplitude distributions under ten additional detections
+is quite weak compared to the aligned/isotropic distinction in Figure
+\ref{fig:O2-predictions}.  The differences in the $\chieff$
+distribution between $q = 1$ and $q = 0.5$ are even smaller than the
+differences between the different magnitude distributions.
+
+\begin{figure}
+  \plotone{../plots/chi-eff-distributions-q05}
+  \caption{Distributions of $\chieff$ assuming all merging black holes
+    have equal masses ($q=1$) or a 2:1 mass ratio ($q = 0.5$).  The
+    details of the distribution are sensitive to the mass ratio, but
+    in our analysis we are primarily sensitive to the changing
+    \emph{sign} of $\chieff$ under the isotropic models.  This latter
+    property is unchanged under changing mass ratio.}
+  \label{fig:mass-ratio-sensitivity}
+\end{figure}
 

### 2017-03-27 paper/AlignedIsotropicGW.tex
All except for Cole's difficult-to-address comment about spin distribution discussion.

@@
 
-\affiliation{Department of Astronomy and Maryland Center for Theory
-  and Computation, University of Maryland, College Park MD
-  20742, United States}
+\affiliation{Department of Astronomy and Joint Space-Science
+  Institute, University of Maryland, College Park, MD 20742--2421,
+  United States}
 
@@
   environment is the angular distribution of the \ac{BH} spins;
-  systems formed through dynamical interacions among already-compact
-  objects are expected to have isotropic spin orientations while
+  systems formed through dynamical interactions among already-compact
+  objects are expected to have isotropic spin orientations whereas
   binaries formed from pairs of stars born together are more likely to
@@
   effect of an additional 10 detections drawn from the various models
-  in the suite we show that such an agumented data set would enable at
-  least a $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
-  odds ratio) distinction between the isotropic and aligned models,
-  and in most cases better than $5\sigma$ ($2.9 \times 10^{-7}$ odds
-  ratio).  The existing data strongly suggest a dynamical formation
-  secnario for the observed systems, and this will be confirmed (or
-  overturned) confidently in the near future by subsequent
-  observations.
+  in the suite we show that if all observations come from a single
+  population such an agumented data set would enable at least a
+  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
+  ratio) distinction between the isotropic and aligned models, and in
+  most cases better than $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).
+  The existing preference for a dynamical formation scenario for the
+  observed systems will be confirmed (or overturned) confidently in
+  the near future by subsequent observations.
 \end{abstract}
@@
 holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
-and $a_{1,2}$ are the corresponding dimensionless spin parameters
-\citep{2016PhRvL.116x1102A}.
+and $0 \leq a_{1,2} \leq 1$ are the corresponding dimensionless spin
+parameters \citep{2016PhRvL.116x1102A}.
 
@@
     reported in \citet{O1-BBH}.  It is notable that none of the
-    $\chieff$ posteriors extends to the high spins in
-    electromagnetically-detected stellar-mass black hole systems
-    \citep{2015PhR...548....1M}.}
+    $\chieff$ posteriors extends to the high spin magnitudes inferred
+    from several electromagnetically-detected stellar-mass black hole
+    systems (see \citet{2015PhR...548....1M} for a summary of such
+    measurements).}
 \end{figure}
@@
 spins or larger spins whose direction is mis-aligned with the orbital
-plane of the binary (i.e.\ spin vectors with with small
-$z$-components).  Mis-alignment is capable of producing
-\emph{negative} values of $\chieff$, however, while small, aligned
-spins will always have $\chieff \geq 0$.  To quantify the degree of
-support for these two alternate explanations of small $\chieff$ values
-in the merging \ac{BBH} population, we compared the Bayesian evidence
-for various simple models of the spin population using the \ac{GW}
-data set.
+plane of the binary (i.e.\ spin vectors with small $z$-components).
+Mis-alignment is capable of producing \emph{negative} values of
+$\chieff$, however, while small, aligned spins will always have
+$\chieff \geq 0$.  To quantify the degree of support for these two
+alternate explanations of small $\chieff$ values in the merging
+\ac{BBH} population, we compared the Bayesian evidence for various
+simple models of the spin population using the \ac{GW} data set.
 
@@
     1 & \textnormal{``flat''} \\
-    2\left(a-1\right) & \textnormal{``high''}
+    2 a & \textnormal{``high''}
   \end{cases},
@@
 
+The distributions in Eq.\ \eqref{eq:magnitude-dists} are not meant to
+represent any particular physical model, but rather to capture our
+uncertainty about the spin magnitude distribution, discussed in detail
+in Section \ref{sec:discussion}; neither observations nor population
+synthesis codes can at this point authoritatively suggest {\it any}
+particular spin distribution.  Our models, however, allow us to see
+how sensitive the $\chieff$ distribution is to spin alignment given
+uncertainties about the spin magnitudes.
+
 We fit hierarchical models of the three LIGO O1 observations using
@@
 distributions, the mixture models' posterior on the mixing fraction
-peaked at 100\% isotropic, explaining why the zero-parameter,
+peaks at 100\% isotropic, which explains why the zero-parameter,
 pure-isotropic models are preferred over the single-parameter mixture
@@
 aligned and isotropic models is $\OOneOddsIsoAligned$, or
-$\OOneSigmaIsoAligned\sigma$; thus the data strongly favour isotropic
-spins among our suite of models.
+$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
+among our suite of models.
 
@@
 Figure \ref{fig:O2-predictions}.  We find that most scenarios with an
-additional 10 detections allow to distinguish the correct angular
-distribution with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
-credibility; and in the worst case the distinction is
+additional 10 detections allow the the correct angular distribution to
+be distinguished with greater than $5\sigma$ ($2.9 \times 10^{-7}$
+odds) credibility; and in the worst case the distinction is
 $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
 number of observations.  As noted below, however, spins observed in
-X-ray binaries are generally large, and most models of isolated binary
+X-ray binaries are typically large, and most models of isolated binary
 evolution do not produce systems with spins uniformly bounded to small
@@
 
-With this in mind, we performed our analysis using three straw-man
-spin distributions for spin parameters from $a/M=0$ to 1: flat
-($P(a/M)=1$); rising ($P(a/M)=2a/M$); and falling ($P(a/M)=2(1-a/M)$).
-These are not meant to represent any particular physical model, and
-indeed neither observations nor population synthesis codes can
-authoritatively suggest {\it any} particular spin distribution.  This
-does, however, allow us to see how sensitive the $\chieff$
-distribution is to spin alignment models given uncertainties about the
-spin magnitudes.
-
 Although BH spin-orbit alignment is assumed in continuum flux
 measurements \citep{2015PhR...548....1M}, some BH XRBs, including GRO
-J1655-40 \citep{Martin:2008}, 4U 1543-47 \citep{MorningstarMiller:2014} and particularly V4641 Sgr \citep{Orosz:2001,Martin:2008b} appear to indicate that the micro
+J1655-40 \citep{Martin:2008}, 4U 1543-47
+\citep{MorningstarMiller:2014} and particularly V4641 Sgr
+\citep{Orosz:2001,Martin:2008b} appear to indicate that the micro
 quasar jet, presumably aligned with the BH spin axis, is misaligned
 with the orbit.  Moreover, initial stellar spins in massive binaries
-have been observed to be misaligned \citep[e.g.,][]{Albrecht:2009}, though there are
+have been observed to be misaligned
+\citep[e.g.,][]{Albrecht:2009,2014ApJ...785...83A}, though there are
 opportunities for realignment during binary evolution, as described
@@
 The spin directions of binary black holes formed dynamically through
-interactions in dense stellar environments \citep{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016} 
-are expected to be isotropic given the absence
-of a preferred direction \citep[e.g.,][]{2016ApJ...832L...2R} and the
-persistence of an isotropic distribution through post-Newtonian
-evolution \citep{2004PhRvD..70l4020S}.
+interactions in dense stellar environments
+\citep{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+are expected to be isotropic given the absence of a preferred
+direction \citep[e.g.,][]{2016ApJ...832L...2R} and the persistence of
+an isotropic distribution through post-Newtonian evolution
+\citep{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
 
@@
 \item The GW (marginal) likelihood, $p\left(d \mid \chieff\right)$.
-  ``Marginal'' because we are (implicitly) integrating over all
-  parameters of the signal but $\chieff$.  Note that it is the
-  likelihood, not the posterior that matters for the hierarchical
-  analysis; if we are given posterior distributions or posterior
-  samples, we need to re-weight to "remove" the prior and obtain the
-  likelihood.
+  Here we use ``marginal'' because we are (implicitly) integrating
+  over all parameters of the signal but $\chieff$.  Note that it is
+  the likelihood rather than the posterior that matters for the
+  hierarchical analysis; if we are given posterior distributions or
+  posterior samples, we need to re-weight to "remove" the prior and
+  obtain the likelihood.
 \item The population distribution for $\chieff$,
@@
 \end{itemize}
-If we don't care about the individual event $\chieff$ paremetrs, we
+If we do not care about the individual event $\chieff$ paremetrs, we
 can integrate them out, obtaining
@@
 analysis using a prior $p\left( \chieff \right)$, then we can
-approximate the integral by an re-weighted average of the population
+approximate the integral by a re-weighted average of the population
 distribution over the samples (here $p\left( \chieff^{ij} \right)$ is

### 2017-03-27 paper/AlignedIsotropicGW.tex
Some text polishing and discussion of difference between measuring angular and magnitude distribution

@@
 \newcommand{\OOneOddsIsoAligned}{\checkme{0.08}}
-\newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{3.0}}
-\newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0012}}
+\newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{2.9}}
+\newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0035}}
 
@@
 against the observed data to obtain a posterior distribution on the
-parameter space that describes the compact binary source.  The spin
-parameter with the largest effect on waveforms, and a correspondingly
-tight constraint from the data, is a mass-weighted, aligned,
-``effective spin,''
+parameter space that describes the compact binary source.  The model
+waveforms incorporate approximations to the true behaviour of the
+merging systems that are expected to break down for sufficiently high
+mis-aligned spins; detailed comparisons with numerical relativity
+computations using no approximations to the dynamics
+\citep{2016PhRvD..94f4035A} show no evidence of such a breakdown in
+GW150914's analysis.  Given the preference of our data for smaller
+spins (discussed below) we assume here that measurements made now and
+in the future are not dominated by systematic errors.
+
+The spin parameter with the largest effect on waveforms, and a
+correspondingly tight constraint from the data, is a mass-weighted,
+aligned, ``effective spin,''
 \begin{equation}
@@
 $\chieff \gtrsim 0.5$ as observed in the majority of the
-electromagnetically-detected stellar-mass black hole population.  The
-hierarchical modelling is relatively insensitive to the precise
-details of the posterior distributions, unlike single-source analysis;
-notably our approximation does permit $\chieff = 0$ for GW151226 while
-the true posterior rules this out at high confidence
+electromagnetically-detected stellar-mass black hole population (see
+Section \ref{sec:discussion}).  The analysis here is relatively
+insensitive to the precise details of the posterior distributions;
+other conclusions are more sensitive.  In particular, our
+approximation does permit $\chieff = 0$ for GW151226 while the true
+posterior rules this out at high confidence
 \citep{2016PhRvL.116x1103A,O1-BBH}.
@@
 $\chieff$, however, while small, aligned spins will always have
-$\chieff \geq 0$.  To quantify the degree of support for these two
-alternate explanations of small $\chieff$ values in the merging
-\ac{BBH} population, we compared the Bayesian evidence for various
-simple models of the spin population using the \ac{GW} data set.
+$\chieff \geq 0$.  This difference provides strong discriminating
+power between the two angular distributions, even without good
+information about the magnitude distribution; to the extent that data
+support negative $\chieff$ they weigh heavily against aligned models.
+To quantify the degree of support for these two alternate explanations
+of small $\chieff$ values in the merging \ac{BBH} population, we
+compared the Bayesian evidence for various simple models of the spin
+population using the \ac{GW} data set.
 
@@
 $\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
-among our suite of models.
+among our suite of models.  While the data favour spin amplitude
+distributions with small spin magnitudes, note that a model with all
+\ac{BBH} systems having zero spin is ruled out by the GW151226
+measurements, which bound $\chieff \gtrsim 0.2$ at 90\% credibility
+\citep{2016PhRvL.116x1103A}.
 
@@
 $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
+Note that such future detections should permit a confident distinction
+between \emph{angular} distributions, we will remain very uncertain
+about the \emph{magnitude} distribution until we have a much larger
+number of observations.  In Figure \ref{fig:O2-predictions}, the odds
+ratio between different magnitude distributions with the same angular
+distribution is much closer to one than the odds ratio between angular
+distributions.
 
@@
     68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
-    separate draws from the model distribution.  The closest ratio
-    between the most-favoured isotropic model and the most-favoured
-    aligned model is $\OTwoOddsIsoAlignedMin$, corresponding to
-    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
-    distribution; most models result in more than $5\sigma$ preference
-    for the correct angular distribution.  Because the three
-    observations from Section \ref{sec:O1} are included in each data
-    set the ``correct'' model is not necessarily preferred over the
-    others, particularly when that model uses the ``high'' magnitude
-    distribution, which is strongly dis-favoured from the O1
-    observations alone.}
+    separate draws from the model distribution.  The closest median
+    ratio between the most-favoured isotropic model and the
+    most-favoured aligned model is $\OTwoOddsIsoAlignedMin$,
+    corresponding to $\OTwoSigmaIsoAlignedMin\sigma$ preference for
+    the correct angular distribution; most models result in more than
+    $5\sigma$ preference for the correct angular distribution.
+    Because the three observations from Section \ref{sec:O1} are
+    included in each data set the ``correct'' model is not necessarily
+    preferred over the others, particularly when that model uses the
+    ``high'' magnitude distribution, which is strongly dis-favoured
+    from the O1 observations alone.}
 \end{figure}
@@
 evolution do not produce systems with spins uniformly bounded to small
-values.
+values.  Additionally, models which do not permit \emph{some} spins
+with $\chieff \gtrsim 0.2$ are ruled out by the GW151226 observations
+\citep{2016PhRvL.116x1103A}.
 
@@
 axis, i.e.\ $\hat{S}\cdot \hat{L} \simeq \pm 1$, in contrast to our
-aligned models.  Anti-parallel spins are strongly disfavoured in
-models of isolated binary evolution \citep[e.g.,]{Kalogera:2000}.
+aligned models; thus the aligned/isotropic discrimitating power in
+\citet{2017CQGra..34cLT01V} relies effectively on measurements of the
+detailed shape of the $\chieff$ distribution and, as noted above,
+measurements are not very sensitive to the spin magnitude distribution
+that controls the shape of $\chieff$.  Note that anti-parallel spins
+are strongly disfavoured in models of isolated binary evolution (see,
+for example, \citet{Kalogera:2000}).
 
@@
 distribution of the dimensionless spin parameters of stellar-mass
-black holes.  Data for such model selection is sparse.
+black holes.  Observational data for such model selection is sparse.
 \citet{2015PhR...548....1M} give current estimates of the spin
 parameters for stellar-mass black holes, obtained using disk
-reflection and/or disk continuum methods.
-
-Most of the systems studied are low-mass X-ray binaries (i.e., the
-donor star has a mass $M_{\rm donor}<1~M_\odot$) rather than the
-high-mass X-ray binaries that are likely to be the progenitors of
-double black hole binaries.  In addition, there are substantial
-systematic errors that can complicate either type of analysis (see
-\citealt{2015PhR...548....1M} for a discussion). Nonetheless, if we
-take the reported spin magnitudes as representative then we find that
-there is a preference for high spins; for example, 14 of the 19
-systems with reported spins have spin parameters in excess of 0.5.
-The masses and spin parameters of stellar-mass black holes are
-unlikely to be altered significantly by accretion (low-mass donors
-simply do not have enough mass and high-mass donors have a very short
-phase in which they transfer mass; a variant of this long-standing
-argument was presented by \citet{1999MNRAS.305..654K}).  Thus the
-current spin parameters probably are close to their values upon core
-collapse.  We emphasize, however, that the specific processes involved
-in the production of black hole binaries could alter the spin
-distribution of those holes: for example, close tidal interactions
-could spin up the core, or stripping of the envelope could reduce the
-available angular momentum.
+reflection and/or disk continuum methods.  Most of the systems studied
+are low-mass X-ray binaries (i.e., the donor star has a mass
+$M_{\rm donor}<1~M_\odot$) rather than the high-mass X-ray binaries
+that are likely to be the progenitors of double black hole binaries.
+In addition, there are substantial systematic errors that can
+complicate either type of analysis (see \citealt{2015PhR...548....1M}
+for a discussion). Nonetheless, if we take the reported spin
+magnitudes as representative then we find that there is a preference
+for high spins; for example, 14 of the 19 systems with reported spins
+have spin parameters in excess of 0.5.  The masses and spin parameters
+of stellar-mass black holes are unlikely to be altered significantly
+by accretion (low-mass donors simply do not have enough mass and
+high-mass donors have a very short phase in which they transfer mass;
+a variant of this long-standing argument was presented by
+\citet{1999MNRAS.305..654K}).  Thus the current spin parameters
+probably are close to their values upon core collapse; however, the
+specific processes involved in the production of black hole binaries
+could alter the spin distribution of those holes.  For example, close
+tidal interactions could spin up the core, or stripping of the
+envelope could reduce the available angular momentum.
 
@@
 There are a few modelling constrains on the predicted spin
-distribution.  We separate these into a discussion of spin magnitudes
-and spin directions (spin-orbit misalignment angles) as in the
-preceding analysis; however, it is worth pointing out that, contrary
-to our simplified assumptions, the magnitude and direction
-distributions may well be coupled.
-
-Several models for binary black hole formation have been proposed
-\citep[e.g.,][]{2016ApJ...818L..22A}, yielding potentially different spin
-distributions.
+distribution (see, for example, \citep{2016ApJ...818L..22A}).  We
+separate these into a discussion of spin magnitudes and spin
+directions (spin-orbit misalignment angles) as in the preceding
+analysis; however, it is worth pointing out that, contrary to our
+simplified assumptions, the magnitude and direction distributions may
+well be coupled.
 
@@
 more probable that spins are aligned and non-zero after evolution
-through this channel than not.
+through this channel than not, while an isotropic distribution of
+non-zero spins is the natural outcome of dynamical formation
+processes.
+
+\acknowledgements
+
+We thank Richard O’Shaughnessy for comments on this work.
 

### 2017-03-27 paper/AlignedIsotropicGW.tex
Spellcheck.

@@
 
-\newcommand{\OOneSigmaIsoAligned}{\checkme{1.8}}
-\newcommand{\OOneOddsIsoAligned}{\checkme{0.08}}
+\newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
+\newcommand{\OOneOddsIsoAligned}{\checkme{0.087}}
 \newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{2.9}}
@@
   in the suite we show that if all observations come from a single
-  population such an agumented data set would enable at least a
+  population such an augmented data set would enable at least a
   $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
@@
 \begin{figure}
-  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models.png}
+  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models}
   \caption{Odds ratios among our models using the approximations to
@@
 axis, i.e.\ $\hat{S}\cdot \hat{L} \simeq \pm 1$, in contrast to our
-aligned models; thus the aligned/isotropic discrimitating power in
+aligned models; thus the aligned/isotropic discriminating power in
 \citet{2017CQGra..34cLT01V} relies effectively on measurements of the
@@
 We use the `uniform', `increasing' and `decreasing' distributions introduced in Section X, assuming all BHs have their spin magnitude drawn from the same distribution.
-We calculate the posterior given by EquationX plot the posterior distribution/cumulative posterior in Figure~\ref{fig:mixture_fraction_posterior}.
+We calculate the posterior given by Equation \ref{eq:magnitude-dists} plot the posterior distribution/cumulative posterior in Figure~\ref{fig:mixture_fraction_posterior}.
 We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the `decreasing', `uniform' and `increasing' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
@@
 results in a distribution for the true effective spin (and all other
-paremeters describing the system) consistent with the data.
+parameters describing the system) consistent with the data.
 Combining, the joint posterior on each system's $\chieff^i$ parameters
@@
 \end{itemize}
-If we do not care about the individual event $\chieff$ paremetrs, we
+If we do not care about the individual event $\chieff$ parameters, we
 can integrate them out, obtaining

### 2017-03-28 paper/AlignedIsotropicGW.tex
Added waveform systematics paper to references and moved w.f. systematics discussion to appendix.

@@
 against the observed data to obtain a posterior distribution on the
-parameter space that describes the compact binary source.  The model
-waveforms incorporate approximations to the true behaviour of the
-merging systems that are expected to break down for sufficiently high
-mis-aligned spins; detailed comparisons with numerical relativity
-computations using no approximations to the dynamics
-\citep{2016PhRvD..94f4035A} show no evidence of such a breakdown in
-GW150914's analysis.  Given the preference of our data for smaller
-spins (discussed below) we assume here that measurements made now and
-in the future are not dominated by systematic errors.
+parameter space that describes the compact binary source.
 
@@
 
+\section{Approximaions in the Gravitational Waveform}
+The model waveforms used to infer the $\chieff$ of the three events
+incorporate approximations to the true behaviour of the merging
+systems that are expected to break down for sufficiently high
+mis-aligned spins.  The effect of these approximations has been
+investigated in detail for the inference on the parameters describing
+GW150914 \citep{2016arXiv161107531T}.  For this source, statistical
+uncertainties dominate over any waveform systematics.  Detailed
+comparisons with numerical relativity computations using no
+approximations to the dynamics \citep{2016PhRvD..94f4035A} also
+suggest that statistical uncertainties dominate the systematics for
+this system.  \citet{2016arXiv161107531T} suggests that systematics
+may dominate for signals with this large SNR ($\simeq 23$) when the
+source is edge-on or with high spin.  The other two events discussed
+in this paper are at much lower SNR, with correspondingly larger
+statistical uncertainties, and are probably similarly oriented and
+with similarly small spins, so we do not expect systematic
+uncertanties to dominate.
+
+Given the preference of our data for a smaller spin population we
+assume here that measurements made in the future are not dominated by
+systematic errors, but this assumption would need to be revisited for
+high-SNR, edge-on, or high-spin sources detected in the future.
+
 \bibliography{AlignedIsotropicGW}

### 2017-03-28 paper/AlignedIsotropicGW.tex
Added Christopher Berry to acknowledgements.
Changed title to make shorter (~90 characters based on Nature limits).

@@
 
-\title{Distinguishing Between Spin-Aligned and Isotropic Binary Black
-  Hole Populations Using Gravitational Wave Observations}
+\title{Distinguishing Spin-Aligned and Isotropic Black Hole
+  Populations With Gravitational Waves}
 
@@
 
-We thank Richard O’Shaughnessy for comments on this work.
+We thank Richard O’Shaughnessy and Christopher Berry for discussions
+comments on this work.
 

### 2017-03-28 paper/AlignedIsotropicGW.tex
Cole ack. IAS visiting fellows program.
Added appendix on rough, order-of-magnitude calculation.

@@
 We thank Richard O’Shaughnessy and Christopher Berry for discussions
-comments on this work.
+comments on this work.  MCM acknowledges support of the University of
+Birmingham Institute for Advanced Study Distinguished Visiting Fellows
+program.
 
@@
 
+\subsection{Order of Magnitude Calculation}
+\label{sec:om-odds-ratio}
+
+It is possible to estimate at an order-of-magnitude level the rate at
+which evidence accumulates in favour of or against the isotropic
+models as more systems are detected.  Based on Figure
+\ref{fig:chieff-distribution-models}, approximate the isotropic
+population $\chieff$ distribution as uniform on
+$\chieff \in \left[ -0.25, 0.25 \right]$ and the aligned population
+$\chieff$ distribution as uniform on
+$\chieff \in \left[0, 0.5\right]$.  Then the odds ratio between the
+isotropic and aligned models for each event is approximately
+\begin{equation}
+  \label{eq:approx-odds}
+  \frac{p\left( d \mid I \right)}{p\left( d \mid A \right)} \simeq
+  \frac{P\left( -0.25 \leq \chieff \leq 0.25 \right)}{P\left( 0 \leq \chieff \leq 0.5 \right) },
+\end{equation}
+where $P\left( A \leq \chieff \leq B \right)$ is the posterior
+probability that $\chieff$ is between $A$ and $B$.  Using our
+approximations to the $\chieff$ posteriors described above, this gives
+an odds ratio of $5$ in favour of the isotropic models, which is about
+a factor of two smaller than the ratio in the more careful calculation
+described in Section \ref{sec:O1}.  This is a satisfactory answer at
+an order-of-magnitude level.
+
+If the true distribution is isotropic and follows this simple model,
+and our measurement uncertainties on $\chieff$ are $\simeq 0.2$, then
+each subsequent measurement contributes on average a factor of
+$\sim 2.5$ to the overall odds.  (There is a 1/3 chance each of the
+critical value of $\chieff = 0$ lying at $-1\sigma$, $0$, and
+$+1\sigma$ with respect to the centre of the posterior.)  After ten
+additional events, then, the odds ratio becomes
+$5 \times 2.5^{10} \simeq 5 \times 10^{5}$, or $4.3 \sigma$,
+consistent with the results of the more detailed calculation described
+above.
+
 \section{Mass Ratio}

### 2017-03-28 paper/AlignedIsotropicGW.tex
Better calculations in the O.M. section.

@@
 If the true distribution is isotropic and follows this simple model,
-and our measurement uncertainties on $\chieff$ are $\simeq 0.2$, then
+and our measurement uncertainties on $\chieff$ are $\simeq 0.1$, then
 each subsequent measurement contributes on average a factor of
-$\sim 2.5$ to the overall odds.  (There is a 1/3 chance each of the
-critical value of $\chieff = 0$ lying at $-1\sigma$, $0$, and
-$+1\sigma$ with respect to the centre of the posterior.)  After ten
-additional events, then, the odds ratio becomes
-$5 \times 2.5^{10} \simeq 5 \times 10^{5}$, or $4.3 \sigma$,
-consistent with the results of the more detailed calculation described
-above.
+$\sim 3$ to the overall odds.  After ten additional events, then, the
+odds ratio becomes $5 \times 3^{10} \simeq 3 \times 10^{5}$, or
+$4.6 \sigma$, consistent with the results of the more detailed
+calculation described above.  If the true distribution of spins
+becomesf half as wide ($\chieff \in [-0.125, 0.125]$ for isotropic and
+$\chieff \in [0, 0.25]$ for aligned spins), with the same
+uncertainties, then the existing odds ratio becomes $1.08$, and each
+subsequent event drawn from the isotropic distribution contributes on
+average a factor of $1.6$.  In this case, after 10 additional events,
+the odds ratio becomes $150$, or $2.7\sigma$.  With small spin
+magnitudes, our angular resolving power vanishes.
 

### 2017-03-29 paper/AlignedIsotropicGW.tex
Made changes to presentation of nuances.

@@
   Birmingham, B15 2TT, United Kingdom}
+\affiliation{Kavli Institute of Theoretical Physics, University of California, Santa Barbara}
 
@@
   Birmingham, B15 2TT, United Kingdom}
+ \affiliation{Kavli Institute of Theoretical Physics, University of California, Santa Barbara}
 
@@
   binaries formed from pairs of stars born together are more likely to
-  have spins aligned with the binary orbit as a consequence of their
-  joint evolution toward a \ac{BBH} system.  By considering existing
+  have spins \ilya{preferentially} aligned with the binary orbit as a consequence of their
+  joint evolution toward a \ac{BBH} system.  We consider existing
   \ac{GW} measurements of $\chieff$, the best-measured combination of
   spin parameters, in the three likely binary black hole detections
-  GW150914, LVT151012, and GW151226, we show that the data already
+  GW150914, LVT151012, and GW151226.
+  \ilya{If binary black hole spin magnitudes extend to high values, as is
+suggested by observations of black hole X-ray binaries,} we show that the data already
   exhibit a $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
-  ratio) preference for an isotropic angular distribution amongst a
-  suite of models for the spin distribution.  By considering the
+  ratio) preference for an isotropic angular distribution.
+  By considering the
   effect of an additional 10 detections drawn from the various models
@@
   $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
-  ratio) distinction between the isotropic and aligned models, and in
+  ratio) distinction between the isotropic and aligned models \ilya{for the assumed spin magnitude distributions}, and in
   most cases better than $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).
-  The existing preference for a dynamical formation scenario for the
-  observed systems will be confirmed (or overturned) confidently in
-  the near future by subsequent observations.
+  The existing preference for \ilya{either an isotropic spin distribution or low spin magntiudes} for the observed systems will be confirmed (or overturned) confidently in the near future by subsequent observations.
 \end{abstract}
@@
 $\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
-spins are intrinsically very small, with $a \lesssim 0.1$ always, then
+spins are intrinsically very small, with $a \lesssim \ilya{0.3}$, \ilya{[I changed the number to be consistent with Simon's analysis, which shows that you can no longer tell the difference for a mean spin of $\sim 0.17$]} then
 it is no longer possible to resolve the negative spin with a small
 number of observations.  As noted below, however, spins observed in
-X-ray binaries are typically large, and most models of isolated binary
-evolution do not produce systems with spins uniformly bounded to small
-values.  Additionally, models which do not permit \emph{some} spins
-with $\chieff \gtrsim 0.2$ are ruled out by the GW151226 observations
+X-ray binaries are typically large. 
+\sout{and most models of isolated binary evolution do not produce systems with spins uniformly bounded to small values}.  
+Additionally, models which do not permit \emph{some} spins
+with $\chieff \gtrsim \ilya{0.1}$ \ilya{[Changed to be consistent with 90\% lower bound from cited paper]} are ruled out by the GW151226 observations
 \citep{2016PhRvL.116x1103A}.
 
-\citet{2017CQGra..34cLT01V} also studied the possibility of
-distinguishing aligned and isotropic angular distributions of \ac{BBH}
-spins, but concluded that several hundred sources would be required to
-adequately separate the models.  The ``aligned'' population in that
-study permitted spins nearly parallel or anti-parallel to the orbital
-axis, i.e.\ $\hat{S}\cdot \hat{L} \simeq \pm 1$, in contrast to our
-aligned models; thus the aligned/isotropic discriminating power in
-\citet{2017CQGra..34cLT01V} relies effectively on measurements of the
-detailed shape of the $\chieff$ distribution and, as noted above,
-measurements are not very sensitive to the spin magnitude distribution
-that controls the shape of $\chieff$.  Note that anti-parallel spins
-are strongly disfavoured in models of isolated binary evolution (see,
-for example, \citet{Kalogera:2000}).
-
 In order to perform our analysis we need to select models for the
-distribution of the dimensionless spin parameters of stellar-mass
+distribution of the \ilya{spin magnitudes} of stellar-mass
 black holes.  Observational data for such model selection is sparse.
@@
 reflection and/or disk continuum methods.  Most of the systems studied
-are low-mass X-ray binaries (i.e., the donor star has a mass
-$M_{\rm donor}<1~M_\odot$) rather than the high-mass X-ray binaries
+are low-mass X-ray binaries rather than the high-mass X-ray binaries
 that are likely to be the progenitors of double black hole binaries.
@@
 complicate either type of analysis (see \citealt{2015PhR...548....1M}
-for a discussion). Nonetheless, if we take the reported spin
+for a discussion), \ilya{and observational selection effects could yield a biased distribution}. Nonetheless, if we take the reported spin
 magnitudes as representative then we find that there is a preference
 for high spins; for example, 14 of the 19 systems with reported spins
-have spin parameters in excess of 0.5.  The masses and spin parameters
+have dimensionless spin parameters in excess of 0.5.  \ilya{It is usually argued that} the masses and spin parameters
 of stellar-mass black holes are unlikely to be altered significantly
-by accretion (low-mass donors simply do not have enough mass and
+by accretion (low-mass donors may not have enough mass and
 high-mass donors have a very short phase in which they transfer mass;
 a variant of this long-standing argument was presented by
-\citet{1999MNRAS.305..654K}).  Thus the current spin parameters
-probably are close to their values upon core collapse; however, the
-specific processes involved in the production of black hole binaries
-could alter the spin distribution of those holes.  For example, close
+\citet{1999MNRAS.305..654K}), \ilya{but see \citet{Podsiadlowski:2003,FragosMcClintock:2015}}.  Thus the current spin parameters
+probably are close to their values upon core collapse.  However, the
+specific processes involved in the production of black hole binaries \ilya{from isolated binaries}
+could alter the spin \ilya{magnitude} distribution of those holes \ilya{relative to the X-ray binary systems}; for example, close
 tidal interactions could spin up the core, or stripping of the
-envelope could reduce the available angular momentum.
-
-Although BH spin-orbit alignment is assumed in continuum flux
-measurements \citep{2015PhR...548....1M}, some BH XRBs, including GRO
-J1655-40 \citep{Martin:2008}, 4U 1543-47
-\citep{MorningstarMiller:2014} and particularly V4641 Sgr
-\citep{Orosz:2001,Martin:2008b} appear to indicate that the micro
-quasar jet, presumably aligned with the BH spin axis, is misaligned
-with the orbit.  Moreover, initial stellar spins in massive binaries
-have been observed to be misaligned
-\citep[e.g.,][]{Albrecht:2009,2014ApJ...785...83A}, though there are
-opportunities for realignment during binary evolution, as described
-below.
-
-There are a few modelling constrains on the predicted spin
-distribution (see, for example, \citep{2016ApJ...818L..22A}).  We
-separate these into a discussion of spin magnitudes and spin
-directions (spin-orbit misalignment angles) as in the preceding
-analysis; however, it is worth pointing out that, contrary to our
-simplified assumptions, the magnitude and direction distributions may
-well be coupled.
+envelope could reduce the available angular momentum \citep{2016MNRAS.462..844K,Zaldarriaga:2017,HotokezakaPiran:2017}.
 
@@
 
-Chemically homogeneous evolution is a possible BBH formation channel
-specific to rapidly spinning massive stars aligned by tidal effects
-\citep{MandeldeMink:2016,Marchant:2016}.  However, wind-driven
-mass loss can spin down even some of these stars to the point that
-values of $\chi_\textrm{eff}$ as low as $\sim 0.2$ are possible,
-though alignment should be preserved \citep{Marchant:2016}.
+\ilya{The spin directions in isolated binaries, whether evolving via the classical channel through a common-envelope phase \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017} or through chemically homogeneous evolution \citep{MandeldeMink:2016,Marchant:2016} are generally expected to be preferentially aligned.
+Despite  observed spin misalignments in both massive stellar binaries \citep{Albrecht:2009,2014ApJ...785...83A} and BH XRBs \citep{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014}, mass accretion and tidal interactions will tend to realign the binary.
+On the other hand, a supernova natal kick (if any) can change the orbital plane and misalign the binary \citep{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can also tilt the spin angle, as in the double pulsar J0737-3039 \citep{2011ApJ...742...81F}; and a variety of uncertain processes, such as wind-driven mass loss and post-collapse fallback, can couple the spin magnitude and direction distributions, contrary to our simplified assumptions.}
 
-The classical isolated binary evolution channel \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,Dominik:2015} through the
-common-envelope (CE) phase \citep{Ivanova:2013} 
-yields the following typical formation channel for
-binary black holes: possible mass transfer from the primary to the
-secondary; the collapse of the primary into a BH; the formation and
-ejection of a CE as a consequence of dynamically unstable mass
-transfer from the secondary to the primary; possible further mass
-transfer from remaining secondary core to primary; collapse of the
-secondary into a BH; a merger induced by GW emission
-\citep[e.g.,][]{2016Natur.534..512B,Stevenson:2017}.  Interaction
-before the collapse of the primary could realign the binary, but the
-primary's spin in the wide (separation $\gtrsim 10$ AU) binary can
-easily be misaligned by even moderate \citep{Mandel:2015kicks}
-natal kicks.  Realignment of the primary through subsequent mass
-accretion is unlikely given the small amount of accreted mass, except
-perhaps during hyper-critical accretion during CE \citep{2005ApJ...632.1035O}, though even this accretion is likely to be small according to recent results \citep{MacLeodRamirezRuiz:2015}; realignment through dissipation in the possibly warped
-accretion disk \citep{1975ApJ...195L..65B} is possible, but calculations
-suggest it is unlikely for stellar-mass binaries \citep{KLOP}.
+\ilya{\citet{2017CQGra..34cLT01V} also studied the possibility of distinguishing aligned and isotropic angular distributions of \ac{BBH} spins, but concluded that several hundred sources would be required to adequately separate models which included both parallel and anti-parallel spins for 'aligned' population.
+In contrast, and in agreement with the present study, \citet{Stevenson:2017spin} found that only $\sim 5$ observations of rapidly spinning black holes would be necessary to distinguish isotropic and aligned spin distributions, though tens of detections would be required for more nuanced admixture models.
+Meanwhile, recent studies by \citet{Fishbach:2017,Gerosa:2017} focused on inference on spin magnitudes.}
 
-The secondary should tidally align with the orbit before the onset of
-mass transfer
-\citep[e.g.,][]{2000ApJ...541..319K,2013PhRvD..87j4028G}; however,
-most of its angular momentum would be lost to the envelope, which will
-be ejected during the CE phase.  The remnant core may or may not be
-efficiently re-aligned, depending on the separation between the
-secondary's core and the BH immediately after the CE is ejected
-\citep{2016MNRAS.462..844K}.  The secondary's natal kick will be small
-relative to the orbital velocity of the close post-CE binary, so any
-further kick-induced misalignment will be moderate
-\citep{2016ApJ...832L...2R}.
+\ilya{In summary}, the angular distribution of spins in \acp{BBH} formed through isolated binary evolution is uncertain; nevertheless, it is more probable that spins are \ilya{preferentially} aligned \sout{and non-zero} after evolution through this channel than not, while an isotropic distribution of \sout{non-zero} spins is the natural outcome of dynamical formation processes.
+\ilya{Therefore, if the current observational trend for low $\chieff$ continues with future gravitational-wave observations, it will be possible to either confirm that \ac{BH} spins are sometimes large in magnitude and isotropic in direction, yielding a strong indication of a dynamical formation origin; or that \ac{BH} spins are overwhelmingly small in magnitude, yielding a notable contradiction with the claimed X-ray binary spin measurements, particularly those for high-mass X-ray binaries.}
 
-The analysis above assumed the conservation of angular momentum during
-core collapse. However, we know that this may not always be the case.
-For example, the magnitude of the angular momentum cannot possibly be
-conserved during the collapse to a BH if the pre-collapse
-dimensionless spin exceeds $1$.  The direction of the angular momentum
-may also change; for example, the misalignment of the secondary pulsar
-in the double pulsar J0737-3039 by $\sim 135^{\circ}$ relative to
-the orbit in the apparent absence of a kick indicated that, at least
-for slowly spinning objects, a spin tilt during a supernova may be
-possible \citep{2011ApJ...742...81F}.  The dependence of the magnitude
-and direction of the spin on its pre-collapse values depends on the
-stellar structure (i.e., the level of angular momentum coupling
-between the collapsing and ejected material), the supernova explosion
-(if any), and the details of the post-collapse fallback process.
-
-In short, the angular distribution of spins in \acp{BBH} formed
-through isolated binary evolution is uncertain; nevertheless, it is
-more probable that spins are aligned and non-zero after evolution
-through this channel than not, while an isotropic distribution of
-non-zero spins is the natural outcome of dynamical formation
-processes.
+%\citet{2017CQGra..34cLT01V} also studied the possibility of distinguishing aligned and isotropic angular distributions of \ac{BBH} spins, but concluded that several hundred sources would be required to adequately separate the models.
+The ``aligned'' population in that study permitted spins nearly parallel or anti-parallel to the orbital axis, i.e.\ $\hat{S}\cdot \hat{L} \simeq \pm 1$, in contrast to our aligned models; thus the aligned/isotropic discriminating power in \citet{2017CQGra..34cLT01V} relies effectively on measurements of the detailed shape of the $\chieff$ distribution and, as noted above, measurements are not very sensitive to the spin magnitude distribution that controls the shape of $\chieff$.
+Note that anti-parallel spins are strongly disfavoured in models of isolated binary evolution (see, for example, \citet{Kalogera:2000}).
 
@@
 
-We thank Richard O’Shaughnessy and Christopher Berry for discussions
-comments on this work.  MCM acknowledges support of the University of
+We thank Richard O'Shaughnessy and Christopher Berry for discussions and
+comments on this work.  WF, SS, IM and AV were supported in part by the STFC.  MCM acknowledges support of the University of
 Birmingham Institute for Advanced Study Distinguished Visiting Fellows
-program.
+program.   SS and IM acknowledge support from the National Science Foundation under Grant No. NSF PHY11-25915.
 
@@
 
-As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
-The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.
+\sout{As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
+The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.}
 
-Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
-In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.
+\sout{Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
+In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.}
 
-In this section we assume that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).
+\ilya{While we carried out Bayesian comparisons between isotropic and aligned spin distributions under various assumptions, a preference for one of the considered models over the others does not necessarily indicate that it is the correct model.
+All of the considered models could be inaccurate for the actual distribution, especially since all of the considered models are based on a number of additional assumptions, such as decoupled spin magnitude and spin misalignment angle distributions and identical distributions for primary and secondary spins.}
+\ilya{We now partly relax the simplified assumptions made earlier by considering the possibility that the} true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries \ilya{with aligned spins and binaries with isotropic spins}.
+\sout{formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).}
 
@@
 The lower 90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the prior of 0.1.
-Thus, irrespective of our ignorance of the BH spin magnitude distribution, we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
+Thus, \ilya{for these spin magnitude distributions} 
+\sout{irrespective of our ignorance of the BH spin magnitude distribution}
+we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
 
-If we assume that the subpopulation of BBHs with isotropically distributed spin orientations corresponds to a subpopulation of dynamically formed BBHs, our result suggests that the majority of merging BBHs observed by LIGO are formed dynamically, rather than through isolated binary evolution.
+\sout{If we assume that the subpopulation of BBHs with isotropically distributed spin orientations corresponds to a subpopulation of dynamically formed BBHs, our result suggests that the majority of merging BBHs observed by LIGO are formed dynamically, rather than through isolated binary evolution.}
 
@@
 calculation described above.  If the true distribution of spins
-becomesf half as wide ($\chieff \in [-0.125, 0.125]$ for isotropic and
+becomes half as wide ($\chieff \in [-0.125, 0.125]$ for isotropic and
 $\chieff \in [0, 0.25]$ for aligned spins), with the same
@@
 
-\section{Approximaions in the Gravitational Waveform}
+\section{Approximations in the Gravitational Waveform}
 The model waveforms used to infer the $\chieff$ of the three events
@@
 with similarly small spins, so we do not expect systematic
-uncertanties to dominate.
+uncertainties to dominate.
 

### 2017-03-29 paper/AlignedIsotropicGW.tex

--- 2017-03-28 paper/AlignedIsotropicGW.tex
@@
 
-As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
+%As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
 The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.
+%
+%Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
+In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.
+%
+%In this section we assume that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).
 
-Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
-In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.
-
-In this section we assume that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).
+In this section we assume that the true distribution of
+\acp{BBH} observed by LIGO is a mixture of binaries formed
+through isolated binary evolution (with aligned spins) and
+dynamical formation (with isotropic spins). Rate uncertainties
+for both of these formation scenarios are large enough that we
+do not know a priori which will dominate the LIGO detection
+rate.
 
@@
 \includegraphics[width=0.45\textwidth]{../plots/posterior_on_isotropic_fraction.png}
-\caption{\textbf{Fraction of the BBH population coming from an isotropic distribution} The dotted (blue) line shows the flat prior on the fraction of BBHs coming from an isotropic distribution $f_i$.
-The solid (red) line shows the posterior after O1, assuming that all BHs have their spin magnitude drawn from a uniform distribution.
-The dashed line assumes a `increasing' distribution $p(a) = 2a$ for BH spin magnitudes, whilst the dot-dash line assumes a `decreasing' distribution $p(a) = 2(1-a)$.
-We see that regardless of our assumption regarding BH spin magnitudes, there is a preference for a large fraction coming from an isotropic distribution.}
+\caption{\textbf{Fraction of the BBH population coming from an
+isotropic distribution} The dotted line shows the flat
+prior on the fraction of BBHs coming from an isotropic
+distribution $f_i$ under the mixture model. The 3 red lines show
+the posterior on $f_i$ after O1 under with our various
+assumptions regarding BH spin magnitudes.
+The solid line shows the posterior assuming that all BHs have
+their spin magnitude drawn
+from the ``flat'' distribution. The dashed line assumes the
+``high'' BH spin magnitude distribution $p(a) = 2a$. The
+dot-dash line assumes the ``low'' distribution $p(a) = 2(1-a)$.
+We see that for a wide range of assumptions regarding BH spin
+magnitudes, the fraction coming from an isotropic distribution
+$f_i$ peaks at 1.}
 \label{fig:mixture_fraction_posterior}
@@
 
-We fit a mixture model (labelled model 'M') where a fraction $f_i$ ($f_i=\lambda$ in the derivations above) of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
+Following \citet{Stevenson:2017spin},  we fit a mixture model (labelled model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
 We assume a flat prior on the fraction $f_i$.
 To test the robustness of our result, we vary the distribution we assume for BH spin magnitude distributions as with the aligned and isotropic models.
-We use the `uniform', `increasing' and `decreasing' distributions introduced in Section X, assuming all BHs have their spin magnitude drawn from the same distribution.
-We calculate the posterior given by Equation \ref{eq:magnitude-dists} plot the posterior distribution/cumulative posterior in Figure~\ref{fig:mixture_fraction_posterior}.
-We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the `decreasing', `uniform' and `increasing' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
+We use the ``flat'', ``high'' and ``low'' distributions (Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin magnitude drawn from the same distribution for both the aligned and isotropic populations.
+We calculate and plot the posterior on $f_i$ given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in the derivation) in Figure~\ref{fig:mixture_fraction_posterior}.
+
+We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high'' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
 The lower 90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the prior of 0.1.
-Thus, irrespective of our ignorance of the BH spin magnitude distribution, we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
+In all cases, the posterior peaks at $f_i = 1$.
+Thus, for a range of assumptions regarding the the BH spin magnitude distribution, we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
 
-If we assume that the subpopulation of BBHs with isotropically distributed spin orientations corresponds to a subpopulation of dynamically formed BBHs, our result suggests that the majority of merging BBHs observed by LIGO are formed dynamically, rather than through isolated binary evolution.
-
-The evidence ratios of these models to the isotropic distribution with uniform spin magnitudes are 0.85, 0.39 and 0.19. Thus we cannot rule out a mixture with the current data.
+The evidence ratios of these models to the isotropic distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for the ``low'', ``flat'' and ``high'' spin magnitude models.
+Thus we cannot rule out a mixture with the current data.
 
@@
   p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \frac{1}{N_i} \sum_{j=1}^{N_i} \frac{p\left( \chieff^{ij} \mid \lambda \right)}{p\left( \chieff^{ij} \right)} \right] p\left(\lambda\right).
+  \label{eq:hierarchical-posterior}
 \end{equation}

### 2017-03-29 paper/AlignedIsotropicGW.tex

--- 2017-03-29 paper/AlignedIsotropicGW.tex
@@
 
+%As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
+The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.
+%
+%Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
+In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.
+%
+%In this section we assume that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).
 \sout{As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
@@
 
+In this section we assume that the true distribution of
+\acp{BBH} observed by LIGO is a mixture of binaries formed
+through isolated binary evolution (with aligned spins) and
+dynamical formation (with isotropic spins). Rate uncertainties
+for both of these formation scenarios are large enough that we
+do not know a priori which will dominate the LIGO detection
+rate.
 \ilya{While we carried out Bayesian comparisons between isotropic and aligned spin distributions under various assumptions, a preference for one of the considered models over the others does not necessarily indicate that it is the correct model.
@@
 
+Following \citet{Stevenson:2017spin},  we fit a mixture model (labelled model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
+We assume a flat prior on the fraction $f_i$.
+To test the robustness of our result, we vary the distribution we assume for BH spin magnitude distributions as with the aligned and isotropic models.
+We use the ``flat'', ``high'' and ``low'' distributions (Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin magnitude drawn from the same distribution for both the aligned and isotropic populations.
+We calculate and plot the posterior on $f_i$ given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in the derivation) in Figure~\ref{fig:mixture_fraction_posterior}.
 We fit a mixture model (labelled model 'M') where a fraction $f_i$ ($f_i=\lambda$ in the derivations above) of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
@@
 
+We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high'' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
+The lower 90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the prior of 0.1.
+In all cases, the posterior peaks at $f_i = 1$.
+Thus, for a range of assumptions regarding the the BH spin magnitude distribution, we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
 \sout{If we assume that the subpopulation of BBHs with isotropically distributed spin orientations corresponds to a subpopulation of dynamically formed BBHs, our result suggests that the majority of merging BBHs observed by LIGO are formed dynamically, rather than through isolated binary evolution.}
 
-The evidence ratios of these models to the isotropic distribution with uniform spin magnitudes are 0.85, 0.39 and 0.19. Thus we cannot rule out a mixture with the current data.
+The evidence ratios of these models to the isotropic distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for the ``low'', ``flat'' and ``high'' spin magnitude models.
+Thus we cannot rule out a mixture with the current data.
 

### 2017-03-29 paper/AlignedIsotropicGW.tex
Added section on effect of small spin magnitudes -- can be cut if not agreed upon, included to make decision easier

@@
 
+\section{Effect of small spin magnitudes}
+\label{sec:smallspins}
+
+In the main text we considered three models for BH spin magnitudes: ``low'', ``flat'' and ``high''. These were intended to capture some of the uncertainty regarding the BH spin magnitude distribution.
+
+Here we extend the ``low'' model as:
+%
+\begin{equation}
+p(a) \propto (1 - a)^{\alpha}
+\label{eq:lowspinalpha}
+\end{equation}
+%
+
+When $\alpha = 0$, this recovers the ``flat'' distribution, whilst $\alpha = 1$
+recovers the ``low'' distribution. For higher values of $\alpha$, this distribution 
+becomes more peaked towards $a = 0$.
+
+%
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{../plots/sigma_v_alpha.png}
+\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
+The top axis shows the mean spin magnitude $a$
+that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
+\label{fig:smallspinsalpha}
+\end{figure}
+%
+
+In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of aligned to isotropic distributions (plotted as the equivalent sigma) with spin magnitudes given by this model with $\alpha$ in the range $0$ -- $6$.
+The top axis I shows the mean spin magnitude
+that value of alpha corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
+We see that if typical BH spins are $\lesssim 0.2$ we have no evidence for one model over the other.
+
 \section{Hierarchical Modelling} 

### 2017-03-29 paper/AlignedIsotropicGW.tex
Fixed conflicts between mine and Ilyas edits to mixture model section (mostly kept Ilyas)

@@
 
-%As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
-The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.
-%
-%Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
-In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.
-%
-%In this section we assume that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).
-\sout{As discussed above, BBHs can form as the end point of isolated binary evolution or dynamically in dense stellar environments.
-The spin misalignments of BBHs is a signature of their formation mechanism; binaries formed through isolated binary evolution are expected to have their spins aligned with the orbital angular momentum, whilst those formed dynamically are expected to have an isotropic distribution.}
+While we carried out Bayesian comparisons between isotropic and aligned spin distributions under various assumptions, a preference for one of the considered models over the others does not necessarily indicate that it is the correct model.
+All of the considered models could be inaccurate for the actual distribution, especially since all of the considered models are based on a number of additional assumptions, such as decoupled spin magnitude and spin misalignment angle distributions and identical distributions for primary and secondary spins.
 
-\sout{Rate uncertainties for both of these formation scenarios are large enough that we do not know a priori which will dominate the LIGO detection rate.
-In the previous section we have shown that a model with isotropic spin directions and 'decreasing' spin magnitudes is favoured by the data.}
-
-In this section we assume that the true distribution of
-\acp{BBH} observed by LIGO is a mixture of binaries formed
-through isolated binary evolution (with aligned spins) and
-dynamical formation (with isotropic spins). Rate uncertainties
-for both of these formation scenarios are large enough that we
-do not know a priori which will dominate the LIGO detection
-rate.
-\ilya{While we carried out Bayesian comparisons between isotropic and aligned spin distributions under various assumptions, a preference for one of the considered models over the others does not necessarily indicate that it is the correct model.
-All of the considered models could be inaccurate for the actual distribution, especially since all of the considered models are based on a number of additional assumptions, such as decoupled spin magnitude and spin misalignment angle distributions and identical distributions for primary and secondary spins.}
-\ilya{We now partly relax the simplified assumptions made earlier by considering the possibility that the} true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries \ilya{with aligned spins and binaries with isotropic spins}.
-\sout{formed through isolated binary evolution (with aligned spins) and dynamical formation (with isotropic spins).}
+We now partly relax the simplified assumptions made earlier by considering the possibility that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries with aligned spins and binaries with isotropic spins.
 
@@
 We calculate and plot the posterior on $f_i$ given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in the derivation) in Figure~\ref{fig:mixture_fraction_posterior}.
-We fit a mixture model (labelled model 'M') where a fraction $f_i$ ($f_i=\lambda$ in the derivations above) of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
-We assume a flat prior on the fraction $f_i$.
-To test the robustness of our result, we vary the distribution we assume for BH spin magnitude distributions as with the aligned and isotropic models.
-We use the `uniform', `increasing' and `decreasing' distributions introduced in Section X, assuming all BHs have their spin magnitude drawn from the same distribution.
-We calculate the posterior given by Equation \ref{eq:magnitude-dists} plot the posterior distribution/cumulative posterior in Figure~\ref{fig:mixture_fraction_posterior}.
-We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the `decreasing', `uniform' and `increasing' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
-The lower 90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the prior of 0.1.
-Thus, \ilya{for these spin magnitude distributions} 
-\sout{irrespective of our ignorance of the BH spin magnitude distribution}
-we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
-
 We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high'' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
@@
 In all cases, the posterior peaks at $f_i = 1$.
-Thus, for a range of assumptions regarding the the BH spin magnitude distribution, we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
-\sout{If we assume that the subpopulation of BBHs with isotropically distributed spin orientations corresponds to a subpopulation of dynamically formed BBHs, our result suggests that the majority of merging BBHs observed by LIGO are formed dynamically, rather than through isolated binary evolution.}
-
+Thus, for these spin magnitude distributions we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
 The evidence ratios of these models to the isotropic distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for the ``low'', ``flat'' and ``high'' spin magnitude models.
@@
 In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of aligned to isotropic distributions (plotted as the equivalent sigma) with spin magnitudes given by this model with $\alpha$ in the range $0$ -- $6$.
-The top axis I shows the mean spin magnitude
+The top axis shows the mean spin magnitude
 that value of alpha corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).

### 2017-03-29 paper/AlignedIsotropicGW.tex
fixed issue with sigma plot

@@
 \includegraphics[width=0.45\textwidth]{../plots/sigma_v_alpha.png}
-\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
-The top axis shows the mean spin magnitude $a$
-that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
+\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma $\sigma_\mathrm{A/I}$) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
+The top axis shows the mean spin magnitude $\bar{a}$
+that the value of $\alpha$ corresponds to.
+$\alpha = 0$ corresponds to the ``flat'' spin magnitude model, $\alpha = 1$ corresponds to the ``low'' model.
+We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
 \label{fig:smallspinsalpha}

### 2017-03-29 paper/AlignedIsotropicGW.tex
Implemented Ilya's 'red' text.

--- 2017-03-29 paper/AlignedIsotropicGW.tex
@@
   binaries formed from pairs of stars born together are more likely to
-  have spins \ilya{preferentially} aligned with the binary orbit as a consequence of their
-  joint evolution toward a \ac{BBH} system.  We consider existing
-  \ac{GW} measurements of $\chieff$, the best-measured combination of
-  spin parameters, in the three likely binary black hole detections
-  GW150914, LVT151012, and GW151226.
-  \ilya{If binary black hole spin magnitudes extend to high values, as is
-suggested by observations of black hole X-ray binaries,} we show that the data already
-  exhibit a $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
-  ratio) preference for an isotropic angular distribution.
-  By considering the
-  effect of an additional 10 detections drawn from the various models
-  in the suite we show that if all observations come from a single
-  population such an augmented data set would enable at least a
-  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
-  ratio) distinction between the isotropic and aligned models \ilya{for the assumed spin magnitude distributions}, and in
-  most cases better than $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).
-  The existing preference for \ilya{either an isotropic spin distribution or low spin magntiudes} for the observed systems will be confirmed (or overturned) confidently in the near future by subsequent observations.
+  have spins preferentially aligned with the binary orbit as a
+  consequence of their joint evolution toward a \ac{BBH} system.  We
+  consider existing \ac{GW} measurements of $\chieff$, the
+  best-measured combination of spin parameters, in the three likely
+  binary black hole detections GW150914, LVT151012, and GW151226.  If
+  binary black hole spin magnitudes extend to high values, as is
+  suggested by observations of black hole X-ray binaries, we show that
+  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
+  ($\OOneOddsIsoAligned$ odds ratio) preference for an isotropic
+  angular distribution.  By considering the effect of an additional 10
+  detections drawn from the various models in the suite we show that
+  if all observations come from a single population such an augmented
+  data set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
+  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
+  isotropic and aligned models for the assumed spin magnitude
+  distributions, and in most cases better than $5\sigma$
+  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
+  either an isotropic spin distribution or low spin magntiudes for the
+  observed systems will be confirmed (or overturned) confidently in
+  the near future by subsequent observations.
 \end{abstract}
@@
 
-Small values of $\chieff$ can result from either intrinsically small
-spins or larger spins whose direction is mis-aligned with the orbital
-plane of the binary (i.e.\ spin vectors with small $z$-components).
-Mis-alignment is capable of producing \emph{negative} values of
-$\chieff$, however, while small, aligned spins will always have
-$\chieff \geq 0$.  This difference provides strong discriminating
-power between the two angular distributions, even without good
-information about the magnitude distribution; to the extent that data
-support negative $\chieff$ they weigh heavily against aligned models.
-To quantify the degree of support for these two alternate explanations
-of small $\chieff$ values in the merging \ac{BBH} population, we
-compared the Bayesian evidence for various simple models of the spin
-population using the \ac{GW} data set.
+Small values of $\chieff$ as exhibited in these systems can result
+from either intrinsically small spins or larger spins whose direction
+is mis-aligned with the orbital plane of the binary (i.e.\ spin
+vectors with small $z$-components).  Mis-alignment is capable of
+producing \emph{negative} values of $\chieff$, however, while small,
+aligned spins will always have $\chieff \geq 0$.  This difference
+provides strong discriminating power between the two angular
+distributions, even without good information about the magnitude
+distribution; to the extent that data support negative $\chieff$ they
+weigh heavily against aligned models.  To quantify the degree of
+support for these two alternate explanations of small $\chieff$ values
+in the merging \ac{BBH} population, we compared the Bayesian evidence
+for various simple models of the spin population using the \ac{GW}
+data set.
 
@@
 systems, the most-favoured model among those with an isotropic angular
-distribution has the ``decreasing'' magnitude distribution; the most
-favoured model among those with an aligned distribution also has the
-``decreasing'' magnitude distribution.  The odds ratio between the
-aligned and isotropic models is $\OOneOddsIsoAligned$, or
+distribution has the ``low'' magnitude distribution; the most favoured
+model among those with an aligned distribution also has the ``low''
+magnitude distribution.  The odds ratio between the aligned and
+isotropic models is $\OOneOddsIsoAligned$, or
 $\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
@@
 be distinguished with greater than $5\sigma$ ($2.9 \times 10^{-7}$
-odds) credibility; and in the worst case the distinction is
-$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
-Note that such future detections should permit a confident distinction
-between \emph{angular} distributions, we will remain very uncertain
-about the \emph{magnitude} distribution until we have a much larger
-number of observations.  In Figure \ref{fig:O2-predictions}, the odds
-ratio between different magnitude distributions with the same angular
-distribution is much closer to one than the odds ratio between angular
-distributions.
+odds) credibility; and in the most pessisimistic case the distinction
+is typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
+odds ratio).  While such future detections should permit a confident
+distinction between \emph{angular} distributions, we will remain very
+uncertain about the \emph{magnitude} distribution until we have a much
+larger number of observations.  In Figure \ref{fig:O2-predictions},
+the odds ratio between different magnitude distributions with the same
+angular distribution is much closer to one than the odds ratio between
+angular distributions.
 
@@
 $\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
-spins are intrinsically very small, with $a \lesssim \ilya{0.3}$, \ilya{[I changed the number to be consistent with Simon's analysis, which shows that you can no longer tell the difference for a mean spin of $\sim 0.17$]} then
-it is no longer possible to resolve the negative spin with a small
-number of observations.  As noted below, however, spins observed in
-X-ray binaries are typically large. 
-\sout{and most models of isolated binary evolution do not produce systems with spins uniformly bounded to small values}.  
-Additionally, models which do not permit \emph{some} spins
-with $\chieff \gtrsim \ilya{0.1}$ \ilya{[Changed to be consistent with 90\% lower bound from cited paper]} are ruled out by the GW151226 observations
-\citep{2016PhRvL.116x1103A}.
+spins are intrinsically very small, with $a \lesssim 0.3$, then it is
+no longer possible to resolve the negative spin with a small number of
+observations.  As noted below, however, spins observed in X-ray
+binaries are typically large.  Additionally, models which do not
+permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
+the GW151226 observations \citep{2016PhRvL.116x1103A}.
 
 In order to perform our analysis we need to select models for the
-distribution of the \ilya{spin magnitudes} of stellar-mass
-black holes.  Observational data for such model selection is sparse.
+distribution of the spin magnitudes of stellar-mass black holes.
+Observational data for such model selection is sparse.
 \citet{2015PhR...548....1M} give current estimates of the spin
@@
 complicate either type of analysis (see \citealt{2015PhR...548....1M}
-for a discussion), \ilya{and observational selection effects could yield a biased distribution}. Nonetheless, if we take the reported spin
-magnitudes as representative then we find that there is a preference
-for high spins; for example, 14 of the 19 systems with reported spins
-have dimensionless spin parameters in excess of 0.5.  \ilya{It is usually argued that} the masses and spin parameters
-of stellar-mass black holes are unlikely to be altered significantly
-by accretion (low-mass donors may not have enough mass and
-high-mass donors have a very short phase in which they transfer mass;
-a variant of this long-standing argument was presented by
-\citet{1999MNRAS.305..654K}), \ilya{but see \citet{Podsiadlowski:2003,FragosMcClintock:2015}}.  Thus the current spin parameters
-probably are close to their values upon core collapse.  However, the
-specific processes involved in the production of black hole binaries \ilya{from isolated binaries}
-could alter the spin \ilya{magnitude} distribution of those holes \ilya{relative to the X-ray binary systems}; for example, close
-tidal interactions could spin up the core, or stripping of the
-envelope could reduce the available angular momentum \citep{2016MNRAS.462..844K,Zaldarriaga:2017,HotokezakaPiran:2017}.
+for a discussion), and selection effects could yield a biased
+distribution. Nonetheless, if we take the reported spin magnitudes as
+representative then we find that there is a preference for high spins;
+for example, 14 of the 19 systems with reported spins have
+dimensionless spin parameters in excess of 0.5.  It is usually argued
+that the masses and spin parameters of stellar-mass black holes are
+unlikely to be altered significantly by accretion (low-mass donors may
+not have enough mass and high-mass donors have a very short phase in
+which they transfer mass; a variant of this long-standing argument was
+presented by \citet{1999MNRAS.305..654K}), but see
+\citet{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current
+spin parameters probably are close to their values upon core collapse.
+However, the specific processes involved in the production of black
+hole binaries from isolated binaries could alter the spin magnitude
+distribution of those holes relative to the X-ray binary systems; for
+example, close tidal interactions could spin up the core, or stripping
+of the envelope could reduce the available angular momentum
+\citep{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
 
@@
 
-\ilya{The spin directions in isolated binaries, whether evolving via the classical channel through a common-envelope phase \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017} or through chemically homogeneous evolution \citep{MandeldeMink:2016,Marchant:2016} are generally expected to be preferentially aligned.
-Despite  observed spin misalignments in both massive stellar binaries \citep{Albrecht:2009,2014ApJ...785...83A} and BH XRBs \citep{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014}, mass accretion and tidal interactions will tend to realign the binary.
-On the other hand, a supernova natal kick (if any) can change the orbital plane and misalign the binary \citep{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can also tilt the spin angle, as in the double pulsar J0737-3039 \citep{2011ApJ...742...81F}; and a variety of uncertain processes, such as wind-driven mass loss and post-collapse fallback, can couple the spin magnitude and direction distributions, contrary to our simplified assumptions.}
+The spin directions in isolated binaries, whether evolving via the
+classical channel through a common-envelope phase
+\citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
+or through chemically homogeneous evolution
+\citep{MandeldeMink:2016,Marchant:2016} are generally expected to be
+preferentially aligned.  Despite observed spin misalignments in both
+massive stellar binaries \citep{Albrecht:2009,2014ApJ...785...83A} and
+BH XRBs
+\citep{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
+mass accretion and tidal interactions will tend to realign the binary.
+On the other hand, a supernova natal kick (if any) can change the
+orbital plane and misalign the binary
+\citep{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can
+also tilt the spin angle, as in the double pulsar J0737-3039
+\citep{2011ApJ...742...81F}; and a variety of uncertain processes,
+such as wind-driven mass loss and post-collapse fallback, can couple
+the spin magnitude and direction distributions, contrary to our
+simplified assumptions.
 
-\ilya{\citet{2017CQGra..34cLT01V} also studied the possibility of distinguishing aligned and isotropic angular distributions of \ac{BBH} spins, but concluded that several hundred sources would be required to adequately separate models which included both parallel and anti-parallel spins for 'aligned' population.
-In contrast, and in agreement with the present study, \citet{Stevenson:2017spin} found that only $\sim 5$ observations of rapidly spinning black holes would be necessary to distinguish isotropic and aligned spin distributions, though tens of detections would be required for more nuanced admixture models.
-Meanwhile, recent studies by \citet{Fishbach:2017,Gerosa:2017} focused on inference on spin magnitudes.}
+\citet{2017CQGra..34cLT01V} also studied the possibility of
+distinguishing aligned and isotropic angular distributions of \ac{BBH}
+spins, but concluded that several hundred sources would be required to
+adequately separate models which included both parallel and
+anti-parallel spins for an ``aligned'' population.  In contrast, and
+in agreement with the present study, \citet{Stevenson:2017spin} found
+that only $\sim 5$ observations of rapidly spinning black holes would
+be necessary to distinguish isotropic and aligned spin distributions,
+though tens of detections would be required for more nuanced admixture
+models.  Meanwhile, recent studies by
+\citet{2017arXiv170306869F,2017arXiv170306223G} focused on inference
+on spin magnitudes.
 
-\ilya{In summary}, the angular distribution of spins in \acp{BBH} formed through isolated binary evolution is uncertain; nevertheless, it is more probable that spins are \ilya{preferentially} aligned \sout{and non-zero} after evolution through this channel than not, while an isotropic distribution of \sout{non-zero} spins is the natural outcome of dynamical formation processes.
-\ilya{Therefore, if the current observational trend for low $\chieff$ continues with future gravitational-wave observations, it will be possible to either confirm that \ac{BH} spins are sometimes large in magnitude and isotropic in direction, yielding a strong indication of a dynamical formation origin; or that \ac{BH} spins are overwhelmingly small in magnitude, yielding a notable contradiction with the claimed X-ray binary spin measurements, particularly those for high-mass X-ray binaries.}
+In summary, the angular distribution of spins in \acp{BBH} formed
+through isolated binary evolution is uncertain; nevertheless, it is
+more probable that spins are preferentially aligned after evolution
+through this channel than not, while an isotropic distribution of
+spins is the natural outcome of dynamical formation processes.
+Therefore, if the current observational trend for low $\chieff$
+continues with future gravitational-wave observations, it will be
+possible to either confirm that \ac{BH} spins are sometimes large in
+magnitude and isotropic in direction, yielding a strong indication of
+a dynamical formation origin; or that \ac{BH} spins are overwhelmingly
+small in magnitude, yielding a notable contradiction with the claimed
+X-ray binary spin measurements, particularly those for high-mass X-ray
+binaries.
 
@@
 Thus we cannot rule out a mixture with the current data.
-
-\section{Effect of small spin magnitudes}
-\label{sec:smallspins}
-
-In the main text we considered three models for BH spin magnitudes: ``low'', ``flat'' and ``high''. These were intended to capture some of the uncertainty regarding the BH spin magnitude distribution.
-
-Here we extend the ``low'' model as:
-%
-\begin{equation}
-p(a) \propto (1 - a)^{\alpha}
-\label{eq:lowspinalpha}
-\end{equation}
-%
-
-When $\alpha = 0$, this recovers the ``flat'' distribution, whilst $\alpha = 1$
-recovers the ``low'' distribution. For higher values of $\alpha$, this distribution 
-becomes more peaked towards $a = 0$.
-
-%
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{../plots/sigma_v_alpha.png}
-\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
-The top axis shows the mean spin magnitude $a$
-that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
-\label{fig:smallspinsalpha}
-\end{figure}
-%
-
-In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of aligned to isotropic distributions (plotted as the equivalent sigma) with spin magnitudes given by this model with $\alpha$ in the range $0$ -- $6$.
-The top axis shows the mean spin magnitude
-that value of alpha corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
-We see that if typical BH spins are $\lesssim 0.2$ we have no evidence for one model over the other.
 
@@
 the odds ratio becomes $150$, or $2.7\sigma$.  With small spin
-magnitudes, our angular resolving power vanishes.
+magnitudes, our angular resolving power vanishes, as discussed in more
+detail in Section \ref{sec:smallspins}.
+
+\section{Effect of small spin magnitudes}
+\label{sec:smallspins}
+
+In the main text we considered three models for BH spin magnitudes: ``low'', ``flat'' and ``high''. These were intended to capture some of the uncertainty regarding the BH spin magnitude distribution.
+
+Here we extend the ``low'' model as:
+%
+\begin{equation}
+p(a) \propto (1 - a)^{\alpha}
+\label{eq:lowspinalpha}
+\end{equation}
+%
+
+When $\alpha = 0$, this recovers the ``flat'' distribution, whilst $\alpha = 1$
+recovers the ``low'' distribution. For higher values of $\alpha$, this distribution 
+becomes more peaked towards $a = 0$.
+
+%
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{../plots/sigma_v_alpha.png}
+\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
+The top axis shows the mean spin magnitude $a$
+that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
+\label{fig:smallspinsalpha}
+\end{figure}
+%
+
+In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of aligned to isotropic distributions (plotted as the equivalent sigma) with spin magnitudes given by this model with $\alpha$ in the range $0$ -- $6$.
+The top axis shows the mean spin magnitude
+that value of alpha corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
+We see that if typical BH spins are $\lesssim 0.2$ we have no evidence for one model over the other.
 

### 2017-03-29 paper/AlignedIsotropicGW.tex
Ilya's requested small changes.

@@
   the data already exhibit a $\OOneSigmaIsoAligned\sigma$
-  ($\OOneOddsIsoAligned$ odds ratio) preference for an isotropic
-  angular distribution.  By considering the effect of an additional 10
-  detections drawn from the various models in the suite we show that
-  if all observations come from a single population such an augmented
-  data set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
+  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
+    $r \ll 1$ is equivalent to $x \sigma$ with
+    $x = \Phi^{-1}\left( 1 - r \right)$, where $\Phi$ is the unit
+    normal CDF.})  preference for an isotropic angular distribution.
+  By considering the effect of an additional 10 detections drawn from
+  the various models in the suite we show that if all observations
+  come from a single population such an augmented data set would
+  enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
   ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
@@
   ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
-  either an isotropic spin distribution or low spin magntiudes for the
+  either an isotropic spin distribution or low spin magnitudes for the
   observed systems will be confirmed (or overturned) confidently in
@@
 \citep{2015PhRvD..91d2003V} compare model gravitational waveforms
-\citep{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
+\citep[e.g.\
+][]{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
 against the observed data to obtain a posterior distribution on the
-parameter space that describes the compact binary source.
+parameters that describe the compact binary source.
 
@@
   \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
-  \right) \cdot \frac{\vec{L}}{\hat{L}} = \frac{1}{M} \left( m_1 a_1^z + m_2 a_2^z \right),
+  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} = \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
 \end{equation}
@@
 holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
-and $0 \leq a_{1,2} \leq 1$ are the corresponding dimensionless spin
-parameters \citep{2016PhRvL.116x1102A}.
+and $0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
+projections of the individual \ac{BH} spins
+\citep{2016PhRvL.116x1102A}.
 
@@
 in \citet{O1-BBH}.  There is essentially no posterior support for
-$\chieff \gtrsim 0.5$ as observed in the majority of the
+$\chi \gtrsim 0.5$ as observed in the majority of the
 electromagnetically-detected stellar-mass black hole population (see
@@
 vectors with small $z$-components).  Mis-alignment is capable of
-producing \emph{negative} values of $\chieff$, however, while small,
-aligned spins will always have $\chieff \geq 0$.  This difference
-provides strong discriminating power between the two angular
-distributions, even without good information about the magnitude
-distribution; to the extent that data support negative $\chieff$ they
-weigh heavily against aligned models.  To quantify the degree of
-support for these two alternate explanations of small $\chieff$ values
-in the merging \ac{BBH} population, we compared the Bayesian evidence
-for various simple models of the spin population using the \ac{GW}
-data set.
+producing \emph{negative} values of $\chieff$, however, while aligned
+spins will always have $\chieff \geq 0$.  This difference provides
+strong discriminating power between the two angular distributions,
+even without good information about the magnitude distribution; to the
+extent that data favour negative $\chieff$ they weigh heavily against
+aligned models.  To quantify the degree of support for these two
+alternate explanations of small $\chieff$ values in the merging
+\ac{BBH} population, we compared the Bayesian evidence for various
+simple models of the spin population using the \ac{GW} data set.
 
@@
 assume that the population spin distribution factorises into a
-distribution for the spin magnitude, $a$, and a distribution for the
+distribution for the spin magnitude $a$ and a distribution for the
 spin angles.  Finally, we assume that the distribution of spins is
@@
     we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$ and
-    that the same distribution of spin vectors obtains for each
-    component.  The ``flat'' (blue lines), ``high,'' (green
-    lines), and ``low'' (red lines) magnitude distributions are
-    defined in Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the
-    $\chieff$ distribution under the assumption that the orientations
-    of the spins are isotropic; dashed lines give the distribution
-    under the assumption that both objects' spins are aligned with the
-    orbital angular momentum.  The isotropic distributions are readily
+    that the distribution of spin vectors is the same for each
+    component.  The ``flat'' (blue lines), ``high,'' (green lines),
+    and ``low'' (red lines) magnitude distributions are defined in
+    Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the $\chieff$
+    distribution under the assumption that the orientations of the
+    spins are isotropic; dashed lines give the distribution under the
+    assumption that both objects' spins are aligned with the orbital
+    angular momentum.  The isotropic distributions are readily
     distinguished from the aligned by the production of negative
@@
 model among those with an aligned distribution also has the ``low''
-magnitude distribution.  The odds ratio between the aligned and
-isotropic models is $\OOneOddsIsoAligned$, or
+magnitude distribution.  The odds ratio between the ``low'' aligned
+and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
 $\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
@@
     and low (``L'') spin magnitude distributions (see Eq.\
-    \eqref{eq:magnitude-dists}) are paired with isotropic, aligned
-    angular distributions (``$\mathrm{I}$'', ``$\mathrm{A}$''), as
-    well as a mixture model of the two (``$\mathrm{M}$'').  The
-    most-favoured models have a decreasing distribution of spin
-    magnitudes.  The odds ratio between these models is
-    $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.  For all
-    magnitude distributions the pure-isotropic models are preferred
-    over the mixture models; correspondingly, the posterior on the
-    mixture fraction peaks at 100\% isotropic.}
+    \eqref{eq:magnitude-dists}) are paired with isotropic (``I'') and
+    aligned (``A'') angular distributions, as well as a mixture model
+    of the two (``$\mathrm{M}$'').  The most-favoured models have the
+    ``low'' distribution of spin magnitudes.  The odds ratio between
+    these models is $\OOneOddsIsoAligned$, or
+    $\OOneSigmaIsoAligned\sigma$.  For all magnitude distributions the
+    pure-isotropic models are preferred over the mixture models;
+    correspondingly, the posterior on the mixture fraction peaks at
+    100\% isotropic.}
   \label{fig:O1-odds}
@@
 Estimates of the rate of \ac{BBH} coalescences give a reasonable
-chance of 10 additional \ac{BBH} detections in the near future
+chance of 10 additional \ac{BBH} detections in the next two years
 \citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming an
-additional 10 detections drawn from each of our six zero-parameter
-models for the spin distribution, we find the odds ratios shown in
-Figure \ref{fig:O2-predictions}.  We find that most scenarios with an
-additional 10 detections allow the the correct angular distribution to
-be distinguished with greater than $5\sigma$ ($2.9 \times 10^{-7}$
-odds) credibility; and in the most pessisimistic case the distinction
-is typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
-odds ratio).  While such future detections should permit a confident
-distinction between \emph{angular} distributions, we will remain very
-uncertain about the \emph{magnitude} distribution until we have a much
-larger number of observations.  In Figure \ref{fig:O2-predictions},
-the odds ratio between different magnitude distributions with the same
-angular distribution is much closer to one than the odds ratio between
-angular distributions.
+additional 10 detections with similar observational uncertainties
+drawn from each of our six zero-parameter models for the spin
+distribution in addition to the three existing detections from O1, we
+find the odds ratios shown in Figure \ref{fig:O2-predictions}.  We
+find that most scenarios with an additional 10 detections allow the
+correct angular distribution to be distinguished with greater than
+$5\sigma$ ($2.9 \times 10^{-7}$ odds) credibility; and in the most
+pessimistic case the distinction is typically
+$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
+While such future detections should permit a confident distinction
+between \emph{angular} distributions, we will remain much less certain
+about the \emph{magnitude} distribution until we have a larger number
+of observations.  In Figure \ref{fig:O2-predictions}, the odds ratio
+between different magnitude distributions with the same angular
+distribution is much closer to one than the odds ratio between angular
+distributions.
 
@@
 \citep{MandeldeMink:2016,Marchant:2016} are generally expected to be
-preferentially aligned.  Despite observed spin misalignments in both
-massive stellar binaries \citep{Albrecht:2009,2014ApJ...785...83A} and
-BH XRBs
+preferentially aligned.  Despite observed spin-orbit misalignments in
+both massive stellar binaries
+\citep{Albrecht:2009,2014ApJ...785...83A} and BH XRBs
 \citep{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
@@
 \caption{\textbf{Fraction of the BBH population coming from an
-isotropic distribution} The dotted line shows the flat
-prior on the fraction of BBHs coming from an isotropic
-distribution $f_i$ under the mixture model. The 3 red lines show
-the posterior on $f_i$ after O1 under with our various
-assumptions regarding BH spin magnitudes.
-The solid line shows the posterior assuming that all BHs have
-their spin magnitude drawn
-from the ``flat'' distribution. The dashed line assumes the
-``high'' BH spin magnitude distribution $p(a) = 2a$. The
-dot-dash line assumes the ``low'' distribution $p(a) = 2(1-a)$.
-We see that for a wide range of assumptions regarding BH spin
-magnitudes, the fraction coming from an isotropic distribution
-$f_i$ peaks at 1.}
+    isotropic distribution under a mixture model.} The dotted line
+  shows the flat prior on the fraction of BBHs coming from an
+  isotropic distribution, $f_i$, under the mixture model. The 3 red
+  lines show the posterior on $f_i$ after O1 with our various
+  assumptions regarding BH spin magnitudes.  The solid line shows the
+  posterior assuming that all BHs have their spin magnitude drawn from
+  the ``flat'' distribution. The dashed line assumes the ``high'' BH
+  spin magnitude distribution $p(a) = 2a$. The dot-dash line assumes
+  the ``low'' distribution $p(a) = 2(1-a)$.  We see that for a wide
+  range of assumptions regarding BH spin magnitudes, the fraction
+  coming from an isotropic distribution $f_i$ peaks at 1.}
 \label{fig:mixture_fraction_posterior}
@@
 
-Following \citet{Stevenson:2017spin},  we fit a mixture model (labelled model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have spins drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have their spins aligned with the orbital angular momentum.
-We assume a flat prior on the fraction $f_i$.
-To test the robustness of our result, we vary the distribution we assume for BH spin magnitude distributions as with the aligned and isotropic models.
-We use the ``flat'', ``high'' and ``low'' distributions (Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin magnitude drawn from the same distribution for both the aligned and isotropic populations.
-We calculate and plot the posterior on $f_i$ given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in the derivation) in Figure~\ref{fig:mixture_fraction_posterior}.
-We find the mean fraction of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high'' distributions for spin magnitudes respectively, compared to the prior mean of 0.5.
-The lower 90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the prior of 0.1.
-In all cases, the posterior peaks at $f_i = 1$.
-Thus, for these spin magnitude distributions we find that the current O1 LIGO observations constrain the majority of BBHs to have their spins drawn from an isotropic distribution.
-The evidence ratios of these models to the isotropic distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for the ``low'', ``flat'' and ``high'' spin magnitude models.
-Thus we cannot rule out a mixture with the current data.
+Following \citet{Stevenson:2017spin}, we fit a mixture model (labelled
+model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs
+have spins drawn from an isotropic distribution, whilst a fraction
+$1 - f_i$ have their spins aligned with the orbital angular
+momentum. We assume a flat prior on the fraction $f_i$. To test the
+robustness of our result, we vary the distribution we assume for BH
+spin magnitude distributions as with the aligned and isotropic
+models. We use the ``flat'', ``high'' and ``low'' distributions
+(Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin
+magnitude drawn from the same distribution for both the aligned and
+isotropic populations. We calculate and plot the posterior on $f_i$
+given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in
+the derivation) in Figure~\ref{fig:mixture_fraction_posterior}. We
+find the mean fraction of BBHs coming from an isotropic distribution
+is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high''
+distributions for spin magnitudes respectively, compared to the prior
+mean of 0.5. The lower 90\% limits are 0.26, 0.39 and 0.52
+respectively, compared to the prior of 0.1. In all cases, the
+posterior peaks at $f_i = 1$. Thus, for these spin magnitude
+distributions we find that the current O1 LIGO observations constrain
+the majority of BBHs to have their spins drawn from an isotropic
+distribution. The evidence ratios of these models to the isotropic
+distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for
+the ``low'', ``flat'' and ``high'' spin magnitude models. Thus we
+cannot rule out a mixture with the current data.
 
@@
 mass ratio.  As an example, the distinction between the three
-different spin amplitude distributions under ten additional detections
+different spin amplitude distributions after ten additional detections
 is quite weak compared to the aligned/isotropic distinction in Figure

### 2017-03-29 paper/AlignedIsotropicGW.tex
As submitted to DCC.
Back to red "checkme" numbers now that DCC submission has been generated.

@@
 \newcommand{\checkme}[1]{\textcolor{red}{#1}}
+%\newcommand{\checkme}[1]{ #1 }
 
@@
 synthesis codes can at this point authoritatively suggest {\it any}
-particular spin distribution.  Our models, however, allow us to see
-how sensitive the $\chieff$ distribution is to spin alignment given
-uncertainties about the spin magnitudes.
+particular spin distribution \citep{2015PhR...548....1M}.  Our models,
+however, allow us to see how sensitive the $\chieff$ distribution is
+to spin alignment given uncertainties about the spin magnitudes.
 

### 2017-03-29 paper/AlignedIsotropicGW.tex
Minor tweaks (address).

@@
   Birmingham, B15 2TT, United Kingdom}
-\affiliation{Kavli Institute of Theoretical Physics, University of California, Santa Barbara}
+\affiliation{Kavli Institute for Theoretical Physics, Santa Barbara, CA 93106}
 
@@
   Birmingham, B15 2TT, United Kingdom}
- \affiliation{Kavli Institute of Theoretical Physics, University of California, Santa Barbara}
+\affiliation{Kavli Institute for Theoretical Physics, Santa Barbara, CA 93106}
 
@@
 small in magnitude, yielding a notable contradiction with the claimed
-X-ray binary spin measurements, particularly those for high-mass X-ray
+\ac{BH} X-ray binary spin measurements, particularly those for high-mass X-ray
 binaries.

### 2017-03-30 paper/AlignedIsotropicGW.tex
DCC v2 with chi_eff bound bug fixed.
Back to red 'checkme' text after DCC posting.
Resized Figs 5 and 6

@@
 \ac{BBH} systems having zero spin is ruled out by the GW151226
-measurements, which bound $\chieff \gtrsim 0.2$ at 90\% credibility
+measurements, which bound $\chieff \gtrsim 0.1$ at 90\% credibility
 \citep{2016PhRvL.116x1103A}.
@@
 \centering
-\includegraphics[width=0.45\textwidth]{../plots/posterior_on_isotropic_fraction.png}
+\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
 \caption{\textbf{Fraction of the BBH population coming from an
@@
 \centering
-\includegraphics[width=0.45\textwidth]{../plots/sigma_v_alpha.png}
+\includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
 \caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.

### 2017-03-30 paper/AlignedIsotropicGW.tex
latexed an alpha

@@
 The top axis shows the mean spin magnitude
-that value of alpha corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
+that value of $\alpha$ corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
 We see that if typical BH spins are $\lesssim 0.2$ we have no evidence for one model over the other.

### 2017-03-30 paper/AlignedIsotropicGW.tex
Fixed #4.

@@
 \ac{BBH} systems having zero spin is ruled out by the GW151226
-measurements, which bound $\chieff \gtrsim 0.1$ at 90\% credibility
-\citep{2016PhRvL.116x1103A}.
+measurements, which bound at least one black hole to have spin
+magnitude $\geq 0.2$ at 90\% credibility \citep{2016PhRvL.116x1103A}.
 

### 2017-04-05 paper/AlignedIsotropicGW.tex
Update AlignedIsotropicGW.tex
Very small changes; added Kulkarni, Hut, and McMillan 1993 to references regarding black hole dynamics in dense stellar environments

@@
 where the $\hat{z}$ axis is aligned with the orbital angular momentum
-of the binary, $m_{1,2}$ are the masses of the more-massive (1) and
+of the binary, $m_{1,2}$ are the gravitational masses of the more-massive (1) and
 less-massive (2) components, $M = m_1 + m_2$ is the total mass,
@@
 in \citet{O1-BBH}.  There is essentially no posterior support for
-$\chi \gtrsim 0.5$ as observed in the majority of the
+$\chi \gtrsim 0.5$ as inferred for the majority of the
 electromagnetically-detected stellar-mass black hole population (see
@@
 insensitive to the precise details of the posterior distributions;
-other conclusions are more sensitive.  In particular, our
+other conclusions are more sensitive.  In particular, our Gaussian
 approximation does permit $\chieff = 0$ for GW151226 while the true
@@
 vectors with small $z$-components).  Mis-alignment is capable of
-producing \emph{negative} values of $\chieff$, however, while aligned
+producing \emph{negative} values of $\chieff$, however, whereas aligned
 spins will always have $\chieff \geq 0$.  This difference provides
@@
     angular momentum.  The isotropic distributions are readily
-    distinguished from the aligned by the production of negative
+    distinguished from the aligned distributions by the production of negative
     $\chieff$ values, while the distinction between the three models
@@
 between different magnitude distributions with the same angular
-distribution is much closer to one than the odds ratio between angular
+distribution is much closer to unity than the odds ratio between angular
 distributions.
@@
 interactions in dense stellar environments
-\citep{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+\citep{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
 are expected to be isotropic given the absence of a preferred
@@
 or through chemically homogeneous evolution
-\citep{MandeldeMink:2016,Marchant:2016} are generally expected to be
+\citep{MandeldeMink:2016,Marchant:2016} are usually expected to be
 preferentially aligned.  Despite observed spin-orbit misalignments in

### 2017-04-11 paper/AlignedIsotropicGW.tex
Commented on abstract, intro.

@@
 
-\email{w.farr@bham.ac.uk,simons@star.sr.bham.ac.uk,miller@astro.umd.edu,imandel@star.sr.bham.ac.uk,av@star.sr.bham.ac.uk}
+\email{w.farr@bham.ac.uk, simons@star.sr.bham.ac.uk, miller@astro.umd.edu,\\imandel@star.sr.bham.ac.uk, av@star.sr.bham.ac.uk}
 
@@
   The first direct detections of \acp{GW} from merging \acp{BBH} open
-  a unique window into the formation environment of massive stars and
-  their compact remnants.  One promising signature of the formation
+  a unique window into the \ilya{\ac{BBH}} formation environment.  \sout{of massive stars and
+  their compact remnants}  One promising signature of the formation
   environment is the angular distribution of the \ac{BH} spins;
@@
     $x = \Phi^{-1}\left( 1 - r \right)$, where $\Phi$ is the unit
-    normal CDF.})  preference for an isotropic angular distribution.
+    normal CDF. \ilya{[Two-tailed vs.~one-tailed]}})  preference for an isotropic angular distribution.
   By considering the effect of an additional 10 detections drawn from
@@
   \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
-  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} = \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
+  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \ilya{\equiv} \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
 \end{equation}
-where the $\hat{z}$ axis is aligned with the orbital angular momentum
-of the binary, $m_{1,2}$ are the gravitational masses of the more-massive (1) and
+where \sout{the $\hat{z}$ axis is aligned with the orbital angular momentum
+of the binary,} $m_{1,2}$ are the gravitational masses of the more-massive (1) and
 less-massive (2) components, $M = m_1 + m_2$ is the total mass,
 $\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
-holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
+holes in the binary, $\vec{L}$ is the orbital angular momentum vector \ilya{assumed to point in the $\hat{z}$ direction},
 and $0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
@@
 electromagnetically-detected stellar-mass black hole population (see
-Section \ref{sec:discussion}).  The analysis here is relatively
+Section \ref{sec:discussion}).
+\ilya{[Figure 1 shows an absence of support for $\chieff \gtrsim 0.5$.
+It does not directly say anything about $\chi$.
+In fact, Figure 3 makes it quite clear that $\chi$ can have support beyond 0.5 as long as spins are isotropic.
+Meanwhile, the EM observations describe $\chi$, not $\chieff$, since only one component is a BH in those X-ray binaries.
+In a way, this whole paper is about the difference between $\chi$ and $\chieff$ (depending on the mis-alignment angles), so conflating them here and in the caption of Figure 1 is very confusing.]}   The analysis here is relatively
 insensitive to the precise details of the posterior distributions;
@@
     from several electromagnetically-detected stellar-mass black hole
-    systems (see \citet{2015PhR...548....1M} for a summary of such
+    systems \ilya{[Again, $\chi$ and $\chieff$ are being conflated; see comment on 2nd paragraph of intro.]} (see \citet{2015PhR...548....1M} for a summary of such
     measurements).}
@@
 from either intrinsically small spins or larger spins whose direction
-is mis-aligned with the orbital plane of the binary (i.e.\ spin
+is mis-aligned with the orbital \sout{plane} \ilya{angular momentum} of the binary (i.e.\ spin
 vectors with small $z$-components).  Mis-alignment is capable of
@@
 assume that the population spin distribution factorises into a
-distribution for the spin magnitude $a$ and a distribution for the
+distribution for the spin magnitude \ilya{$\chi$} and a distribution for the
 spin angles.  Finally, we assume that the distribution of spins is
@@
   \label{eq:magnitude-dists}
-  p(a) = \begin{cases}
-    2\left(1-a\right) & \textnormal{``low''} \\
+  p(\ilya{\chi}) = \begin{cases}
+    2\left(1-\chi \right) & \textnormal{``low''} \\
     1 & \textnormal{``flat''} \\
-    2 a & \textnormal{``high''}
+    2 \chi & \textnormal{``high''}
   \end{cases},
@@
 chance of 10 additional \ac{BBH} detections in the next two years
-\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming an
-additional 10 detections with similar observational uncertainties
+\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming \ilya{10} \sout{an}
+additional \sout{10} detections with similar observational uncertainties
 drawn from each of our six zero-parameter models for the spin
@@
 find that most scenarios with an additional 10 detections allow the
-correct angular distribution to be distinguished with greater than
+\sout{correct} \ilya{simulated} angular distribution to be \sout{distinguished} \ilya{inferred} with greater than
 $5\sigma$ ($2.9 \times 10^{-7}$ odds) credibility; and in the most
@@
 While such future detections should permit a confident distinction
-between \emph{angular} distributions, we will remain much less certain
-about the \emph{magnitude} distribution until we have a larger number
+between \emph{angular} distributions, we would remain much less certain
+about the \emph{magnitude} distribution \ilya{among the three options considered here} until we have a larger number
 of observations.  In Figure \ref{fig:O2-predictions}, the odds ratio

### 2017-04-11 paper/AlignedIsotropicGW.tex
Finished edits.

@@
 \citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming \ilya{10} \sout{an}
-additional \sout{10} detections with similar observational uncertainties
+additional \sout{10} detections with similar observational uncertainties \ilya{[I think it would be useful to specify (in an appendix?) what model we use for the observational uncertainty on $\chieff$.]}
 drawn from each of our six zero-parameter models for the spin
@@
 $\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
-spins are intrinsically very small, with $a \lesssim 0.3$, then it is
-no longer possible to resolve the negative spin with a small number of
+spins are intrinsically very small, with $\ilya{\chi} \lesssim 0.3$, then it is
+no longer possible to resolve the negative \ilya{effective} spin with a small number of
 observations.  As noted below, however, spins observed in X-ray
@@
 \citet{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current
-spin parameters probably are close to their values upon core collapse.
+spin parameters probably are close to their values upon core collapse, \ilya{at least in high-mass X-ray binaries}.
 However, the specific processes involved in the production of black
@@
 both massive stellar binaries
-\citep{Albrecht:2009,2014ApJ...785...83A} and BH XRBs
+\citep{Albrecht:2009,2014ApJ...785...83A} and BH X-ray binaries
 \citep{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
@@
 models.  Meanwhile, recent studies by
-\citet{2017arXiv170306869F,2017arXiv170306223G} focused on inference
+\citet{2017arXiv170306869F} and \citet{2017arXiv170306223G} focused on inference
 on spin magnitudes.
@@
 continues with future gravitational-wave observations, it will be
-possible to either confirm that \ac{BH} spins are sometimes large in
-magnitude and isotropic in direction, yielding a strong indication of
+possible to either confirm that \ac{BH} spins are isotropic in direction \ilya{and} sometimes large in
+magnitude, yielding a strong indication of
 a dynamical formation origin; or that \ac{BH} spins are overwhelmingly
@@
   hierarchical analysis; if we are given posterior distributions or
-  posterior samples, we need to re-weight to "remove" the prior and
+  posterior samples, we need to re-weight to ``remove'' the prior and
   obtain the likelihood.
@@
 where $P\left( A \leq \chieff \leq B \right)$ is the posterior
-probability that $\chieff$ is between $A$ and $B$.  Using our
+probability \ilya{[Should really be likelihood, not posterior, right?  You are approximating the first term in the product in (B1)...]} that $\chieff$ is between $A$ and $B$.  Using our
 approximations to the $\chieff$ posteriors described above, this gives
@@
 each subsequent measurement contributes on average a factor of
-$\sim 3$ to the overall odds.  After ten additional events, then, the
+$\sim 3$ to the overall odds.
+\ilya{[This ranges from $\sim$1 to $\sim$50 (2.5 $\sigma$).
+Not clear what an average of 3 means here (geometric mean, I hope?).
+I am still not a huge fan of appendix B; it raises more questions than it answers, as far as I am concerned.]}  After ten additional events, then, the
 odds ratio becomes $5 \times 3^{10} \simeq 3 \times 10^{5}$, or
@@
 \begin{equation}
-p(a) \propto (1 - a)^{\alpha}
+p(\chi) \propto (1 - \chi)^{\alpha}
 \label{eq:lowspinalpha}
@@
 recovers the ``low'' distribution. For higher values of $\alpha$, this distribution 
-becomes more peaked towards $a = 0$.
+becomes more peaked towards $\ilya{\chi} = 0$.
 
@@
 \includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
-\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
-The top axis shows the mean spin magnitude $a$
-that the value of alpha corresponds to. We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.}
+\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models.} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
+The top axis shows the mean spin magnitude $\ilya{\chi}$
+ \sout{that the value of} \ilya{which this $\alpha$} corresponds to.
+We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.
+\ilya{[Change $\bar{a}$ to $\bar{\chi}$ on upper abscissa label.]}}
 \label{fig:smallspinsalpha}
@@
 The top axis shows the mean spin magnitude
-that value of $\alpha$ corresponds to (i.e., for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
+that value of $\alpha$ corresponds to (\ilya{e.g.}, for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
 We see that if typical BH spins are $\lesssim 0.2$ we have no evidence for one model over the other.
@@
 \section{Approximations in the Gravitational Waveform}
-The model waveforms used to infer the $\chieff$ of the three events
+The model waveforms used to infer the $\chieff$ of the three LIGO events
 incorporate approximations to the true behaviour of the merging
 systems that are expected to break down for sufficiently high
-mis-aligned spins.  The effect of these approximations has been
-investigated in detail for the inference on the parameters describing
-GW150914 \citep{2016arXiv161107531T}.  For this source, statistical
+mis-aligned spins.  The effect of these approximations \ilya{on inference}  
+on the parameters describing
+GW150914 has been
+investigated in detail \citep{2016arXiv161107531T} \ilya{[Should be an Abbott et al. ref like the others?]}.  For this source, statistical
 uncertainties dominate over any waveform systematics.  Detailed
@@
 may dominate for signals with this large SNR ($\simeq 23$) when the
-source is edge-on or with high spin.  The other two events discussed
+source is edge-on or \ilya{has} high spins.  The other two events discussed
 in this paper are at much lower SNR, with correspondingly larger
@@
 
-Given the preference of our data for a smaller spin population we
+\sout{Given the preference of our data for a smaller spin population} We
 assume here that measurements made in the future are not dominated by
 systematic errors, but this assumption would need to be revisited for
-high-SNR, edge-on, or high-spin sources detected in the future.
+high-SNR, edge-on, or high-spin sources detected in the future.  \ilya{[I thought we argued that there is no strong preference for small spin magnitudes.  We should be consistent...]}
 

### 2017-04-12 paper/AlignedIsotropicGW.tex
Fix #5 and other minor Ilya edits.

@@
   The first direct detections of \acp{GW} from merging \acp{BBH} open
-  a unique window into the \ilya{\ac{BBH}} formation environment.  \sout{of massive stars and
-  their compact remnants}  One promising signature of the formation
-  environment is the angular distribution of the \ac{BH} spins;
-  systems formed through dynamical interactions among already-compact
-  objects are expected to have isotropic spin orientations whereas
-  binaries formed from pairs of stars born together are more likely to
-  have spins preferentially aligned with the binary orbit as a
-  consequence of their joint evolution toward a \ac{BBH} system.  We
-  consider existing \ac{GW} measurements of $\chieff$, the
-  best-measured combination of spin parameters, in the three likely
-  binary black hole detections GW150914, LVT151012, and GW151226.  If
-  binary black hole spin magnitudes extend to high values, as is
-  suggested by observations of black hole X-ray binaries, we show that
-  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
-  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
-    $r \ll 1$ is equivalent to $x \sigma$ with
-    $x = \Phi^{-1}\left( 1 - r \right)$, where $\Phi$ is the unit
-    normal CDF. \ilya{[Two-tailed vs.~one-tailed]}})  preference for an isotropic angular distribution.
-  By considering the effect of an additional 10 detections drawn from
-  the various models in the suite we show that if all observations
-  come from a single population such an augmented data set would
-  enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
+  a unique window into the \ac{BBH} formation environment.  One
+  promising signature of the formation environment is the angular
+  distribution of the \ac{BH} spins; systems formed through dynamical
+  interactions among already-compact objects are expected to have
+  isotropic spin orientations whereas binaries formed from pairs of
+  stars born together are more likely to have spins preferentially
+  aligned with the binary orbit as a consequence of their joint
+  evolution toward a \ac{BBH} system.  We consider existing \ac{GW}
+  measurements of $\chieff$, the best-measured combination of spin
+  parameters, in the three likely binary black hole detections
+  GW150914, LVT151012, and GW151226.  If binary black hole spin
+  magnitudes extend to high values, as is suggested by observations of
+  black hole X-ray binaries, we show that the data already exhibit a
+  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
+  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
+    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
+    $\Phi$ is the unit normal CDF.})  preference for an isotropic
+  angular distribution.  By considering the effect of an additional 10
+  detections drawn from the various models in the suite we show that
+  if all observations come from a single population such an augmented
+  data set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
   ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
@@
   \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
-  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \ilya{\equiv} \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
+  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
 \end{equation}
-where \sout{the $\hat{z}$ axis is aligned with the orbital angular momentum
-of the binary,} $m_{1,2}$ are the gravitational masses of the more-massive (1) and
-less-massive (2) components, $M = m_1 + m_2$ is the total mass,
+where $m_{1,2}$ are the gravitational masses of the more-massive (1)
+and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
 $\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
-holes in the binary, $\vec{L}$ is the orbital angular momentum vector \ilya{assumed to point in the $\hat{z}$ direction},
-and $0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
+holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
+assumed to point in the $\hat{z}$ direction, and
+$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
 projections of the individual \ac{BH} spins
@@
 electromagnetically-detected stellar-mass black hole population (see
-Section \ref{sec:discussion}).
-\ilya{[Figure 1 shows an absence of support for $\chieff \gtrsim 0.5$.
-It does not directly say anything about $\chi$.
-In fact, Figure 3 makes it quite clear that $\chi$ can have support beyond 0.5 as long as spins are isotropic.
-Meanwhile, the EM observations describe $\chi$, not $\chieff$, since only one component is a BH in those X-ray binaries.
-In a way, this whole paper is about the difference between $\chi$ and $\chieff$ (depending on the mis-alignment angles), so conflating them here and in the caption of Figure 1 is very confusing.]}   The analysis here is relatively
-insensitive to the precise details of the posterior distributions;
-other conclusions are more sensitive.  In particular, our Gaussian
-approximation does permit $\chieff = 0$ for GW151226 while the true
-posterior rules this out at high confidence
+Section \ref{sec:discussion}).  \ilya{[Figure 1 shows an absence of
+  support for $\chieff \gtrsim 0.5$.  It does not directly say
+  anything about $\chi$.  In fact, Figure 3 makes it quite clear that
+  $\chi$ can have support beyond 0.5 as long as spins are isotropic.
+  Meanwhile, the EM observations describe $\chi$, not $\chieff$, since
+  only one component is a BH in those X-ray binaries.  In a way, this
+  whole paper is about the difference between $\chi$ and $\chieff$
+  (depending on the mis-alignment angles), so conflating them here and
+  in the caption of Figure 1 is very confusing.]}  The analysis here
+is relatively insensitive to the precise details of the posterior
+distributions; other conclusions are more sensitive.  In particular,
+our Gaussian approximation does permit $\chieff = 0$ for GW151226
+while the true posterior rules this out at high confidence
 \citep{2016PhRvL.116x1103A,O1-BBH}.

### 2017-04-12 paper/AlignedIsotropicGW.tex
Fix #6: chi, chieff, a confusion.

@@
 distribution with the same mean and 90\% credible interval as quoted
-in \citet{O1-BBH}.  There is essentially no posterior support for
-$\chi \gtrsim 0.5$ as inferred for the majority of the
+in \citet{O1-BBH}.  Since all three mergers involve black holes of
+comparable masses, there is essentially no posterior support for both
+black holes' $\chi \gtrsim 0.5$, in contrast to the majority of the
 electromagnetically-detected stellar-mass black hole population (see
-Section \ref{sec:discussion}).  \ilya{[Figure 1 shows an absence of
-  support for $\chieff \gtrsim 0.5$.  It does not directly say
-  anything about $\chi$.  In fact, Figure 3 makes it quite clear that
-  $\chi$ can have support beyond 0.5 as long as spins are isotropic.
-  Meanwhile, the EM observations describe $\chi$, not $\chieff$, since
-  only one component is a BH in those X-ray binaries.  In a way, this
-  whole paper is about the difference between $\chi$ and $\chieff$
-  (depending on the mis-alignment angles), so conflating them here and
-  in the caption of Figure 1 is very confusing.]}  The analysis here
-is relatively insensitive to the precise details of the posterior
-distributions; other conclusions are more sensitive.  In particular,
-our Gaussian approximation does permit $\chieff = 0$ for GW151226
-while the true posterior rules this out at high confidence
+Section \ref{sec:discussion}).  The analysis here is relatively
+insensitive to the precise details of the posterior distributions;
+other conclusions are more sensitive.  In particular, our Gaussian
+approximation does permit $\chieff = 0$ for GW151226 while the true
+posterior rules this out at high confidence
 \citep{2016PhRvL.116x1103A,O1-BBH}.
@@
     reported in \citet{O1-BBH}.  It is notable that none of the
-    $\chieff$ posteriors extends to the high spin magnitudes inferred
-    from several electromagnetically-detected stellar-mass black hole
-    systems \ilya{[Again, $\chi$ and $\chieff$ are being conflated; see comment on 2nd paragraph of intro.]} (see \citet{2015PhR...548....1M} for a summary of such
-    measurements).}
+    $\chieff$ posteriors extends to the regime of high spin magnitudes
+    inferred from several electromagnetically-detected stellar-mass
+    black hole systems (see \citet{2015PhR...548....1M} for a summary
+    of such measurements).}
 \end{figure}
@@
 from either intrinsically small spins or larger spins whose direction
-is mis-aligned with the orbital \sout{plane} \ilya{angular momentum} of the binary (i.e.\ spin
-vectors with small $z$-components).  Mis-alignment is capable of
-producing \emph{negative} values of $\chieff$, however, whereas aligned
-spins will always have $\chieff \geq 0$.  This difference provides
-strong discriminating power between the two angular distributions,
-even without good information about the magnitude distribution; to the
-extent that data favour negative $\chieff$ they weigh heavily against
-aligned models.  To quantify the degree of support for these two
-alternate explanations of small $\chieff$ values in the merging
-\ac{BBH} population, we compared the Bayesian evidence for various
-simple models of the spin population using the \ac{GW} data set.
+is mis-aligned with the orbital angular momentum of the binary (i.e.\
+spin vectors with small $z$-components).  Mis-alignment is capable of
+producing \emph{negative} values of $\chieff$, however, whereas
+aligned spins will always have $\chieff \geq 0$.  This difference
+provides strong discriminating power between the two angular
+distributions, even without good information about the magnitude
+distribution; to the extent that data favour negative $\chieff$ they
+weigh heavily against aligned models.  To quantify the degree of
+support for these two alternate explanations of small $\chieff$ values
+in the merging \ac{BBH} population, we compared the Bayesian evidence
+for various simple models of the spin population using the \ac{GW}
+data set.
 
@@
 assume that the population spin distribution factorises into a
-distribution for the spin magnitude \ilya{$\chi$} and a distribution for the
+distribution for the spin magnitude $a$ and a distribution for the
 spin angles.  Finally, we assume that the distribution of spins is
@@
   \label{eq:magnitude-dists}
-  p(\ilya{\chi}) = \begin{cases}
-    2\left(1-\chi \right) & \textnormal{``low''} \\
+  p(a) = \begin{cases}
+    2\left(1-a \right) & \textnormal{``low''} \\
     1 & \textnormal{``flat''} \\
-    2 \chi & \textnormal{``high''}
+    2 a & \textnormal{``high''}
   \end{cases},
@@
 in Section \ref{sec:discussion}; neither observations nor population
-synthesis codes can at this point authoritatively suggest {\it any}
+synthesis codes can at this point authoritatively suggest \emph{any}
 particular spin distribution \citep{2015PhR...548....1M}.  Our models,
@@
 chance of 10 additional \ac{BBH} detections in the next two years
-\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming \ilya{10} \sout{an}
-additional \sout{10} detections with similar observational uncertainties \ilya{[I think it would be useful to specify (in an appendix?) what model we use for the observational uncertainty on $\chieff$.]}
+\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
+additional detections with similar observational uncertainties
+\ilya{[I think it would be useful to specify (in an appendix?) what
+  model we use for the observational uncertainty on $\chieff$.]}
 drawn from each of our six zero-parameter models for the spin
@@
 find that most scenarios with an additional 10 detections allow the
-\sout{correct} \ilya{simulated} angular distribution to be \sout{distinguished} \ilya{inferred} with greater than
+simulated angular distribution to be inferred with greater than
 $5\sigma$ ($2.9 \times 10^{-7}$ odds) credibility; and in the most
@@
 While such future detections should permit a confident distinction
-between \emph{angular} distributions, we would remain much less certain
-about the \emph{magnitude} distribution \ilya{among the three options considered here} until we have a larger number
-of observations.  In Figure \ref{fig:O2-predictions}, the odds ratio
-between different magnitude distributions with the same angular
-distribution is much closer to unity than the odds ratio between angular
-distributions.
+between \emph{angular} distributions, we would remain much less
+certain about the \emph{magnitude} distribution among the three
+options considered here until we have a larger number of observations.
+In Figure \ref{fig:O2-predictions}, the odds ratio between different
+magnitude distributions with the same angular distribution is much
+closer to unity than the odds ratio between angular distributions.
 
@@
 $\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
-spins are intrinsically very small, with $\ilya{\chi} \lesssim 0.3$, then it is
-no longer possible to resolve the negative \ilya{effective} spin with a small number of
-observations.  As noted below, however, spins observed in X-ray
-binaries are typically large.  Additionally, models which do not
+spins are intrinsically very small, with $a \lesssim 0.3$, then it is
+no longer possible to resolve the negative effective spin with a small
+number of observations.  As noted below, however, spins observed in
+X-ray binaries are typically large.  Additionally, models which do not
 permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
@@
 \citet{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current
-spin parameters probably are close to their values upon core collapse, \ilya{at least in high-mass X-ray binaries}.
-However, the specific processes involved in the production of black
-hole binaries from isolated binaries could alter the spin magnitude
-distribution of those holes relative to the X-ray binary systems; for
-example, close tidal interactions could spin up the core, or stripping
-of the envelope could reduce the available angular momentum
+spin parameters probably are close to their values upon core collapse,
+at least in high-mass X-ray binaries.  However, the specific processes
+involved in the production of black hole binaries from isolated
+binaries could alter the spin magnitude distribution of those holes
+relative to the X-ray binary systems; for example, close tidal
+interactions could spin up the core, or stripping of the envelope
+could reduce the available angular momentum
 \citep{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
@@
 continues with future gravitational-wave observations, it will be
-possible to either confirm that \ac{BH} spins are isotropic in direction \ilya{and} sometimes large in
-magnitude, yielding a strong indication of
-a dynamical formation origin; or that \ac{BH} spins are overwhelmingly
-small in magnitude, yielding a notable contradiction with the claimed
-\ac{BH} X-ray binary spin measurements, particularly those for high-mass X-ray
-binaries.
-
-%\citet{2017CQGra..34cLT01V} also studied the possibility of distinguishing aligned and isotropic angular distributions of \ac{BBH} spins, but concluded that several hundred sources would be required to adequately separate the models.
-The ``aligned'' population in that study permitted spins nearly parallel or anti-parallel to the orbital axis, i.e.\ $\hat{S}\cdot \hat{L} \simeq \pm 1$, in contrast to our aligned models; thus the aligned/isotropic discriminating power in \citet{2017CQGra..34cLT01V} relies effectively on measurements of the detailed shape of the $\chieff$ distribution and, as noted above, measurements are not very sensitive to the spin magnitude distribution that controls the shape of $\chieff$.
-Note that anti-parallel spins are strongly disfavoured in models of isolated binary evolution (see, for example, \citet{Kalogera:2000}).
+possible to either confirm that \ac{BH} spins are isotropic in
+direction and sometimes large in magnitude, yielding a strong
+indication of a dynamical formation origin; or that \ac{BH} spins are
+overwhelmingly small in magnitude, yielding a notable contradiction
+with the claimed \ac{BH} X-ray binary spin measurements, particularly
+those for high-mass X-ray binaries.
 
@@
 where $P\left( A \leq \chieff \leq B \right)$ is the posterior
-probability \ilya{[Should really be likelihood, not posterior, right?  You are approximating the first term in the product in (B1)...]} that $\chieff$ is between $A$ and $B$.  Using our
-approximations to the $\chieff$ posteriors described above, this gives
-an odds ratio of $5$ in favour of the isotropic models, which is about
-a factor of two smaller than the ratio in the more careful calculation
-described in Section \ref{sec:O1}.  This is a satisfactory answer at
-an order-of-magnitude level.
+probability (here used to approximate the likelihood) that $\chieff$
+is between $A$ and $B$.  Using our approximations to the $\chieff$
+posteriors described above, this gives an odds ratio of $5$ in favour
+of the isotropic models, which is about a factor of two smaller than
+the ratio in the more careful calculation described in Section
+\ref{sec:O1}.  This is a satisfactory answer at an order-of-magnitude
+level.
 
@@
 and our measurement uncertainties on $\chieff$ are $\simeq 0.1$, then
-each subsequent measurement contributes on average a factor of
-$\sim 3$ to the overall odds.
-\ilya{[This ranges from $\sim$1 to $\sim$50 (2.5 $\sigma$).
-Not clear what an average of 3 means here (geometric mean, I hope?).
-I am still not a huge fan of appendix B; it raises more questions than it answers, as far as I am concerned.]}  After ten additional events, then, the
+the geometric mean of each subsequent measurement's contribution to
+the overall odds is $\sim 3$.  After ten additional events, then, the
 odds ratio becomes $5 \times 3^{10} \simeq 3 \times 10^{5}$, or
@@
 
-In the main text we considered three models for BH spin magnitudes: ``low'', ``flat'' and ``high''. These were intended to capture some of the uncertainty regarding the BH spin magnitude distribution.
+In the main text we considered three models for BH spin magnitudes:
+``low'', ``flat'' and ``high''. These were intended to capture some of
+the uncertainty regarding the BH spin magnitude distribution.
 
@@
 \begin{equation}
-p(\chi) \propto (1 - \chi)^{\alpha}
+p(a) \propto (1 - a)^{\alpha}
 \label{eq:lowspinalpha}
@@
 
-When $\alpha = 0$, this recovers the ``flat'' distribution, whilst $\alpha = 1$
-recovers the ``low'' distribution. For higher values of $\alpha$, this distribution 
-becomes more peaked towards $\ilya{\chi} = 0$.
+When $\alpha = 0$, this recovers the ``flat'' distribution, whilst
+$\alpha = 1$ recovers the ``low'' distribution. For higher values of
+$\alpha$, this distribution becomes more peaked towards
+$a = 0$.
 
@@
 \includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
-\caption{\textbf{Effect of small spins on evidence ratio of aligned against isotropic models.} The blue line shows the evidence ratio (plotted as the equivalent sigma) between a model where all systems are aligned, versus one where all systems are from an isotropic distribution as a function of the power law $\alpha$ corresponding to Equation~\ref{eq:lowspinalpha}.
-The top axis shows the mean spin magnitude $\ilya{\chi}$
- \sout{that the value of} \ilya{which this $\alpha$} corresponds to.
-We see that for mean spin magnitudes $\lesssim 0.2$ we find no evidence for either distribution over the other.
-\ilya{[Change $\bar{a}$ to $\bar{\chi}$ on upper abscissa label.]}}
+\caption{\textbf{Effect of small spins on evidence ratio of aligned
+    against isotropic models.} The blue line shows the evidence ratio
+  (plotted as the equivalent sigma) between a model where all systems
+  are aligned, versus one where all systems are from an isotropic
+  distribution as a function of the power law $\alpha$ corresponding
+  to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
+  magnitude $\bar{a}$ which this $\alpha$ corresponds to. We see that
+  for mean spin magnitudes $\lesssim 0.2$ we find no evidence for
+  either distribution over the other.}
 \label{fig:smallspinsalpha}
@@
 
-In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of aligned to isotropic distributions (plotted as the equivalent sigma) with spin magnitudes given by this model with $\alpha$ in the range $0$ -- $6$.
-The top axis shows the mean spin magnitude
-that value of $\alpha$ corresponds to (\ilya{e.g.}, for the ``flat'' distribution $\alpha = 0$, the mean spin magnitude is 0.5).
-We see that if typical BH spins are $\lesssim 0.2$ we have no evidence for one model over the other.
+In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of
+aligned to isotropic distributions (plotted as the equivalent sigma)
+with spin magnitudes given by this model with $\alpha$ in the range
+$0$ -- $6$. The top axis shows the mean spin magnitude that value of
+$\alpha$ corresponds to (e.g., for the ``flat'' distribution
+$\alpha = 0$, the mean spin magnitude is 0.5). We see that if typical
+BH spins are $\lesssim 0.2$ we have no evidence for one model over the
+other.
 
@@
 \section{Approximations in the Gravitational Waveform}
-The model waveforms used to infer the $\chieff$ of the three LIGO events
-incorporate approximations to the true behaviour of the merging
+The model waveforms used to infer the $\chieff$ of the three LIGO
+events incorporate approximations to the true behaviour of the merging
 systems that are expected to break down for sufficiently high
-mis-aligned spins.  The effect of these approximations \ilya{on inference}  
-on the parameters describing
-GW150914 has been
-investigated in detail \citep{2016arXiv161107531T} \ilya{[Should be an Abbott et al. ref like the others?]}.  For this source, statistical
-uncertainties dominate over any waveform systematics.  Detailed
-comparisons with numerical relativity computations using no
-approximations to the dynamics \citep{2016PhRvD..94f4035A} also
-suggest that statistical uncertainties dominate the systematics for
-this system.  \citet{2016arXiv161107531T} suggests that systematics
-may dominate for signals with this large SNR ($\simeq 23$) when the
-source is edge-on or \ilya{has} high spins.  The other two events discussed
-in this paper are at much lower SNR, with correspondingly larger
-statistical uncertainties, and are probably similarly oriented and
-with similarly small spins, so we do not expect systematic
-uncertainties to dominate.
+mis-aligned spins.  The effect of these approximations on inference on
+the parameters describing GW150914 has been investigated in detail
+\citep{2016arXiv161107531T} \ilya{[Should be an Abbott et al. ref like
+  the others?]}.  For this source, statistical uncertainties dominate
+over any waveform systematics.  Detailed comparisons with numerical
+relativity computations using no approximations to the dynamics
+\citep{2016PhRvD..94f4035A} also suggest that statistical
+uncertainties dominate the systematics for this system.
+\citet{2016arXiv161107531T} suggests that systematics may dominate for
+signals with this large SNR ($\simeq 23$) when the source is edge-on
+or has high spins.  The other two events discussed in this paper are
+at much lower SNR, with correspondingly larger statistical
+uncertainties, and are probably similarly oriented and with similarly
+small spins, so we do not expect systematic uncertainties to dominate.
 
-\sout{Given the preference of our data for a smaller spin population} We
-assume here that measurements made in the future are not dominated by
-systematic errors, but this assumption would need to be revisited for
-high-SNR, edge-on, or high-spin sources detected in the future.  \ilya{[I thought we argued that there is no strong preference for small spin magnitudes.  We should be consistent...]}
+We assume here that measurements made in the future are not dominated
+by systematic errors, but this assumption would need to be revisited
+for high-SNR, edge-on, or high-spin sources detected in the future.
 

### 2017-04-12 paper/AlignedIsotropicGW.tex
Fix #3 and Fix #9

@@
 \acrodef{GW}{gravitational wave}
+\acrodef{O1}{first observing}
 \acrodef{PE}{parameter estimation}
@@
 inferred on $\chieff$ for the three likely \ac{GW} detections
-GW150914, GW151226 and LVT151012 from \citet{O1-BBH}.  Because samples
-drawn from the posterior on $\chieff$ are not publicly released at
-this time, we have approximated the posterior as a Gaussian
-distribution with the same mean and 90\% credible interval as quoted
-in \citet{O1-BBH}.  Since all three mergers involve black holes of
-comparable masses, there is essentially no posterior support for both
-black holes' $\chi \gtrsim 0.5$, in contrast to the majority of the
-electromagnetically-detected stellar-mass black hole population (see
-Section \ref{sec:discussion}).  The analysis here is relatively
-insensitive to the precise details of the posterior distributions;
-other conclusions are more sensitive.  In particular, our Gaussian
-approximation does permit $\chieff = 0$ for GW151226 while the true
-posterior rules this out at high confidence
+GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
+run (O1) \citep{O1-BBH}.  Because samples drawn from the posterior on
+$\chieff$ are not publicly released at this time, we have approximated
+the posterior as a Gaussian distribution with the same mean and 90\%
+credible interval as quoted in \citet{O1-BBH}.  Since all three
+mergers involve black holes of comparable masses, there is essentially
+no posterior support for both black holes' $\chi \gtrsim 0.5$, in
+contrast to the majority of the electromagnetically-detected
+stellar-mass black hole population (see Section \ref{sec:discussion}).
+The analysis here is relatively insensitive to the precise details of
+the posterior distributions; other conclusions are more sensitive.  In
+particular, our Gaussian approximation does permit $\chieff = 0$ for
+GW151226 while the true posterior rules this out at high confidence
 \citep{2016PhRvL.116x1103A,O1-BBH}.
@@
   \caption{\label{fig:O1-posteriors} Approximate posteriors on
-    $\chieff$ from the LIGO O1 observations in \citet{O1-BBH}.  We
-    approximate the posteriors reported in \citet{O1-BBH} using
-    Gaussians with the same median and 90\% credible interval as
-    reported in \citet{O1-BBH}.  It is notable that none of the
-    $\chieff$ posteriors extends to the regime of high spin magnitudes
-    inferred from several electromagnetically-detected stellar-mass
-    black hole systems (see \citet{2015PhR...548....1M} for a summary
-    of such measurements).}
+    $\chieff$ from the Advanced LIGO O1 run observations in
+    \citet{O1-BBH}.  We approximate the posteriors reported in
+    \citet{O1-BBH} using Gaussians with the same median and 90\%
+    credible interval as reported in \citet{O1-BBH}.  It is notable
+    that none of the $\chieff$ posteriors extends to the regime of
+    high spin magnitudes inferred from several
+    electromagnetically-detected stellar-mass black hole systems (see
+    \citet{2015PhR...548....1M} for a summary of such measurements).}
 \end{figure}
@@
 drawn from each of our six zero-parameter models for the spin
-distribution in addition to the three existing detections from O1, we
-find the odds ratios shown in Figure \ref{fig:O2-predictions}.  We
-find that most scenarios with an additional 10 detections allow the
-simulated angular distribution to be inferred with greater than
-$5\sigma$ ($2.9 \times 10^{-7}$ odds) credibility; and in the most
-pessimistic case the distinction is typically
-$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
-While such future detections should permit a confident distinction
-between \emph{angular} distributions, we would remain much less
-certain about the \emph{magnitude} distribution among the three
+distribution in addition to the three existing detections from
+O1, we find the odds ratios shown in Figure
+\ref{fig:O2-predictions}.  We find that most scenarios with an
+additional 10 detections allow the simulated angular distribution to
+be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
+credibility; and in the most pessimistic case the distinction is
+typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
+odds ratio).  While such future detections should permit a confident
+distinction between \emph{angular} distributions, we would remain much
+less certain about the \emph{magnitude} distribution among the three
 options considered here until we have a larger number of observations.

### 2017-04-12 paper/AlignedIsotropicGW.tex
Added Gerosa & B. Farr to acknowledgments
Ben moves from acknowledgements to co-author.

@@
 
+\author[0000-0002-2916-9200]{Ben Farr}
+
+\affiliation{Enrico Fermi Institute, University of Chicago, Chicago, IL 60637, USA}
+
 \author[0000-0002-6254-1617]{Alberto Vecchio}
@@
 
-\email{w.farr@bham.ac.uk, simons@star.sr.bham.ac.uk, miller@astro.umd.edu,\\imandel@star.sr.bham.ac.uk, av@star.sr.bham.ac.uk}
+\email{w.farr@bham.ac.uk, simons@star.sr.bham.ac.uk, miller@astro.umd.edu,\\imandel@star.sr.bham.ac.uk, farr@uchicago.edu, av@star.sr.bham.ac.uk}
 
@@
 
-We thank Richard O'Shaughnessy and Christopher Berry for discussions and
-comments on this work.  WF, SS, IM and AV were supported in part by the STFC.  MCM acknowledges support of the University of
-Birmingham Institute for Advanced Study Distinguished Visiting Fellows
-program.   SS and IM acknowledge support from the National Science Foundation under Grant No. NSF PHY11-25915.
+We thank Richard O'Shaughnessy, Christopher Berry, and Davide Gerosa
+for discussions and comments on this work.  WF, SS, IM and AV were
+supported in part by the STFC.  MCM acknowledges support of the
+University of Birmingham Institute for Advanced Study Distinguished
+Visiting Fellows program.  SS and IM acknowledge support from the
+National Science Foundation under Grant No. NSF PHY11-25915.
 

### 2017-04-12 paper/AlignedIsotropicGW.tex
Fix #8 about misalignment increasing during inspiral.
Update Ben's affiliation

@@
 
-\affiliation{Enrico Fermi Institute, University of Chicago, Chicago, IL 60637, USA}
+\affiliation{Enrico Fermi Institute and Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, United States}
 
@@
 the spin magnitude and direction distributions, contrary to our
-simplified assumptions.
+simplified assumptions.  A small misalignment at wide separation can
+also evolve to a more significant misalignment in component spins as
+the binary shrinks from \ac{GW} emission \citep{2015PhRvD..92f4016G},
+but $\chieff$ is approximately conserved through this evolution.
 

### 2017-04-12 paper/AlignedIsotropicGW.tex
Updating chi text.

@@
 the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval as quoted in \citet{O1-BBH}.  Since all three
+credible interval as quoted in \citet{O1-BBH}.
+\ilya{None of the $\chieff$ posteriors are consistent with two black holes with large aligned spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred for the majority of black holes in X-ray binaries  (see Section \ref{sec:discussion}).}
+\sout{Since all three
 mergers involve black holes of comparable masses, there is essentially
@@
 contrast to the majority of the electromagnetically-detected
-stellar-mass black hole population (see Section \ref{sec:discussion}).
+stellar-mass black hole population (see Section \ref{sec:discussion}).}
 The analysis here is relatively insensitive to the precise details of

### 2017-04-12 paper/AlignedIsotropicGW.tex
Caption 1 tweak.

@@
     credible interval as reported in \citet{O1-BBH}.  It is notable
-    that none of the $\chieff$ posteriors extends to the regime of
-    high spin magnitudes inferred from several
-    electromagnetically-detected stellar-mass black hole systems (see
+    that none of the $\chieff$ posteriors \ilya{supports high \ac{BH}
+    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries} (see
     \citet{2015PhR...548....1M} for a summary of such measurements).}
@@
 the parameters describing GW150914 has been investigated in detail
-\citep{2016arXiv161107531T} \ilya{[Should be an Abbott et al. ref like
-  the others?]}.  For this source, statistical uncertainties dominate
+\citep{2016arXiv161107531T}.  For this source, statistical uncertainties dominate
 over any waveform systematics.  Detailed comparisons with numerical

### 2017-04-12 paper/AlignedIsotropicGW.tex
Removing a bit of red.
Fix #10, effect of other parameters on chi_eff measurements.

@@
 
-\newcommand{\OOneSigmaIsoAligned}{\checkme{1.7}}
-\newcommand{\OOneOddsIsoAligned}{\checkme{0.087}}
-\newcommand{\OTwoSigmaIsoAlignedMin}{\checkme{2.9}}
-\newcommand{\OTwoOddsIsoAlignedMin}{\checkme{0.0035}}
+\newcommand{\OOneSigmaIsoAligned}{1.7}
+\newcommand{\OOneOddsIsoAligned}{0.087}
+\newcommand{\OTwoSigmaIsoAlignedMin}{2.9}
+\newcommand{\OTwoOddsIsoAlignedMin}{0.0035}
 
@@
 \citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
-additional detections with similar observational uncertainties
-\ilya{[I think it would be useful to specify (in an appendix?) what
-  model we use for the observational uncertainty on $\chieff$.]}
-drawn from each of our six zero-parameter models for the spin
-distribution in addition to the three existing detections from
-O1, we find the odds ratios shown in Figure
+additional detections drawn from each of our six zero-parameter models
+for the spin distribution in addition to the three existing detections
+from O1, with observational uncertainties drawn randomly from the
+three Gaussian widths used to approximate the $\chieff$ posteriors in
+Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
+  $\chieff$ depends on the other parameters of the merging \ac{BBH}
+  system, particularly on the mass ratio.  Our assumption about future
+  observational uncertainties is appropriate if the parameters of the
+  existing three events are representative of the parameters of future
+  detections.}, we find the odds ratios shown in Figure
 \ref{fig:O2-predictions}.  We find that most scenarios with an

### 2017-04-12 paper/AlignedIsotropicGW.tex
Adopt Ilya's suggested phrasing for posteriors inconsistent with large, aligned spins.

@@
 the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval as quoted in \citet{O1-BBH}.
-\ilya{None of the $\chieff$ posteriors are consistent with two black holes with large aligned spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred for the majority of black holes in X-ray binaries  (see Section \ref{sec:discussion}).}
-\sout{Since all three
-mergers involve black holes of comparable masses, there is essentially
-no posterior support for both black holes' $\chi \gtrsim 0.5$, in
-contrast to the majority of the electromagnetically-detected
-stellar-mass black hole population (see Section \ref{sec:discussion}).}
-The analysis here is relatively insensitive to the precise details of
-the posterior distributions; other conclusions are more sensitive.  In
-particular, our Gaussian approximation does permit $\chieff = 0$ for
-GW151226 while the true posterior rules this out at high confidence
+credible interval as quoted in \citet{O1-BBH}.  None of the $\chieff$
+posteriors are consistent with two black holes with large aligned
+spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
+inferred for the majority of black holes in X-ray binaries (see
+Section \ref{sec:discussion}).  The analysis here is relatively
+insensitive to the precise details of the posterior distributions;
+other conclusions are more sensitive.  In particular, our Gaussian
+approximation does permit $\chieff = 0$ for GW151226 while the true
+posterior rules this out at high confidence
 \citep{2016PhRvL.116x1103A,O1-BBH}.

### 2017-04-12 paper/AlignedIsotropicGW.tex
Citations, oh so many citations, in the abstract.

@@
 \begin{abstract}
-  The first direct detections of \acp{GW} from merging \acp{BBH} open
-  a unique window into the \ac{BBH} formation environment.  One
-  promising signature of the formation environment is the angular
-  distribution of the \ac{BH} spins; systems formed through dynamical
-  interactions among already-compact objects are expected to have
-  isotropic spin orientations whereas binaries formed from pairs of
-  stars born together are more likely to have spins preferentially
-  aligned with the binary orbit as a consequence of their joint
-  evolution toward a \ac{BBH} system.  We consider existing \ac{GW}
-  measurements of $\chieff$, the best-measured combination of spin
-  parameters, in the three likely binary black hole detections
-  GW150914, LVT151012, and GW151226.  If binary black hole spin
-  magnitudes extend to high values, as is suggested by observations of
-  black hole X-ray binaries, we show that the data already exhibit a
+  The first direct detections of \acp{GW}
+  \citep{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from merging
+  \acp{BBH} open a unique window into the \ac{BBH} formation
+  environment.  One promising signature of the formation environment
+  is the angular distribution of the \ac{BH} spins; systems formed
+  through dynamical interactions among already-compact objects are
+  expected to have isotropic spin orientations
+  \citep{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
+  whereas binaries formed from pairs of stars born together are more
+  likely to have spins preferentially aligned with the binary orbit as
+  a consequence of their joint evolution toward a \ac{BBH} system
+  \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
+  We consider existing \ac{GW} measurements of $\chieff$, the
+  best-measured combination of spin parameters
+  \citep{2016PhRvL.116x1102A,O1-BBH}, in the three likely binary black
+  hole detections GW150914, LVT151012, and GW151226.  If binary black
+  hole spin magnitudes extend to high values, as is suggested by
+  observations of black hole X-ray binaries
+  \citep{2015PhR...548....1M}, we show that the data already exhibit a
   $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
@@
   angular distribution.  By considering the effect of an additional 10
-  detections drawn from the various models in the suite we show that
-  if all observations come from a single population such an augmented
-  data set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
-  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
-  isotropic and aligned models for the assumed spin magnitude
-  distributions, and in most cases better than $5\sigma$
-  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
-  either an isotropic spin distribution or low spin magnitudes for the
-  observed systems will be confirmed (or overturned) confidently in
-  the near future by subsequent observations.
+  detections \citep{2016ApJ...833L...1A} drawn from the various models
+  in the suite we show that if all observations come from a single
+  population such an augmented data set would enable at least a
+  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
+  ratio) distinction between the isotropic and aligned models for the
+  assumed spin magnitude distributions, and in most cases better than
+  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
+  preference for either an isotropic spin distribution or low spin
+  magnitudes for the observed systems will be confirmed (or
+  overturned) confidently in the near future by subsequent
+  observations.
 \end{abstract}

### 2017-04-12 paper/AlignedIsotropicGW.tex
Removed bold first sentences in figures.

@@
 \includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
-\caption{\textbf{Fraction of the BBH population coming from an
-    isotropic distribution under a mixture model.} The dotted line
-  shows the flat prior on the fraction of BBHs coming from an
-  isotropic distribution, $f_i$, under the mixture model. The 3 red
-  lines show the posterior on $f_i$ after O1 with our various
-  assumptions regarding BH spin magnitudes.  The solid line shows the
-  posterior assuming that all BHs have their spin magnitude drawn from
-  the ``flat'' distribution. The dashed line assumes the ``high'' BH
-  spin magnitude distribution $p(a) = 2a$. The dot-dash line assumes
-  the ``low'' distribution $p(a) = 2(1-a)$.  We see that for a wide
-  range of assumptions regarding BH spin magnitudes, the fraction
-  coming from an isotropic distribution $f_i$ peaks at 1.}
+\caption{Fraction of the BBH population coming from an isotropic
+  distribution under a mixture model. The dotted line shows the flat
+  prior on the fraction of BBHs coming from an isotropic distribution,
+  $f_i$, under the mixture model. The 3 red lines show the posterior
+  on $f_i$ after O1 with our various assumptions regarding BH spin
+  magnitudes.  The solid line shows the posterior assuming that all
+  BHs have their spin magnitude drawn from the ``flat''
+  distribution. The dashed line assumes the ``high'' BH spin magnitude
+  distribution $p(a) = 2a$. The dot-dash line assumes the ``low''
+  distribution $p(a) = 2(1-a)$.  We see that for a wide range of
+  assumptions regarding BH spin magnitudes, the fraction coming from
+  an isotropic distribution $f_i$ peaks at 1.}
 \label{fig:mixture_fraction_posterior}
@@
 \includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
-\caption{\textbf{Effect of small spins on evidence ratio of aligned
-    against isotropic models.} The blue line shows the evidence ratio
-  (plotted as the equivalent sigma) between a model where all systems
-  are aligned, versus one where all systems are from an isotropic
-  distribution as a function of the power law $\alpha$ corresponding
-  to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
+\caption{Effect of small spins on evidence ratio of aligned against
+  isotropic models. The blue line shows the evidence ratio (plotted as
+  the equivalent sigma) between a model where all systems are aligned,
+  versus one where all systems are from an isotropic distribution as a
+  function of the power law $\alpha$ corresponding to
+  Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
   magnitude $\bar{a}$ which this $\alpha$ corresponds to. We see that

### 2017-04-12 paper/AlignedIsotropicGW.tex
Binary spirals in.

--- 2017-04-12 paper/AlignedIsotropicGW.tex
@@
 the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval as quoted in \citet{O1-BBH}.  None of the $\chieff$
-posteriors are consistent with two black holes with large aligned
-spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
-inferred for the majority of black holes in X-ray binaries (see
-Section \ref{sec:discussion}).  The analysis here is relatively
-insensitive to the precise details of the posterior distributions;
-other conclusions are more sensitive.  In particular, our Gaussian
-approximation does permit $\chieff = 0$ for GW151226 while the true
-posterior rules this out at high confidence
+credible interval as quoted in \citet{O1-BBH}.
+\ilya{None of the $\chieff$ posteriors are consistent with two black holes with large aligned spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred for the majority of black holes in X-ray binaries  (see Section \ref{sec:discussion}).}
+\sout{Since all three
+mergers involve black holes of comparable masses, there is essentially
+no posterior support for both black holes' $\chi \gtrsim 0.5$, in
+contrast to the majority of the electromagnetically-detected
+stellar-mass black hole population (see Section \ref{sec:discussion}).}
+The analysis here is relatively insensitive to the precise details of
+the posterior distributions; other conclusions are more sensitive.  In
+particular, our Gaussian approximation does permit $\chieff = 0$ for
+GW151226 while the true posterior rules this out at high confidence
 \citep{2016PhRvL.116x1103A,O1-BBH}.
@@
 \citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
-additional detections drawn from each of our six zero-parameter models
-for the spin distribution in addition to the three existing detections
-from O1, with observational uncertainties drawn randomly from the
-three Gaussian widths used to approximate the $\chieff$ posteriors in
-Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
-  $\chieff$ depends on the other parameters of the merging \ac{BBH}
-  system, particularly on the mass ratio.  Our assumption about future
-  observational uncertainties is appropriate if the parameters of the
-  existing three events are representative of the parameters of future
-  detections.}, we find the odds ratios shown in Figure
+additional detections with similar observational uncertainties
+\ilya{[I think it would be useful to specify (in an appendix?) what
+  model we use for the observational uncertainty on $\chieff$.]}
+drawn from each of our six zero-parameter models for the spin
+distribution in addition to the three existing detections from
+O1, we find the odds ratios shown in Figure
 \ref{fig:O2-predictions}.  We find that most scenarios with an
@@
 also evolve to a more significant misalignment in component spins as
-the binary shrinks from \ac{GW} emission \citep{2015PhRvD..92f4016G},
+the binary spirals in through \ac{GW} emission \citep{2015PhRvD..92f4016G},
 but $\chieff$ is approximately conserved through this evolution.

### 2017-04-12 paper/AlignedIsotropicGW.tex
Removed last red.

@@
 spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
-inferred for the majority of black holes in X-ray binaries (see
+inferred for the majority of black holes in X-ray binaries with claimed spin measurements (see
 Section \ref{sec:discussion}).  The analysis here is relatively
@@
     credible interval as reported in \citet{O1-BBH}.  It is notable
-    that none of the $\chieff$ posteriors \ilya{supports high \ac{BH}
-    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries} (see
+    that none of the $\chieff$ posteriors support high \ac{BH}
+    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries (see
     \citet{2015PhR...548....1M} for a summary of such measurements).}
@@
   observational uncertainties is appropriate if the parameters of the
-  existing three events are representative of the parameters of future
+  three detected events are representative of the parameters of future
   detections.}, we find the odds ratios shown in Figure

### 2017-04-12 paper/AlignedIsotropicGW.tex
Formatting
Added Nature statements about author contributions, etc.

@@
 
-\email{w.farr@bham.ac.uk, simons@star.sr.bham.ac.uk, miller@astro.umd.edu,\\imandel@star.sr.bham.ac.uk, farr@uchicago.edu, av@star.sr.bham.ac.uk}
+\email{w.farr@bham.ac.uk, simons@star.sr.bham.ac.uk, miller@astro.umd.edu,\\
+imandel@star.sr.bham.ac.uk, farr@uchicago.edu, av@star.sr.bham.ac.uk}
 
@@
 
+Supplementary Information is linked to the online version of the paper
+at \url{http://www.nature.com/nature}.
+
 We thank Richard O'Shaughnessy, Christopher Berry, and Davide Gerosa
@@
 National Science Foundation under Grant No. NSF PHY11-25915.
+
+All authors contributed to the work presented in this paper.
+
+Reprints and permissions information is available at
+\url{http://www.nature.com/reprints}.
+
+The authors declare no competing financial interests.
+
+Correspondence and requests for materials should be addressed to
+\href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
 

### 2017-04-12 paper/CoverLetter.tex
Cover letter.

% !TEX TS-program = pdflatexmk

\documentclass{letter}

\usepackage{hyperref}

\signature{Will M. Farr}

\address{Birmingham Institute for Gravitational Wave Astronomy\\School
  of Physics and Astronomy\\University of Birmingham\\Birmingham\\B15
  2TT\\United
  Kingdom\\\href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}}

\begin{document}

\begin{letter}{Dr.\ Leslie Sage\\968 National Press Building\\529 14th Street NW\\Washington DC 20045-1938\\United States}

\opening{Dear Dr. Sage:}

Please find enclosed a submission to be considered for a Nature
Letter.  We think these results are particularly exciting because they
are the first time the spin measurements in the existing three likely
gravitational wave events (GW150914, LVT151012, and GW151226) are
analysed together to constrain the formation scenario of the observed
population of merging binary black holes, a long-standing problem in
gravitational wave astronomy.  Additionally, the paper explains the
origin of and corrects a misconception in the existing literature
(particularly the Vitale, et. al (2017) paper referenced in our
submission) that $\mathcal{O}(100)$ detections will be required to
distinguish between different formation scenario in these objects.

Our submission has been formatted using the AASTeX LaTeX package.  We
estimate that the main body of the text, excluding the abstract, is
$\sim 1600$ words long, with four figures, each single-panel, and is
therefore within or very close to the Nature Letter limit.  The
abstract is $\sim 240$ words long (not counting references, which are
``Harvard style'' in the AASTeX package).  The document contains five
sections in an Appendix that we would like to be supplemental
information to the Letter.

\closing{Sincerely,}

\end{letter}

\end{document}

### 2017-04-12 paper/AlignedIsotropicGW.tex

@@
     $\Phi$ is the unit normal CDF.})  preference for an isotropic
-  angular distribution.  By considering the effect of an additional 10
+  angular distribution.  By considering the effect of 10 additional
   detections \citep{2016ApJ...833L...1A} drawn from the various models

### 2017-04-20 paper/AlignedIsotropicGW.tex
90% -> 99%

@@
 measurements, which bound at least one black hole to have spin
-magnitude $\geq 0.2$ at 90\% credibility \citep{2016PhRvL.116x1103A}.
+magnitude $\geq 0.2$ at 99\% credibility \citep{2016PhRvL.116x1103A}.
 

### 2017-05-08 paper/AlignedIsotropicGW.tex
Realised arXiver will only upload take three figures in a comment.
Add section on param-dependent chi_eff constraints

@@
 % arXiver will snag the right plots so they show up on Twitter :)
-%@arxiver{chi-eff-distributions.pdf,Wills_evidence_ratio_figure_with_mixture_models.png,chi-eff-mock-posteriors.pdf,six-way-O2-model-selection.pdf}
+%@arxiver{chi-eff-distributions.pdf,Wills_evidence_ratio_figure_with_mixture_models.png,six-way-O2-model-selection.pdf}
 
@@
   three detected events are representative of the parameters of future
-  detections.}, we find the odds ratios shown in Figure
+  detections. See Section \ref{sec:chi-eff-precision}
+for further discussion.}, we find the odds ratios shown in Figure
 \ref{fig:O2-predictions}.  We find that most scenarios with an
@@
 
+\section{Precision of $\chieff$ measurements}
+\label{sec:chi-eff-precision}
+Throughout this work we have made the simplifying assumption that the
+precision to which $\chieff$ can be constrained for individual binaries
+is independent of the binary's properties.  In practice, our ability to
+constrain $\chieff$ \emph{is} dependent on the system's properties, in
+particular its true $\chieff$ and mass ratio, which we illustrate in
+figure \ref{fig:chi-eff-constraints}.
+
+For this figure a detected population\footnote{We qualify a system as
+``detected'' if it produces a SNR above $8$ in the second-loudest
+detector to select only coincident events.} of 500 binaries was selected
+from a population with component masses distributed uniformly between
+$1$ and $30~\mathrm{M}_\odot$ with $m_1 + m_2 < 30~\mathrm{M}_\odot$,
+locations distributed uniformly in volume, and orientations distributed
+isotropically.  Data were simulated for each binary, and posteriors were
+estimated using the LIGO-Virgo parameter estimation library
+\texttt{LALInference}~\citep{2015PhRvD..91d2003V} using inspiral-only
+waveform models (merger and ringdown effects can provide additional
+information for some binaries, but we ignore those effects here).
+$\chieff$ is better constrained for binaries with low (negative)
+effective spins and low mass ratios.
+
+By neglecting these effects, We do not expect to qualitatively affect
+out conclusions, though they could affect predictions for the total
+number of detections necessary to constrain the population.  For
+example, if the universe preferentially forms asymmetric binaries with
+low mass ratios, individual $\chieff$ constraints will be systematically
+better, reducing the number of binaries needed to infer the properties
+of the population.
+
+\begin{figure}
+  \plotone{../plots/chi-eff-90cl}
+  \caption{Widths of the 90\% credible intervals for $\chieff$ for 500
+    binaries in a simulated detected population. $\chieff$ is better
+    constrained for systems with low $\chieff$ and low mass ratio.}
+  \label{fig:chi-eff-constraints}
+\end{figure}
+
 \bibliography{AlignedIsotropicGW}

### 2017-05-10 paper/AlignedIsotropicGW.tex
Initial transition to nature formatting, some problems remain.

@@
 
-\documentclass[modern,linenumbers]{aastex61}
+\documentclass{nature}
 
@@
 \usepackage{amssymb}
+\usepackage{hyperref}
+\usepackage{graphicx}
 
@@
 
+\newcommand{\plotone}[1]{\includegraphics[width=\columnwidth]{#1}}
+
+% Journal Abbreviations
+\newcommand{\aap}{Astronomy and Astrophysics}
+\newcommand{\apj}{Astrophysical Journal}
+\newcommand{\apjl}{Astrophysical Journal Letters}
+\newcommand{\apjs}{Astrophysical Journal Supplements}
+\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
+\newcommand{\nat}{Nature}
+\newcommand{\physrep}{Physics Reports}
+\newcommand{\prd}{Physical Review D}
+
 \begin{document}
@@
 
-\author[0000-0003-1540-8562]{Will M. Farr}
+\author{Will M. Farr$^1$, Simon Stevenson$^1$, M. Coleman Miller$^2$,
+  Ilya Mandel$^{1,3}$, Ben Farr$^4$ \& Alberto Vecchio$^1$}
 
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
+\maketitle
+
+\begin{affiliations}
+\item Birmingham Institute for Gravitational Wave Astronomy and
   School of Physics and Astronomy, University of Birmingham,
-  Birmingham, B15 2TT, United Kingdom}
-
-\author[0000-0002-6100-537X]{Simon Stevenson}
-
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
-  School of Physics and Astronomy, University of Birmingham,
-  Birmingham, B15 2TT, United Kingdom}
-\affiliation{Kavli Institute for Theoretical Physics, Santa Barbara, CA 93106}
-
-\author{M. Coleman Miller}
-
-\affiliation{Department of Astronomy and Joint Space-Science
+  Birmingham, B15 2TT, United Kingdom
+\item Department of Astronomy and Joint Space-Science
   Institute, University of Maryland, College Park, MD 20742--2421,
-  United States}
-
-\author[0000-0002-6134-8946]{Ilya Mandel}
-
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
-  School of Physics and Astronomy, University of Birmingham,
-  Birmingham, B15 2TT, United Kingdom}
-\affiliation{Kavli Institute for Theoretical Physics, Santa Barbara, CA 93106}
-
-\author[0000-0002-2916-9200]{Ben Farr}
-
-\affiliation{Enrico Fermi Institute and Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, United States}
-
-\author[0000-0002-6254-1617]{Alberto Vecchio}
-
-\affiliation{Birmingham Institute for Gravitational Wave Astronomy and
-  School of Physics and Astronomy, University of Birmingham,
-  Birmingham, B15 2TT, United Kingdom}
-
-\email{w.farr@bham.ac.uk, simons@star.sr.bham.ac.uk, miller@astro.umd.edu,\\
-imandel@star.sr.bham.ac.uk, farr@uchicago.edu, av@star.sr.bham.ac.uk}
+  United States
+\item Kavli Institute for Theoretical Physics, Santa Barbara, CA 93106
+\item Enrico Fermi Institute and Kavli Institute for Cosmological
+  Physics, University of Chicago, Chicago, IL 60637, United States
+\end{affiliations}
 
@@
   The first direct detections of \acp{GW}
-  \citep{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from merging
+  \cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from merging
   \acp{BBH} open a unique window into the \ac{BBH} formation
@@
   expected to have isotropic spin orientations
-  \citep{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
+  \cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
   whereas binaries formed from pairs of stars born together are more
@@
   a consequence of their joint evolution toward a \ac{BBH} system
-  \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
+  \cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
   We consider existing \ac{GW} measurements of $\chieff$, the
   best-measured combination of spin parameters
-  \citep{2016PhRvL.116x1102A,O1-BBH}, in the three likely binary black
+  \cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely binary black
   hole detections GW150914, LVT151012, and GW151226.  If binary black
@@
   observations of black hole X-ray binaries
-  \citep{2015PhR...548....1M}, we show that the data already exhibit a
+  \cite{2015PhR...548....1M}, we show that the data already exhibit a
   $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
@@
   angular distribution.  By considering the effect of 10 additional
-  detections \citep{2016ApJ...833L...1A} drawn from the various models
+  detections \cite{2016ApJ...833L...1A} drawn from the various models
   in the suite we show that if all observations come from a single
@@
 Following the detection of a merging \ac{BBH} system, \ac{PE} tools
-\citep{2015PhRvD..91d2003V} compare model gravitational waveforms
-\citep[e.g.\
-][]{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
+\cite{2015PhRvD..91d2003V} compare model gravitational waveforms
+\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
 against the observed data to obtain a posterior distribution on the
@@
 projections of the individual \ac{BH} spins
-\citep{2016PhRvL.116x1102A}.
+\cite{2016PhRvL.116x1102A}.
 
@@
 GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
-run (O1) \citep{O1-BBH}.  Because samples drawn from the posterior on
+run (O1) \cite{O1-BBH}.  Because samples drawn from the posterior on
 $\chieff$ are not publicly released at this time, we have approximated
 the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval as quoted in \citet{O1-BBH}.  None of the $\chieff$
+credible interval as quoted in \cite{O1-BBH}.  None of the $\chieff$
 posteriors are consistent with two black holes with large aligned
@@
 posterior rules this out at high confidence
-\citep{2016PhRvL.116x1103A,O1-BBH}.
+\cite{2016PhRvL.116x1103A,O1-BBH}.
 
@@
     $\chieff$ from the Advanced LIGO O1 run observations in
-    \citet{O1-BBH}.  We approximate the posteriors reported in
-    \citet{O1-BBH} using Gaussians with the same median and 90\%
-    credible interval as reported in \citet{O1-BBH}.  It is notable
+    \cite{O1-BBH}.  We approximate the posteriors reported in
+    \cite{O1-BBH} using Gaussians with the same median and 90\%
+    credible interval as reported in \cite{O1-BBH}.  It is notable
     that none of the $\chieff$ posteriors support high \ac{BH}
     spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries (see
-    \citet{2015PhR...548....1M} for a summary of such measurements).}
+    \cite{2015PhR...548....1M} for a summary of such measurements).}
 \end{figure}
@@
 that the merging black holes are of equal mass (this is marginally
-consistent with the three observations \citep{O1-BBH}, and the
+consistent with the three observations \cite{O1-BBH}, and the
 $\chieff$ distribution is not particularly sensitive to the mass ratio
@@
 synthesis codes can at this point authoritatively suggest \emph{any}
-particular spin distribution \citep{2015PhR...548....1M}.  Our models,
+particular spin distribution \cite{2015PhR...548....1M}.  Our models,
 however, allow us to see how sensitive the $\chieff$ distribution is
@@
 measurements, which bound at least one black hole to have spin
-magnitude $\geq 0.2$ at 99\% credibility \citep{2016PhRvL.116x1103A}.
+magnitude $\geq 0.2$ at 99\% credibility \cite{2016PhRvL.116x1103A}.
 
@@
 chance of 10 additional \ac{BBH} detections in the next two years
-\citep{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
+\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
 additional detections drawn from each of our six zero-parameter models
@@
 permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
-the GW151226 observations \citep{2016PhRvL.116x1103A}.
+the GW151226 observations \cite{2016PhRvL.116x1103A}.
 
@@
 Observational data for such model selection is sparse.
-\citet{2015PhR...548....1M} give current estimates of the spin
+\cite{2015PhR...548....1M} give current estimates of the spin
 parameters for stellar-mass black holes, obtained using disk
@@
 In addition, there are substantial systematic errors that can
-complicate either type of analysis (see \citealt{2015PhR...548....1M}
-for a discussion), and selection effects could yield a biased
+complicate either type of analysis (see \cite{2015PhR...548....1M} for
+a discussion), and selection effects could yield a biased
 distribution. Nonetheless, if we take the reported spin magnitudes as
@@
 which they transfer mass; a variant of this long-standing argument was
-presented by \citet{1999MNRAS.305..654K}), but see
-\citet{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current
-spin parameters probably are close to their values upon core collapse,
-at least in high-mass X-ray binaries.  However, the specific processes
+presented by \cite{1999MNRAS.305..654K}), but see
+\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current spin
+parameters probably are close to their values upon core collapse, at
+least in high-mass X-ray binaries.  However, the specific processes
 involved in the production of black hole binaries from isolated
@@
 could reduce the available angular momentum
-\citep{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
+\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
 
@@
 interactions in dense stellar environments
-\citep{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
 are expected to be isotropic given the absence of a preferred
-direction \citep[e.g.,][]{2016ApJ...832L...2R} and the persistence of
-an isotropic distribution through post-Newtonian evolution
-\citep{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
+direction \cite{2016ApJ...832L...2R} and the persistence of an
+isotropic distribution through post-Newtonian evolution
+\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
 
@@
 classical channel through a common-envelope phase
-\citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
+\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
 or through chemically homogeneous evolution
-\citep{MandeldeMink:2016,Marchant:2016} are usually expected to be
+\cite{MandeldeMink:2016,Marchant:2016} are usually expected to be
 preferentially aligned.  Despite observed spin-orbit misalignments in
 both massive stellar binaries
-\citep{Albrecht:2009,2014ApJ...785...83A} and BH X-ray binaries
-\citep{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
+\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray binaries
+\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
 mass accretion and tidal interactions will tend to realign the binary.
@@
 orbital plane and misalign the binary
-\citep{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can
+\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can
 also tilt the spin angle, as in the double pulsar J0737-3039
-\citep{2011ApJ...742...81F}; and a variety of uncertain processes,
+\cite{2011ApJ...742...81F}; and a variety of uncertain processes,
 such as wind-driven mass loss and post-collapse fallback, can couple
@@
 also evolve to a more significant misalignment in component spins as
-the binary spirals in through \ac{GW} emission \citep{2015PhRvD..92f4016G},
+the binary spirals in through \ac{GW} emission \cite{2015PhRvD..92f4016G},
 but $\chieff$ is approximately conserved through this evolution.
 
-\citet{2017CQGra..34cLT01V} also studied the possibility of
+\cite{2017CQGra..34cLT01V} also studied the possibility of
 distinguishing aligned and isotropic angular distributions of \ac{BBH}
@@
 anti-parallel spins for an ``aligned'' population.  In contrast, and
-in agreement with the present study, \citet{Stevenson:2017spin} found
+in agreement with the present study, \cite{Stevenson:2017spin} found
 that only $\sim 5$ observations of rapidly spinning black holes would
@@
 models.  Meanwhile, recent studies by
-\citet{2017arXiv170306869F} and \citet{2017arXiv170306223G} focused on inference
+\cite{2017arXiv170306869F} and \cite{2017arXiv170306223G} focused on inference
 on spin magnitudes.
@@
 
-\acknowledgements
-
-Supplementary Information is linked to the online version of the paper
-at \url{http://www.nature.com/nature}.
-
-We thank Richard O'Shaughnessy, Christopher Berry, and Davide Gerosa
-for discussions and comments on this work.  WF, SS, IM and AV were
-supported in part by the STFC.  MCM acknowledges support of the
+\begin{addendum}
+\item We thank Richard O'Shaughnessy, Christopher Berry, and Davide
+  Gerosa for discussions and comments on this work.  WF, SS, IM and AV
+  were supported in part by the STFC.  MCM acknowledges support of the
   University of Birmingham Institute for Advanced Study Distinguished
@@
   National Science Foundation under Grant No. NSF PHY11-25915.
+\item[Supplementary Information] Supplementary Information is linked
+  to the online version of the paper at
+  \url{http://www.nature.com/nature}.
+\item[Author Contributions] All authors contributed to the work
+  presented in this paper.
+\item[Author Information] Reprints and permissions information is
+  available at \url{http://www.nature.com/reprints}.  The authors
+  declare no competing financial interests.  Correspondence and
+  requests for materials should be addressed to
+  \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
+\end{addendum}
 
-All authors contributed to the work presented in this paper.
-
-Reprints and permissions information is available at
-\url{http://www.nature.com/reprints}.
-
-The authors declare no competing financial interests.
-
-Correspondence and requests for materials should be addressed to
-\href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
-
-\appendix
+\begin{methods}
 
@@
 
-Following \citet{Stevenson:2017spin}, we fit a mixture model (labelled
+Following \cite{Stevenson:2017spin}, we fit a mixture model (labelled
 model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs
@@
 measurement uncertainty in our analysis; thus our analysis must be
-\emph{hierarchical} \citep{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
+\emph{hierarchical} \cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
 In a hierarchical analysis, we assume that each event has a true, but
@@
 the parameters describing GW150914 has been investigated in detail
-\citep{2016arXiv161107531T}.  For this source, statistical uncertainties dominate
+\cite{2016arXiv161107531T}.  For this source, statistical uncertainties dominate
 over any waveform systematics.  Detailed comparisons with numerical
 relativity computations using no approximations to the dynamics
-\citep{2016PhRvD..94f4035A} also suggest that statistical
+\cite{2016PhRvD..94f4035A} also suggest that statistical
 uncertainties dominate the systematics for this system.
-\citet{2016arXiv161107531T} suggests that systematics may dominate for
+\cite{2016arXiv161107531T} suggests that systematics may dominate for
 signals with this large SNR ($\simeq 23$) when the source is edge-on
@@
 estimated using the LIGO-Virgo parameter estimation library
-\texttt{LALInference}~\citep{2015PhRvD..91d2003V} using inspiral-only
+\texttt{LALInference}~\cite{2015PhRvD..91d2003V} using inspiral-only
 waveform models (merger and ringdown effects can provide additional
@@
 
+\end{methods}
+
+\bibliographystyle{naturemag}
 \bibliography{AlignedIsotropicGW}

### 2017-05-18 paper/AlignedIsotropicGW.tex
First pass fixing reference styles.  Fixes #21.

@@
 \usepackage{amssymb}
+\usepackage{color}
 \usepackage{hyperref}
@@
 \begin{abstract}
-  The first direct detections of \acp{GW}
-  \cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from merging
-  \acp{BBH} open a unique window into the \ac{BBH} formation
+  The first direct detections of
+  \acp{GW}\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
+  merging \acp{BBH} open a unique window into the \ac{BBH} formation
   environment.  One promising signature of the formation environment
@@
   through dynamical interactions among already-compact objects are
-  expected to have isotropic spin orientations
-  \cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
+  expected to have isotropic spin
+  orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
   whereas binaries formed from pairs of stars born together are more
   likely to have spins preferentially aligned with the binary orbit as
-  a consequence of their joint evolution toward a \ac{BBH} system
-  \cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
+  a consequence of their joint evolution toward a \ac{BBH}
+  system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
   We consider existing \ac{GW} measurements of $\chieff$, the
-  best-measured combination of spin parameters
-  \cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely binary black
-  hole detections GW150914, LVT151012, and GW151226.  If binary black
-  hole spin magnitudes extend to high values, as is suggested by
-  observations of black hole X-ray binaries
-  \cite{2015PhR...548....1M}, we show that the data already exhibit a
-  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
+  best-measured combination of spin
+  parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
+  binary black hole detections GW150914, LVT151012, and GW151226.  If
+  binary black hole spin magnitudes extend to high values, as is
+  suggested by observations of black hole X-ray
+  binaries\cite{2015PhR...548....1M}, we show that the data already
+  exhibit a $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
   ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
@@
   angular distribution.  By considering the effect of 10 additional
-  detections \cite{2016ApJ...833L...1A} drawn from the various models
+  detections\cite{2016ApJ...833L...1A} drawn from the various models
   in the suite we show that if all observations come from a single
@@
 
-Following the detection of a merging \ac{BBH} system, \ac{PE} tools
-\cite{2015PhRvD..91d2003V} compare model gravitational waveforms
-\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
+Following the detection of a merging \ac{BBH} system, \ac{PE}
+tools\cite{2015PhRvD..91d2003V} compare model gravitational
+waveforms\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
 against the observed data to obtain a posterior distribution on the
@@
 $0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
-projections of the individual \ac{BH} spins
-\cite{2016PhRvL.116x1102A}.
+projections of the individual \ac{BH} spins\cite{2016PhRvL.116x1102A}.
 
@@
 GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
-run (O1) \cite{O1-BBH}.  Because samples drawn from the posterior on
+run (O1)\cite{O1-BBH}.  Because samples drawn from the posterior on
 $\chieff$ are not publicly released at this time, we have approximated
 the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval as quoted in \cite{O1-BBH}.  None of the $\chieff$
-posteriors are consistent with two black holes with large aligned
-spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
-inferred for the majority of black holes in X-ray binaries with claimed spin measurements (see
-Section \ref{sec:discussion}).  The analysis here is relatively
-insensitive to the precise details of the posterior distributions;
-other conclusions are more sensitive.  In particular, our Gaussian
-approximation does permit $\chieff = 0$ for GW151226 while the true
-posterior rules this out at high confidence
-\cite{2016PhRvL.116x1103A,O1-BBH}.
+credible interval.  None of the $\chieff$ posteriors are consistent
+with two black holes with large aligned spins,
+$\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred
+for the majority of black holes in X-ray binaries with claimed spin
+measurements (see Section \ref{sec:discussion}).  The analysis here is
+relatively insensitive to the precise details of the posterior
+distributions; other conclusions are more sensitive.  In particular,
+our Gaussian approximation does permit $\chieff = 0$ for GW151226
+while the true posterior rules this out at high
+confidence\cite{2016PhRvL.116x1103A,O1-BBH}.
 
@@
   \caption{\label{fig:O1-posteriors} Approximate posteriors on
-    $\chieff$ from the Advanced LIGO O1 run observations in
-    \cite{O1-BBH}.  We approximate the posteriors reported in
-    \cite{O1-BBH} using Gaussians with the same median and 90\%
-    credible interval as reported in \cite{O1-BBH}.  It is notable
-    that none of the $\chieff$ posteriors support high \ac{BH}
-    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries (see
-    \cite{2015PhR...548....1M} for a summary of such measurements).}
+    $\chieff$ from the Advanced LIGO O1 run observations\cite{O1-BBH}.
+    We approximate the posteriors reported using Gaussians with the
+    same median and 90\% credible interval.  It is notable that none
+    of the $\chieff$ posteriors support high \ac{BH} spin magnitudes
+    with aligned spins, suggested by observations of stellar-mass
+    black holes in X-ray binaries\cite{2015PhR...548....1M}.}
 \end{figure}
@@
 that the merging black holes are of equal mass (this is marginally
-consistent with the three observations \cite{O1-BBH}, and the
+consistent with the three observations\cite{O1-BBH}, and the
 $\chieff$ distribution is not particularly sensitive to the mass ratio
@@
 synthesis codes can at this point authoritatively suggest \emph{any}
-particular spin distribution \cite{2015PhR...548....1M}.  Our models,
+particular spin distribution\cite{2015PhR...548....1M}.  Our models,
 however, allow us to see how sensitive the $\chieff$ distribution is
@@
 measurements, which bound at least one black hole to have spin
-magnitude $\geq 0.2$ at 99\% credibility \cite{2016PhRvL.116x1103A}.
+magnitude $\geq 0.2$ at 99\% credibility\cite{2016PhRvL.116x1103A}.
 
@@
 Estimates of the rate of \ac{BBH} coalescences give a reasonable
-chance of 10 additional \ac{BBH} detections in the next two years
-\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
-additional detections drawn from each of our six zero-parameter models
-for the spin distribution in addition to the three existing detections
-from O1, with observational uncertainties drawn randomly from the
-three Gaussian widths used to approximate the $\chieff$ posteriors in
-Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
-  $\chieff$ depends on the other parameters of the merging \ac{BBH}
-  system, particularly on the mass ratio.  Our assumption about future
-  observational uncertainties is appropriate if the parameters of the
-  three detected events are representative of the parameters of future
-  detections. See Section \ref{sec:chi-eff-precision}
-for further discussion.}, we find the odds ratios shown in Figure
-\ref{fig:O2-predictions}.  We find that most scenarios with an
-additional 10 detections allow the simulated angular distribution to
-be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
-credibility; and in the most pessimistic case the distinction is
-typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
-odds ratio).  While such future detections should permit a confident
-distinction between \emph{angular} distributions, we would remain much
-less certain about the \emph{magnitude} distribution among the three
-options considered here until we have a larger number of observations.
-In Figure \ref{fig:O2-predictions}, the odds ratio between different
-magnitude distributions with the same angular distribution is much
-closer to unity than the odds ratio between angular distributions.
+chance of 10 additional \ac{BBH} detections in the next two
+years\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming
+10 additional detections drawn from each of our six zero-parameter
+models for the spin distribution in addition to the three existing
+detections from O1, with observational uncertainties drawn randomly
+from the three Gaussian widths used to approximate the $\chieff$
+posteriors in Figure \ref{fig:O1-posteriors}\footnote{The measurement
+  uncertainty in $\chieff$ depends on the other parameters of the
+  merging \ac{BBH} system, particularly on the mass ratio.  Our
+  assumption about future observational uncertainties is appropriate
+  if the parameters of the three detected events are representative of
+  the parameters of future detections. See Section
+  \ref{sec:chi-eff-precision} for further discussion.}, we find the
+odds ratios shown in Figure \ref{fig:O2-predictions}.  We find that
+most scenarios with an additional 10 detections allow the simulated
+angular distribution to be inferred with greater than $5\sigma$
+($2.9 \times 10^{-7}$ odds) credibility; and in the most pessimistic
+case the distinction is typically $\OTwoSigmaIsoAlignedMin\sigma$
+($\OTwoOddsIsoAlignedMin$ odds ratio).  While such future detections
+should permit a confident distinction between \emph{angular}
+distributions, we would remain much less certain about the
+\emph{magnitude} distribution among the three options considered here
+until we have a larger number of observations.  In Figure
+\ref{fig:O2-predictions}, the odds ratio between different magnitude
+distributions with the same angular distribution is much closer to
+unity than the odds ratio between angular distributions.
 
@@
 permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
-the GW151226 observations \cite{2016PhRvL.116x1103A}.
+the GW151226 observations\cite{2016PhRvL.116x1103A}.
 
@@
 distribution of the spin magnitudes of stellar-mass black holes.
-Observational data for such model selection is sparse.
-\cite{2015PhR...548....1M} give current estimates of the spin
-parameters for stellar-mass black holes, obtained using disk
-reflection and/or disk continuum methods.  Most of the systems studied
-are low-mass X-ray binaries rather than the high-mass X-ray binaries
-that are likely to be the progenitors of double black hole binaries.
-In addition, there are substantial systematic errors that can
-complicate either type of analysis (see \cite{2015PhR...548....1M} for
-a discussion), and selection effects could yield a biased
-distribution. Nonetheless, if we take the reported spin magnitudes as
-representative then we find that there is a preference for high spins;
-for example, 14 of the 19 systems with reported spins have
-dimensionless spin parameters in excess of 0.5.  It is usually argued
-that the masses and spin parameters of stellar-mass black holes are
-unlikely to be altered significantly by accretion (low-mass donors may
-not have enough mass and high-mass donors have a very short phase in
-which they transfer mass; a variant of this long-standing argument was
-presented by \cite{1999MNRAS.305..654K}), but see
-\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current spin
-parameters probably are close to their values upon core collapse, at
-least in high-mass X-ray binaries.  However, the specific processes
-involved in the production of black hole binaries from isolated
-binaries could alter the spin magnitude distribution of those holes
-relative to the X-ray binary systems; for example, close tidal
+Observational data for such model selection is
+sparse\cite{2015PhR...548....1M}.  Most of the systems studied are
+low-mass X-ray binaries rather than the high-mass X-ray binaries that
+are likely to be the progenitors of double black hole binaries.  In
+addition, there are substantial systematic errors that can complicate
+either type of analysis\cite{2015PhR...548....1M} and selection
+effects could yield a biased distribution. Nonetheless, if we take the
+reported spin magnitudes as representative then we find that there is
+a preference for high spins; for example, 14 of the 19 systems with
+reported spins have dimensionless spin parameters in excess of 0.5.
+It is usually argued that the masses and spin parameters of
+stellar-mass black holes are unlikely to be altered significantly by
+accretion (low-mass donors may not have enough mass and high-mass
+donors have a very short phase in which they transfer
+mass\cite{1999MNRAS.305..654K}), but this may not be true for all
+systems\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the
+current spin parameters probably are close to their values upon core
+collapse, at least in high-mass X-ray binaries.  However, the specific
+processes involved in the production of black hole binaries from
+isolated binaries could alter the spin magnitude distribution of those
+holes relative to the X-ray binary systems; for example, close tidal
 interactions could spin up the core, or stripping of the envelope
-could reduce the available angular momentum
-\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
+could reduce the available angular
+momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
 
 The spin directions of binary black holes formed dynamically through
-interactions in dense stellar environments
-\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+interactions in dense stellar
+environments\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
 are expected to be isotropic given the absence of a preferred
-direction \cite{2016ApJ...832L...2R} and the persistence of an
-isotropic distribution through post-Newtonian evolution
-\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
+direction\cite{2016ApJ...832L...2R} and the persistence of an
+isotropic distribution through post-Newtonian
+evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
 
 The spin directions in isolated binaries, whether evolving via the
-classical channel through a common-envelope phase
-\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
-or through chemically homogeneous evolution
-\cite{MandeldeMink:2016,Marchant:2016} are usually expected to be
-preferentially aligned.  Despite observed spin-orbit misalignments in
-both massive stellar binaries
-\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray binaries
-\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
+classical channel through a common-envelope
+phase\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
+or through chemically homogeneous
+evolution\cite{MandeldeMink:2016,Marchant:2016} are usually expected
+to be preferentially aligned.  Despite observed spin-orbit
+misalignments in both massive stellar
+binaries\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray
+binaries\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
 mass accretion and tidal interactions will tend to realign the binary.
 On the other hand, a supernova natal kick (if any) can change the
-orbital plane and misalign the binary
-\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can
-also tilt the spin angle, as in the double pulsar J0737-3039
-\cite{2011ApJ...742...81F}; and a variety of uncertain processes,
-such as wind-driven mass loss and post-collapse fallback, can couple
-the spin magnitude and direction distributions, contrary to our
-simplified assumptions.  A small misalignment at wide separation can
-also evolve to a more significant misalignment in component spins as
-the binary spirals in through \ac{GW} emission \cite{2015PhRvD..92f4016G},
-but $\chieff$ is approximately conserved through this evolution.
+orbital plane and misalign the
+binary\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova
+can also tilt the spin angle, as in the double pulsar
+J0737-3039\cite{2011ApJ...742...81F}; and a variety of uncertain
+processes, such as wind-driven mass loss and post-collapse fallback,
+can couple the spin magnitude and direction distributions, contrary to
+our simplified assumptions.  A small misalignment at wide separation
+can also evolve to a more significant misalignment in component spins
+as the binary spirals in through \ac{GW}
+emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
+conserved through this evolution.
 
-\cite{2017CQGra..34cLT01V} also studied the possibility of
-distinguishing aligned and isotropic angular distributions of \ac{BBH}
-spins, but concluded that several hundred sources would be required to
-adequately separate models which included both parallel and
-anti-parallel spins for an ``aligned'' population.  In contrast, and
-in agreement with the present study, \cite{Stevenson:2017spin} found
-that only $\sim 5$ observations of rapidly spinning black holes would
-be necessary to distinguish isotropic and aligned spin distributions,
-though tens of detections would be required for more nuanced admixture
-models.  Meanwhile, recent studies by
-\cite{2017arXiv170306869F} and \cite{2017arXiv170306223G} focused on inference
-on spin magnitudes.
+\fixme{\cite{2017CQGra..34cLT01V} also studied the possibility of
+  distinguishing aligned and isotropic angular distributions of
+  \ac{BBH} spins, but concluded that several hundred sources would be
+  required to adequately separate models which included both parallel
+  and anti-parallel spins for an ``aligned'' population.  In contrast,
+  and in agreement with the present study,\cite{Stevenson:2017spin}
+  found that only $\sim 5$ observations of rapidly spinning black
+  holes would be necessary to distinguish isotropic and aligned spin
+  distributions, though tens of detections would be required for more
+  nuanced admixture models.  Meanwhile, recent studies
+  by\cite{2017arXiv170306869F} and\cite{2017arXiv170306223G} focused
+  on inference on spin magnitudes.}
 
@@
 
-Following \cite{Stevenson:2017spin}, we fit a mixture model (labelled
-model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs
-have spins drawn from an isotropic distribution, whilst a fraction
+We also fit a mixture model\cite{Stevenson:2017spin} (labelled model
+'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have
+spins drawn from an isotropic distribution, whilst a fraction
 $1 - f_i$ have their spins aligned with the orbital angular
@@
 measurement uncertainty in our analysis; thus our analysis must be
-\emph{hierarchical} \cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
+\emph{hierarchical}\cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
 In a hierarchical analysis, we assume that each event has a true, but
@@
 mis-aligned spins.  The effect of these approximations on inference on
-the parameters describing GW150914 has been investigated in detail
-\cite{2016arXiv161107531T}.  For this source, statistical uncertainties dominate
-over any waveform systematics.  Detailed comparisons with numerical
-relativity computations using no approximations to the dynamics
-\cite{2016PhRvD..94f4035A} also suggest that statistical
-uncertainties dominate the systematics for this system.
-\cite{2016arXiv161107531T} suggests that systematics may dominate for
-signals with this large SNR ($\simeq 23$) when the source is edge-on
-or has high spins.  The other two events discussed in this paper are
-at much lower SNR, with correspondingly larger statistical
-uncertainties, and are probably similarly oriented and with similarly
-small spins, so we do not expect systematic uncertainties to dominate.
+the parameters describing GW150914 has been investigated in
+detail\cite{2016arXiv161107531T}.  For this source, statistical
+uncertainties dominate over any waveform systematics.  Detailed
+comparisons with numerical relativity computations using no
+approximations to the dynamics\cite{2016PhRvD..94f4035A} also suggest
+that statistical uncertainties dominate the systematics for this
+system.  Systematics may dominate for signals with this large SNR
+($\simeq 23$) when the source is edge-on or has high
+spins\cite{2016arXiv161107531T}.  The other two events discussed in
+this paper are at much lower SNR, with correspondingly larger
+statistical uncertainties, and are probably similarly oriented and
+with similarly small spins, so we do not expect systematic
+uncertainties to dominate.
 
@@
   ``detected'' if it produces a SNR above $8$ in the second-loudest
-detector to select only coincident events.} of 500 binaries was selected
-from a population with component masses distributed uniformly between
-$1$ and $30~\mathrm{M}_\odot$ with $m_1 + m_2 < 30~\mathrm{M}_\odot$,
-locations distributed uniformly in volume, and orientations distributed
-isotropically.  Data were simulated for each binary, and posteriors were
-estimated using the LIGO-Virgo parameter estimation library
-\texttt{LALInference}~\cite{2015PhRvD..91d2003V} using inspiral-only
+  detector to select only coincident events.} of 500 binaries was
+selected from a population with component masses distributed uniformly
+between $1$ and $30~\mathrm{M}_\odot$ with
+$m_1 + m_2 < 30~\mathrm{M}_\odot$, locations distributed uniformly in
+volume, and orientations distributed isotropically.  Data were
+simulated for each binary, and posteriors were estimated using the
+LIGO-Virgo parameter estimation library
+\texttt{LALInference}\cite{2015PhRvD..91d2003V} using inspiral-only
 waveform models (merger and ringdown effects can provide additional

### 2017-05-24 paper/AlignedIsotropicGW.tex
Eliminate abbreviations in abstract, add in bibliography journals.

@@
 % Journal Abbreviations
-\newcommand{\aap}{Astronomy and Astrophysics}
-\newcommand{\apj}{Astrophysical Journal}
-\newcommand{\apjl}{Astrophysical Journal Letters}
-\newcommand{\apjs}{Astrophysical Journal Supplements}
-\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
-\newcommand{\nat}{Nature}
-\newcommand{\physrep}{Physics Reports}
-\newcommand{\prd}{Physical Review D}
+\newcommand{\aap}{Astron. and Astrophys.}
+\newcommand{\apj}{Astrophys. J.}
+\newcommand{\apjl}{Astrophys. J. Lett.}
+\newcommand{\apjs}{Astrophys. J. Supp.}
+\newcommand{\mnras}{Mon. Not. R. Astron. Soc.}
+\newcommand{\nat}{Nat.}
+\newcommand{\physrep}{Phys. Rep.}
+\newcommand{\prd}{Phys. Rev. D}
 
@@
 \begin{abstract}
-  The first direct detections of
-  \acp{GW}\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
-  merging \acp{BBH} open a unique window into the \ac{BBH} formation
-  environment.  One promising signature of the formation environment
-  is the angular distribution of the \ac{BH} spins; systems formed
-  through dynamical interactions among already-compact objects are
-  expected to have isotropic spin
+  The first direct detections of gravitational
+  waves\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
+  merging binary black holes open a unique window into the binary
+  black hole formation environment.  One promising signature of the
+  formation environment is the angular distribution of the black hole
+  spins; systems formed through dynamical interactions among
+  already-compact objects are expected to have isotropic spin
   orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
@@
   likely to have spins preferentially aligned with the binary orbit as
-  a consequence of their joint evolution toward a \ac{BBH}
+  a consequence of their joint evolution toward a binary black hole
   system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
-  We consider existing \ac{GW} measurements of $\chieff$, the
-  best-measured combination of spin
+  We consider existing gravitational wave measurements of $\chieff$,
+  the best-measured combination of spin
   parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely

### 2017-05-24 paper/AlignedIsotropicGW.tex
Better wording in abstract, closes #15.

@@
   system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
-  We consider existing gravitational wave measurements of $\chieff$,
-  the best-measured combination of spin
+  We consider existing gravitational wave measurements of the binary
+  effective spin, the best-measured combination of spin
   parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
@@
   angular distribution.  By considering the effect of 10 additional
-  detections\cite{2016ApJ...833L...1A} drawn from the various models
-  in the suite we show that if all observations come from a single
-  population such an augmented data set would enable at least a
-  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
-  ratio) distinction between the isotropic and aligned models for the
-  assumed spin magnitude distributions, and in most cases better than
-  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
-  preference for either an isotropic spin distribution or low spin
-  magnitudes for the observed systems will be confirmed (or
-  overturned) confidently in the near future by subsequent
-  observations.
+  detections\cite{2016ApJ...833L...1A} of systems, we show that if all
+  observations come from a single population such an augmented data
+  set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
+  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
+  isotropic and aligned models for the assumed spin magnitude
+  distributions, and in most cases better than $5\sigma$
+  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
+  either an isotropic spin distribution or low spin magnitudes for the
+  observed systems will be confirmed (or overturned) confidently in
+  the near future by subsequent observations.
 \end{abstract}

### 2017-05-24 paper/AlignedIsotropicGW.tex
Removed section headings from main body.

@@
 
-\section{\ac{GW} Spin Measurements and Model Selection}
-\label{sec:O1}
-
 Following the detection of a merging \ac{BBH} system, \ac{PE}
@@
 for the majority of black holes in X-ray binaries with claimed spin
-measurements (see Section \ref{sec:discussion}).  The analysis here is
+measurements (see below).  The analysis here is
 relatively insensitive to the precise details of the posterior
@@
 uncertainty about the spin magnitude distribution, discussed in detail
-in Section \ref{sec:discussion}; neither observations nor population
-synthesis codes can at this point authoritatively suggest \emph{any}
-particular spin distribution\cite{2015PhR...548....1M}.  Our models,
-however, allow us to see how sensitive the $\chieff$ distribution is
-to spin alignment given uncertainties about the spin magnitudes.
+below; neither observations nor population synthesis codes can at this
+point authoritatively suggest \emph{any} particular spin
+distribution\cite{2015PhR...548....1M}.  Our models, however, allow us
+to see how sensitive the $\chieff$ distribution is to spin alignment
+given uncertainties about the spin magnitudes.
 
@@
     predicted with 10 additional observations above the three
-    discussed in Section \ref{sec:O1}.  Each panel corresponds to
-    additional observations drawn from one of the $\chieff$
-    distribution models.  The model from which the additional
-    observations are drawn is outlined in red.  The height of the blue
-    bar gives the median odds ratio relative to the model from which
-    the additional observations are drawn; the green line gives the
-    68\% ($1 \sigma$) symmetric interval of odds ratios over 1000
-    separate draws from the model distribution.  The closest median
-    ratio between the most-favoured isotropic model and the
-    most-favoured aligned model is $\OTwoOddsIsoAlignedMin$,
-    corresponding to $\OTwoSigmaIsoAlignedMin\sigma$ preference for
-    the correct angular distribution; most models result in more than
-    $5\sigma$ preference for the correct angular distribution.
-    Because the three observations from Section \ref{sec:O1} are
-    included in each data set the ``correct'' model is not necessarily
-    preferred over the others, particularly when that model uses the
-    ``high'' magnitude distribution, which is strongly dis-favoured
-    from the O1 observations alone.}
+    discussed above.  Each panel corresponds to additional
+    observations drawn from one of the $\chieff$ distribution models.
+    The model from which the additional observations are drawn is
+    outlined in red.  The height of the blue bar gives the median odds
+    ratio relative to the model from which the additional observations
+    are drawn; the green line gives the 68\% ($1 \sigma$) symmetric
+    interval of odds ratios over 1000 separate draws from the model
+    distribution.  The closest median ratio between the most-favoured
+    isotropic model and the most-favoured aligned model is
+    $\OTwoOddsIsoAlignedMin$, corresponding to
+    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
+    distribution; most models result in more than $5\sigma$ preference
+    for the correct angular distribution.  Because the three existing
+    observations are included in each data set the ``correct'' model
+    is not necessarily preferred over the others, particularly when
+    that model uses the ``high'' magnitude distribution, which is
+    strongly dis-favoured from the O1 observations alone.}
 \end{figure}
-
-\section{Discussion}
-\label{sec:discussion}
 
@@
 of the isotropic models, which is about a factor of two smaller than
-the ratio in the more careful calculation described in Section
-\ref{sec:O1}.  This is a satisfactory answer at an order-of-magnitude
-level.
+the ratio in the more careful calculation described above.  This is a
+satisfactory answer at an order-of-magnitude level.
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #16, moving equations to methods section.

@@
 The spin parameter with the largest effect on waveforms, and a
-correspondingly tight constraint from the data, is a mass-weighted,
-aligned, ``effective spin,''
-\begin{equation}
-  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
-  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
-\end{equation}
-where $m_{1,2}$ are the gravitational masses of the more-massive (1)
-and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
-$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
-holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
-assumed to point in the $\hat{z}$ direction, and
-$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
-projections of the individual \ac{BH} spins\cite{2016PhRvL.116x1102A}.
+correspondingly tight constraint from the
+data\cite{2016PhRvL.116x1102A}, is a mass-weighted combination of the
+components of the dimensionless spin vectors of the two black holes
+that are aligned with the orbital axis, the ``effective spin,''
+$-1 \leq \chieff \leq 1$.
 
@@
 that the merging black holes are of equal mass (this is marginally
-consistent with the three observations\cite{O1-BBH}, and the
-$\chieff$ distribution is not particularly sensitive to the mass ratio
-between the merging objects---see Section \ref{sec:mass-ratio}).  We
-assume that the population spin distribution factorises into a
-distribution for the spin magnitude $a$ and a distribution for the
-spin angles.  Finally, we assume that the distribution of spins is
-common to each component in a merging binary.  Choosing one of three
-magnitude distributions
-\begin{equation}
-  \label{eq:magnitude-dists}
-  p(a) = \begin{cases}
-    2\left(1-a \right) & \textnormal{``low''} \\
-    1 & \textnormal{``flat''} \\
-    2 a & \textnormal{``high''}
-  \end{cases},
-\end{equation}
-and pairing with either an isotropic angular distribution or a
+consistent with the three observations\cite{O1-BBH}, and the $\chieff$
+distribution is not particularly sensitive to the mass ratio between
+the merging objects---see Section \ref{sec:mass-ratio}).  We assume
+that the population spin distribution factorises into a distribution
+for the spin magnitude $a$ and a distribution for the spin angles.
+Finally, we assume that the distribution of spins is common to each
+component in a merging binary.  Choosing one of three magnitude
+distributions (see Methods Section
+\ref{methsec:chieff-spin-magnitude}), ``low'' (mean $a = 0.33$,
+standard deviation $0.24$), ``flat'' (mean $a = 0.5$, standard
+deviation $0.29$), ``high'' (mean $a = 0.67$, standard deviation
+$0.24$) and pairing with either an isotropic angular distribution or a
 distribution that generates perfect alignment with the positive $z$
@@
   \caption{\label{fig:chieff-distribution-models} The models for the
-    distribution of $\chieff$ considered in this paper.  In all models
-    we assume that the binary mass ratio $q \equiv m_1/m_2 = 1$ and
-    that the distribution of spin vectors is the same for each
-    component.  The ``flat'' (blue lines), ``high,'' (green lines),
-    and ``low'' (red lines) magnitude distributions are defined in
-    Eq.\ \eqref{eq:magnitude-dists}.  Solid lines give the $\chieff$
-    distribution under the assumption that the orientations of the
-    spins are isotropic; dashed lines give the distribution under the
-    assumption that both objects' spins are aligned with the orbital
-    angular momentum.  The isotropic distributions are readily
-    distinguished from the aligned distributions by the production of negative
-    $\chieff$ values, while the distinction between the three models
-    for the spin magnitude distribution is less sharp.}
+    population distribution of $\chieff$ considered in this paper.  In
+    all models we assume that the binary mass ratio
+    $q \equiv m_1/m_2 = 1$ and that the distribution of spin vectors
+    is the same for each component.  The ``flat'' (blue lines),
+    ``high,'' (green lines), and ``low'' (red lines) magnitude
+    distributions are defined in Eq.\ \eqref{eq:magnitude-dists}.
+    Solid lines give the $\chieff$ distribution under the assumption
+    that the orientations of the spins are isotropic; dashed lines
+    give the distribution under the assumption that both objects'
+    spins are aligned with the orbital angular momentum.  The
+    isotropic distributions are readily distinguished from the aligned
+    distributions by the production of negative $\chieff$ values,
+    while the distinction between the three models for the spin
+    magnitude distribution is less sharp.}
 \end{figure}
 
-The distributions in Eq.\ \eqref{eq:magnitude-dists} are not meant to
-represent any particular physical model, but rather to capture our
-uncertainty about the spin magnitude distribution, discussed in detail
-below; neither observations nor population synthesis codes can at this
-point authoritatively suggest \emph{any} particular spin
+These magnitude distributions not meant to represent any particular
+physical model, but rather to capture our uncertainty about the spin
+magnitude distribution, discussed in detail below; neither
+observations nor population synthesis codes can at this point
+authoritatively suggest \emph{any} particular spin
 distribution\cite{2015PhR...548....1M}.  Our models, however, allow us
@@
 \end{figure}
-
-\subsection{Future Spin Measurements}
-\label{subsec:future}
 
@@
 \begin{methods}
+
+\section{Effective Spin and Spin Magnitude Distributions}
+\label{methsec:chieff-spin-magnitude}
+
+The effective spin is defined by\cite{2016PhRvL.116x1102A} 
+\begin{equation}
+  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
+  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
+\end{equation}
+where $m_{1,2}$ are the gravitational masses of the more-massive (1)
+and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
+$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
+holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
+assumed to point in the $\hat{z}$ direction, and
+$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
+projections of the individual \ac{BH} spins.
+
+We form the population distributions of $\chieff$ shown in Figure
+\ref{fig:chieff-distribution-models} by assuming that each black hole
+in a binary has a dimensionless spin magnitude drawn from one of three
+distributions,
+\begin{equation}
+  \label{eq:magnitude-dists}
+  p(a) = \begin{cases}
+    2\left(1-a \right) & \textnormal{``low''} \\
+    1 & \textnormal{``flat''} \\
+    2 a & \textnormal{``high''}
+  \end{cases},
+\end{equation}
+referred to as ``low,'' ``flat,'' and ``high'' in the text above.
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #18, remove summary paragraph.

@@
 
-In summary, the angular distribution of spins in \acp{BBH} formed
-through isolated binary evolution is uncertain; nevertheless, it is
-more probable that spins are preferentially aligned after evolution
-through this channel than not, while an isotropic distribution of
-spins is the natural outcome of dynamical formation processes.
-Therefore, if the current observational trend for low $\chieff$
-continues with future gravitational-wave observations, it will be
-possible to either confirm that \ac{BH} spins are isotropic in
-direction and sometimes large in magnitude, yielding a strong
-indication of a dynamical formation origin; or that \ac{BH} spins are
-overwhelmingly small in magnitude, yielding a notable contradiction
-with the claimed \ac{BH} X-ray binary spin measurements, particularly
-those for high-mass X-ray binaries.
-
 \begin{addendum}

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #25; comments on Salvo and Simon's papers distributed through the text.
Where we discuss the "power" of chieff > 0, reference to Salvo's paper, with 10--20 sources required to get 3\sigma opposed to our 10 for >> 5 sigma.

Where we discuss difficulty of determining spin magnitudes, refer to recent studies in support.

Previous para devoted to Salvo, Simon, and spin magnitudes now only cautions that if multiple populations present to significant degree, may required 10s to 100s of detections, citing Salvo and Simon, depending on magnitudes.

@@
 \emph{magnitude} distribution among the three options considered here
-until we have a larger number of observations.  In Figure
+until we have a larger number of
+observations\cite{2017arXiv170306869F,2017arXiv170306223G}.  In Figure
 \ref{fig:O2-predictions}, the odds ratio between different magnitude
@@
 permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
-the GW151226 observations\cite{2016PhRvL.116x1103A}.
+the GW151226 observations\cite{2016PhRvL.116x1103A}.  An ``aligned''
+model with spin magnitudes from our ``flat'' distribution but
+permitting spin vectors oriented anti-parallel to the orbital angular
+momentum (leading to the possibility of positive \emph{or} negative
+$\chieff$) can only be distinguised from an isotropic true population
+at $\sim 3 \sigma$ after 10--20
+observations\cite{2017CQGra..34cLT01V}; our flat aligned model can be
+distinguished from such a population at better than $5\sigma$ (odds
+$< 10^{-8}$) after 10 observations, emphasising the information
+content of the bound $\chieff > 0$ for our aligned models.
 
@@
 conserved through this evolution.
-
-\fixme{\cite{2017CQGra..34cLT01V} also studied the possibility of
-  distinguishing aligned and isotropic angular distributions of
-  \ac{BBH} spins, but concluded that several hundred sources would be
-  required to adequately separate models which included both parallel
-  and anti-parallel spins for an ``aligned'' population.  In contrast,
-  and in agreement with the present study,\cite{Stevenson:2017spin}
-  found that only $\sim 5$ observations of rapidly spinning black
-  holes would be necessary to distinguish isotropic and aligned spin
-  distributions, though tens of detections would be required for more
-  nuanced admixture models.  Meanwhile, recent studies
-  by\cite{2017arXiv170306869F} and\cite{2017arXiv170306223G} focused
-  on inference on spin magnitudes.}
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Salvo added to acknowledgements for comments.
Close #17, tightening language in 'summary' paragraphs.

@@
 
-In order to perform our analysis we need to select models for the
-distribution of the spin magnitudes of stellar-mass black holes.
-Observational data for such model selection is
+Observational data on spin magnitudes in black hole systems is
 sparse\cite{2015PhR...548....1M}.  Most of the systems studied are
@@
 addition, there are substantial systematic errors that can complicate
-either type of analysis\cite{2015PhR...548....1M} and selection
-effects could yield a biased distribution. Nonetheless, if we take the
-reported spin magnitudes as representative then we find that there is
-a preference for high spins; for example, 14 of the 19 systems with
-reported spins have dimensionless spin parameters in excess of 0.5.
-It is usually argued that the masses and spin parameters of
-stellar-mass black holes are unlikely to be altered significantly by
-accretion (low-mass donors may not have enough mass and high-mass
-donors have a very short phase in which they transfer
-mass\cite{1999MNRAS.305..654K}), but this may not be true for all
+these analyses\cite{2015PhR...548....1M} and selection effects could
+yield a biased distribution. Nonetheless, if we take the reported spin
+magnitudes as representative then we find that there is a preference
+for high spins; for example, 14 of the 19 systems with reported spins
+have dimensionless spin parameters in excess of 0.5.  It is usually
+argued that the masses and spin parameters of stellar-mass black holes
+are unlikely to be altered significantly by
+accretion\cite{1999MNRAS.305..654K}, but this may not be true for all
 systems\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the
-current spin parameters probably are close to their values upon core
-collapse, at least in high-mass X-ray binaries.  However, the specific
-processes involved in the production of black hole binaries from
-isolated binaries could alter the spin magnitude distribution of those
-holes relative to the X-ray binary systems; for example, close tidal
-interactions could spin up the core, or stripping of the envelope
-could reduce the available angular
+current spin parameters are probably are close to their values upon
+core collapse, at least in high-mass X-ray binaries.  However, the
+specific processes involved in the production of black hole binaries
+from isolated binaries could alter the spin magnitude distribution of
+those holes relative to the X-ray binary systems; for example, close
+tidal interactions could spin up the core, or stripping of the
+envelope could reduce the available angular
 momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
@@
 
-The spin directions in isolated binaries, whether evolving via the
-classical channel through a common-envelope
-phase\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
-or through chemically homogeneous
-evolution\cite{MandeldeMink:2016,Marchant:2016} are usually expected
-to be preferentially aligned.  Despite observed spin-orbit
-misalignments in both massive stellar
+The spin directions in isolated
+binaries\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
+are usually expected to be preferentially aligned.  Despite observed
+spin-orbit misalignments in both massive stellar
 binaries\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray
@@
 can also tilt the spin angle, as in the double pulsar
-J0737-3039\cite{2011ApJ...742...81F}; and a variety of uncertain
-processes, such as wind-driven mass loss and post-collapse fallback,
-can couple the spin magnitude and direction distributions, contrary to
-our simplified assumptions.  A small misalignment at wide separation
-can also evolve to a more significant misalignment in component spins
-as the binary spirals in through \ac{GW}
-emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
-conserved through this evolution.
+J0737-3039\cite{2011ApJ...742...81F}; and evolutionary processes, such
+as wind-driven mass loss and post-collapse fallback, can couple the
+spin magnitude and direction distributions, contrary to our simplified
+assumptions.  A small misalignment at wide separation can also evolve
+to a more significant misalignment in component spins as the binary
+spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
+$\chieff$ is approximately conserved through this evolution.
 
 \begin{addendum}
-\item We thank Richard O'Shaughnessy, Christopher Berry, and Davide
-  Gerosa for discussions and comments on this work.  WF, SS, IM and AV
-  were supported in part by the STFC.  MCM acknowledges support of the
-  University of Birmingham Institute for Advanced Study Distinguished
-  Visiting Fellows program.  SS and IM acknowledge support from the
-  National Science Foundation under Grant No. NSF PHY11-25915.
+\item We thank Richard O'Shaughnessy, Christopher Berry, Davide
+  Gerosa, and Salvatore Vitale for discussions and comments on this
+  work.  WF, SS, IM and AV were supported in part by the STFC.  MCM
+  acknowledges support of the University of Birmingham Institute for
+  Advanced Study Distinguished Visiting Fellows program.  SS and IM
+  acknowledge support from the National Science Foundation under Grant
+  No. NSF PHY11-25915.
 \item[Supplementary Information] Supplementary Information is linked

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #22, code availability statement.

@@
 
+\section{Code Availability}
+
+This analysis used the Julia language\cite{Julia}, Python libraries
+NumPy and SciPy\cite{NumPy,SciPy}, the plotting library
+Matplotlib\cite{Matplotlib}, and performed computations in IPython
+notebooks\cite{IPython}.  A repository containing the code and
+notebooks used for this analysis, together with the \LaTeX{} source
+for this document, can be found under an open-source ``MIT'' license
+at \url{https://github.com/farr/AlignedVersusIsoSpin}.
+
 \section{Effective Spin and Spin Magnitude Distributions}

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #23, figure captions formatting.

@@
   \plotone{../plots/chi-eff-mock-posteriors}
-  \caption{\label{fig:O1-posteriors} Approximate posteriors on
-    $\chieff$ from the Advanced LIGO O1 run observations\cite{O1-BBH}.
-    We approximate the posteriors reported using Gaussians with the
-    same median and 90\% credible interval.  It is notable that none
-    of the $\chieff$ posteriors support high \ac{BH} spin magnitudes
-    with aligned spins, suggested by observations of stellar-mass
-    black holes in X-ray binaries\cite{2015PhR...548....1M}.}
+  \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
+      $\chieff$ from the Advanced LIGO O1 run
+      observations\cite{O1-BBH}.}  We approximate the posteriors
+    reported using Gaussians with the same median and 90\% credible
+    interval.  It is notable that none of the $\chieff$ posteriors
+    support high \ac{BH} spin magnitudes with aligned spins, suggested
+    by observations of stellar-mass black holes in X-ray
+    binaries\cite{2015PhR...548....1M}.}
 \end{figure}
@@
   \plotone{../plots/chi-eff-distributions}
-  \caption{\label{fig:chieff-distribution-models} The models for the
-    population distribution of $\chieff$ considered in this paper.  In
+  \caption{\label{fig:chieff-distribution-models} \textbf{The models for the
+    population distribution of $\chieff$ considered in this paper.}  In
     all models we assume that the binary mass ratio
@@
   \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models}
-  \caption{Odds ratios among our models using the approximations to
-    the posteriors on $\chieff$ from the O1 observations shown in
-    Figure \ref{fig:O1-posteriors}.  The flat (``F''), high (``H''),
-    and low (``L'') spin magnitude distributions (see Eq.\
-    \eqref{eq:magnitude-dists}) are paired with isotropic (``I'') and
-    aligned (``A'') angular distributions, as well as a mixture model
-    of the two (``$\mathrm{M}$'').  The most-favoured models have the
-    ``low'' distribution of spin magnitudes.  The odds ratio between
-    these models is $\OOneOddsIsoAligned$, or
-    $\OOneSigmaIsoAligned\sigma$.  For all magnitude distributions the
-    pure-isotropic models are preferred over the mixture models;
-    correspondingly, the posterior on the mixture fraction peaks at
-    100\% isotropic.}
+  \caption{\textbf{Odds ratios among our models using the
+      approximations to the posteriors on $\chieff$ from the O1
+      observations shown in Figure \ref{fig:O1-posteriors}.}  The flat
+    (``F''), high (``H''), and low (``L'') spin magnitude
+    distributions (see Eq.\ \eqref{eq:magnitude-dists}) are paired
+    with isotropic (``I'') and aligned (``A'') angular distributions,
+    as well as a mixture model of the two (``$\mathrm{M}$'').  The
+    most-favoured models have the ``low'' distribution of spin
+    magnitudes.  The odds ratio between these models is
+    $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.  For all
+    magnitude distributions the pure-isotropic models are preferred
+    over the mixture models; correspondingly, the posterior on the
+    mixture fraction peaks at 100\% isotropic.}
   \label{fig:O1-odds}
@@
   \plotone{../plots/six-way-O2-model-selection}
-  \caption{\label{fig:O2-predictions} Distribution of odds ratios
-    predicted with 10 additional observations above the three
-    discussed above.  Each panel corresponds to additional
+  \caption{\label{fig:O2-predictions} \textbf{Distribution of odds
+      ratios predicted with 10 additional observations above the three
+      discussed above.}  Each panel corresponds to additional
     observations drawn from one of the $\chieff$ distribution models.
@@
 \includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
-\caption{Fraction of the BBH population coming from an isotropic
-  distribution under a mixture model. The dotted line shows the flat
+\caption{\textbf{Fraction of the BBH population coming from an isotropic
+  distribution under a mixture model.} The dotted line shows the flat
   prior on the fraction of BBHs coming from an isotropic distribution,
@@
 \includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
-\caption{Effect of small spins on evidence ratio of aligned against
-  isotropic models. The blue line shows the evidence ratio (plotted as
-  the equivalent sigma) between a model where all systems are aligned,
-  versus one where all systems are from an isotropic distribution as a
-  function of the power law $\alpha$ corresponding to
-  Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
+\caption{\textbf{Effect of small spins on evidence ratio of aligned
+    against isotropic models.} The blue line shows the evidence ratio
+  (plotted as the equivalent sigma) between a model where all systems
+  are aligned, versus one where all systems are from an isotropic
+  distribution as a function of the power law $\alpha$ corresponding
+  to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
   magnitude $\bar{a}$ which this $\alpha$ corresponds to. We see that
@@
   \plotone{../plots/chi-eff-distributions-q05}
-  \caption{Distributions of $\chieff$ assuming all merging black holes
-    have equal masses ($q=1$) or a 2:1 mass ratio ($q = 0.5$).  The
-    details of the distribution are sensitive to the mass ratio, but
-    in our analysis we are primarily sensitive to the changing
-    \emph{sign} of $\chieff$ under the isotropic models.  This latter
-    property is unchanged under changing mass ratio.}
+  \caption{\textbf{Distributions of $\chieff$ assuming all merging
+      black holes have equal masses ($q=1$) or a 2:1 mass ratio
+      ($q = 0.5$).}  The details of the distribution are sensitive to
+    the mass ratio, but in our analysis we are primarily sensitive to
+    the changing \emph{sign} of $\chieff$ under the isotropic models.
+    This latter property is unchanged under changing mass ratio.}
   \label{fig:mass-ratio-sensitivity}
@@
   \plotone{../plots/chi-eff-90cl}
-  \caption{Widths of the 90\% credible intervals for $\chieff$ for 500
-    binaries in a simulated detected population. $\chieff$ is better
-    constrained for systems with low $\chieff$ and low mass ratio.}
+  \caption{\textbf{Widths of the 90\% credible intervals for $\chieff$
+      for 500 binaries in a simulated detected population.} $\chieff$
+    is better constrained for systems with low $\chieff$ and low mass
+    ratio.}
   \label{fig:chi-eff-constraints}

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #24, Figure --> Extended Data Figure.
Close #26 with comment on selection effects.
Close #29
Removed statement about XRBs from abstract.  It is too hard to explain the
context within the abstract, and we have an extensive discussion later in the
paper.

@@
   binary black hole detections GW150914, LVT151012, and GW151226.  If
-  binary black hole spin magnitudes extend to high values, as is
-  suggested by observations of black hole X-ray
-  binaries\cite{2015PhR...548....1M}, we show that the data already
-  exhibit a $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
-  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
-    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
-    $\Phi$ is the unit normal CDF.})  preference for an isotropic
-  angular distribution.  By considering the effect of 10 additional
+  binary black hole spin magnitudes extend to high values we show that
+  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
+  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
+    $r \ll 1$ is equivalent to $x \sigma$ with
+    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
+    normal CDF.})  preference for an isotropic angular distribution.
+  By considering the effect of 10 additional
   detections\cite{2016ApJ...833L...1A} of systems, we show that if all
@@
 against the observed data to obtain a posterior distribution on the
-parameters that describe the compact binary source.
-
-The spin parameter with the largest effect on waveforms, and a
-correspondingly tight constraint from the
-data\cite{2016PhRvL.116x1102A}, is a mass-weighted combination of the
-components of the dimensionless spin vectors of the two black holes
-that are aligned with the orbital axis, the ``effective spin,''
-$-1 \leq \chieff \leq 1$.
+parameters that describe the compact binary source.  The spin
+parameter with the largest effect on waveforms, and a correspondingly
+tight constraint from the data\cite{2016PhRvL.116x1102A}, is a
+mass-weighted combination of the components of the dimensionless spin
+vectors of the two black holes that are aligned with the orbital axis,
+the ``effective spin,'' $-1 \leq \chieff \leq 1$.
 
@@
 These models are shown in Figure \ref{fig:chieff-distribution-models}.
+While the Advanced LIGO searches use spin-aligned templates they are
+efficient in detecting misaligned binary black hole
+systems\cite{2016PhRvD..93l2003A}; we assume here that the $\chieff$
+distribution of observed sources follows the true population.
 
@@
 \begin{methods}
+
+% Reset the Figure Counter
+\setcounter{figure}{0}
+\renewcommand{\figurename}{Extended Data Figure}
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #30, discussion of spin magnitude bounds and definition.

@@
 vectors of the two black holes that are aligned with the orbital axis,
-the ``effective spin,'' $-1 \leq \chieff \leq 1$.
+the ``effective spin,'' $-1 < \chieff < 1$ (see Methods Section
+\ref{methsec:chieff-spin-magnitude}).
 
@@
 $0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
-projections of the individual \ac{BH} spins.
+projections of the individual \ac{BH} spins.  Because the
+dimensionless spin parameter,
+\begin{equation}
+  \label{eq:a-def}
+  a_{1,2} = \frac{c}{G m_{1,2}^2} \left|\vec{S}_{1,2} \right|,
+\end{equation}
+of each black hole is bounded by $0 \leq a_{1,2} < 1$, the projections
+along the orbital axis are bounded by $-1 < \chi_{1,2} < 1$, and
+$-1 < \chieff < 1$.
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #31, truncated Gaussian.

@@
 the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval.  None of the $\chieff$ posteriors are consistent
-with two black holes with large aligned spins,
-$\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred
-for the majority of black holes in X-ray binaries with claimed spin
-measurements (see below).  The analysis here is
+credible interval, truncated to $-1 < \chieff < 1$.  None of the
+$\chieff$ posteriors are consistent with two black holes with large
+aligned spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large
+spins inferred for the majority of black holes in X-ray binaries with
+claimed spin measurements (see below).  The analysis here is
 relatively insensitive to the precise details of the posterior

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #34; this was already discussed in main text, but needs to use the appropriate label ('Section' -> 'Methods Section').

@@
 distribution is not particularly sensitive to the mass ratio between
-the merging objects---see Section \ref{sec:mass-ratio}).  We assume
-that the population spin distribution factorises into a distribution
-for the spin magnitude $a$ and a distribution for the spin angles.
-Finally, we assume that the distribution of spins is common to each
-component in a merging binary.  Choosing one of three magnitude
-distributions (see Methods Section
+the merging objects---see Methods Section \ref{sec:mass-ratio}).  We
+assume that the population spin distribution factorises into a
+distribution for the spin magnitude $a$ and a distribution for the
+spin angles.  Finally, we assume that the distribution of spins is
+common to each component in a merging binary.  Choosing one of three
+magnitude distributions (see Methods Section
 \ref{methsec:chieff-spin-magnitude}), ``low'' (mean $a = 0.33$,

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #33, shared P(a).

@@
 spin angles.  Finally, we assume that the distribution of spins is
-common to each component in a merging binary.  Choosing one of three
-magnitude distributions (see Methods Section
-\ref{methsec:chieff-spin-magnitude}), ``low'' (mean $a = 0.33$,
-standard deviation $0.24$), ``flat'' (mean $a = 0.5$, standard
-deviation $0.29$), ``high'' (mean $a = 0.67$, standard deviation
-$0.24$) and pairing with either an isotropic angular distribution or a
-distribution that generates perfect alignment with the positive $z$
-axis yields six different models for the $\chieff$ distribution.
-These models are shown in Figure \ref{fig:chieff-distribution-models}.
-While the Advanced LIGO searches use spin-aligned templates they are
-efficient in detecting misaligned binary black hole
-systems\cite{2016PhRvD..93l2003A}; we assume here that the $\chieff$
-distribution of observed sources follows the true population.
+common to each component in a merging binary (in reality the
+distributions of spin for each component in the binary may differ
+systematically due to different formation and accretion histories, but
+our analysis is not sensitive to the magnitude distribution at this
+level; see below).  Choosing one of three magnitude distributions (see
+Methods Section \ref{methsec:chieff-spin-magnitude}), ``low'' (mean
+$a = 0.33$, standard deviation $0.24$), ``flat'' (mean $a = 0.5$,
+standard deviation $0.29$), ``high'' (mean $a = 0.67$, standard
+deviation $0.24$) and pairing with either an isotropic angular
+distribution or a distribution that generates perfect alignment with
+the positive $z$ axis yields six different models for the $\chieff$
+distribution.  These models are shown in Figure
+\ref{fig:chieff-distribution-models}.  While the Advanced LIGO
+searches use spin-aligned templates they are efficient in detecting
+misaligned binary black hole systems\cite{2016PhRvD..93l2003A}; we
+assume here that the $\chieff$ distribution of observed sources
+follows the true population.
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Added p(a) plot, addressing #32.

@@
 referred to as ``low,'' ``flat,'' and ``high'' in the text above.
+These distributions are shown in Extended Data Figure \ref{fig:pa}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{../plots/pa}
+  \caption{\textbf{Distributions of spin magnitudes.}  See Eq.\
+    \eqref{eq:magnitude-dists} for the definition of the low (blue
+    line), flat (green line), and high (red line) magnitude
+    distributions used here.  The distributions have mean spin $0.33$,
+    $0.5$, and $0.67$ and standard deviations $0.24$, $0.29$, and
+    $0.24$.}
+  \label{fig:pa}
+\end{figure}
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #32 by adding forward reference to generalised 'low' distribution.

@@
 no longer possible to resolve the negative effective spin with a small
-number of observations.  As noted below, however, spins observed in
-X-ray binaries are typically large.  Additionally, models which do not
-permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
-the GW151226 observations\cite{2016PhRvL.116x1103A}.  An ``aligned''
-model with spin magnitudes from our ``flat'' distribution but
-permitting spin vectors oriented anti-parallel to the orbital angular
-momentum (leading to the possibility of positive \emph{or} negative
-$\chieff$) can only be distinguised from an isotropic true population
-at $\sim 3 \sigma$ after 10--20
-observations\cite{2017CQGra..34cLT01V}; our flat aligned model can be
-distinguished from such a population at better than $5\sigma$ (odds
-$< 10^{-8}$) after 10 observations, emphasising the information
-content of the bound $\chieff > 0$ for our aligned models.
+number of observations (see Methods Section \ref{sec:smallspins}).  As
+noted below, however, spins observed in X-ray binaries are typically
+large.  Additionally, models which do not permit \emph{some} spins
+with $\chieff \gtrsim 0.1$ are ruled out by the GW151226
+observations\cite{2016PhRvL.116x1103A}.  An ``aligned'' model with
+spin magnitudes from our ``flat'' distribution but permitting spin
+vectors oriented anti-parallel to the orbital angular momentum
+(leading to the possibility of positive \emph{or} negative $\chieff$)
+can only be distinguised from an isotropic true population at
+$\sim 3 \sigma$ after 10--20 observations\cite{2017CQGra..34cLT01V};
+our flat aligned model can be distinguished from such a population at
+better than $5\sigma$ (odds $< 10^{-8}$) after 10 observations,
+emphasising the information content of the bound $\chieff > 0$ for our
+aligned models.
 

### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #35, added references suggested around kicks and commented on correlation between rates and kicks.

@@
 can also tilt the spin angle, as in the double pulsar
-J0737-3039\cite{2011ApJ...742...81F}; and evolutionary processes, such
-as wind-driven mass loss and post-collapse fallback, can couple the
-spin magnitude and direction distributions, contrary to our simplified
-assumptions.  A small misalignment at wide separation can also evolve
-to a more significant misalignment in component spins as the binary
-spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
-$\chieff$ is approximately conserved through this evolution.
+J0737-3039\cite{2011ApJ...742...81F}.  Large kicks are probably
+associated with reduced merger rates as more potential merging systems
+are distrupted by a supernova at black hole
+formation\cite{1995MNRAS.277L..35B,2002ApJ...572..962S,2003CQGra..20S..73F,2004ApJ...616..414W,2016ApJ...832L...2R}. Evolutionary
+processes, such as wind-driven mass loss and post-collapse fallback,
+can couple the spin magnitude and direction distributions, contrary to
+our simplified assumptions.  A small misalignment at wide separation
+can also evolve to a more significant misalignment in component spins
+as the binary spirals in through \ac{GW}
+emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
+conserved through this evolution.
 

### 2017-05-25 paper/AlignedIsotropicGW.tex
Typo fix.

@@
 J0737-3039\cite{2011ApJ...742...81F}.  Large kicks are probably
-associated with reduced merger rates as more potential merging systems
-are distrupted by a supernova at black hole
+associated with reduced merger rates as more potentially merging binaries 
+are disrupted by a supernova at black hole
 formation\cite{1995MNRAS.277L..35B,2002ApJ...572..962S,2003CQGra..20S..73F,2004ApJ...616..414W,2016ApJ...832L...2R}. Evolutionary
@@
 the ``low'', ``flat'' and ``high'' spin magnitude models. Thus we
-cannot rule out a mixture with the current data.
+cannot rule out a mixture with the current data.  If several different
+components contribute significantly to the true spin distribution it
+may take tens to hundreds of detections to accurately determined the
+mixing fraction, depending on the distribution of spin
+magnitudes\cite{2017CQGra..34cLT01V,Stevenson:2017spin}.
 

### 2017-05-29 paper/AlignedIsotropicGW.tex
readded simons KITP affiliation

@@
 
-\author{Will M. Farr$^1$, Simon Stevenson$^1$, M. Coleman Miller$^2$,
+\author{Will M. Farr$^1$, Simon Stevenson$^{1,3}$, M. Coleman Miller$^2$,
   Ilya Mandel$^{1,3}$, Ben Farr$^4$ \& Alberto Vecchio$^1$}
@@
 \includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
-\caption{\textbf{Effect of small spins on evidence ratio of aligned
-    against isotropic models.} The blue line shows the evidence ratio
+\caption{\textbf{Effect of small spins on evidence ratio of isotropic against
+  aligned models.} The blue line shows the evidence ratio
   (plotted as the equivalent sigma) between a model where all systems
-  are aligned, versus one where all systems are from an isotropic
-  distribution as a function of the power law $\alpha$ corresponding
+  are from an isotropic distribution, versus one where all systems are
+  aligned, as a function of the power law $\alpha$ corresponding
   to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
@@
 In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of
-aligned to isotropic distributions (plotted as the equivalent sigma)
+isotropic to aligned distributions (plotted as the equivalent sigma)
 with spin magnitudes given by this model with $\alpha$ in the range
-$0$ -- $6$. The top axis shows the mean spin magnitude that value of
+$0$--$6$. The top axis shows the mean spin magnitude that value of
 $\alpha$ corresponds to (e.g., for the ``flat'' distribution

### 2017-05-30 paper/AlignedIsotropicGW.tex
Correct language re spin constraints

@@
 information for some binaries, but we ignore those effects here).
-$\chieff$ is better constrained for binaries with low (negative)
-effective spins and low mass ratios.
+$\chieff$ is better constrained for binaries with high
+effective spins and high ($\sim$equal) mass ratios.
 
-By neglecting these effects, We do not expect to qualitatively affect
+By neglecting these effects we do not expect to qualitatively affect
 out conclusions, though they could affect predictions for the total
@@
 low mass ratios, individual $\chieff$ constraints will be systematically
-better, reducing the number of binaries needed to infer the properties
+worse, requiring more binaries to infer the properties
 of the population.
@@
       for 500 binaries in a simulated detected population.} $\chieff$
-    is better constrained for systems with low $\chieff$ and low mass
+    is better constrained for systems with high $\chieff$ and high mass
     ratio.}

### 2017-06-01 paper/AlignedIsotropicGW.tex
Close (for real, this time) #35
Close #37

@@
   observations come from a single population such an augmented data
-  set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
-  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
-  isotropic and aligned models for the assumed spin magnitude
-  distributions, and in most cases better than $5\sigma$
-  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
-  either an isotropic spin distribution or low spin magnitudes for the
-  observed systems will be confirmed (or overturned) confidently in
-  the near future by subsequent observations.
+  set would typically enable at least a
+  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
+  ratio) distinction between the isotropic and aligned models for the
+  assumed spin magnitude distributions, and in most cases better than
+  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
+  preference for either an isotropic spin distribution or low spin
+  magnitudes for the observed systems will be confirmed (or
+  overturned) confidently in the near future by subsequent
+  observations.
 \end{abstract}
@@
 can also tilt the spin angle, as in the double pulsar
-J0737-3039\cite{2011ApJ...742...81F}.  Large kicks are probably
-associated with reduced merger rates as more potentially merging binaries 
-are disrupted by a supernova at black hole
-formation\cite{1995MNRAS.277L..35B,2002ApJ...572..962S,2003CQGra..20S..73F,2004ApJ...616..414W,2016ApJ...832L...2R}. Evolutionary
-processes, such as wind-driven mass loss and post-collapse fallback,
-can couple the spin magnitude and direction distributions, contrary to
-our simplified assumptions.  A small misalignment at wide separation
-can also evolve to a more significant misalignment in component spins
-as the binary spirals in through \ac{GW}
-emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
-conserved through this evolution.
+J0737-3039\cite{2011ApJ...742...81F}.  Evolutionary processes, such as
+wind-driven mass loss and post-collapse fallback, can couple the spin
+magnitude and direction distributions, contrary to our simplified
+assumptions.  A small misalignment at wide separation can also evolve
+to a more significant misalignment in component spins as the binary
+spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
+$\chieff$ is approximately conserved through this evolution.
 

### 2017-06-01 paper/AlignedIsotropicGW.tex
Close #38.

@@
 distributions of spin for each component in the binary may differ
-systematically due to different formation and accretion histories, but
-our analysis is not sensitive to the magnitude distribution at this
-level; see below).  Choosing one of three magnitude distributions (see
-Methods Section \ref{methsec:chieff-spin-magnitude}), ``low'' (mean
-$a = 0.33$, standard deviation $0.24$), ``flat'' (mean $a = 0.5$,
-standard deviation $0.29$), ``high'' (mean $a = 0.67$, standard
-deviation $0.24$) and pairing with either an isotropic angular
-distribution or a distribution that generates perfect alignment with
-the positive $z$ axis yields six different models for the $\chieff$
-distribution.  These models are shown in Figure
-\ref{fig:chieff-distribution-models}.  While the Advanced LIGO
-searches use spin-aligned templates they are efficient in detecting
-misaligned binary black hole systems\cite{2016PhRvD..93l2003A}; we
-assume here that the $\chieff$ distribution of observed sources
-follows the true population.
+systematically due to different formation histories.  Choosing one of
+three magnitude distributions (see Methods Section
+\ref{methsec:chieff-spin-magnitude}), ``low'' (mean $a = 0.33$,
+standard deviation $0.24$), ``flat'' (mean $a = 0.5$, standard
+deviation $0.29$), ``high'' (mean $a = 0.67$, standard deviation
+$0.24$) and pairing with either an isotropic angular distribution or a
+distribution that generates perfect alignment with the positive $z$
+axis yields six different models for the $\chieff$ distribution.
+These models are shown in Figure \ref{fig:chieff-distribution-models}.
+While the Advanced LIGO searches use spin-aligned templates they are
+efficient in detecting misaligned binary black hole
+systems\cite{2016PhRvD..93l2003A}; we assume here that the $\chieff$
+distribution of observed sources follows the true population.
 

### 2017-06-01 paper/AlignedIsotropicGW.tex
Consistency in 'Methods Section XX'.  Close #39

@@
 these six different, zero-parameter population distributions (see
-Section \ref{sec:hierarchical}).  We also fit three mixture models for
-the population, where the spin magnitude distribution is fixed but the
-angular distribution is a weighted sum of the isotropic and aligned
-distribution.  The evidence, or marginal likelihood, for each of the
-models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
-distributions, the mixture models' posterior on the mixing fraction
-peaks at 100\% isotropic, which explains why the zero-parameter,
-pure-isotropic models are preferred over the single-parameter mixture
-models for every magnitude distribution with this data set.  Not
-surprisingly, given the small $\chieff$ values in the three detected
-systems, the most-favoured model among those with an isotropic angular
-distribution has the ``low'' magnitude distribution; the most favoured
-model among those with an aligned distribution also has the ``low''
-magnitude distribution.  The odds ratio between the ``low'' aligned
-and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
-$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
-among our suite of models.  While the data favour spin amplitude
-distributions with small spin magnitudes, note that a model with all
-\ac{BBH} systems having zero spin is ruled out by the GW151226
-measurements, which bound at least one black hole to have spin
-magnitude $\geq 0.2$ at 99\% credibility\cite{2016PhRvL.116x1103A}.
+Methods Section \ref{sec:hierarchical}).  We also fit three mixture
+models for the population, where the spin magnitude distribution is
+fixed but the angular distribution is a weighted sum of the isotropic
+and aligned distribution.  The evidence, or marginal likelihood, for
+each of the models is shown in Figure \ref{fig:O1-odds}.  For all
+three magnitude distributions, the mixture models' posterior on the
+mixing fraction peaks at 100\% isotropic, which explains why the
+zero-parameter, pure-isotropic models are preferred over the
+single-parameter mixture models for every magnitude distribution with
+this data set.  Not surprisingly, given the small $\chieff$ values in
+the three detected systems, the most-favoured model among those with
+an isotropic angular distribution has the ``low'' magnitude
+distribution; the most favoured model among those with an aligned
+distribution also has the ``low'' magnitude distribution.  The odds
+ratio between the ``low'' aligned and ``low'' isotropic models is
+$\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$; thus the data
+favour isotropic spins among our suite of models.  While the data
+favour spin amplitude distributions with small spin magnitudes, note
+that a model with all \ac{BBH} systems having zero spin is ruled out
+by the GW151226 measurements, which bound at least one black hole to
+have spin magnitude $\geq 0.2$ at 99\%
+credibility\cite{2016PhRvL.116x1103A}.
 
@@
   if the parameters of the three detected events are representative of
-  the parameters of future detections. See Section
+  the parameters of future detections. See Methods Section
   \ref{sec:chi-eff-precision} for further discussion.}, we find the
@@
 magnitudes, our angular resolving power vanishes, as discussed in more
-detail in Section \ref{sec:smallspins}.
+detail in Methods Section \ref{sec:smallspins}.
 

### 2017-06-01 paper/AlignedIsotropicGW.tex
Close #41; more specific about models that give 1.7sigma.
Close #42, moving paragraph on dynamical formation out from between isolated formation 'graphs.
Close #43, make clear 'isolated binary black holes' not just 'binaries'.

@@
     most-favoured models have the ``low'' distribution of spin
-    magnitudes.  The odds ratio between these models is
-    $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.  For all
-    magnitude distributions the pure-isotropic models are preferred
-    over the mixture models; correspondingly, the posterior on the
-    mixture fraction peaks at 100\% isotropic.}
+    magnitudes.  The odds ratio between the best aligned and best
+    isotropic models is $\OOneOddsIsoAligned$, or
+    $\OOneSigmaIsoAligned\sigma$.  For all magnitude distributions the
+    pure-isotropic models are preferred over the mixture models;
+    correspondingly, the posterior on the mixture fraction peaks at
+    100\% isotropic.}
   \label{fig:O1-odds}
@@
 
-The spin directions of binary black holes formed dynamically through
-interactions in dense stellar
-environments\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
-are expected to be isotropic given the absence of a preferred
-direction\cite{2016ApJ...832L...2R} and the persistence of an
-isotropic distribution through post-Newtonian
-evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
-
-The spin directions in isolated
-binaries\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
+The spin directions in isolated binary black
+holes\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
 are usually expected to be preferentially aligned.  Despite observed
@@
 $\chieff$ is approximately conserved through this evolution.
+
+The spin directions of binary black holes formed dynamically through
+interactions in dense stellar
+environments\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+are expected to be isotropic given the absence of a preferred
+direction\cite{2016ApJ...832L...2R} and the persistence of an
+isotropic distribution through post-Newtonian
+evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
 

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #44.

@@
 are usually expected to be preferentially aligned.  Despite observed
-spin-orbit misalignments in both massive stellar
-binaries\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray
-binaries\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
-mass accretion and tidal interactions will tend to realign the binary.
-On the other hand, a supernova natal kick (if any) can change the
-orbital plane and misalign the
+spin-orbit misalignments in massive stellar
+binaries\cite{Albrecht:2009,2014ApJ...785...83A}, mass transfer and
+tidal interactions will tend to realign the binary.  On the other
+hand, there is some evidence of spin-orbit misalignment in black hole
+X-ray
+binaries\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014}.
+This is consistent with the expectation that a supernova natal kick
+(if any) can change the orbital plane and misalign the
 binary\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova
-can also tilt the spin angle, as in the double pulsar
-J0737-3039\cite{2011ApJ...742...81F}.  Evolutionary processes, such as
-wind-driven mass loss and post-collapse fallback, can couple the spin
-magnitude and direction distributions, contrary to our simplified
-assumptions.  A small misalignment at wide separation can also evolve
-to a more significant misalignment in component spins as the binary
-spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
-$\chieff$ is approximately conserved through this evolution.
+can also tilt the spin angle\cite{2011ApJ...742...81F}.  Evolutionary
+processes, such as wind-driven mass loss and post-collapse fallback,
+can couple the spin magnitude and direction distributions, contrary to
+our simplified assumptions.  A small misalignment at wide separation
+can also evolve to a more significant misalignment in component spins
+as the binary spirals in through \ac{GW}
+emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
+conserved through this evolution.
 

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #45, clarify spins parallel to angular momentum.

@@
   whereas binaries formed from pairs of stars born together are more
-  likely to have spins preferentially aligned with the binary orbit as
-  a consequence of their joint evolution toward a binary black hole
+  likely to have spins preferentially aligned with the binary orbital
+  angular momentum as a consequence of their joint evolution toward a
+  binary black hole
   system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #46.

@@
     $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
-    normal CDF.})  preference for an isotropic angular distribution.
-  By considering the effect of 10 additional
+    normal CDF.})  preference for an isotropic angular distribution
+  over an aligned one.  By considering the effect of 10 additional
   detections\cite{2016ApJ...833L...1A} of systems, we show that if all
@@
   $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
-  ratio) distinction between the isotropic and aligned models for the
-  assumed spin magnitude distributions, and in most cases better than
-  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
-  preference for either an isotropic spin distribution or low spin
-  magnitudes for the observed systems will be confirmed (or
-  overturned) confidently in the near future by subsequent
-  observations.
+  ratio) distinction, and in most cases better than $5\sigma$
+  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
+  either an isotropic spin distribution or low spin magnitudes for the
+  observed systems will be confirmed (or overturned) confidently in
+  the near future by subsequent observations.
 \end{abstract}

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #47, three, not two, years to 10 detections.

@@
 Estimates of the rate of \ac{BBH} coalescences give a reasonable
-chance of 10 additional \ac{BBH} detections in the next two
+chance of 10 additional \ac{BBH} detections in the next three
 years\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #49, too many 'are's.
Close #50, typo in chi_1 chi_2 bounds.

@@
 systems\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the
-current spin parameters are probably are close to their values upon
-core collapse, at least in high-mass X-ray binaries.  However, the
-specific processes involved in the production of black hole binaries
-from isolated binaries could alter the spin magnitude distribution of
-those holes relative to the X-ray binary systems; for example, close
-tidal interactions could spin up the core, or stripping of the
-envelope could reduce the available angular
+current spin parameters are probably close to their values upon core
+collapse, at least in high-mass X-ray binaries.  However, the specific
+processes involved in the production of black hole binaries from
+isolated binaries could alter the spin magnitude distribution of those
+holes relative to the X-ray binary systems; for example, close tidal
+interactions could spin up the core, or stripping of the envelope
+could reduce the available angular
 momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
@@
 assumed to point in the $\hat{z}$ direction, and
-$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
+$-1 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
 projections of the individual \ac{BH} spins.  Because the

### 2017-06-02 paper/AlignedIsotropicGW.tex
Close #52, 0.3 --> 0.2 in 'small spin' discussion.

@@
 $\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
-spins are intrinsically very small, with $a \lesssim 0.3$, then it is
+spins are intrinsically very small, with $a \lesssim 0.2$, then it is
 no longer possible to resolve the negative effective spin with a small

### 2017-06-02 paper/AlignedIsotropicGW.tex
Incorporation of GW170104 final values.

@@
 
-\newcommand{\OOneSigmaIsoAligned}{1.7}
-\newcommand{\OOneOddsIsoAligned}{0.087}
-\newcommand{\OTwoSigmaIsoAlignedMin}{2.9}
-\newcommand{\OTwoOddsIsoAlignedMin}{0.0035}
+\newcommand{\OOneSigmaIsoAligned}{2.4}
+\newcommand{\OOneOddsIsoAligned}{0.015}
+\newcommand{\OTwoSigmaIsoAlignedMin}{2.4}
+\newcommand{\OTwoOddsIsoAlignedMin}{0.016}
 
@@
   The first direct detections of gravitational
-  waves\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
-  merging binary black holes open a unique window into the binary
+  waves\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH,PhysRevLett.118.221101}
+  from merging binary black holes open a unique window into the binary
   black hole formation environment.  One promising signature of the
@@
   effective spin, the best-measured combination of spin
-  parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
-  binary black hole detections GW150914, LVT151012, and GW151226.  If
-  binary black hole spin magnitudes extend to high values we show that
-  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
-  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
-    $r \ll 1$ is equivalent to $x \sigma$ with
-    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
-    normal CDF.})  preference for an isotropic angular distribution
-  over an aligned one.  By considering the effect of 10 additional
-  detections\cite{2016ApJ...833L...1A} of systems, we show that if all
-  observations come from a single population such an augmented data
-  set would typically enable at least a
+  parameters\cite{2016PhRvL.116x1102A,O1-BBH,PhysRevLett.118.221101},
+  in the four likely binary black hole detections GW150914, LVT151012,
+  GW151226, and GW170104.  If binary black hole spin magnitudes extend
+  to high values we show that the data already exhibit a
+  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
+  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
+    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
+    $\Phi$ is the unit normal CDF.})  preference for an isotropic
+  angular distribution over an aligned one.  By considering the effect
+  of 10 additional detections\cite{2016ApJ...833L...1A} of systems, we
+  show that if all observations come from a single population such an
+  augmented data set would typically enable at least a
   $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
@@
 Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
-inferred on $\chieff$ for the three likely \ac{GW} detections
-GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
-run (O1)\cite{O1-BBH}.  Because samples drawn from the posterior on
-$\chieff$ are not publicly released at this time, we have approximated
-the posterior as a Gaussian distribution with the same mean and 90\%
-credible interval, truncated to $-1 < \chieff < 1$.  None of the
-$\chieff$ posteriors are consistent with two black holes with large
-aligned spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large
-spins inferred for the majority of black holes in X-ray binaries with
-claimed spin measurements (see below).  The analysis here is
-relatively insensitive to the precise details of the posterior
-distributions; other conclusions are more sensitive.  In particular,
-our Gaussian approximation does permit $\chieff = 0$ for GW151226
-while the true posterior rules this out at high
+inferred on $\chieff$ for the four likely \ac{GW} detections GW150914,
+GW151226, GW170104, and LVT151012 from Advanced LIGO's first and
+second observing runs (O1 and O2)\cite{O1-BBH,PhysRevLett.118.221101}.
+Because samples drawn from the posterior on $\chieff$ are not publicly
+released at this time, we have approximated the posterior as a
+Gaussian distribution with the same mean and 90\% credible interval,
+truncated to $-1 < \chieff < 1$.  None of the $\chieff$ posteriors are
+consistent with two black holes with large aligned spins,
+$\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred
+for the majority of black holes in X-ray binaries with claimed spin
+measurements (see below).  The analysis here is relatively insensitive
+to the precise details of the posterior distributions; other
+conclusions are more sensitive.  In particular, our Gaussian
+approximation does permit $\chieff = 0$ for GW151226 while the true
+posterior rules this out at high
 confidence\cite{2016PhRvL.116x1103A,O1-BBH}.

### 2017-06-02 paper/AlignedIsotropicGW.tex
O1 --> O1 and O2.

@@
 
-Following the detection of a merging \ac{BBH} system, \ac{PE}
-tools\cite{2015PhRvD..91d2003V} compare model gravitational
-waveforms\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
-against the observed data to obtain a posterior distribution on the
-parameters that describe the compact binary source.  The spin
-parameter with the largest effect on waveforms, and a correspondingly
-tight constraint from the data\cite{2016PhRvL.116x1102A}, is a
-mass-weighted combination of the components of the dimensionless spin
-vectors of the two black holes that are aligned with the orbital axis,
-the ``effective spin,'' $-1 < \chieff < 1$ (see Methods Section
+Following the detection of a merging binary black hole system,
+parameter estimation tools\cite{2015PhRvD..91d2003V} compare model
+gravitational waveforms against the observed data to obtain a
+posterior distribution on the parameters that describe the compact
+binary source.  The spin parameter with the largest effect on
+waveforms, and a correspondingly tight constraint from the
+data\cite{2016PhRvL.116x1102A}, is a mass-weighted combination of the
+components of the dimensionless spin vectors of the two black holes
+that are aligned with the orbital axis, the ``effective spin,''
+$-1 < \chieff < 1$ (see Methods Section
 \ref{methsec:chieff-spin-magnitude}).
@@
 support for these two alternate explanations of small $\chieff$ values
-in the merging \ac{BBH} population, we compared the Bayesian evidence
-for various simple models of the spin population using the \ac{GW}
-data set.
+in the merging binary black hole population, we compared the Bayesian
+evidence for various simple models of the spin population using the
+\ac{GW} data set.
 
-Each of our models for the merging \ac{BBH} spin population assumes
-that the merging black holes are of equal mass (this is marginally
-consistent with the three observations\cite{O1-BBH}, and the $\chieff$
+Each of our models for the merging binary black hole spin population
+assumes that the merging black holes are of equal mass (this is
+marginally consistent with the
+observations\cite{O1-BBH,PhysRevLett.118.221101}, and the $\chieff$
 distribution is not particularly sensitive to the mass ratio between
@@
 These models are shown in Figure \ref{fig:chieff-distribution-models}.
-While the Advanced LIGO searches use spin-aligned templates they are
-efficient in detecting misaligned binary black hole
-systems\cite{2016PhRvD..93l2003A}; we assume here that the $\chieff$
-distribution of observed sources follows the true population.
 
@@
 
-We fit hierarchical models of the three LIGO O1 observations using
-these six different, zero-parameter population distributions (see
-Methods Section \ref{sec:hierarchical}).  We also fit three mixture
-models for the population, where the spin magnitude distribution is
-fixed but the angular distribution is a weighted sum of the isotropic
-and aligned distribution.  The evidence, or marginal likelihood, for
-each of the models is shown in Figure \ref{fig:O1-odds}.  For all
-three magnitude distributions, the mixture models' posterior on the
-mixing fraction peaks at 100\% isotropic, which explains why the
-zero-parameter, pure-isotropic models are preferred over the
-single-parameter mixture models for every magnitude distribution with
-this data set.  Not surprisingly, given the small $\chieff$ values in
-the three detected systems, the most-favoured model among those with
-an isotropic angular distribution has the ``low'' magnitude
-distribution; the most favoured model among those with an aligned
-distribution also has the ``low'' magnitude distribution.  The odds
-ratio between the ``low'' aligned and ``low'' isotropic models is
-$\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$; thus the data
-favour isotropic spins among our suite of models.  While the data
-favour spin amplitude distributions with small spin magnitudes, note
-that a model with all \ac{BBH} systems having zero spin is ruled out
-by the GW151226 measurements, which bound at least one black hole to
-have spin magnitude $\geq 0.2$ at 99\%
+We fit hierarchical models of the four existing LIGO O1 and O2
+observations using these six different, zero-parameter population
+distributions (see Methods Section \ref{sec:hierarchical}).  We also
+fit three mixture models for the population, where the spin magnitude
+distribution is fixed but the angular distribution is a weighted sum
+of the isotropic and aligned distribution.  The evidence, or marginal
+likelihood, for each of the models is shown in Figure
+\ref{fig:O1-odds}.  For all three magnitude distributions, the mixture
+models' posterior on the mixing fraction peaks at 100\% isotropic,
+which explains why the zero-parameter, pure-isotropic models are
+preferred over the single-parameter mixture models for every magnitude
+distribution with this data set.  Not surprisingly, given the small
+$\chieff$ values in the three detected systems, the most-favoured
+model among those with an isotropic angular distribution has the
+``low'' magnitude distribution; the most favoured model among those
+with an aligned distribution also has the ``low'' magnitude
+distribution.  The odds ratio between the ``low'' aligned and ``low''
+isotropic models is $\OOneOddsIsoAligned$, or
+$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
+among our suite of models.  While the data favour spin amplitude
+distributions with small spin magnitudes, note that a model with all
+binary black hole systems having zero spin is ruled out by the
+GW151226 measurements, which bound at least one black hole to have
+spin magnitude $\geq 0.2$ at 99\%
 credibility\cite{2016PhRvL.116x1103A}.
@@
   \caption{\textbf{Odds ratios among our models using the
-      approximations to the posteriors on $\chieff$ from the O1
+      approximations to the posteriors on $\chieff$ from the O1 and O2
       observations shown in Figure \ref{fig:O1-posteriors}.}  The flat
@@
 
-Estimates of the rate of \ac{BBH} coalescences give a reasonable
-chance of 10 additional \ac{BBH} detections in the next three
-years\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming
-10 additional detections drawn from each of our six zero-parameter
-models for the spin distribution in addition to the three existing
-detections from O1, with observational uncertainties drawn randomly
-from the three Gaussian widths used to approximate the $\chieff$
-posteriors in Figure \ref{fig:O1-posteriors}\footnote{The measurement
-  uncertainty in $\chieff$ depends on the other parameters of the
-  merging \ac{BBH} system, particularly on the mass ratio.  Our
-  assumption about future observational uncertainties is appropriate
-  if the parameters of the three detected events are representative of
-  the parameters of future detections. See Methods Section
+Estimates of the rate of binary black hole coalescences give a
+reasonable chance of 10 additional binary black hole detections in the
+next three years\cite{O1-BBH,2016ApJ...833L...1A}.  Assuming 10
+additional detections drawn from each of our six zero-parameter models
+for the spin distribution in addition to the four existing detections
+from O1 and O2, with observational uncertainties drawn randomly from
+the three Gaussian widths used to approximate the $\chieff$ posteriors
+in Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty
+  in $\chieff$ depends on the other parameters of the merging binary
+  black hole system, particularly on the mass ratio.  Our assumption
+  about future observational uncertainties is appropriate if the
+  parameters of the three detected events are representative of the
+  parameters of future detections. See Methods Section
   \ref{sec:chi-eff-precision} for further discussion.}, we find the
@@
     that model uses the ``high'' magnitude distribution, which is
-    strongly dis-favoured from the O1 observations alone.}
+    strongly dis-favoured from the O1 and O2 observations alone.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
-\caption{\textbf{Fraction of the BBH population coming from an isotropic
-  distribution under a mixture model.} The dotted line shows the flat
-  prior on the fraction of BBHs coming from an isotropic distribution,
-  $f_i$, under the mixture model. The 3 red lines show the posterior
-  on $f_i$ after O1 with our various assumptions regarding BH spin
-  magnitudes.  The solid line shows the posterior assuming that all
-  BHs have their spin magnitude drawn from the ``flat''
-  distribution. The dashed line assumes the ``high'' BH spin magnitude
-  distribution $p(a) = 2a$. The dot-dash line assumes the ``low''
-  distribution $p(a) = 2(1-a)$.  We see that for a wide range of
-  assumptions regarding BH spin magnitudes, the fraction coming from
-  an isotropic distribution $f_i$ peaks at 1.}
+\caption{\textbf{Fraction of the BBH population coming from an
+    isotropic distribution under a mixture model.} The dotted line
+  shows the flat prior on the fraction of BBHs coming from an
+  isotropic distribution, $f_i$, under the mixture model. The 3 red
+  lines show the posterior on $f_i$ after O1 and O2 with our various
+  assumptions regarding BH spin magnitudes.  The solid line shows the
+  posterior assuming that all BHs have their spin magnitude drawn from
+  the ``flat'' distribution. The dashed line assumes the ``high'' BH
+  spin magnitude distribution $p(a) = 2a$. The dot-dash line assumes
+  the ``low'' distribution $p(a) = 2(1-a)$.  We see that for a wide
+  range of assumptions regarding BH spin magnitudes, the fraction
+  coming from an isotropic distribution $f_i$ peaks at 1.}
 \label{fig:mixture_fraction_posterior}
@@
 posterior peaks at $f_i = 1$. Thus, for these spin magnitude
-distributions we find that the current O1 LIGO observations constrain
-the majority of BBHs to have their spins drawn from an isotropic
-distribution. The evidence ratios of these models to the isotropic
-distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for
-the ``low'', ``flat'' and ``high'' spin magnitude models. Thus we
-cannot rule out a mixture with the current data.  If several different
-components contribute significantly to the true spin distribution it
-may take tens to hundreds of detections to accurately determined the
-mixing fraction, depending on the distribution of spin
+distributions we find that the current O1 and O2 LIGO observations
+constrain the majority of BBHs to have their spins drawn from an
+isotropic distribution. The evidence ratios of these models to the
+isotropic distribution with ``flat'' spin magnitudes are 0.85, 0.39
+and 0.19 for the ``low'', ``flat'' and ``high'' spin magnitude
+models. Thus we cannot rule out a mixture with the current data.  If
+several different components contribute significantly to the true spin
+distribution it may take tens to hundreds of detections to accurately
+determined the mixing fraction, depending on the distribution of spin
 magnitudes\cite{2017CQGra..34cLT01V,Stevenson:2017spin}.
@@
 
-\section{Approximations in the Gravitational Waveform}
+\section{Approximations in the Gravitational Waveform and Selection Effects}
+While the Advanced LIGO searches use spin-aligned templates they are
+efficient in detecting misaligned binary black hole
+systems\cite{2016PhRvD..93l2003A}; we assume here that the $\chieff$
+distribution of observed sources follows the true population.
+
 The model waveforms used to infer the $\chieff$ of the three LIGO

### 2017-06-03 paper/AlignedIsotropicGW.tex
Reduced words in abstract.

@@
   from merging binary black holes open a unique window into the binary
-  black hole formation environment.  One promising signature of the
-  formation environment is the angular distribution of the black hole
-  spins; systems formed through dynamical interactions among
-  already-compact objects are expected to have isotropic spin
+  black hole formation environment.  One promising environmental
+  signature is the angular distribution of the black hole spins;
+  systems formed through dynamical interactions among already-compact
+  objects are expected to have isotropic spin
   orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
@@
   likely to have spins preferentially aligned with the binary orbital
-  angular momentum as a consequence of their joint evolution toward a
-  binary black hole
-  system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
+  angular
+  momentum\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
   We consider existing gravitational wave measurements of the binary
@@
   GW151226, and GW170104.  If binary black hole spin magnitudes extend
-  to high values we show that the data already exhibit a
+  to high values we show that the data exhibit a
   $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
@@
   angular distribution over an aligned one.  By considering the effect
-  of 10 additional detections\cite{2016ApJ...833L...1A} of systems, we
-  show that if all observations come from a single population such an
-  augmented data set would typically enable at least a
+  of 10 additional detections\cite{2016ApJ...833L...1A}, we show that
+  such an augmented data set would typically enable at least a
   $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
   ratio) distinction, and in most cases better than $5\sigma$
-  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
-  either an isotropic spin distribution or low spin magnitudes for the
-  observed systems will be confirmed (or overturned) confidently in
-  the near future by subsequent observations.
+  ($2.9 \times 10^{-7}$ odds ratio).  % The existing preference for
+  % either an isotropic spin distribution or low spin magnitudes for the
+  % observed systems will be confirmed (or overturned) confidently in
+  % the near future by subsequent observations.
 \end{abstract}

### 2017-06-03 paper/AlignedIsotropicGW.tex
Removed some references, but still need to eliminate some more to get to 30.

@@
   angular
-  momentum\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
+  momentum\cite{TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
   We consider existing gravitational wave measurements of the binary
@@
 The spin directions in isolated binary black
-holes\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
+holes\cite{TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
 are usually expected to be preferentially aligned.  Despite observed
 spin-orbit misalignments in massive stellar
-binaries\cite{Albrecht:2009,2014ApJ...785...83A}, mass transfer and
-tidal interactions will tend to realign the binary.  On the other
-hand, there is some evidence of spin-orbit misalignment in black hole
-X-ray
-binaries\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014}.
-This is consistent with the expectation that a supernova natal kick
-(if any) can change the orbital plane and misalign the
-binary\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova
-can also tilt the spin angle\cite{2011ApJ...742...81F}.  Evolutionary
-processes, such as wind-driven mass loss and post-collapse fallback,
-can couple the spin magnitude and direction distributions, contrary to
-our simplified assumptions.  A small misalignment at wide separation
-can also evolve to a more significant misalignment in component spins
-as the binary spirals in through \ac{GW}
-emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
-conserved through this evolution.
+binaries\cite{Albrecht:2009}, mass transfer and tidal interactions
+will tend to realign the binary.  On the other hand, there is some
+evidence of spin-orbit misalignment in black hole X-ray
+binaries\cite{Orosz:2001,Martin:2008b,MorningstarMiller:2014}.  This
+is consistent with the expectation that a supernova natal kick (if
+any) can change the orbital plane and misalign the
+binary\cite{2000ApJ...541..319K}; the supernova can also tilt the spin
+angle\cite{2011ApJ...742...81F}.  Evolutionary processes, such as
+wind-driven mass loss and post-collapse fallback, can couple the spin
+magnitude and direction distributions, contrary to our simplified
+assumptions.  A small misalignment at wide separation can also evolve
+to a more significant misalignment in component spins as the binary
+spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
+$\chieff$ is approximately conserved through this evolution.
 

### 2017-06-03 paper/AlignedIsotropicGW.tex
Some typo fixes, and a few places where three has become four after GW170104.

@@
   \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
-      $\chieff$ from the Advanced LIGO O1 run
-      observations\cite{O1-BBH}.}  We approximate the posteriors
-    reported using Gaussians with the same median and 90\% credible
-    interval.  It is notable that none of the $\chieff$ posteriors
-    support high \ac{BH} spin magnitudes with aligned spins, suggested
-    by observations of stellar-mass black holes in X-ray
-    binaries\cite{2015PhR...548....1M}.}
+      $\chieff$ from the Advanced LIGO O1 and O2
+      observations\cite{O1-BBH,PhysRevLett.118.221101}.}  We
+    approximate the posteriors reported using Gaussians with the same
+    median and 90\% credible interval.  It is notable that none of the
+    $\chieff$ posteriors support high \ac{BH} spin magnitudes with
+    aligned spins, suggested by observations of stellar-mass black
+    holes in X-ray binaries\cite{2015PhR...548....1M}.}
 \end{figure}
@@
 distributions of spin for each component in the binary may differ
-systematically due to different formation histories.  Choosing one of
+systematically due to different formation histories).  Choosing one of
 three magnitude distributions (see Methods Section
@@
   \caption{\label{fig:O2-predictions} \textbf{Distribution of odds
-      ratios predicted with 10 additional observations above the three
+      ratios predicted with 10 additional observations above the four
       discussed above.}  Each panel corresponds to additional
@@
     distribution; most models result in more than $5\sigma$ preference
-    for the correct angular distribution.  Because the three existing
+    for the correct angular distribution.  Because the four existing
     observations are included in each data set the ``correct'' model
@@
 given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in
-the derivation) in Figure~\ref{fig:mixture_fraction_posterior}. We
-find the mean fraction of BBHs coming from an isotropic distribution
-is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high''
-distributions for spin magnitudes respectively, compared to the prior
-mean of 0.5. The lower 90\% limits are 0.26, 0.39 and 0.52
-respectively, compared to the prior of 0.1. In all cases, the
-posterior peaks at $f_i = 1$. Thus, for these spin magnitude
-distributions we find that the current O1 and O2 LIGO observations
-constrain the majority of BBHs to have their spins drawn from an
-isotropic distribution. The evidence ratios of these models to the
-isotropic distribution with ``flat'' spin magnitudes are 0.85, 0.39
-and 0.19 for the ``low'', ``flat'' and ``high'' spin magnitude
-models. Thus we cannot rule out a mixture with the current data.  If
-several different components contribute significantly to the true spin
-distribution it may take tens to hundreds of detections to accurately
-determined the mixing fraction, depending on the distribution of spin
+the derivation) in Extended Data
+Figure~\ref{fig:mixture_fraction_posterior}. We find the mean fraction
+of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78
+assuming the ``low'', ``flat'' and ``high'' distributions for spin
+magnitudes respectively, compared to the prior mean of 0.5. The lower
+90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the
+prior of 0.1. In all cases, the posterior peaks at $f_i = 1$. Thus,
+for these spin magnitude distributions we find that the current O1 and
+O2 LIGO observations constrain the majority of BBHs to have their
+spins drawn from an isotropic distribution. The evidence ratios of
+these models to the isotropic distribution with ``flat'' spin
+magnitudes are 0.85, 0.39 and 0.19 for the ``low'', ``flat'' and
+``high'' spin magnitude models. Thus we cannot rule out a mixture with
+the current data.  If several different components contribute
+significantly to the true spin distribution it may take tens to
+hundreds of detections to accurately determined the mixing fraction,
+depending on the distribution of spin
 magnitudes\cite{2017CQGra..34cLT01V,Stevenson:2017spin}.
@@
 
-In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of
-isotropic to aligned distributions (plotted as the equivalent sigma)
-with spin magnitudes given by this model with $\alpha$ in the range
-$0$--$6$. The top axis shows the mean spin magnitude that value of
-$\alpha$ corresponds to (e.g., for the ``flat'' distribution
+In Extended Data Figure~\ref{fig:smallspinsalpha} we plot the evidence
+ratio of isotropic to aligned distributions (plotted as the equivalent
+sigma) with spin magnitudes given by this model with $\alpha$ in the
+range $0$--$6$. The top axis shows the mean spin magnitude that value
+of $\alpha$ corresponds to (e.g., for the ``flat'' distribution
 $\alpha = 0$, the mean spin magnitude is 0.5). We see that if typical
@@
 
-Figure \ref{fig:mass-ratio-sensitivity} shows the distributions of
-$\chieff$ that would obtain with a mass ratio $q = m_2/m_1 = 0.5$
-compared to the distributions with $q = 1$ used above.  The details of
-the distribution are sensitive to the mass ratio, but in our analysis
-we are primarily sensitive to the changing \emph{sign} of $\chieff$
-under the isotropic models.  This latter property is insensitive to
-mass ratio.  As an example, the distinction between the three
-different spin amplitude distributions after ten additional detections
-is quite weak compared to the aligned/isotropic distinction in Figure
-\ref{fig:O2-predictions}.  The differences in the $\chieff$
-distribution between $q = 1$ and $q = 0.5$ are even smaller than the
-differences between the different magnitude distributions.
+Extended Data Figure \ref{fig:mass-ratio-sensitivity} shows the
+distributions of $\chieff$ that would obtain with a mass ratio
+$q = m_2/m_1 = 0.5$ compared to the distributions with $q = 1$ used
+above.  The details of the distribution are sensitive to the mass
+ratio, but in our analysis we are primarily sensitive to the changing
+\emph{sign} of $\chieff$ under the isotropic models.  This latter
+property is insensitive to mass ratio.  As an example, the distinction
+between the three different spin amplitude distributions after ten
+additional detections is quite weak compared to the aligned/isotropic
+distinction in Figure \ref{fig:O2-predictions}.  The differences in
+the $\chieff$ distribution between $q = 1$ and $q = 0.5$ are even
+smaller than the differences between the different magnitude
+distributions.
 

### 2017-06-03 paper/AlignedIsotropicGW.tex
Minor text fix in section 8.

@@
 particular its true $\chieff$ and mass ratio, which we illustrate in
-figure \ref{fig:chi-eff-constraints}.
+Extended Data Figure \ref{fig:chi-eff-constraints}.
 
@@
 
-By neglecting these effects we do not expect to qualitatively affect
+We do not expect these effects to qualitatively affect
 out conclusions, though they could affect predictions for the total

### 2017-06-03 paper/AlignedIsotropicGW.tex
Added table of evidence accumulation to supplementary materials

@@
 \centering
-\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
+\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction_with_GW170104.png}
 \caption{\textbf{Fraction of the BBH population coming from an
@@
 
-We also fit a mixture model\cite{Stevenson:2017spin} (labelled model
+We fit a mixture model\cite{Stevenson:2017spin} (labelled model
 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have
@@
 
+\subsection{Accumulation of evidence}
+
+In Table~\ref{tab:accumulation} we show how the evidence for an isotropic distribution increases when including; only the 2 confirmed events---GW150914 and GW151226---from O1; all O1 events (including LVT151012); and all 4 likely binary black hole mergers, including GW170104.
+
+\begin{table}
+\begin{centering}
+\begin{tabular}{ c |  c  |  c  | c  }
+  \hline 
+  \hline
+Events & $\sigma_{I/A}$ &  $\sigma_{I/A}$ & $\sigma_{I/A}$ \\ 
+ & ``Low'' & ``Flat'' & ``High'' \\ \hline
+GW150914 and GW151226 &  1.3 & 2.2 & 3.7 \\ %\hline
+All O1 events & 1.7 & 2.7 & 4.4 \\ %\hline
+All O1 events and GW170104 & \textbf{2.4} & 3.6 & 5.4 \\ \hline
+\end{tabular}
+\caption[]{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
+The number in bold is the result we quote in the main text.}
+\label{tab:accumulation}
+\end{centering}
+\end{table}
+
 \section{Effect of small spin magnitudes}
@@
 \centering
-\includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
+\includegraphics[width=\textwidth]{../plots/sigma_v_alpha_GW170104.png}
 \caption{\textbf{Effect of small spins on evidence ratio of isotropic against

### 2017-06-03 paper/AlignedIsotropicGW.tex
Edited numbers in Methods section 3

@@
 Figure~\ref{fig:mixture_fraction_posterior}. We find the mean fraction
-of BBHs coming from an isotropic distribution is 0.63, 0.71 and 0.78
+of BBHs coming from an isotropic distribution is 0.70, 0.77 and 0.81
 assuming the ``low'', ``flat'' and ``high'' distributions for spin
 magnitudes respectively, compared to the prior mean of 0.5. The lower
-90\% limits are 0.26, 0.39 and 0.52 respectively, compared to the
+90\% limits are 0.38, 0.51 and 0.60 respectively, compared to the
 prior of 0.1. In all cases, the posterior peaks at $f_i = 1$. Thus,
@@
 O2 LIGO observations constrain the majority of BBHs to have their
-spins drawn from an isotropic distribution. The evidence ratios of
-these models to the isotropic distribution with ``flat'' spin
-magnitudes are 0.85, 0.39 and 0.19 for the ``low'', ``flat'' and
-``high'' spin magnitude models. Thus we cannot rule out a mixture with
+spins drawn from an isotropic distribution. The evidence ratios of these mixture models to the isotropic distribution with ``low'' spin magnitudes are 0.43, 0.20 and 0.10 for the ``low'', ``flat''
+and ``high'' spin magnitude models. Thus we cannot rule out a mixture with
 the current data.  If several different components contribute

### 2017-06-04 paper/AlignedIsotropicGW.tex
Alberto's remaining comments, and restore the 'caveats' sentence to abstract.

@@
   of 10 additional detections\cite{2016ApJ...833L...1A}, we show that
-  such an augmented data set would typically enable at least a
-  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
-  ratio) distinction, and in most cases better than $5\sigma$
-  ($2.9 \times 10^{-7}$ odds ratio).  % The existing preference for
-  % either an isotropic spin distribution or low spin magnitudes for the
-  % observed systems will be confirmed (or overturned) confidently in
-  % the near future by subsequent observations.
+  such an augmented data set would enable in most cases a prefernce
+  stronger than $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The
+  existing preference for either an isotropic spin distribution or low
+  spin magnitudes for the observed systems will be confirmed (or
+  overturned) confidently in the near future.
 \end{abstract}
@@
 for the majority of black holes in X-ray binaries with claimed spin
-measurements (see below).  The analysis here is relatively insensitive
-to the precise details of the posterior distributions; other
-conclusions are more sensitive.  In particular, our Gaussian
-approximation does permit $\chieff = 0$ for GW151226 while the true
-posterior rules this out at high
+measurements\cite{2015PhR...548....1M} (see below).  The analysis here
+is relatively insensitive to the precise details of the posterior
+distributions; other conclusions are more sensitive.  In particular,
+our Gaussian approximation does permit $\chieff = 0$ for GW151226
+while the true posterior rules this out at high
 confidence\cite{2016PhRvL.116x1103A,O1-BBH}.
@@
 
-These magnitude distributions not meant to represent any particular
-physical model, but rather to capture our uncertainty about the spin
-magnitude distribution, discussed in detail below; neither
+These magnitude distributions are not meant to represent any
+particular physical model, but rather to capture our uncertainty about
+the spin magnitude distribution, discussed in detail below; neither
 observations nor population synthesis codes can at this point

### 2017-06-04 paper/AlignedIsotropicGW.tex
Correct chi_{1,2} bounds.

@@
 holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
-assumed to point in the $\hat{z}$ direction, and
-$-1 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
-projections of the individual \ac{BH} spins.  Because the
-dimensionless spin parameter,
+assumed to point in the $\hat{z}$ direction, and $\chi_{1,2}$ are the
+corresponding dimensionless projections of the individual \ac{BH}
+spins.  Because the dimensionless spin parameter,
 \begin{equation}

### 2017-06-04 paper/AlignedIsotropicGW.tex
O2 --> GW170104 where appropriate.

@@
   \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
-      $\chieff$ from the Advanced LIGO O1 and O2
+      $\chieff$ from the Advanced LIGO O1 and GW170104
       observations\cite{O1-BBH,PhysRevLett.118.221101}.}  We
@@
 
-We fit hierarchical models of the four existing LIGO O1 and O2
+We fit hierarchical models of the four existing LIGO O1 and GW170104
 observations using these six different, zero-parameter population
@@
   \caption{\textbf{Odds ratios among our models using the
-      approximations to the posteriors on $\chieff$ from the O1 and O2
+      approximations to the posteriors on $\chieff$ from the O1 and GW170104
       observations shown in Figure \ref{fig:O1-posteriors}.}  The flat
@@
 for the spin distribution in addition to the four existing detections
-from O1 and O2, with observational uncertainties drawn randomly from
-the three Gaussian widths used to approximate the $\chieff$ posteriors
-in Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty
-  in $\chieff$ depends on the other parameters of the merging binary
-  black hole system, particularly on the mass ratio.  Our assumption
-  about future observational uncertainties is appropriate if the
-  parameters of the three detected events are representative of the
-  parameters of future detections. See Methods Section
-  \ref{sec:chi-eff-precision} for further discussion.}, we find the
-odds ratios shown in Figure \ref{fig:O2-predictions}.  We find that
-most scenarios with an additional 10 detections allow the simulated
-angular distribution to be inferred with greater than $5\sigma$
-($2.9 \times 10^{-7}$ odds) credibility; and in the most pessimistic
-case the distinction is typically $\OTwoSigmaIsoAlignedMin\sigma$
-($\OTwoOddsIsoAlignedMin$ odds ratio).  While such future detections
-should permit a confident distinction between \emph{angular}
-distributions, we would remain much less certain about the
-\emph{magnitude} distribution among the three options considered here
-until we have a larger number of
+from O1 and GW170104, with observational uncertainties drawn randomly
+from the three Gaussian widths used to approximate the $\chieff$
+posteriors in Figure \ref{fig:O1-posteriors}\footnote{The measurement
+  uncertainty in $\chieff$ depends on the other parameters of the
+  merging binary black hole system, particularly on the mass ratio.
+  Our assumption about future observational uncertainties is
+  appropriate if the parameters of the three detected events are
+  representative of the parameters of future detections. See Methods
+  Section \ref{sec:chi-eff-precision} for further discussion.}, we
+find the odds ratios shown in Figure \ref{fig:O2-predictions}.  We
+find that most scenarios with an additional 10 detections allow the
+simulated angular distribution to be inferred with greater than
+$5\sigma$ ($2.9 \times 10^{-7}$ odds) credibility; and in the most
+pessimistic case the distinction is typically
+$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
+While such future detections should permit a confident distinction
+between \emph{angular} distributions, we would remain much less
+certain about the \emph{magnitude} distribution among the three
+options considered here until we have a larger number of
 observations\cite{2017arXiv170306869F,2017arXiv170306223G}.  In Figure
@@
     that model uses the ``high'' magnitude distribution, which is
-    strongly dis-favoured from the O1 and O2 observations alone.}
+    strongly dis-favoured from the O1 and GW170104 observations
+    alone.}
 \end{figure}
@@
   isotropic distribution, $f_i$, under the mixture model. The 3 red
-  lines show the posterior on $f_i$ after O1 and O2 with our various
-  assumptions regarding BH spin magnitudes.  The solid line shows the
-  posterior assuming that all BHs have their spin magnitude drawn from
-  the ``flat'' distribution. The dashed line assumes the ``high'' BH
-  spin magnitude distribution $p(a) = 2a$. The dot-dash line assumes
-  the ``low'' distribution $p(a) = 2(1-a)$.  We see that for a wide
-  range of assumptions regarding BH spin magnitudes, the fraction
-  coming from an isotropic distribution $f_i$ peaks at 1.}
+  lines show the posterior on $f_i$ after O1 and GW170104 with our
+  various assumptions regarding BH spin magnitudes.  The solid line
+  shows the posterior assuming that all BHs have their spin magnitude
+  drawn from the ``flat'' distribution. The dashed line assumes the
+  ``high'' BH spin magnitude distribution $p(a) = 2a$. The dot-dash
+  line assumes the ``low'' distribution $p(a) = 2(1-a)$.  We see that
+  for a wide range of assumptions regarding BH spin magnitudes, the
+  fraction coming from an isotropic distribution $f_i$ peaks at 1.}
 \label{fig:mixture_fraction_posterior}
@@
 
-We fit a mixture model\cite{Stevenson:2017spin} (labelled model
-'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have
-spins drawn from an isotropic distribution, whilst a fraction
-$1 - f_i$ have their spins aligned with the orbital angular
-momentum. We assume a flat prior on the fraction $f_i$. To test the
-robustness of our result, we vary the distribution we assume for BH
-spin magnitude distributions as with the aligned and isotropic
-models. We use the ``flat'', ``high'' and ``low'' distributions
+We fit a mixture model\cite{Stevenson:2017spin} (labelled model 'M' in
+Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have spins
+drawn from an isotropic distribution, whilst a fraction $1 - f_i$ have
+their spins aligned with the orbital angular momentum. We assume a
+flat prior on the fraction $f_i$. To test the robustness of our
+result, we vary the distribution we assume for BH spin magnitude
+distributions as with the aligned and isotropic models. We use the
+``flat'', ``high'' and ``low'' distributions
 (Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin
@@
 for these spin magnitude distributions we find that the current O1 and
-O2 LIGO observations constrain the majority of BBHs to have their
-spins drawn from an isotropic distribution. The evidence ratios of these mixture models to the isotropic distribution with ``low'' spin magnitudes are 0.43, 0.20 and 0.10 for the ``low'', ``flat''
-and ``high'' spin magnitude models. Thus we cannot rule out a mixture with
+GW170104 LIGO observations constrain the majority of BBHs to have
+their spins drawn from an isotropic distribution. The evidence ratios
+of these mixture models to the isotropic distribution with ``low''
+spin magnitudes are 0.43, 0.20 and 0.10 for the ``low'', ``flat'' and
+``high'' spin magnitude models. Thus we cannot rule out a mixture with
 the current data.  If several different components contribute

### 2017-06-04 paper/AlignedIsotropicGW.tex
Punctuation fix.

@@
 
-In Table~\ref{tab:accumulation} we show how the evidence for an isotropic distribution increases when including; only the 2 confirmed events---GW150914 and GW151226---from O1; all O1 events (including LVT151012); and all 4 likely binary black hole mergers, including GW170104.
+In Table~\ref{tab:accumulation} we show how the evidence for an isotropic distribution increases when including: only the 2 confirmed events---GW150914 and GW151226---from O1; all O1 events (including LVT151012); and all 4 likely binary black hole mergers, including GW170104.
 

### 2017-06-04 paper/AlignedIsotropicGW.tex
Some typos fixed, some references eliminated.

@@
   effective spin, the best-measured combination of spin
-  parameters\cite{2016PhRvL.116x1102A,O1-BBH,PhysRevLett.118.221101},
+  parameters\cite{O1-BBH,PhysRevLett.118.221101},
   in the four likely binary black hole detections GW150914, LVT151012,
@@
 waveforms, and a correspondingly tight constraint from the
-data\cite{2016PhRvL.116x1102A}, is a mass-weighted combination of the
-components of the dimensionless spin vectors of the two black holes
-that are aligned with the orbital axis, the ``effective spin,''
-$-1 < \chieff < 1$ (see Methods Section
-\ref{methsec:chieff-spin-magnitude}).
+data\cite{O1-BBH}, is a mass-weighted combination of the components of
+the dimensionless spin vectors of the two black holes that are aligned
+with the orbital axis, the ``effective spin,'' $-1 < \chieff < 1$ (see
+Methods Section \ref{methsec:chieff-spin-magnitude}).
 
@@
 
-We fit hierarchical models of the four existing LIGO O1 and GW170104
+We fit hierarchical models of the three existing LIGO O1 and GW170104
 observations using these six different, zero-parameter population
@@
 could reduce the available angular
-momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
+momentum\cite{2016MNRAS.462..844K,2017arXiv170203952H}.
 

### 2017-06-04 paper/AlignedIsotropicGW.tex
Temporarily commented out 'nature.com' addresses for arXiv submission---it's not available there yet :).
More tweaks for arXiv submission.

@@
 \begin{figure}
-  \plotone{../plots/chi-eff-mock-posteriors}
+  \plotone{plots/chi-eff-mock-posteriors}
   \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
@@
 \begin{figure}
-  \plotone{../plots/chi-eff-distributions}
+  \plotone{plots/chi-eff-distributions}
   \caption{\label{fig:chieff-distribution-models} \textbf{The models for the
@@
 \begin{figure}
-  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models}
+  \plotone{plots/Wills_evidence_ratio_figure_with_mixture_models}
   \caption{\textbf{Odds ratios among our models using the
@@
 \begin{figure}
-  \plotone{../plots/six-way-O2-model-selection}
+  \plotone{plots/six-way-O2-model-selection}
   \caption{\label{fig:O2-predictions} \textbf{Distribution of odds
@@
   No. NSF PHY11-25915.
-\item[Supplementary Information] Supplementary Information is linked
-  to the online version of the paper at
-  \url{http://www.nature.com/nature}.
+% \item[Supplementary Information] Supplementary Information is linked
+%   to the online version of the paper at
+%   \url{http://www.nature.com/nature}.
 \item[Author Contributions] All authors contributed to the work
   presented in this paper.
-\item[Author Information] Reprints and permissions information is
-  available at \url{http://www.nature.com/reprints}.  The authors
-  declare no competing financial interests.  Correspondence and
-  requests for materials should be addressed to
+\item[Author Information] %Reprints and permissions information is
+  %available at \url{http://www.nature.com/reprints}.  
+  The authors declare no competing financial interests.
+  Correspondence and requests for materials should be addressed to
   \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{../plots/pa}
+  \includegraphics[width=\columnwidth]{plots/pa}
   \caption{\textbf{Distributions of spin magnitudes.}  See Eq.\
@@
 \centering
-\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction_with_GW170104.png}
+\includegraphics[width=\textwidth]{plots/posterior_on_isotropic_fraction_with_GW170104.png}
 \caption{\textbf{Fraction of the BBH population coming from an
@@
 %\centering
-%\includegraphics[width=0.45\textwidth]{../plots/isotropic_fraction_cumulative_posterior.png}
+%\includegraphics[width=0.45\textwidth]{plots/isotropic_fraction_cumulative_posterior.png}
 %\caption{\textbf{Fraction of the BBH population coming from an isotropic distribution} The dotted (blue) line shows the cumulative distribution for the flat prior on the fraction of BBHs coming from an isotropic distribution $f_i$.
@@
 \centering
-\includegraphics[width=\textwidth]{../plots/sigma_v_alpha_GW170104.png}
+\includegraphics[width=\textwidth]{plots/sigma_v_alpha_GW170104.png}
 \caption{\textbf{Effect of small spins on evidence ratio of isotropic against
@@
 \begin{figure}
-  \plotone{../plots/chi-eff-distributions-q05}
+  \plotone{plots/chi-eff-distributions-q05}
   \caption{\textbf{Distributions of $\chieff$ assuming all merging
@@
 \begin{figure}
-  \plotone{../plots/chi-eff-90cl}
+  \plotone{plots/chi-eff-90cl}
   \caption{\textbf{Widths of the 90\% credible intervals for $\chieff$

### 2017-06-05 paper/AlignedIsotropicGW.tex
Revert "Temporarily commented out 'nature.com' addresses for arXiv submission---it's not available there yet :)."
This reverts commit 98d1def41d58150cd88d5267f238130b7c5314de.
More references on the chopping block.

--- 2017-06-04 paper/AlignedIsotropicGW.tex
@@
   objects are expected to have isotropic spin
-  orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
+  orientations\cite{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
   whereas binaries formed from pairs of stars born together are more
@@
   angular
-  momentum\cite{TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
+  momentum\cite{TutukovYungelson:1993,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
   We consider existing gravitational wave measurements of the binary
   effective spin, the best-measured combination of spin
-  parameters\cite{O1-BBH,PhysRevLett.118.221101},
-  in the four likely binary black hole detections GW150914, LVT151012,
-  GW151226, and GW170104.  If binary black hole spin magnitudes extend
-  to high values we show that the data exhibit a
-  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
-  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
-    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
-    $\Phi$ is the unit normal CDF.})  preference for an isotropic
-  angular distribution over an aligned one.  By considering the effect
-  of 10 additional detections\cite{2016ApJ...833L...1A}, we show that
-  such an augmented data set would enable in most cases a prefernce
-  stronger than $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The
-  existing preference for either an isotropic spin distribution or low
-  spin magnitudes for the observed systems will be confirmed (or
+  parameters\cite{O1-BBH,PhysRevLett.118.221101}, in the four likely
+  binary black hole detections GW150914, LVT151012, GW151226, and
+  GW170104.  If binary black hole spin magnitudes extend to high
+  values we show that the data exhibit a $\OOneSigmaIsoAligned\sigma$
+  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
+    $r \ll 1$ is equivalent to $x \sigma$ with
+    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
+    normal CDF.})  preference for an isotropic angular distribution
+  over an aligned one.  By considering the effect of 10 additional
+  detections\cite{2016ApJ...833L...1A}, we show that such an augmented
+  data set would enable in most cases a prefernce stronger than
+  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
+  preference for either an isotropic spin distribution or low spin
+  magnitudes for the observed systems will be confirmed (or
   overturned) confidently in the near future.
@@
 Following the detection of a merging binary black hole system,
-parameter estimation tools\cite{2015PhRvD..91d2003V} compare model
-gravitational waveforms against the observed data to obtain a
-posterior distribution on the parameters that describe the compact
-binary source.  The spin parameter with the largest effect on
-waveforms, and a correspondingly tight constraint from the
-data\cite{O1-BBH}, is a mass-weighted combination of the components of
-the dimensionless spin vectors of the two black holes that are aligned
-with the orbital axis, the ``effective spin,'' $-1 < \chieff < 1$ (see
-Methods Section \ref{methsec:chieff-spin-magnitude}).
+parameter estimation tools compare model gravitational waveforms
+against the observed data to obtain a posterior distribution on the
+parameters that describe the compact binary source.  The spin
+parameter with the largest effect on waveforms, and a correspondingly
+tight constraint from the data\cite{O1-BBH}, is a mass-weighted
+combination of the components of the dimensionless spin vectors of the
+two black holes that are aligned with the orbital axis, the
+``effective spin,'' $-1 < \chieff < 1$ (see Methods Section
+\ref{methsec:chieff-spin-magnitude}).
 
@@
 certain about the \emph{magnitude} distribution among the three
-options considered here until we have a larger number of
-observations\cite{2017arXiv170306869F,2017arXiv170306223G}.  In Figure
-\ref{fig:O2-predictions}, the odds ratio between different magnitude
-distributions with the same angular distribution is much closer to
-unity than the odds ratio between angular distributions.
+options considered here until we have a larger number of observations.
+In Figure \ref{fig:O2-predictions}, the odds ratio between different
+magnitude distributions with the same angular distribution is much
+closer to unity than the odds ratio between angular distributions.
 
@@
 accretion\cite{1999MNRAS.305..654K}, but this may not be true for all
-systems\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the
-current spin parameters are probably close to their values upon core
-collapse, at least in high-mass X-ray binaries.  However, the specific
-processes involved in the production of black hole binaries from
-isolated binaries could alter the spin magnitude distribution of those
-holes relative to the X-ray binary systems; for example, close tidal
-interactions could spin up the core, or stripping of the envelope
-could reduce the available angular
-momentum\cite{2016MNRAS.462..844K,2017arXiv170203952H}.
+systems\cite{2015ApJ...800...17F}.  Thus the current spin parameters
+are probably close to their values upon core collapse, at least in
+high-mass X-ray binaries.  However, the specific processes involved in
+the production of black hole binaries from isolated binaries could
+alter the spin magnitude distribution of those holes relative to the
+X-ray binary systems; for example, close tidal interactions could spin
+up the core, or stripping of the envelope could reduce the available
+angular
+momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
 
 The spin directions in isolated binary black
-holes\cite{TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
+holes\cite{TutukovYungelson:1993,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
 are usually expected to be preferentially aligned.  Despite observed
@@
 evidence of spin-orbit misalignment in black hole X-ray
-binaries\cite{Orosz:2001,Martin:2008b,MorningstarMiller:2014}.  This
-is consistent with the expectation that a supernova natal kick (if
-any) can change the orbital plane and misalign the
+binaries\cite{Martin:2008b,MorningstarMiller:2014}.  This is
+consistent with the expectation that a supernova natal kick (if any)
+can change the orbital plane and misalign the
 binary\cite{2000ApJ...541..319K}; the supernova can also tilt the spin
@@
 interactions in dense stellar
-environments\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+environments\cite{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
 are expected to be isotropic given the absence of a preferred
@@
 LIGO measures $\chieff$ better than any other spin parameter, but
-still with significant uncertainty, so we need to properly incorporate
-measurement uncertainty in our analysis; thus our analysis must be
-\emph{hierarchical}\cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
-In a hierarchical analysis, we assume that each event has a true, but
+still with significant uncertainty\cite{2015PhRvD..91d2003V}, so we
+need to properly incorporate measurement uncertainty in our analysis;
+thus our analysis must be
+\emph{hierarchical}\cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.  In
+a hierarchical analysis, we assume that each event has a true, but
 unknown, value of the effective spin, drawn from the population
@@
 ``low'', ``flat'' and ``high''. These were intended to capture some of
-the uncertainty regarding the BH spin magnitude distribution.
+the uncertainty regarding the BH spin magnitude distribution.  We may
+remain observationally uncertain about the spin magnitude distribution
+until we have $\mathcal{O}(100)$
+observations\cite{2017arXiv170306869F,2017arXiv170306223G}.
 

### 2017-06-05 paper/response.txt
As re-submitted.

Thanks very much for the referee reports and editorial comments on our
submitted manuscript.  We are re-submitting, having addressed all the
comments provided.  There have been a very large number of changes to
the text and references; below we list the changes made to address
each point individually.  If you would like, we could also provide a
diff against the original submitted LaTeX that would collect all the
changes together (without the logical ordering by point).

Editor's Comments
=================

Before addressing the comments in-line below, we should note that
GW170104 has been announced between receiving referee reports and this
re-submission.  We have, at the request of the editor, incorporated it
into our calculations.  This changes the odds ratio between the
isotropic and aligned models from 1.7 sigma to 2.4 sigma, and also the
expected minimum "distinguishing power" between models with ten
additional detections from 2.9 sigma to 2.4 sigma (not a
typo---increased confidence in the isotropic models means that more
detections from the "aligned" distributions are required to invert the
odds ratio in favour of these latter models).  Very little other text
has been changed as a result of the new detection (we have added
"GW170104" in appropriate places, changed "three" to "four," etc); the
central conclusions of the paper, and the caveats to the conclusions
about low spins / kick misalignments remain basically unchanged.

Item-by-item comments follow below:

> We estimate the current length of your paper to be ~1800 words,
  which exceeds our usual limit of 1500 words for Letters with four
  display items. Keep in mind that important technical details that
  are not central to the main message of the paper can be moved into
  the Methods section (right now, titled Appendix -- it is far below
  our general guideline of 3000 words).

Through addressing the other comments provided and moving material
into the Methods section (see below), we have significantly shortened
the main text.

> Letters begin with a fully referenced paragraph, ideally of about
  200 words, aimed at readers in other disciplines. This paragraph
  starts with a 2- to 3-sentence, basic introduction to the field that
  conveys to readers the ‘state of the art’ immediately before this
  work. It continues with a 1-sentence statement of the main findings
  starting 'Here we report’ or an equivalent phrase; and finally,
  concludes with 2 to 3 sentences putting the main findings into
  general context so it is clear how the results described in the
  paper have moved the field forward. Please do not introduce or use
  *any* abbreviations/acronyms in the first paragraph, and severely
  restrict their use elsewhere, as they make the paper difficult for
  non-specialists to read. The spin parameter should certainly be
  defined when first used, but it is probably not necessary to use it
  in the first para.

We have removed the abbreviations and the spin parameter frome the
abstract, and shortened it to be below 200 words.

> Further introductory material in the main text of the paper should
  not be necessary. Please note that Letters do not have
  sections/subsections/section headings.

We have removed any section headings from the letter portion of the
text (our Methods section retains section headings).

> Each display equation uses 30-40 word equivalents, and our style is
  to avoid them in the main text *unless* they are newly derived (not
  relevant here) or crucial to the overall logical flow of the
  text. The definition (eq 1) of effective spin probably could be
  shifted to the Methods, with a brief description in words in the
  main text. I am open to persuasion that eq 2 needs to be in the main
  text, but probably it too can be shifted to the Methods.

Both equations have now been shifted to the appendix.

> The "Discussion" section is rather long-winded and inconclusive. We
  recommend shifting that to the Methods, with a brief summary
  paragraph in the main text.

We have shortened and re-arranged the discussion section
significantly.  However, we would prefer not to move what remains to
Methods; we feel that the "caveats," particularly surrounding spin
magnitudes and alignment in "field" binaries, are scientifically
important.  For example, even one of our referees over-interpreted our
results, claiming that we had shown that dynamical formation scenarios
are strongly preferred over "field" formation---but in fact, it is
certainly possible that field formation of BBH mergers produces small
spins, in which case we cannot distinguish the two cases.  Moving the
remaining discussion paragraphs (there are three, and one is quite
short) to the Methods section would de-emphasise these caveats further
and, we feel, could lead to more confusion.

> Any discussion at the end of the paper should also be brief, and not
  repeat what is already written in the initial summary
  paragraph. (The final paragraph should be omitted -- Nature papers
  do not end with summaries.)

We removed the "summary" pargraph.

> At the end of the main text document (after the main figure
  legends), there should be a section entitled "Methods", which
  provides a more detailed discussion of the additional methodological
  information that would allow other researchers to replicate the
  results (we define “Methods” quite broadly). The Methods section
  will not appear in the print version but will be fully copy-edited
  and appear online in the full-text HTML and PDF versions. The
  Methods section should be written as concisely as possible but
  should contain all elements necessary to allow interpretation and
  reproduction of the results. If there are additional references in
  the Methods section, their numbering should continue from the last
  reference in the main paper, and the list should follow the Methods
  section. If the methods require chemical structures, figures or
  tables, these should be supplied as Extended Data (see below). For
  mathematically complex methods, or methods that require an unusually
  large number of figures or tables (beyond what can be accommodated
  as Extended Data), the entire Methods section should instead be
  supplied as a separate Supplementary Information.

Our previous "Appendix" has been incorporated in a "Methods" section.
We have moved material from the main text to this section (as
indicated above and below), and have also added sections on code
availability, the accumulation of evidence with events (see below re:
GW170104), and precision of chi_eff measurements (including Extended
Data Figure 5).

> Letters can have up to 30 references in the main text; additional
  references can be cited in (and listed after) the Methods section,
  as detailed above. Each reference should have its own number, and
  you should please include final page numbers, as well as initial
  ones, unless the first ‘page’ is an article number (i.e. A&A
  beginning in 2010, ApJ beginning in 2011). Our abbreviations for the
  common astronomy journals are: Astrophys. J., Astron. J.,
  Astron. Astrophys., Mon. Not. R. Astron. Soc.. Authors' last names
  should come first, followed by their initials. Any reference to work
  in preparation should be made in the main text, and not numbered in
  the reference list. As we are now printing the titles of all
  references cited, we would appreciate it if you could include them
  in the revised version of your text. The year of publication should
  be in parentheses at the end of the line.

We have re-formatted the bibliography in line with Nature
expectations.  We now have exactly 30 references in the main text,
having eliminated some citations that were superfluous and moved
others that were technical in nature to the Methods section.  The
references eliminated (in ADS notation) and the justification are

* 1993Natur.364..421K (Kulkarni, Hut & McMillan 1993): redundant with
  Siggurdsson & Hernquist (1993).

* 1973NInfo..27...70T (Tutukov & Yungelson 1973): redundant with
  Tutukov & Yungelson 1993.

* 1997MNRAS.288..245L (Lipunov, Postnov & Prokhorov 1997): redundant
  with Tutukov & Yungelson (1993) and Belczynski, et al. (2016).

* 2016PhRvL.116x1102A (Abbott, et al. 2016, properties of merger
  GW150914): redundant with Abbott, et al. (2016), 2016PhRvX...6d1015A
  (Black holes during O1 observing run).

* 2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H: These
  waveform technical references are not relevant to our argument or
  discussion.

* 2016ApJS..227...14A (Abbott, et al. 2016, GW150914 rates paper
  supplemental information): gives technical information not relevant
  to our argument, and therefore redundant with 2016ApJ...833L...1A
  (GW150914 rates paper).

* 2003MNRAS.341..385P (Podsiadlowski, Rappaport & Han 2003): redundant
  with 2015ApJ...800...17F (Fragos & McClintock 2015).

> We are now requiring authors to make a statement in the Methods
  section regarding code availability, under the subheading "Code
  availability". Authors can specify that "The code used to generate x
  is available at y", or say that "We have opted not to make the code
  available because ...".  This does not apply to standard data
  reduction software (which should be named), but it does apply to
  anything that you have modelled.

We now make such a statement in Methods Section 1.

> Your Figures 5-7 should be converted to Extended Data Figure 1, ED
  Figure 2 and ED Figure 3. They should be referred to in that way in
  the text.

Done.

Referee 1
=========

> The analysis presented in the manuscript is timely, well explained
  and convincing (but see a few minor points below). The authors use
  (estimates of) the measurement uncertainties published by LIGO for
  the three events to determine the probability of observing the
  measured parallel spin values from different underlying
  populations. Multiple spin magnitude distributions are considered,
  which is useful given our very limited information on the true
  distribution. In fact the authors are likely conservative with
  allowing negative spins with respect to the orbital axis for the
  aligned case.

[This statement is part of the "summary pargraphs" at the beginning of
the report.]  In fact, the referee is confused on this point; because
we *do not* allow negative spins, we obtain such distinguishing
power.  We have tried to increase the prevalence of the text pointing
this out, in light of the referee's confusion; the relevant paragraph
is found on page 10 of the main body.

> My chief concern with the manuscript is related to the novelty of
  the results. The analysis of the existing 3 (2 + 1 likely)
  detections is indeed performed here for the first time, and it is an
  interesting result. However, the calculation regarding the number of
  events needed for a 3-5 sigma constraint is similar to two previous
  works, both referenced in the manuscript. Stevenson, Berry and
  Mandel (1703.06873) carried out a very similar analysis to the one
  in the manuscript. They established that about 5 observations would
  be sufficient to distinguish between the two extreme cases of fully
  spin-aligned or fully isotropic spin distributions, the identical
  case considered in the present manuscript. While Stevenson et
  al. use simulations as opposed to relying on the reconstructed spin
  distributions by LIGO, and they consider a constant spin (0.7) as
  opposed to the distributions adopted in the present manuscript, the
  methods and conclusions are essentially the same.

It is true that some of our conclusions are the same as Stevenson et
al., but here (as the referee points out) we have for the first time
applied such an analysis to actual detections; also, we have explored
how the distinguishability between aligned and isotropic distributions
behaves when the spin = 0.7 (optimistic) assumption of Stevenson et al
is weakened.  Finally, we explore the limits of this method when spins
are intrinsically small (< 0.2), and discuss this possibility in the
context of field binary formation theory.  We feel this work, unlike
Stevenson et al, gives the modelling community a clear target: for
field fromation scenarios to work going forward, either kicks must be
extreme enough to render the distribution approximately isotropic *or*
spins must be typically < 0.2 to escape the constraints we have placed
from the existing observations.

> Regarding the above publications [Stevenson, et al. & Vitale et
  al.], while they are cited and briefly discussed in the manuscript,
  their their presentation is somewhat lacking. Mainly, the result of
  Vitale et al. is said to carry out a similar analysis, but concludes
  that several hundred sources would be needed (see lines 166-169). In
  reality, Vitale et al. focuses on measuring the precise
  contributions of different channels, e.g. 70% vs. 20%, as opposed to
  ruling out the extreme cases in which all binaries originate from
  the same channel (as is considered in the present paper). The
  several hundred measurements will be needed for such precision
  measurements, not for ruling out extreme cases.

We have re-organised the discussion of Vitale, et al. (now Ref. 17).
Rather than occupying its own paragraph, the different conclusions
from Vitale et al are addressed where appropriate in the text.  On
page 10, where we highlight that our "resolving power" comes from the
constraint on our aligned models that chi_eff > 0, we now note that
this constraint was not imposed on Vitale's "aligned" models, and
highlight that in our work we find much higher confidence in isotropy
after 10 detections as a result (>> 5 sigma versus ~3 sigma).  And
then, on page 17 (Methods Section), where we discuss our mixture model
and the possibility that there may be multiple channels in operation,
we note that Vitale+ found that it would take O(100) detections to pin
down the mixing fraction.

> One additional concern with the analysis is that it considers
  LVT151012 in the sample of observations along with the two 5sigma
  detections. I note that it is mostly standard in the literature to
  treat this candidate as a real event, on paper it has 10% chance of
  being not astrophysical. Taking this 10% false alarm into account
  may significantly reduce the 1.7sigma limit presented in the
  manuscript.

Taking the 10% probability of being "terrestrial" in origin into
account to a first approximation reduces the accumulated evidence from
LVT151012 by about 10%, so the odds ratio changes by a similar amount;
this is not a significant change.  Incorportaing a "background"
mixture model into our analysis would significantly complicate it,
without changing the final result appreciably.

> Another minor concern is that the template based searches used by
  LIGO effectively use priors for the underlying spin distributions,
  and for the searches finding the three events only spin-aligned
  templates were used. It would be useful for the reader to briefly
  comment on whether this can alter the observed spin distribution
  from the underlying distribution.

We now comment on this possibility in Methods Section 7.

Referee 2
==========

> The paper as written is too technical and jargon laden for Nature.
  The topic and result are, in my judgement, of interest to Nature
  readers, and I recommend consideration for publication with a
  significant rewrite of the abstract, introduction and conclusion to
  make it more acceptable to a general audience.

As you will have seen, we have re-written much of the introduction and
conclusion to make it less technical and/or jargon laden.

> The effective spin, χeff, appears in the abstract without context.
  This is not appropriate for a Nature paper, whose abstract is
  targeted for the general reader.

Removed.

> The abstract suggests binary black hole spins extend to high values
  - this statement is strictly limited to stellar mass black holes,
  and is derived for single black holes accreting from a main sequence
  star or giant star.

We still note in the abstract that our conclusions are dependent on
the BBH spins extending to high values, but more completely discuss
this caveat on page 10 and 11 (final pargraph of page 10, extending
onto 11).

> Section 1: the discussion of the effective spin needs to be expanded
  and made clearer for the general reader.  There is notational
  confusion with both "a" and "χ" used to refer to spin

We have moved the equation and discussion referred to here to Methods
Section 2, and now highlight the distinctions between "a" (spin
magnitude), "chi" (spin projection on z-axis), and "chi_eff"
(mass-weighted average of chis).

> It is really necessary, for a Nature paper, to have a very brief
  discussion of the spin parametrization and that there is a true
  maximum spin for a black hole.

We give this discussion and the associated limits now in Methods
Section 2.

> Approximating the posterior distribution as gaussian is strictly
  non-physical. I believe that a (truncated) gaussian is an adequate
  approximation to the distribution, but this requires some
  justification in text.

We have noted that our distributions are actually truncated in the
main text; since no distribution extends near +/- 1, truncation has no
effect on our results.

> The motivation for the choice of P(a) needs to be expanded a bit,
  the toy models used are not physically motivated and that needs to
  be made clear. I'd even suggest a figure (for the general reader) of
  P(a) for the three distributions, if there is room. Some mention of
  the more general "low" distribution in Appendix C needs to be made
  in the main text.

We state in the text (this was also present in the last submitted
version of the paper): "These magnitude distributions are not meant to
represent any particular physical model, but rather to capture our
uncertainty about the spin magnitude distribution; neither
observations nor population synthesis codes can at this point
authoritatively suggest any particular spin distribution.  Our models,
however, allow us to see how sensitive the chi_eff distribution is to
spin alignment given uncertainties about the spin magnitudes."  We
have added a figure (Extended Data Figure 1) and some discussion
(Methods Section 2, bottom of page 13) on the models.  We now mention
"Appendix C" (now Methods Section 5) in the text on page 10, when
discussing how our model selection would degrade if spin magnitudes
are small.

> I infer that the same P(a) is used for both members of the binary?
  For binaries formed in the field, the mass accretion history for
  each binary member is somewhat different in a systematic way and any
  spin due to accretion, as opposed to natal spin, is likely to be
  systematically different, to some extent, for each component. This
  may be inconsequential, but may be prudent to mention.

We now clearly state that p(a) is the same for both members in the
binary and that this may not be true in the field due to
systematically different evolutionary histories of the binary
members.  Page 4, final pargraph.

> There needs to a mention in the main text of the mass ratio
  discussion in Appendix D.

We now mention this in the final paragraph on Page 4, when discussing
the equal-mass assumption we use in the main text.  We have also added
a section dealing with the dependence of chi_eff uncertainty on mass
ratio (Methods Section 8), which is discussed Footnote 2.

> The spin-kick discussion is inadequate: there are discussion of the
  issue in Brandt et al. 1995, Sipior et al 2002, Fryer 2003, Willems
  et al 2005, as well as a number of more recent related papers.  In
  particular, IF there are kicks, they correlate with the probability
  of the binary remaining bound and its time to coalescence, and hence
  the post-kick spin alignment is correlated with probability of
  merger, possibly in a mass-dependent way. Arguably this is beyond
  the scope of the paper, but it ought to be mentioned.

We have re-written the alignment/kick discussion for field binaries to
be more clear and concise.  It now appears in the middle paragraph on
page 11.  We feel, however, that a detailed discussion of the effect
of large kicks on rates is outside the scope of this paper (if we talk
about these details, what about the effect of large kicks in dynamical
formation scenarios---they reduce the rates further here---and what
about all the other uncertainties in binary evolution and their effect
on rates/spin correlations, etc).

> The argument in the conclusion that this paper agrees with Stevenson
  et al 2017, as opposed to Vitale et al, is somewhat misleading, as
  there is substantial overlap in authorship and methodology between
  Stevenson et al and this paper. The consistency of these results
  should not be implied to constitute independent verification of
  them.

We have eliminated this paragraph, and do not make claims about
independent verification in this current draft.

### 2017-06-05 paper/AlignedIsotropicGW.tex
As re-submitted.

@@
   detections\cite{2016ApJ...833L...1A}, we show that such an augmented
-  data set would enable in most cases a prefernce stronger than
+  data set would enable in most cases a preference stronger than
   $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
@@
 observations\cite{O1-BBH,PhysRevLett.118.221101}, and the $\chieff$
-distribution is not particularly sensitive to the mass ratio between
-the merging objects---see Methods Section \ref{sec:mass-ratio}).  We
-assume that the population spin distribution factorises into a
-distribution for the spin magnitude $a$ and a distribution for the
-spin angles.  Finally, we assume that the distribution of spins is
-common to each component in a merging binary (in reality the
-distributions of spin for each component in the binary may differ
-systematically due to different formation histories).  Choosing one of
-three magnitude distributions (see Methods Section
+distribution is not sensitive to the mass ratio---see Methods Section
+\ref{sec:mass-ratio}).  We assume that the population spin
+distribution factorises into a distribution for the spin magnitude $a$
+and a distribution for the spin angles.  Finally, we assume that the
+distribution of spins is common to each component in a merging binary
+(the distributions of spin for each component in the binary could
+differ systematically due to different formation histories).  Choosing
+one of three magnitude distributions (see Methods Section
 \ref{methsec:chieff-spin-magnitude}), ``low'' (mean $a = 0.33$,
@@
 deviation $0.29$), ``high'' (mean $a = 0.67$, standard deviation
-$0.24$) and pairing with either an isotropic angular distribution or a
-distribution that generates perfect alignment with the positive $z$
-axis yields six different models for the $\chieff$ distribution.
-These models are shown in Figure \ref{fig:chieff-distribution-models}.
+$0.24$) and pairing with an isotropic angular distribution or a
+distribution that generates perfect alignment yields six different
+models for the $\chieff$ distribution.  These models are shown in
+Figure \ref{fig:chieff-distribution-models}.
 
@@
 particular physical model, but rather to capture our uncertainty about
-the spin magnitude distribution, discussed in detail below; neither
-observations nor population synthesis codes can at this point
-authoritatively suggest \emph{any} particular spin
-distribution\cite{2015PhR...548....1M}.  Our models, however, allow us
-to see how sensitive the $\chieff$ distribution is to spin alignment
-given uncertainties about the spin magnitudes.
+the spin magnitude distribution; neither observations nor population
+synthesis codes can at this point authoritatively suggest \emph{any}
+particular spin distribution\cite{2015PhR...548....1M}.  Our models,
+however, allow us to see how sensitive the $\chieff$ distribution is
+to spin alignment given uncertainties about the spin magnitudes.
 
@@
 distributions (see Methods Section \ref{sec:hierarchical}).  We also
-fit three mixture models for the population, where the spin magnitude
-distribution is fixed but the angular distribution is a weighted sum
-of the isotropic and aligned distribution.  The evidence, or marginal
-likelihood, for each of the models is shown in Figure
-\ref{fig:O1-odds}.  For all three magnitude distributions, the mixture
-models' posterior on the mixing fraction peaks at 100\% isotropic,
-which explains why the zero-parameter, pure-isotropic models are
-preferred over the single-parameter mixture models for every magnitude
-distribution with this data set.  Not surprisingly, given the small
-$\chieff$ values in the three detected systems, the most-favoured
-model among those with an isotropic angular distribution has the
-``low'' magnitude distribution; the most favoured model among those
-with an aligned distribution also has the ``low'' magnitude
-distribution.  The odds ratio between the ``low'' aligned and ``low''
-isotropic models is $\OOneOddsIsoAligned$, or
-$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
-among our suite of models.  While the data favour spin amplitude
-distributions with small spin magnitudes, note that a model with all
-binary black hole systems having zero spin is ruled out by the
-GW151226 measurements, which bound at least one black hole to have
-spin magnitude $\geq 0.2$ at 99\%
+fit three mixture models for the population, where the angular
+distribution is a weighted sum of the isotropic and aligned
+distributions.  The evidence, or marginal likelihood, for each of the
+models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
+distributions, the mixture models' posterior on the mixing fraction
+peaks at 100\% isotropic.  Not surprisingly, given the small $\chieff$
+values in the three detected systems, the most-favoured model among
+those with an isotropic angular distribution has the ``low'' magnitude
+distribution; the most favoured model among those with an aligned
+distribution also has the ``low'' magnitude distribution.  The odds
+ratio between the ``low'' aligned and ``low'' isotropic models is
+$\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$; thus the data
+favour isotropic spins among our suite of models.  While the data
+favour spin amplitude distributions with small spin magnitudes, note
+that a model with all binary black hole systems having zero spin is
+ruled out by the GW151226 measurements, which bound at least one black
+hole to have spin magnitude $\geq 0.2$ at 99\%
 credibility\cite{2016PhRvL.116x1103A}.
@@
 additional detections drawn from each of our six zero-parameter models
-for the spin distribution in addition to the four existing detections
-from O1 and GW170104, with observational uncertainties drawn randomly
-from the three Gaussian widths used to approximate the $\chieff$
-posteriors in Figure \ref{fig:O1-posteriors}\footnote{The measurement
-  uncertainty in $\chieff$ depends on the other parameters of the
-  merging binary black hole system, particularly on the mass ratio.
-  Our assumption about future observational uncertainties is
-  appropriate if the parameters of the three detected events are
-  representative of the parameters of future detections. See Methods
-  Section \ref{sec:chi-eff-precision} for further discussion.}, we
-find the odds ratios shown in Figure \ref{fig:O2-predictions}.  We
-find that most scenarios with an additional 10 detections allow the
-simulated angular distribution to be inferred with greater than
-$5\sigma$ ($2.9 \times 10^{-7}$ odds) credibility; and in the most
-pessimistic case the distinction is typically
-$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
-While such future detections should permit a confident distinction
-between \emph{angular} distributions, we would remain much less
-certain about the \emph{magnitude} distribution among the three
-options considered here until we have a larger number of observations.
-In Figure \ref{fig:O2-predictions}, the odds ratio between different
-magnitude distributions with the same angular distribution is much
-closer to unity than the odds ratio between angular distributions.
+in addition to the four existing detections from O1 and GW170104, with
+observational uncertainties drawn randomly from the three Gaussian
+widths used to approximate the $\chieff$ posteriors in Figure
+\ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
+  $\chieff$ depends on the other parameters of the merging binary
+  black hole system, particularly on the mass ratio.  Our assumption
+  about future observational uncertainties is appropriate if the
+  parameters of the three detected events are representative of the
+  parameters of future detections. See Methods Section
+  \ref{sec:chi-eff-precision} for further discussion.}, we find the
+odds ratios shown in Figure \ref{fig:O2-predictions}.  We find that
+most scenarios with an additional 10 detections allow the simulated
+angular distribution to be inferred with greater than $5\sigma$
+($2.9 \times 10^{-7}$ odds) credibility. In the most pessimistic case
+the distinction is typically $\OTwoSigmaIsoAlignedMin\sigma$
+($\OTwoOddsIsoAlignedMin$ odds ratio).  While such future detections
+should permit a confident distinction between \emph{angular}
+distributions, we would remain much less certain about the
+\emph{magnitude} distribution among the three options considered here
+until we have a larger number of observations.
 
@@
 (leading to the possibility of positive \emph{or} negative $\chieff$)
-can only be distinguised from an isotropic true population at
+can only be distinguished from an isotropic true population at
 $\sim 3 \sigma$ after 10--20 observations\cite{2017CQGra..34cLT01V};
@@
 better than $5\sigma$ (odds $< 10^{-8}$) after 10 observations,
-emphasising the information content of the bound $\chieff > 0$ for our
+emphasizing the information content of the bound $\chieff > 0$ for our
 aligned models.

### 2017-06-07 paper/AlignedIsotropicGW.tex
Rename style file so arXiv doesn't try to use its own nature.cls which has no figures.
Moved plots file for arXiv submission.
changed rounding of number in table

@@
 
-\documentclass{nature}
+\documentclass{nature-arxiv}
 
@@
 All O1 events & 1.7 & 2.7 & 4.4 \\ %\hline
-All O1 events and GW170104 & \textbf{2.4} & 3.6 & 5.4 \\ \hline
+All O1 events and GW170104 & \textbf{2.4} & 3.5 & 5.4 \\ \hline
 \end{tabular}
@@
 
-\bibliographystyle{naturemag}
+\bibliographystyle{naturemag-arxiv}
 \bibliography{AlignedIsotropicGW}

### 2017-06-29 paper/AlignedIsotropicGW.tex
Removed footnote 1, incorporated others into text.

@@
 
-\documentclass{nature-arxiv}
+\documentclass{nature}
 
@@
   values we show that the data exhibit a $\OOneSigmaIsoAligned\sigma$
-  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
-    $r \ll 1$ is equivalent to $x \sigma$ with
-    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
-    normal CDF.})  preference for an isotropic angular distribution
-  over an aligned one.  By considering the effect of 10 additional
-  detections\cite{2016ApJ...833L...1A}, we show that such an augmented
-  data set would enable in most cases a preference stronger than
-  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
-  preference for either an isotropic spin distribution or low spin
-  magnitudes for the observed systems will be confirmed (or
+  ($\OOneOddsIsoAligned$ odds ratio) preference for an isotropic
+  angular distribution over an aligned one.  By considering the effect
+  of 10 additional detections\cite{2016ApJ...833L...1A}, we show that
+  such an augmented data set would enable in most cases a preference
+  stronger than $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The
+  existing preference for either an isotropic spin distribution or low
+  spin magnitudes for the observed systems will be confirmed (or
   overturned) confidently in the near future.
@@
 widths used to approximate the $\chieff$ posteriors in Figure
-\ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
-  $\chieff$ depends on the other parameters of the merging binary
-  black hole system, particularly on the mass ratio.  Our assumption
-  about future observational uncertainties is appropriate if the
-  parameters of the three detected events are representative of the
-  parameters of future detections. See Methods Section
-  \ref{sec:chi-eff-precision} for further discussion.}, we find the
-odds ratios shown in Figure \ref{fig:O2-predictions}.  We find that
-most scenarios with an additional 10 detections allow the simulated
-angular distribution to be inferred with greater than $5\sigma$
-($2.9 \times 10^{-7}$ odds) credibility. In the most pessimistic case
-the distinction is typically $\OTwoSigmaIsoAlignedMin\sigma$
-($\OTwoOddsIsoAlignedMin$ odds ratio).  While such future detections
-should permit a confident distinction between \emph{angular}
-distributions, we would remain much less certain about the
-\emph{magnitude} distribution among the three options considered here
-until we have a larger number of observations.
+\ref{fig:O1-posteriors}, we find the odds ratios shown in Figure
+\ref{fig:O2-predictions}.  The measurement uncertainty in $\chieff$
+depends on the other parameters of the merging binary black hole
+system, particularly on the mass ratio.  Our assumption about future
+observational uncertainties is appropriate if the parameters of the
+three detected events are representative of the parameters of future
+detections. See Methods Section \ref{sec:chi-eff-precision} for
+further discussion.  We find that most scenarios with an additional 10
+detections allow the simulated angular distribution to be inferred
+with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
+credibility. In the most pessimistic case the distinction is typically
+$\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds ratio).
+While such future detections should permit a confident distinction
+between \emph{angular} distributions, we would remain much less
+certain about the \emph{magnitude} distribution among the three
+options considered here until we have a larger number of observations.
 
@@
 
-For this figure a detected population\footnote{We qualify a system as
-  ``detected'' if it produces a SNR above $8$ in the second-loudest
-  detector to select only coincident events.} of 500 binaries was
-selected from a population with component masses distributed uniformly
-between $1$ and $30~\mathrm{M}_\odot$ with
-$m_1 + m_2 < 30~\mathrm{M}_\odot$, locations distributed uniformly in
-volume, and orientations distributed isotropically.  Data were
-simulated for each binary, and posteriors were estimated using the
-LIGO-Virgo parameter estimation library
-\texttt{LALInference}\cite{2015PhRvD..91d2003V} using inspiral-only
-waveform models (merger and ringdown effects can provide additional
-information for some binaries, but we ignore those effects here).
-$\chieff$ is better constrained for binaries with high
+For this figure a detected population of 500 binaries was selected
+from a population with component masses distributed uniformly between
+$1$ and $30~\mathrm{M}_\odot$ with $m_1 + m_2 < 30~\mathrm{M}_\odot$,
+locations distributed uniformly in volume, and orientations
+distributed isotropically.  We qualify a system as ``detected'' if it
+produces a SNR above $8$ in the second-loudest detector to select only
+coincident events.  Data were simulated for each binary, and
+posteriors were estimated using the LIGO-Virgo parameter estimation
+library \texttt{LALInference}\cite{2015PhRvD..91d2003V} using
+inspiral-only waveform models (merger and ringdown effects can provide
+additional information for some binaries, but we ignore those effects
+here).  $\chieff$ is better constrained for binaries with high
 effective spins and high ($\sim$equal) mass ratios.

### 2017-06-29 paper/AlignedIsotropicGW.tex
Moved bibliographies into text, split between main and methods sections.

@@
 \newcommand{\prd}{Phys. Rev. D}
+
+\newcounter{mainbib}
 
@@
 
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{2016PhRvL.116f1102A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Observation of Gravitational Waves from a Binary
+  Black Hole Merger}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{061102}
+  (\bibinfo{year}{2016}).
+
+\bibitem{2016PhRvL.116x1103A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{GW151226: Observation of Gravitational Waves from a
+  22-Solar-Mass Binary Black Hole Coalescence}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241103}
+  (\bibinfo{year}{2016}).
+
+\bibitem{O1-BBH}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Binary Black Hole Mergers in the First Advanced LIGO
+  Observing Run}}.
+\newblock \emph{\bibinfo{journal}{Physical Review X}}
+  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{041015} (\bibinfo{year}{2016}).
+
+\bibitem{PhysRevLett.118.221101}
+\bibinfo{author}{Abbott, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{Gw170104: Observation of a 50-solar-mass binary black
+  hole coalescence at redshift 0.2}.
+\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
+  \textbf{\bibinfo{volume}{118}}, \bibinfo{pages}{221101}
+  (\bibinfo{year}{2017}).
+\newblock
+  \urlprefix\url{https://link.aps.org/doi/10.1103/PhysRevLett.118.221101}.
+
+\bibitem{SigurdssonHernquist:1993}
+\bibinfo{author}{{Sigurdsson}, S.} \& \bibinfo{author}{{Hernquist}, L.}
+\newblock \bibinfo{title}{{Primordial black holes in globular clusters}}.
+\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{364}},
+  \bibinfo{pages}{423--425} (\bibinfo{year}{1993}).
+
+\bibitem{PZMcMillan:2000}
+\bibinfo{author}{{Portegies Zwart}, S.~F.} \& \bibinfo{author}{McMillan, S.
+  L.~W.}
+\newblock \bibinfo{title}{Black hole mergers in the universe}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{528}},
+  \bibinfo{pages}{L17} (\bibinfo{year}{2000}).
+
+\bibitem{Rodriguez:2015}
+\bibinfo{author}{{Rodriguez}, C.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{{Binary Black Hole Mergers from Globular Clusters:
+  Implications for Advanced LIGO}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{051101}
+  (\bibinfo{year}{2015}).
+
+\bibitem{Stone:2016}
+\bibinfo{author}{{Stone}, N.~C.}, \bibinfo{author}{{Metzger}, B.~D.} \&
+  \bibinfo{author}{{Haiman}, Z.}
+\newblock \bibinfo{title}{{Assisted inspirals of stellar mass black holes
+  embedded in AGN discs: solving the `final au problem'}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{464}},
+  \bibinfo{pages}{946--954} (\bibinfo{year}{2017}).
+
+\bibitem{2016ApJ...832L...2R}
+\bibinfo{author}{{Rodriguez}, C.~L.}, \bibinfo{author}{{Zevin}, M.},
+  \bibinfo{author}{{Pankow}, C.}, \bibinfo{author}{{Kalogera}, V.} \&
+  \bibinfo{author}{{Rasio}, F.~A.}
+\newblock \bibinfo{title}{{Illuminating Black Hole Binary Formation Channels
+  with Spins in Advanced LIGO}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{832}},
+  \bibinfo{pages}{L2} (\bibinfo{year}{2016}).
+
+\bibitem{TutukovYungelson:1993}
+\bibinfo{author}{Tutukov, A.~V.} \& \bibinfo{author}{Yungelson, L.~R.}
+\newblock \bibinfo{title}{The merger rate of neutron star and black hole
+  binaries}.
+\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
+  Society}} \textbf{\bibinfo{volume}{260}}, \bibinfo{pages}{675--678}
+  (\bibinfo{year}{1993}).
+
+\bibitem{2016Natur.534..512B}
+\bibinfo{author}{{Belczynski}, K.}, \bibinfo{author}{{Holz}, D.~E.},
+  \bibinfo{author}{{Bulik}, T.} \& \bibinfo{author}{{O'Shaughnessy}, R.}
+\newblock \bibinfo{title}{{The first gravitational-wave source from the
+  isolated evolution of two stars in the 40-100 solar mass range}}.
+\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{534}},
+  \bibinfo{pages}{512--515} (\bibinfo{year}{2016}).
+
+\bibitem{Stevenson:2017}
+\bibinfo{author}{{Stevenson}, S.} \emph{et~al.}
+\newblock \bibinfo{title}{{Formation of the first three gravitational-wave
+  observations through isolated binary evolution}}.
+\newblock \emph{\bibinfo{journal}{Nat. Commun.}} \textbf{\bibinfo{volume}{8}},
+  \bibinfo{pages}{14906} (\bibinfo{year}{2017}).
+
+\bibitem{MandeldeMink:2016}
+\bibinfo{author}{{Mandel}, I.} \& \bibinfo{author}{{de Mink}, S.~E.}
+\newblock \bibinfo{title}{{Merging binary black holes formed through chemically
+  homogeneous evolution in short-period stellar binaries}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{458}},
+  \bibinfo{pages}{2634--2647} (\bibinfo{year}{2016}).
+
+\bibitem{Marchant:2016}
+\bibinfo{author}{{Marchant}, P.}, \bibinfo{author}{{Langer}, N.},
+  \bibinfo{author}{{Podsiadlowski}, P.}, \bibinfo{author}{{Tauris}, T.~M.} \&
+  \bibinfo{author}{{Moriya}, T.~J.}
+\newblock \bibinfo{title}{{A new route towards merging massive black holes}}.
+\newblock \emph{\bibinfo{journal}{\aap}} \textbf{\bibinfo{volume}{588}},
+  \bibinfo{pages}{A50} (\bibinfo{year}{2016}).
+
+\bibitem{2016ApJ...833L...1A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{The Rate of Binary Black Hole Mergers Inferred from
+  Advanced LIGO Observations Surrounding GW150914}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{833}},
+  \bibinfo{pages}{L1} (\bibinfo{year}{2016}).
+
+\bibitem{2015PhR...548....1M}
+\bibinfo{author}{{Miller}, M.~C.} \& \bibinfo{author}{{Miller}, J.~M.}
+\newblock \bibinfo{title}{{The masses and spins of neutron stars and
+  stellar-mass black holes}}.
+\newblock \emph{\bibinfo{journal}{\physrep}} \textbf{\bibinfo{volume}{548}},
+  \bibinfo{pages}{1--34} (\bibinfo{year}{2015}).
+
+\bibitem{2017CQGra..34cLT01V}
+\bibinfo{author}{{Vitale}, S.}, \bibinfo{author}{{Lynch}, R.},
+  \bibinfo{author}{{Sturani}, R.} \& \bibinfo{author}{{Graff}, P.}
+\newblock \bibinfo{title}{{Use of gravitational waves to probe the formation
+  channels of compact binaries}}.
+\newblock \emph{\bibinfo{journal}{Classical and Quantum Gravity}}
+  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{03LT01}
+  (\bibinfo{year}{2017}).
+
+\bibitem{1999MNRAS.305..654K}
+\bibinfo{author}{{King}, A.~R.} \& \bibinfo{author}{{Kolb}, U.}
+\newblock \bibinfo{title}{{The evolution of black hole mass and angular
+  momentum}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{305}},
+  \bibinfo{pages}{654--660} (\bibinfo{year}{1999}).
+
+\bibitem{2015ApJ...800...17F}
+\bibinfo{author}{{Fragos}, T.} \& \bibinfo{author}{{McClintock}, J.~E.}
+\newblock \bibinfo{title}{{The Origin of Black Hole Spin in Galactic Low-mass
+  X-Ray Binaries}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{800}},
+  \bibinfo{pages}{17} (\bibinfo{year}{2015}).
+
+\bibitem{2016MNRAS.462..844K}
+\bibinfo{author}{{Kushnir}, D.}, \bibinfo{author}{{Zaldarriaga}, M.},
+  \bibinfo{author}{{Kollmeier}, J.~A.} \& \bibinfo{author}{{Waldman}, R.}
+\newblock \bibinfo{title}{{GW150914: spin-based constraints on the merger time
+  of the progenitor system}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{462}},
+  \bibinfo{pages}{844--849} (\bibinfo{year}{2016}).
+
+\bibitem{2017arXiv170200885Z}
+\bibinfo{author}{{Zaldarriaga}, M.}, \bibinfo{author}{{Kushnir}, D.} \&
+  \bibinfo{author}{{Kollmeier}, J.~A.}
+\newblock \bibinfo{title}{{Research Note: The Expected Spins of Gravitational
+  Wave Sources With Isolated Field Binary Progenitors}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1702.00885}.
+
+\bibitem{2017arXiv170203952H}
+\bibinfo{author}{{Hotokezaka}, K.} \& \bibinfo{author}{{Piran}, T.}
+\newblock \bibinfo{title}{{Implications of the low binary black hole spins
+  observed by LIGO}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1702.03952}.
+
+\bibitem{Albrecht:2009}
+\bibinfo{author}{{Albrecht}, S.}, \bibinfo{author}{{Reffert}, S.},
+  \bibinfo{author}{{Snellen}, I.~A.~G.} \& \bibinfo{author}{{Winn}, J.~N.}
+\newblock \bibinfo{title}{{Misaligned spin and orbital axes cause the anomalous
+  precession of DI Herculis}}.
+\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{461}},
+  \bibinfo{pages}{373--376} (\bibinfo{year}{2009}).
+
+\bibitem{Martin:2008b}
+\bibinfo{author}{{Martin}, R.~G.}, \bibinfo{author}{{Reis}, R.~C.} \&
+  \bibinfo{author}{{Pringle}, J.~E.}
+\newblock \bibinfo{title}{{Misalignment of the microquasar V4641 Sgr (SAX
+  J1819.3-2525)}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{391}},
+  \bibinfo{pages}{L15--L18} (\bibinfo{year}{2008}).
+
+\bibitem{MorningstarMiller:2014}
+\bibinfo{author}{{Morningstar}, W.~R.} \& \bibinfo{author}{{Miller}, J.~M.}
+\newblock \bibinfo{title}{{The Spin of the Black Hole 4U 1543-47}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{793}},
+  \bibinfo{pages}{L33} (\bibinfo{year}{2014}).
+
+\bibitem{2000ApJ...541..319K}
+\bibinfo{author}{{Kalogera}, V.}
+\newblock \bibinfo{title}{{Spin-Orbit Misalignment in Close Binaries with Two
+  Compact Objects}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{541}},
+  \bibinfo{pages}{319--328} (\bibinfo{year}{2000}).
+
+\bibitem{2011ApJ...742...81F}
+\bibinfo{author}{{Farr}, W.~M.}, \bibinfo{author}{{Kremer}, K.},
+  \bibinfo{author}{{Lyutikov}, M.} \& \bibinfo{author}{{Kalogera}, V.}
+\newblock \bibinfo{title}{{Spin Tilts in the Double Pulsar Reveal Supernova
+  Spin Angular-momentum Production}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{742}},
+  \bibinfo{pages}{81} (\bibinfo{year}{2011}).
+
+\bibitem{2015PhRvD..92f4016G}
+\bibinfo{author}{{Gerosa}, D.}, \bibinfo{author}{{Kesden}, M.},
+  \bibinfo{author}{{Sperhake}, U.}, \bibinfo{author}{{Berti}, E.} \&
+  \bibinfo{author}{{O'Shaughnessy}, R.}
+\newblock \bibinfo{title}{{Multi-timescale analysis of phase transitions in
+  precessing black-hole binaries}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{92}},
+  \bibinfo{pages}{064016} (\bibinfo{year}{2015}).
+
+\bibitem{2004PhRvD..70l4020S}
+\bibinfo{author}{{Schnittman}, J.~D.}
+\newblock \bibinfo{title}{{Spin-orbit resonance and the evolution of compact
+  binary systems}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{70}},
+  \bibinfo{pages}{124020} (\bibinfo{year}{2004}).
+
+\bibitem{2007ApJ...661L.147B}
+\bibinfo{author}{{Bogdanovi{\'c}}, T.}, \bibinfo{author}{{Reynolds}, C.~S.} \&
+  \bibinfo{author}{{Miller}, M.~C.}
+\newblock \bibinfo{title}{{Alignment of the Spins of Supermassive Black Holes
+  Prior to Coalescence}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{661}},
+  \bibinfo{pages}{L147--L150} (\bibinfo{year}{2007}).
+
+\setcounter{mainbib}{\value{enumiv}}
+\end{thebibliography}
+
 \begin{addendum}
@@
   No. NSF PHY11-25915.
-\item[Supplementary Information] Supplementary Information is linked
-  to the online version of the paper at
-  \url{http://www.nature.com/nature}.
 \item[Author Contributions] All authors contributed to the work
@@
 
-\bibliographystyle{naturemag-arxiv}
-\bibliography{AlignedIsotropicGW}
+\begin{thebibliography}{10}
+\setcounter{enumiv}{\value{mainbib}}
+
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{Julia}
+\bibinfo{author}{Bezanson, J.}, \bibinfo{author}{Karpinski, S.},
+  \bibinfo{author}{Shah, V.~B.} \& \bibinfo{author}{Edelman, A.}
+\newblock \bibinfo{title}{{J}ulia: A fast dynamic language for technical
+  computing}  (\bibinfo{year}{2012}).
+\newblock \eprint{arXiv:1209.5145}.
+
+\bibitem{NumPy}
+\bibinfo{author}{van~der Walt, S.}, \bibinfo{author}{Colbert, S.~C.} \&
+  \bibinfo{author}{Varoquaux, G.}
+\newblock \bibinfo{title}{The numpy array: A structure for efficient numerical
+  computation}.
+\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{22--30}
+  (\bibinfo{year}{2011}).
+
+\bibitem{SciPy}
+\bibinfo{author}{Jones, E.}, \bibinfo{author}{Oliphant, T.},
+  \bibinfo{author}{Peterson, P.} \emph{et~al.}
+\newblock \bibinfo{title}{{SciPy}: Open source scientific tools for {Python}}
+  (\bibinfo{year}{2001--}).
+\newblock \urlprefix\url{http://www.scipy.org/}.
+\newblock \bibinfo{note}{[Online; accessed 24 May 2017]}.
+
+\bibitem{Matplotlib}
+\bibinfo{author}{Hunter, J.~D.}
+\newblock \bibinfo{title}{Matplotlib: A 2d graphics environment}.
+\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
+  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{90--95} (\bibinfo{year}{2007}).
+
+\bibitem{IPython}
+\bibinfo{author}{Pérez, F.} \& \bibinfo{author}{Granger, B.~E.}
+\newblock \bibinfo{title}{Ipython: A system for interactive scientific
+  computing}.
+\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
+  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{21--29} (\bibinfo{year}{2007}).
+\newblock
+  \urlprefix\url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2007.53}.
+
+\bibitem{2016PhRvL.116x1102A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Properties of the Binary Black Hole Merger
+  GW150914}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241102}
+  (\bibinfo{year}{2016}).
+
+\bibitem{Stevenson:2017spin}
+\bibinfo{author}{{Stevenson}, S.}, \bibinfo{author}{{Berry}, C.~P.~L.} \&
+  \bibinfo{author}{{Mandel}, I.}
+\newblock \bibinfo{title}{{Hierarchical analysis of gravitational-wave
+  measurements of binary black hole spin-orbit misalignments}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{1703.06873}.
+
+\bibitem{2015PhRvD..91d2003V}
+\bibinfo{author}{{Veitch}, J.} \emph{et~al.}
+\newblock \bibinfo{title}{{Parameter estimation for compact binaries with
+  ground-based gravitational-wave observations using the LALInference software
+  library}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{91}},
+  \bibinfo{pages}{042003} (\bibinfo{year}{2015}).
+
+\bibitem{2010ApJ...725.2166H}
+\bibinfo{author}{{Hogg}, D.~W.}, \bibinfo{author}{{Myers}, A.~D.} \&
+  \bibinfo{author}{{Bovy}, J.}
+\newblock \bibinfo{title}{{Inferring the Eccentricity Distribution}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{725}},
+  \bibinfo{pages}{2166--2175} (\bibinfo{year}{2010}).
+
+\bibitem{2010PhRvD..81h4029M}
+\bibinfo{author}{{Mandel}, I.}
+\newblock \bibinfo{title}{{Parameter estimation on gravitational waves from
+  multiple coalescing binaries}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{81}},
+  \bibinfo{pages}{084029} (\bibinfo{year}{2010}).
+
+\bibitem{2017arXiv170306869F}
+\bibinfo{author}{{Fishbach}, M.}, \bibinfo{author}{{Holz}, D.} \&
+  \bibinfo{author}{{Farr}, B.}
+\newblock \bibinfo{title}{{Are LIGO's Black Holes Made From Smaller Black
+  Holes?}}
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1703.06869}.
+
+\bibitem{2017arXiv170306223G}
+\bibinfo{author}{{Gerosa}, D.} \& \bibinfo{author}{{Berti}, E.}
+\newblock \bibinfo{title}{{First or second generation? Black hole census with
+  gravitational wave observations}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1703.06223}.
+
+\bibitem{2016PhRvD..93l2003A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{GW150914: First results from the search for binary
+  black hole coalescence with Advanced LIGO}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{93}},
+  \bibinfo{pages}{122003} (\bibinfo{year}{2016}).
+
+\bibitem{2016arXiv161107531T}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Effects of waveform model systematics on the
+  interpretation of GW150914}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2016}).
+\newblock \eprint{arXiv:1611.07531}.
+
+\bibitem{2016PhRvD..94f4035A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Directly comparing GW150914 with numerical solutions
+  of Einstein's equations for binary black hole coalescence}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{94}},
+  \bibinfo{pages}{064035} (\bibinfo{year}{2016}).
+
+\end{thebibliography}
+
+%\bibliographystyle{naturemag-arxiv}
+%\bibliography{AlignedIsotropicGW}
 

### 2017-06-29 paper/AlignedIsotropicGW.tex
Moved code availability statement to end of methods and added data availability.
Moved around captions and table.

@@
 
-\begin{figure}
-  \plotone{../plots/chi-eff-mock-posteriors}
-  \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
-      $\chieff$ from the Advanced LIGO O1 and GW170104
-      observations\cite{O1-BBH,PhysRevLett.118.221101}.}  We
-    approximate the posteriors reported using Gaussians with the same
-    median and 90\% credible interval.  It is notable that none of the
-    $\chieff$ posteriors support high \ac{BH} spin magnitudes with
-    aligned spins, suggested by observations of stellar-mass black
-    holes in X-ray binaries\cite{2015PhR...548....1M}.}
-\end{figure}
-
 Small values of $\chieff$ as exhibited in these systems can result
@@
 Figure \ref{fig:chieff-distribution-models}.
-
-\begin{figure}
-  \plotone{../plots/chi-eff-distributions}
-  \caption{\label{fig:chieff-distribution-models} \textbf{The models for the
-    population distribution of $\chieff$ considered in this paper.}  In
-    all models we assume that the binary mass ratio
-    $q \equiv m_1/m_2 = 1$ and that the distribution of spin vectors
-    is the same for each component.  The ``flat'' (blue lines),
-    ``high,'' (green lines), and ``low'' (red lines) magnitude
-    distributions are defined in Eq.\ \eqref{eq:magnitude-dists}.
-    Solid lines give the $\chieff$ distribution under the assumption
-    that the orientations of the spins are isotropic; dashed lines
-    give the distribution under the assumption that both objects'
-    spins are aligned with the orbital angular momentum.  The
-    isotropic distributions are readily distinguished from the aligned
-    distributions by the production of negative $\chieff$ values,
-    while the distinction between the three models for the spin
-    magnitude distribution is less sharp.}
-\end{figure}
 
@@
 credibility\cite{2016PhRvL.116x1103A}.
-
-\begin{figure}
-  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models}
-  \caption{\textbf{Odds ratios among our models using the
-      approximations to the posteriors on $\chieff$ from the O1 and GW170104
-      observations shown in Figure \ref{fig:O1-posteriors}.}  The flat
-    (``F''), high (``H''), and low (``L'') spin magnitude
-    distributions (see Eq.\ \eqref{eq:magnitude-dists}) are paired
-    with isotropic (``I'') and aligned (``A'') angular distributions,
-    as well as a mixture model of the two (``$\mathrm{M}$'').  The
-    most-favoured models have the ``low'' distribution of spin
-    magnitudes.  The odds ratio between the best aligned and best
-    isotropic models is $\OOneOddsIsoAligned$, or
-    $\OOneSigmaIsoAligned\sigma$.  For all magnitude distributions the
-    pure-isotropic models are preferred over the mixture models;
-    correspondingly, the posterior on the mixture fraction peaks at
-    100\% isotropic.}
-  \label{fig:O1-odds}
-\end{figure}
 
@@
 
+Most of our resolving power for the spin angular distribution is a
+result of the fact that our ``aligned'' models cannot produce
+$\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
+spins are intrinsically very small, with $a \lesssim 0.2$, then it is
+no longer possible to resolve the negative effective spin with a small
+number of observations (see Methods Section \ref{sec:smallspins}).  As
+noted below, however, spins observed in X-ray binaries are typically
+large.  Additionally, models which do not permit \emph{some} spins
+with $\chieff \gtrsim 0.1$ are ruled out by the GW151226
+observations\cite{2016PhRvL.116x1103A}.  An ``aligned'' model with
+spin magnitudes from our ``flat'' distribution but permitting spin
+vectors oriented anti-parallel to the orbital angular momentum
+(leading to the possibility of positive \emph{or} negative $\chieff$)
+can only be distinguished from an isotropic true population at
+$\sim 3 \sigma$ after 10--20 observations\cite{2017CQGra..34cLT01V};
+our flat aligned model can be distinguished from such a population at
+better than $5\sigma$ (odds $< 10^{-8}$) after 10 observations,
+emphasizing the information content of the bound $\chieff > 0$ for our
+aligned models.
+
+Observational data on spin magnitudes in black hole systems is
+sparse\cite{2015PhR...548....1M}.  Most of the systems studied are
+low-mass X-ray binaries rather than the high-mass X-ray binaries that
+are likely to be the progenitors of double black hole binaries.  In
+addition, there are substantial systematic errors that can complicate
+these analyses\cite{2015PhR...548....1M} and selection effects could
+yield a biased distribution. Nonetheless, if we take the reported spin
+magnitudes as representative then we find that there is a preference
+for high spins; for example, 14 of the 19 systems with reported spins
+have dimensionless spin parameters in excess of 0.5.  It is usually
+argued that the masses and spin parameters of stellar-mass black holes
+are unlikely to be altered significantly by
+accretion\cite{1999MNRAS.305..654K}, but this may not be true for all
+systems\cite{2015ApJ...800...17F}.  Thus the current spin parameters
+are probably close to their values upon core collapse, at least in
+high-mass X-ray binaries.  However, the specific processes involved in
+the production of black hole binaries from isolated binaries could
+alter the spin magnitude distribution of those holes relative to the
+X-ray binary systems; for example, close tidal interactions could spin
+up the core, or stripping of the envelope could reduce the available
+angular
+momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
+
+The spin directions in isolated binary black
+holes\cite{TutukovYungelson:1993,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
+are usually expected to be preferentially aligned.  Despite observed
+spin-orbit misalignments in massive stellar
+binaries\cite{Albrecht:2009}, mass transfer and tidal interactions
+will tend to realign the binary.  On the other hand, there is some
+evidence of spin-orbit misalignment in black hole X-ray
+binaries\cite{Martin:2008b,MorningstarMiller:2014}.  This is
+consistent with the expectation that a supernova natal kick (if any)
+can change the orbital plane and misalign the
+binary\cite{2000ApJ...541..319K}; the supernova can also tilt the spin
+angle\cite{2011ApJ...742...81F}.  Evolutionary processes, such as
+wind-driven mass loss and post-collapse fallback, can couple the spin
+magnitude and direction distributions, contrary to our simplified
+assumptions.  A small misalignment at wide separation can also evolve
+to a more significant misalignment in component spins as the binary
+spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
+$\chieff$ is approximately conserved through this evolution.
+
+The spin directions of binary black holes formed dynamically through
+interactions in dense stellar
+environments\cite{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
+are expected to be isotropic given the absence of a preferred
+direction\cite{2016ApJ...832L...2R} and the persistence of an
+isotropic distribution through post-Newtonian
+evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
+
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{2016PhRvL.116f1102A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Observation of Gravitational Waves from a Binary
+  Black Hole Merger}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{061102}
+  (\bibinfo{year}{2016}).
+
+\bibitem{2016PhRvL.116x1103A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{GW151226: Observation of Gravitational Waves from a
+  22-Solar-Mass Binary Black Hole Coalescence}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241103}
+  (\bibinfo{year}{2016}).
+
+\bibitem{O1-BBH}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Binary Black Hole Mergers in the First Advanced LIGO
+  Observing Run}}.
+\newblock \emph{\bibinfo{journal}{Physical Review X}}
+  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{041015} (\bibinfo{year}{2016}).
+
+\bibitem{PhysRevLett.118.221101}
+\bibinfo{author}{Abbott, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{Gw170104: Observation of a 50-solar-mass binary black
+  hole coalescence at redshift 0.2}.
+\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
+  \textbf{\bibinfo{volume}{118}}, \bibinfo{pages}{221101}
+  (\bibinfo{year}{2017}).
+\newblock
+  \urlprefix\url{https://link.aps.org/doi/10.1103/PhysRevLett.118.221101}.
+
+\bibitem{SigurdssonHernquist:1993}
+\bibinfo{author}{{Sigurdsson}, S.} \& \bibinfo{author}{{Hernquist}, L.}
+\newblock \bibinfo{title}{{Primordial black holes in globular clusters}}.
+\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{364}},
+  \bibinfo{pages}{423--425} (\bibinfo{year}{1993}).
+
+\bibitem{PZMcMillan:2000}
+\bibinfo{author}{{Portegies Zwart}, S.~F.} \& \bibinfo{author}{McMillan, S.
+  L.~W.}
+\newblock \bibinfo{title}{Black hole mergers in the universe}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{528}},
+  \bibinfo{pages}{L17} (\bibinfo{year}{2000}).
+
+\bibitem{Rodriguez:2015}
+\bibinfo{author}{{Rodriguez}, C.~L.} \emph{et~al.}
+\newblock \bibinfo{title}{{Binary Black Hole Mergers from Globular Clusters:
+  Implications for Advanced LIGO}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{051101}
+  (\bibinfo{year}{2015}).
+
+\bibitem{Stone:2016}
+\bibinfo{author}{{Stone}, N.~C.}, \bibinfo{author}{{Metzger}, B.~D.} \&
+  \bibinfo{author}{{Haiman}, Z.}
+\newblock \bibinfo{title}{{Assisted inspirals of stellar mass black holes
+  embedded in AGN discs: solving the `final au problem'}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{464}},
+  \bibinfo{pages}{946--954} (\bibinfo{year}{2017}).
+
+\bibitem{2016ApJ...832L...2R}
+\bibinfo{author}{{Rodriguez}, C.~L.}, \bibinfo{author}{{Zevin}, M.},
+  \bibinfo{author}{{Pankow}, C.}, \bibinfo{author}{{Kalogera}, V.} \&
+  \bibinfo{author}{{Rasio}, F.~A.}
+\newblock \bibinfo{title}{{Illuminating Black Hole Binary Formation Channels
+  with Spins in Advanced LIGO}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{832}},
+  \bibinfo{pages}{L2} (\bibinfo{year}{2016}).
+
+\bibitem{TutukovYungelson:1993}
+\bibinfo{author}{Tutukov, A.~V.} \& \bibinfo{author}{Yungelson, L.~R.}
+\newblock \bibinfo{title}{The merger rate of neutron star and black hole
+  binaries}.
+\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
+  Society}} \textbf{\bibinfo{volume}{260}}, \bibinfo{pages}{675--678}
+  (\bibinfo{year}{1993}).
+
+\bibitem{2016Natur.534..512B}
+\bibinfo{author}{{Belczynski}, K.}, \bibinfo{author}{{Holz}, D.~E.},
+  \bibinfo{author}{{Bulik}, T.} \& \bibinfo{author}{{O'Shaughnessy}, R.}
+\newblock \bibinfo{title}{{The first gravitational-wave source from the
+  isolated evolution of two stars in the 40-100 solar mass range}}.
+\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{534}},
+  \bibinfo{pages}{512--515} (\bibinfo{year}{2016}).
+
+\bibitem{Stevenson:2017}
+\bibinfo{author}{{Stevenson}, S.} \emph{et~al.}
+\newblock \bibinfo{title}{{Formation of the first three gravitational-wave
+  observations through isolated binary evolution}}.
+\newblock \emph{\bibinfo{journal}{Nat. Commun.}} \textbf{\bibinfo{volume}{8}},
+  \bibinfo{pages}{14906} (\bibinfo{year}{2017}).
+
+\bibitem{MandeldeMink:2016}
+\bibinfo{author}{{Mandel}, I.} \& \bibinfo{author}{{de Mink}, S.~E.}
+\newblock \bibinfo{title}{{Merging binary black holes formed through chemically
+  homogeneous evolution in short-period stellar binaries}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{458}},
+  \bibinfo{pages}{2634--2647} (\bibinfo{year}{2016}).
+
+\bibitem{Marchant:2016}
+\bibinfo{author}{{Marchant}, P.}, \bibinfo{author}{{Langer}, N.},
+  \bibinfo{author}{{Podsiadlowski}, P.}, \bibinfo{author}{{Tauris}, T.~M.} \&
+  \bibinfo{author}{{Moriya}, T.~J.}
+\newblock \bibinfo{title}{{A new route towards merging massive black holes}}.
+\newblock \emph{\bibinfo{journal}{\aap}} \textbf{\bibinfo{volume}{588}},
+  \bibinfo{pages}{A50} (\bibinfo{year}{2016}).
+
+\bibitem{2016ApJ...833L...1A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{The Rate of Binary Black Hole Mergers Inferred from
+  Advanced LIGO Observations Surrounding GW150914}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{833}},
+  \bibinfo{pages}{L1} (\bibinfo{year}{2016}).
+
+\bibitem{2015PhR...548....1M}
+\bibinfo{author}{{Miller}, M.~C.} \& \bibinfo{author}{{Miller}, J.~M.}
+\newblock \bibinfo{title}{{The masses and spins of neutron stars and
+  stellar-mass black holes}}.
+\newblock \emph{\bibinfo{journal}{\physrep}} \textbf{\bibinfo{volume}{548}},
+  \bibinfo{pages}{1--34} (\bibinfo{year}{2015}).
+
+\bibitem{2017CQGra..34cLT01V}
+\bibinfo{author}{{Vitale}, S.}, \bibinfo{author}{{Lynch}, R.},
+  \bibinfo{author}{{Sturani}, R.} \& \bibinfo{author}{{Graff}, P.}
+\newblock \bibinfo{title}{{Use of gravitational waves to probe the formation
+  channels of compact binaries}}.
+\newblock \emph{\bibinfo{journal}{Classical and Quantum Gravity}}
+  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{03LT01}
+  (\bibinfo{year}{2017}).
+
+\bibitem{1999MNRAS.305..654K}
+\bibinfo{author}{{King}, A.~R.} \& \bibinfo{author}{{Kolb}, U.}
+\newblock \bibinfo{title}{{The evolution of black hole mass and angular
+  momentum}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{305}},
+  \bibinfo{pages}{654--660} (\bibinfo{year}{1999}).
+
+\bibitem{2015ApJ...800...17F}
+\bibinfo{author}{{Fragos}, T.} \& \bibinfo{author}{{McClintock}, J.~E.}
+\newblock \bibinfo{title}{{The Origin of Black Hole Spin in Galactic Low-mass
+  X-Ray Binaries}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{800}},
+  \bibinfo{pages}{17} (\bibinfo{year}{2015}).
+
+\bibitem{2016MNRAS.462..844K}
+\bibinfo{author}{{Kushnir}, D.}, \bibinfo{author}{{Zaldarriaga}, M.},
+  \bibinfo{author}{{Kollmeier}, J.~A.} \& \bibinfo{author}{{Waldman}, R.}
+\newblock \bibinfo{title}{{GW150914: spin-based constraints on the merger time
+  of the progenitor system}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{462}},
+  \bibinfo{pages}{844--849} (\bibinfo{year}{2016}).
+
+\bibitem{2017arXiv170200885Z}
+\bibinfo{author}{{Zaldarriaga}, M.}, \bibinfo{author}{{Kushnir}, D.} \&
+  \bibinfo{author}{{Kollmeier}, J.~A.}
+\newblock \bibinfo{title}{{Research Note: The Expected Spins of Gravitational
+  Wave Sources With Isolated Field Binary Progenitors}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1702.00885}.
+
+\bibitem{2017arXiv170203952H}
+\bibinfo{author}{{Hotokezaka}, K.} \& \bibinfo{author}{{Piran}, T.}
+\newblock \bibinfo{title}{{Implications of the low binary black hole spins
+  observed by LIGO}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1702.03952}.
+
+\bibitem{Albrecht:2009}
+\bibinfo{author}{{Albrecht}, S.}, \bibinfo{author}{{Reffert}, S.},
+  \bibinfo{author}{{Snellen}, I.~A.~G.} \& \bibinfo{author}{{Winn}, J.~N.}
+\newblock \bibinfo{title}{{Misaligned spin and orbital axes cause the anomalous
+  precession of DI Herculis}}.
+\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{461}},
+  \bibinfo{pages}{373--376} (\bibinfo{year}{2009}).
+
+\bibitem{Martin:2008b}
+\bibinfo{author}{{Martin}, R.~G.}, \bibinfo{author}{{Reis}, R.~C.} \&
+  \bibinfo{author}{{Pringle}, J.~E.}
+\newblock \bibinfo{title}{{Misalignment of the microquasar V4641 Sgr (SAX
+  J1819.3-2525)}}.
+\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{391}},
+  \bibinfo{pages}{L15--L18} (\bibinfo{year}{2008}).
+
+\bibitem{MorningstarMiller:2014}
+\bibinfo{author}{{Morningstar}, W.~R.} \& \bibinfo{author}{{Miller}, J.~M.}
+\newblock \bibinfo{title}{{The Spin of the Black Hole 4U 1543-47}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{793}},
+  \bibinfo{pages}{L33} (\bibinfo{year}{2014}).
+
+\bibitem{2000ApJ...541..319K}
+\bibinfo{author}{{Kalogera}, V.}
+\newblock \bibinfo{title}{{Spin-Orbit Misalignment in Close Binaries with Two
+  Compact Objects}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{541}},
+  \bibinfo{pages}{319--328} (\bibinfo{year}{2000}).
+
+\bibitem{2011ApJ...742...81F}
+\bibinfo{author}{{Farr}, W.~M.}, \bibinfo{author}{{Kremer}, K.},
+  \bibinfo{author}{{Lyutikov}, M.} \& \bibinfo{author}{{Kalogera}, V.}
+\newblock \bibinfo{title}{{Spin Tilts in the Double Pulsar Reveal Supernova
+  Spin Angular-momentum Production}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{742}},
+  \bibinfo{pages}{81} (\bibinfo{year}{2011}).
+
+\bibitem{2015PhRvD..92f4016G}
+\bibinfo{author}{{Gerosa}, D.}, \bibinfo{author}{{Kesden}, M.},
+  \bibinfo{author}{{Sperhake}, U.}, \bibinfo{author}{{Berti}, E.} \&
+  \bibinfo{author}{{O'Shaughnessy}, R.}
+\newblock \bibinfo{title}{{Multi-timescale analysis of phase transitions in
+  precessing black-hole binaries}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{92}},
+  \bibinfo{pages}{064016} (\bibinfo{year}{2015}).
+
+\bibitem{2004PhRvD..70l4020S}
+\bibinfo{author}{{Schnittman}, J.~D.}
+\newblock \bibinfo{title}{{Spin-orbit resonance and the evolution of compact
+  binary systems}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{70}},
+  \bibinfo{pages}{124020} (\bibinfo{year}{2004}).
+
+\bibitem{2007ApJ...661L.147B}
+\bibinfo{author}{{Bogdanovi{\'c}}, T.}, \bibinfo{author}{{Reynolds}, C.~S.} \&
+  \bibinfo{author}{{Miller}, M.~C.}
+\newblock \bibinfo{title}{{Alignment of the Spins of Supermassive Black Holes
+  Prior to Coalescence}}.
+\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{661}},
+  \bibinfo{pages}{L147--L150} (\bibinfo{year}{2007}).
+
+\setcounter{mainbib}{\value{enumiv}}
+\end{thebibliography}
+
+\begin{addendum}
+\item We thank Richard O'Shaughnessy, Christopher Berry, Davide
+  Gerosa, and Salvatore Vitale for discussions and comments on this
+  work.  WF, SS, IM and AV were supported in part by the STFC.  MCM
+  acknowledges support of the University of Birmingham Institute for
+  Advanced Study Distinguished Visiting Fellows program.  SS and IM
+  acknowledge support from the National Science Foundation under Grant
+  No. NSF PHY11-25915.
+\item[Author Contributions] All authors contributed to the work
+  presented in this paper.
+\item[Author Information] Reprints and permissions information is
+  available at \url{http://www.nature.com/reprints}.  The authors
+  declare no competing financial interests.  Correspondence and
+  requests for materials should be addressed to
+  \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
+\end{addendum}
+
+\newpage
+
+\begin{figure}
+  \plotone{../plots/chi-eff-mock-posteriors}
+  \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
+      $\chieff$ from the Advanced LIGO O1 and GW170104
+      observations\cite{O1-BBH,PhysRevLett.118.221101}.}  We
+    approximate the posteriors reported using Gaussians with the same
+    median and 90\% credible interval.  It is notable that none of the
+    $\chieff$ posteriors support high \ac{BH} spin magnitudes with
+    aligned spins, suggested by observations of stellar-mass black
+    holes in X-ray binaries\cite{2015PhR...548....1M}.}
+\end{figure}
+
+\begin{figure}
+  \plotone{../plots/chi-eff-distributions}
+  \caption{\label{fig:chieff-distribution-models} \textbf{The models for the
+    population distribution of $\chieff$ considered in this paper.}  In
+    all models we assume that the binary mass ratio
+    $q \equiv m_1/m_2 = 1$ and that the distribution of spin vectors
+    is the same for each component.  The ``flat'' (blue lines),
+    ``high,'' (green lines), and ``low'' (red lines) magnitude
+    distributions are defined in Eq.\ \eqref{eq:magnitude-dists}.
+    Solid lines give the $\chieff$ distribution under the assumption
+    that the orientations of the spins are isotropic; dashed lines
+    give the distribution under the assumption that both objects'
+    spins are aligned with the orbital angular momentum.  The
+    isotropic distributions are readily distinguished from the aligned
+    distributions by the production of negative $\chieff$ values,
+    while the distinction between the three models for the spin
+    magnitude distribution is less sharp.}
+\end{figure}
+
+\begin{figure}
+  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models}
+  \caption{\textbf{Odds ratios among our models using the
+      approximations to the posteriors on $\chieff$ from the O1 and GW170104
+      observations shown in Figure \ref{fig:O1-posteriors}.}  The flat
+    (``F''), high (``H''), and low (``L'') spin magnitude
+    distributions (see Eq.\ \eqref{eq:magnitude-dists}) are paired
+    with isotropic (``I'') and aligned (``A'') angular distributions,
+    as well as a mixture model of the two (``$\mathrm{M}$'').  The
+    most-favoured models have the ``low'' distribution of spin
+    magnitudes.  The odds ratio between the best aligned and best
+    isotropic models is $\OOneOddsIsoAligned$, or
+    $\OOneSigmaIsoAligned\sigma$.  For all magnitude distributions the
+    pure-isotropic models are preferred over the mixture models;
+    correspondingly, the posterior on the mixture fraction peaks at
+    100\% isotropic.}
+  \label{fig:O1-odds}
+\end{figure}
+
 \begin{figure}
@@
 
-Most of our resolving power for the spin angular distribution is a
-result of the fact that our ``aligned'' models cannot produce
-$\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
-spins are intrinsically very small, with $a \lesssim 0.2$, then it is
-no longer possible to resolve the negative effective spin with a small
-number of observations (see Methods Section \ref{sec:smallspins}).  As
-noted below, however, spins observed in X-ray binaries are typically
-large.  Additionally, models which do not permit \emph{some} spins
-with $\chieff \gtrsim 0.1$ are ruled out by the GW151226
-observations\cite{2016PhRvL.116x1103A}.  An ``aligned'' model with
-spin magnitudes from our ``flat'' distribution but permitting spin
-vectors oriented anti-parallel to the orbital angular momentum
-(leading to the possibility of positive \emph{or} negative $\chieff$)
-can only be distinguished from an isotropic true population at
-$\sim 3 \sigma$ after 10--20 observations\cite{2017CQGra..34cLT01V};
-our flat aligned model can be distinguished from such a population at
-better than $5\sigma$ (odds $< 10^{-8}$) after 10 observations,
-emphasizing the information content of the bound $\chieff > 0$ for our
-aligned models.
-
-Observational data on spin magnitudes in black hole systems is
-sparse\cite{2015PhR...548....1M}.  Most of the systems studied are
-low-mass X-ray binaries rather than the high-mass X-ray binaries that
-are likely to be the progenitors of double black hole binaries.  In
-addition, there are substantial systematic errors that can complicate
-these analyses\cite{2015PhR...548....1M} and selection effects could
-yield a biased distribution. Nonetheless, if we take the reported spin
-magnitudes as representative then we find that there is a preference
-for high spins; for example, 14 of the 19 systems with reported spins
-have dimensionless spin parameters in excess of 0.5.  It is usually
-argued that the masses and spin parameters of stellar-mass black holes
-are unlikely to be altered significantly by
-accretion\cite{1999MNRAS.305..654K}, but this may not be true for all
-systems\cite{2015ApJ...800...17F}.  Thus the current spin parameters
-are probably close to their values upon core collapse, at least in
-high-mass X-ray binaries.  However, the specific processes involved in
-the production of black hole binaries from isolated binaries could
-alter the spin magnitude distribution of those holes relative to the
-X-ray binary systems; for example, close tidal interactions could spin
-up the core, or stripping of the envelope could reduce the available
-angular
-momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.
-
-The spin directions in isolated binary black
-holes\cite{TutukovYungelson:1993,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
-are usually expected to be preferentially aligned.  Despite observed
-spin-orbit misalignments in massive stellar
-binaries\cite{Albrecht:2009}, mass transfer and tidal interactions
-will tend to realign the binary.  On the other hand, there is some
-evidence of spin-orbit misalignment in black hole X-ray
-binaries\cite{Martin:2008b,MorningstarMiller:2014}.  This is
-consistent with the expectation that a supernova natal kick (if any)
-can change the orbital plane and misalign the
-binary\cite{2000ApJ...541..319K}; the supernova can also tilt the spin
-angle\cite{2011ApJ...742...81F}.  Evolutionary processes, such as
-wind-driven mass loss and post-collapse fallback, can couple the spin
-magnitude and direction distributions, contrary to our simplified
-assumptions.  A small misalignment at wide separation can also evolve
-to a more significant misalignment in component spins as the binary
-spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
-$\chieff$ is approximately conserved through this evolution.
-
-The spin directions of binary black holes formed dynamically through
-interactions in dense stellar
-environments\cite{SigurdssonHernquist:1993,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
-are expected to be isotropic given the absence of a preferred
-direction\cite{2016ApJ...832L...2R} and the persistence of an
-isotropic distribution through post-Newtonian
-evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.
-
-\begin{thebibliography}{10}
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem{2016PhRvL.116f1102A}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{Observation of Gravitational Waves from a Binary
-  Black Hole Merger}}.
-\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
-  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{061102}
-  (\bibinfo{year}{2016}).
-
-\bibitem{2016PhRvL.116x1103A}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{GW151226: Observation of Gravitational Waves from a
-  22-Solar-Mass Binary Black Hole Coalescence}}.
-\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
-  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241103}
-  (\bibinfo{year}{2016}).
-
-\bibitem{O1-BBH}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{Binary Black Hole Mergers in the First Advanced LIGO
-  Observing Run}}.
-\newblock \emph{\bibinfo{journal}{Physical Review X}}
-  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{041015} (\bibinfo{year}{2016}).
-
-\bibitem{PhysRevLett.118.221101}
-\bibinfo{author}{Abbott, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{Gw170104: Observation of a 50-solar-mass binary black
-  hole coalescence at redshift 0.2}.
-\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
-  \textbf{\bibinfo{volume}{118}}, \bibinfo{pages}{221101}
-  (\bibinfo{year}{2017}).
-\newblock
-  \urlprefix\url{https://link.aps.org/doi/10.1103/PhysRevLett.118.221101}.
-
-\bibitem{SigurdssonHernquist:1993}
-\bibinfo{author}{{Sigurdsson}, S.} \& \bibinfo{author}{{Hernquist}, L.}
-\newblock \bibinfo{title}{{Primordial black holes in globular clusters}}.
-\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{364}},
-  \bibinfo{pages}{423--425} (\bibinfo{year}{1993}).
-
-\bibitem{PZMcMillan:2000}
-\bibinfo{author}{{Portegies Zwart}, S.~F.} \& \bibinfo{author}{McMillan, S.
-  L.~W.}
-\newblock \bibinfo{title}{Black hole mergers in the universe}.
-\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{528}},
-  \bibinfo{pages}{L17} (\bibinfo{year}{2000}).
-
-\bibitem{Rodriguez:2015}
-\bibinfo{author}{{Rodriguez}, C.~L.} \emph{et~al.}
-\newblock \bibinfo{title}{{Binary Black Hole Mergers from Globular Clusters:
-  Implications for Advanced LIGO}}.
-\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
-  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{051101}
-  (\bibinfo{year}{2015}).
-
-\bibitem{Stone:2016}
-\bibinfo{author}{{Stone}, N.~C.}, \bibinfo{author}{{Metzger}, B.~D.} \&
-  \bibinfo{author}{{Haiman}, Z.}
-\newblock \bibinfo{title}{{Assisted inspirals of stellar mass black holes
-  embedded in AGN discs: solving the `final au problem'}}.
-\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{464}},
-  \bibinfo{pages}{946--954} (\bibinfo{year}{2017}).
-
-\bibitem{2016ApJ...832L...2R}
-\bibinfo{author}{{Rodriguez}, C.~L.}, \bibinfo{author}{{Zevin}, M.},
-  \bibinfo{author}{{Pankow}, C.}, \bibinfo{author}{{Kalogera}, V.} \&
-  \bibinfo{author}{{Rasio}, F.~A.}
-\newblock \bibinfo{title}{{Illuminating Black Hole Binary Formation Channels
-  with Spins in Advanced LIGO}}.
-\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{832}},
-  \bibinfo{pages}{L2} (\bibinfo{year}{2016}).
-
-\bibitem{TutukovYungelson:1993}
-\bibinfo{author}{Tutukov, A.~V.} \& \bibinfo{author}{Yungelson, L.~R.}
-\newblock \bibinfo{title}{The merger rate of neutron star and black hole
-  binaries}.
-\newblock \emph{\bibinfo{journal}{Monthly Notices of the Royal Astronomical
-  Society}} \textbf{\bibinfo{volume}{260}}, \bibinfo{pages}{675--678}
-  (\bibinfo{year}{1993}).
-
-\bibitem{2016Natur.534..512B}
-\bibinfo{author}{{Belczynski}, K.}, \bibinfo{author}{{Holz}, D.~E.},
-  \bibinfo{author}{{Bulik}, T.} \& \bibinfo{author}{{O'Shaughnessy}, R.}
-\newblock \bibinfo{title}{{The first gravitational-wave source from the
-  isolated evolution of two stars in the 40-100 solar mass range}}.
-\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{534}},
-  \bibinfo{pages}{512--515} (\bibinfo{year}{2016}).
-
-\bibitem{Stevenson:2017}
-\bibinfo{author}{{Stevenson}, S.} \emph{et~al.}
-\newblock \bibinfo{title}{{Formation of the first three gravitational-wave
-  observations through isolated binary evolution}}.
-\newblock \emph{\bibinfo{journal}{Nat. Commun.}} \textbf{\bibinfo{volume}{8}},
-  \bibinfo{pages}{14906} (\bibinfo{year}{2017}).
-
-\bibitem{MandeldeMink:2016}
-\bibinfo{author}{{Mandel}, I.} \& \bibinfo{author}{{de Mink}, S.~E.}
-\newblock \bibinfo{title}{{Merging binary black holes formed through chemically
-  homogeneous evolution in short-period stellar binaries}}.
-\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{458}},
-  \bibinfo{pages}{2634--2647} (\bibinfo{year}{2016}).
-
-\bibitem{Marchant:2016}
-\bibinfo{author}{{Marchant}, P.}, \bibinfo{author}{{Langer}, N.},
-  \bibinfo{author}{{Podsiadlowski}, P.}, \bibinfo{author}{{Tauris}, T.~M.} \&
-  \bibinfo{author}{{Moriya}, T.~J.}
-\newblock \bibinfo{title}{{A new route towards merging massive black holes}}.
-\newblock \emph{\bibinfo{journal}{\aap}} \textbf{\bibinfo{volume}{588}},
-  \bibinfo{pages}{A50} (\bibinfo{year}{2016}).
-
-\bibitem{2016ApJ...833L...1A}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{The Rate of Binary Black Hole Mergers Inferred from
-  Advanced LIGO Observations Surrounding GW150914}}.
-\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{833}},
-  \bibinfo{pages}{L1} (\bibinfo{year}{2016}).
-
-\bibitem{2015PhR...548....1M}
-\bibinfo{author}{{Miller}, M.~C.} \& \bibinfo{author}{{Miller}, J.~M.}
-\newblock \bibinfo{title}{{The masses and spins of neutron stars and
-  stellar-mass black holes}}.
-\newblock \emph{\bibinfo{journal}{\physrep}} \textbf{\bibinfo{volume}{548}},
-  \bibinfo{pages}{1--34} (\bibinfo{year}{2015}).
-
-\bibitem{2017CQGra..34cLT01V}
-\bibinfo{author}{{Vitale}, S.}, \bibinfo{author}{{Lynch}, R.},
-  \bibinfo{author}{{Sturani}, R.} \& \bibinfo{author}{{Graff}, P.}
-\newblock \bibinfo{title}{{Use of gravitational waves to probe the formation
-  channels of compact binaries}}.
-\newblock \emph{\bibinfo{journal}{Classical and Quantum Gravity}}
-  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{03LT01}
-  (\bibinfo{year}{2017}).
-
-\bibitem{1999MNRAS.305..654K}
-\bibinfo{author}{{King}, A.~R.} \& \bibinfo{author}{{Kolb}, U.}
-\newblock \bibinfo{title}{{The evolution of black hole mass and angular
-  momentum}}.
-\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{305}},
-  \bibinfo{pages}{654--660} (\bibinfo{year}{1999}).
-
-\bibitem{2015ApJ...800...17F}
-\bibinfo{author}{{Fragos}, T.} \& \bibinfo{author}{{McClintock}, J.~E.}
-\newblock \bibinfo{title}{{The Origin of Black Hole Spin in Galactic Low-mass
-  X-Ray Binaries}}.
-\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{800}},
-  \bibinfo{pages}{17} (\bibinfo{year}{2015}).
-
-\bibitem{2016MNRAS.462..844K}
-\bibinfo{author}{{Kushnir}, D.}, \bibinfo{author}{{Zaldarriaga}, M.},
-  \bibinfo{author}{{Kollmeier}, J.~A.} \& \bibinfo{author}{{Waldman}, R.}
-\newblock \bibinfo{title}{{GW150914: spin-based constraints on the merger time
-  of the progenitor system}}.
-\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{462}},
-  \bibinfo{pages}{844--849} (\bibinfo{year}{2016}).
-
-\bibitem{2017arXiv170200885Z}
-\bibinfo{author}{{Zaldarriaga}, M.}, \bibinfo{author}{{Kushnir}, D.} \&
-  \bibinfo{author}{{Kollmeier}, J.~A.}
-\newblock \bibinfo{title}{{Research Note: The Expected Spins of Gravitational
-  Wave Sources With Isolated Field Binary Progenitors}}.
-\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
-\newblock \eprint{arXiv:1702.00885}.
-
-\bibitem{2017arXiv170203952H}
-\bibinfo{author}{{Hotokezaka}, K.} \& \bibinfo{author}{{Piran}, T.}
-\newblock \bibinfo{title}{{Implications of the low binary black hole spins
-  observed by LIGO}}.
-\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
-\newblock \eprint{arXiv:1702.03952}.
-
-\bibitem{Albrecht:2009}
-\bibinfo{author}{{Albrecht}, S.}, \bibinfo{author}{{Reffert}, S.},
-  \bibinfo{author}{{Snellen}, I.~A.~G.} \& \bibinfo{author}{{Winn}, J.~N.}
-\newblock \bibinfo{title}{{Misaligned spin and orbital axes cause the anomalous
-  precession of DI Herculis}}.
-\newblock \emph{\bibinfo{journal}{\nat}} \textbf{\bibinfo{volume}{461}},
-  \bibinfo{pages}{373--376} (\bibinfo{year}{2009}).
-
-\bibitem{Martin:2008b}
-\bibinfo{author}{{Martin}, R.~G.}, \bibinfo{author}{{Reis}, R.~C.} \&
-  \bibinfo{author}{{Pringle}, J.~E.}
-\newblock \bibinfo{title}{{Misalignment of the microquasar V4641 Sgr (SAX
-  J1819.3-2525)}}.
-\newblock \emph{\bibinfo{journal}{\mnras}} \textbf{\bibinfo{volume}{391}},
-  \bibinfo{pages}{L15--L18} (\bibinfo{year}{2008}).
-
-\bibitem{MorningstarMiller:2014}
-\bibinfo{author}{{Morningstar}, W.~R.} \& \bibinfo{author}{{Miller}, J.~M.}
-\newblock \bibinfo{title}{{The Spin of the Black Hole 4U 1543-47}}.
-\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{793}},
-  \bibinfo{pages}{L33} (\bibinfo{year}{2014}).
-
-\bibitem{2000ApJ...541..319K}
-\bibinfo{author}{{Kalogera}, V.}
-\newblock \bibinfo{title}{{Spin-Orbit Misalignment in Close Binaries with Two
-  Compact Objects}}.
-\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{541}},
-  \bibinfo{pages}{319--328} (\bibinfo{year}{2000}).
-
-\bibitem{2011ApJ...742...81F}
-\bibinfo{author}{{Farr}, W.~M.}, \bibinfo{author}{{Kremer}, K.},
-  \bibinfo{author}{{Lyutikov}, M.} \& \bibinfo{author}{{Kalogera}, V.}
-\newblock \bibinfo{title}{{Spin Tilts in the Double Pulsar Reveal Supernova
-  Spin Angular-momentum Production}}.
-\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{742}},
-  \bibinfo{pages}{81} (\bibinfo{year}{2011}).
-
-\bibitem{2015PhRvD..92f4016G}
-\bibinfo{author}{{Gerosa}, D.}, \bibinfo{author}{{Kesden}, M.},
-  \bibinfo{author}{{Sperhake}, U.}, \bibinfo{author}{{Berti}, E.} \&
-  \bibinfo{author}{{O'Shaughnessy}, R.}
-\newblock \bibinfo{title}{{Multi-timescale analysis of phase transitions in
-  precessing black-hole binaries}}.
-\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{92}},
-  \bibinfo{pages}{064016} (\bibinfo{year}{2015}).
-
-\bibitem{2004PhRvD..70l4020S}
-\bibinfo{author}{{Schnittman}, J.~D.}
-\newblock \bibinfo{title}{{Spin-orbit resonance and the evolution of compact
-  binary systems}}.
-\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{70}},
-  \bibinfo{pages}{124020} (\bibinfo{year}{2004}).
-
-\bibitem{2007ApJ...661L.147B}
-\bibinfo{author}{{Bogdanovi{\'c}}, T.}, \bibinfo{author}{{Reynolds}, C.~S.} \&
-  \bibinfo{author}{{Miller}, M.~C.}
-\newblock \bibinfo{title}{{Alignment of the Spins of Supermassive Black Holes
-  Prior to Coalescence}}.
-\newblock \emph{\bibinfo{journal}{\apjl}} \textbf{\bibinfo{volume}{661}},
-  \bibinfo{pages}{L147--L150} (\bibinfo{year}{2007}).
-
-\setcounter{mainbib}{\value{enumiv}}
-\end{thebibliography}
-
-\begin{addendum}
-\item We thank Richard O'Shaughnessy, Christopher Berry, Davide
-  Gerosa, and Salvatore Vitale for discussions and comments on this
-  work.  WF, SS, IM and AV were supported in part by the STFC.  MCM
-  acknowledges support of the University of Birmingham Institute for
-  Advanced Study Distinguished Visiting Fellows program.  SS and IM
-  acknowledge support from the National Science Foundation under Grant
-  No. NSF PHY11-25915.
-\item[Author Contributions] All authors contributed to the work
-  presented in this paper.
-\item[Author Information] Reprints and permissions information is
-  available at \url{http://www.nature.com/reprints}.  The authors
-  declare no competing financial interests.  Correspondence and
-  requests for materials should be addressed to
-  \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
-\end{addendum}
-
 \begin{methods}
@@
 \renewcommand{\figurename}{Extended Data Figure}
-
-\section{Code Availability}
-
-This analysis used the Julia language\cite{Julia}, Python libraries
-NumPy and SciPy\cite{NumPy,SciPy}, the plotting library
-Matplotlib\cite{Matplotlib}, and performed computations in IPython
-notebooks\cite{IPython}.  A repository containing the code and
-notebooks used for this analysis, together with the \LaTeX{} source
-for this document, can be found under an open-source ``MIT'' license
-at \url{https://github.com/farr/AlignedVersusIsoSpin}.
+\renewcommand{\tablename}{Extended Data Table}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{../plots/pa}
-  \caption{\textbf{Distributions of spin magnitudes.}  See Eq.\
-    \eqref{eq:magnitude-dists} for the definition of the low (blue
-    line), flat (green line), and high (red line) magnitude
-    distributions used here.  The distributions have mean spin $0.33$,
-    $0.5$, and $0.67$ and standard deviations $0.24$, $0.29$, and
-    $0.24$.}
-  \label{fig:pa}
-\end{figure}
-
 \section{Mixture model}
@@
 %
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction_with_GW170104.png}
-\caption{\textbf{Fraction of the BBH population coming from an
-    isotropic distribution under a mixture model.} The dotted line
-  shows the flat prior on the fraction of BBHs coming from an
-  isotropic distribution, $f_i$, under the mixture model. The 3 red
-  lines show the posterior on $f_i$ after O1 and GW170104 with our
-  various assumptions regarding BH spin magnitudes.  The solid line
-  shows the posterior assuming that all BHs have their spin magnitude
-  drawn from the ``flat'' distribution. The dashed line assumes the
-  ``high'' BH spin magnitude distribution $p(a) = 2a$. The dot-dash
-  line assumes the ``low'' distribution $p(a) = 2(1-a)$.  We see that
-  for a wide range of assumptions regarding BH spin magnitudes, the
-  fraction coming from an isotropic distribution $f_i$ peaks at 1.}
-\label{fig:mixture_fraction_posterior}
-\end{figure}
 %
@@
 
-In Table~\ref{tab:accumulation} we show how the evidence for an isotropic distribution increases when including: only the 2 confirmed events---GW150914 and GW151226---from O1; all O1 events (including LVT151012); and all 4 likely binary black hole mergers, including GW170104.
-
-\begin{table}
-\begin{centering}
-\begin{tabular}{ c |  c  |  c  | c  }
-  \hline 
-  \hline
-Events & $\sigma_{I/A}$ &  $\sigma_{I/A}$ & $\sigma_{I/A}$ \\ 
- & ``Low'' & ``Flat'' & ``High'' \\ \hline
-GW150914 and GW151226 &  1.3 & 2.2 & 3.7 \\ %\hline
-All O1 events & 1.7 & 2.7 & 4.4 \\ %\hline
-All O1 events and GW170104 & \textbf{2.4} & 3.5 & 5.4 \\ \hline
-\end{tabular}
-\caption[]{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
-The number in bold is the result we quote in the main text.}
-\label{tab:accumulation}
-\end{centering}
-\end{table}
+In Extended Data Table~\ref{tab:accumulation} we show how the evidence for an isotropic distribution increases when including: only the 2 confirmed events---GW150914 and GW151226---from O1; all O1 events (including LVT151012); and all 4 likely binary black hole mergers, including GW170104.
 
@@
 %
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{../plots/sigma_v_alpha_GW170104.png}
-\caption{\textbf{Effect of small spins on evidence ratio of isotropic against
-  aligned models.} The blue line shows the evidence ratio
-  (plotted as the equivalent sigma) between a model where all systems
-  are from an isotropic distribution, versus one where all systems are
-  aligned, as a function of the power law $\alpha$ corresponding
-  to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
-  magnitude $\bar{a}$ which this $\alpha$ corresponds to. We see that
-  for mean spin magnitudes $\lesssim 0.2$ we find no evidence for
-  either distribution over the other.}
-\label{fig:smallspinsalpha}
-\end{figure}
 %
@@
 distributions.
-
-\begin{figure}
-  \plotone{../plots/chi-eff-distributions-q05}
-  \caption{\textbf{Distributions of $\chieff$ assuming all merging
-      black holes have equal masses ($q=1$) or a 2:1 mass ratio
-      ($q = 0.5$).}  The details of the distribution are sensitive to
-    the mass ratio, but in our analysis we are primarily sensitive to
-    the changing \emph{sign} of $\chieff$ under the isotropic models.
-    This latter property is unchanged under changing mass ratio.}
-  \label{fig:mass-ratio-sensitivity}
-\end{figure}
 
@@
 
+\section{Code Availability}
+
+This analysis used the Julia language\cite{Julia}, Python libraries
+NumPy and SciPy\cite{NumPy,SciPy}, the plotting library
+Matplotlib\cite{Matplotlib}, and performed computations in IPython
+notebooks\cite{IPython}.  A repository containing the code and
+notebooks used for this analysis, together with the \LaTeX{} source
+for this document, can be found under an open-source ``MIT'' license
+at \url{https://github.com/farr/AlignedVersusIsoSpin}.
+
+\section{Data Availability}
+
+The datasets generated and analysed during the current study are
+available at \url{https://github.com/farr/AlignedVersusIsoSpin}.
+
+\begin{thebibliography}{10}
+\setcounter{enumiv}{\value{mainbib}}
+
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{Julia}
+\bibinfo{author}{Bezanson, J.}, \bibinfo{author}{Karpinski, S.},
+  \bibinfo{author}{Shah, V.~B.} \& \bibinfo{author}{Edelman, A.}
+\newblock \bibinfo{title}{{J}ulia: A fast dynamic language for technical
+  computing}  (\bibinfo{year}{2012}).
+\newblock \eprint{arXiv:1209.5145}.
+
+\bibitem{NumPy}
+\bibinfo{author}{van~der Walt, S.}, \bibinfo{author}{Colbert, S.~C.} \&
+  \bibinfo{author}{Varoquaux, G.}
+\newblock \bibinfo{title}{The numpy array: A structure for efficient numerical
+  computation}.
+\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{22--30}
+  (\bibinfo{year}{2011}).
+
+\bibitem{SciPy}
+\bibinfo{author}{Jones, E.}, \bibinfo{author}{Oliphant, T.},
+  \bibinfo{author}{Peterson, P.} \emph{et~al.}
+\newblock \bibinfo{title}{{SciPy}: Open source scientific tools for {Python}}
+  (\bibinfo{year}{2001--}).
+\newblock \urlprefix\url{http://www.scipy.org/}.
+\newblock \bibinfo{note}{[Online; accessed 24 May 2017]}.
+
+\bibitem{Matplotlib}
+\bibinfo{author}{Hunter, J.~D.}
+\newblock \bibinfo{title}{Matplotlib: A 2d graphics environment}.
+\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
+  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{90--95} (\bibinfo{year}{2007}).
+
+\bibitem{IPython}
+\bibinfo{author}{Pérez, F.} \& \bibinfo{author}{Granger, B.~E.}
+\newblock \bibinfo{title}{Ipython: A system for interactive scientific
+  computing}.
+\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
+  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{21--29} (\bibinfo{year}{2007}).
+\newblock
+  \urlprefix\url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2007.53}.
+
+\bibitem{2016PhRvL.116x1102A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Properties of the Binary Black Hole Merger
+  GW150914}}.
+\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
+  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241102}
+  (\bibinfo{year}{2016}).
+
+\bibitem{Stevenson:2017spin}
+\bibinfo{author}{{Stevenson}, S.}, \bibinfo{author}{{Berry}, C.~P.~L.} \&
+  \bibinfo{author}{{Mandel}, I.}
+\newblock \bibinfo{title}{{Hierarchical analysis of gravitational-wave
+  measurements of binary black hole spin-orbit misalignments}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{1703.06873}.
+
+\bibitem{2015PhRvD..91d2003V}
+\bibinfo{author}{{Veitch}, J.} \emph{et~al.}
+\newblock \bibinfo{title}{{Parameter estimation for compact binaries with
+  ground-based gravitational-wave observations using the LALInference software
+  library}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{91}},
+  \bibinfo{pages}{042003} (\bibinfo{year}{2015}).
+
+\bibitem{2010ApJ...725.2166H}
+\bibinfo{author}{{Hogg}, D.~W.}, \bibinfo{author}{{Myers}, A.~D.} \&
+  \bibinfo{author}{{Bovy}, J.}
+\newblock \bibinfo{title}{{Inferring the Eccentricity Distribution}}.
+\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{725}},
+  \bibinfo{pages}{2166--2175} (\bibinfo{year}{2010}).
+
+\bibitem{2010PhRvD..81h4029M}
+\bibinfo{author}{{Mandel}, I.}
+\newblock \bibinfo{title}{{Parameter estimation on gravitational waves from
+  multiple coalescing binaries}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{81}},
+  \bibinfo{pages}{084029} (\bibinfo{year}{2010}).
+
+\bibitem{2017arXiv170306869F}
+\bibinfo{author}{{Fishbach}, M.}, \bibinfo{author}{{Holz}, D.} \&
+  \bibinfo{author}{{Farr}, B.}
+\newblock \bibinfo{title}{{Are LIGO's Black Holes Made From Smaller Black
+  Holes?}}
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1703.06869}.
+
+\bibitem{2017arXiv170306223G}
+\bibinfo{author}{{Gerosa}, D.} \& \bibinfo{author}{{Berti}, E.}
+\newblock \bibinfo{title}{{First or second generation? Black hole census with
+  gravitational wave observations}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
+\newblock \eprint{arXiv:1703.06223}.
+
+\bibitem{2016PhRvD..93l2003A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{GW150914: First results from the search for binary
+  black hole coalescence with Advanced LIGO}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{93}},
+  \bibinfo{pages}{122003} (\bibinfo{year}{2016}).
+
+\bibitem{2016arXiv161107531T}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Effects of waveform model systematics on the
+  interpretation of GW150914}}.
+\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2016}).
+\newblock \eprint{arXiv:1611.07531}.
+
+\bibitem{2016PhRvD..94f4035A}
+\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
+\newblock \bibinfo{title}{{Directly comparing GW150914 with numerical solutions
+  of Einstein's equations for binary black hole coalescence}}.
+\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{94}},
+  \bibinfo{pages}{064035} (\bibinfo{year}{2016}).
+
+\end{thebibliography}
+
+\newpage
+
+\begin{table}
+\begin{centering}
+\begin{tabular}{ c |  c  |  c  | c  }
+  \hline 
+  \hline
+Events & $\sigma_{I/A}$ &  $\sigma_{I/A}$ & $\sigma_{I/A}$ \\ 
+ & ``Low'' & ``Flat'' & ``High'' \\ \hline
+GW150914 and GW151226 &  1.3 & 2.2 & 3.7 \\ %\hline
+All O1 events & 1.7 & 2.7 & 4.4 \\ %\hline
+All O1 events and GW170104 & \textbf{2.4} & 3.5 & 5.4 \\ \hline
+\end{tabular}
+\caption[]{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
+The number in bold is the result we quote in the main text.}
+\label{tab:accumulation}
+\end{centering}
+\end{table}
+
+\newpage
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{../plots/pa}
+  \caption{\textbf{Distributions of spin magnitudes.}  See Eq.\
+    \eqref{eq:magnitude-dists} for the definition of the low (blue
+    line), flat (green line), and high (red line) magnitude
+    distributions used here.  The distributions have mean spin $0.33$,
+    $0.5$, and $0.67$ and standard deviations $0.24$, $0.29$, and
+    $0.24$.}
+  \label{fig:pa}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction_with_GW170104.png}
+\caption{\textbf{Fraction of the BBH population coming from an
+    isotropic distribution under a mixture model.} The dotted line
+  shows the flat prior on the fraction of BBHs coming from an
+  isotropic distribution, $f_i$, under the mixture model. The 3 red
+  lines show the posterior on $f_i$ after O1 and GW170104 with our
+  various assumptions regarding BH spin magnitudes.  The solid line
+  shows the posterior assuming that all BHs have their spin magnitude
+  drawn from the ``flat'' distribution. The dashed line assumes the
+  ``high'' BH spin magnitude distribution $p(a) = 2a$. The dot-dash
+  line assumes the ``low'' distribution $p(a) = 2(1-a)$.  We see that
+  for a wide range of assumptions regarding BH spin magnitudes, the
+  fraction coming from an isotropic distribution $f_i$ peaks at 1.}
+\label{fig:mixture_fraction_posterior}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{../plots/sigma_v_alpha_GW170104.png}
+\caption{\textbf{Effect of small spins on evidence ratio of isotropic against
+  aligned models.} The blue line shows the evidence ratio
+  (plotted as the equivalent sigma) between a model where all systems
+  are from an isotropic distribution, versus one where all systems are
+  aligned, as a function of the power law $\alpha$ corresponding
+  to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
+  magnitude $\bar{a}$ which this $\alpha$ corresponds to. We see that
+  for mean spin magnitudes $\lesssim 0.2$ we find no evidence for
+  either distribution over the other.}
+\label{fig:smallspinsalpha}
+\end{figure}
+
+\begin{figure}
+  \plotone{../plots/chi-eff-distributions-q05}
+  \caption{\textbf{Distributions of $\chieff$ assuming all merging
+      black holes have equal masses ($q=1$) or a 2:1 mass ratio
+      ($q = 0.5$).}  The details of the distribution are sensitive to
+    the mass ratio, but in our analysis we are primarily sensitive to
+    the changing \emph{sign} of $\chieff$ under the isotropic models.
+    This latter property is unchanged under changing mass ratio.}
+  \label{fig:mass-ratio-sensitivity}
+\end{figure}
+
 \begin{figure}
@@
 
-\begin{thebibliography}{10}
-\setcounter{enumiv}{\value{mainbib}}
-
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem{Julia}
-\bibinfo{author}{Bezanson, J.}, \bibinfo{author}{Karpinski, S.},
-  \bibinfo{author}{Shah, V.~B.} \& \bibinfo{author}{Edelman, A.}
-\newblock \bibinfo{title}{{J}ulia: A fast dynamic language for technical
-  computing}  (\bibinfo{year}{2012}).
-\newblock \eprint{arXiv:1209.5145}.
-
-\bibitem{NumPy}
-\bibinfo{author}{van~der Walt, S.}, \bibinfo{author}{Colbert, S.~C.} \&
-  \bibinfo{author}{Varoquaux, G.}
-\newblock \bibinfo{title}{The numpy array: A structure for efficient numerical
-  computation}.
-\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
-  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{22--30}
-  (\bibinfo{year}{2011}).
-
-\bibitem{SciPy}
-\bibinfo{author}{Jones, E.}, \bibinfo{author}{Oliphant, T.},
-  \bibinfo{author}{Peterson, P.} \emph{et~al.}
-\newblock \bibinfo{title}{{SciPy}: Open source scientific tools for {Python}}
-  (\bibinfo{year}{2001--}).
-\newblock \urlprefix\url{http://www.scipy.org/}.
-\newblock \bibinfo{note}{[Online; accessed 24 May 2017]}.
-
-\bibitem{Matplotlib}
-\bibinfo{author}{Hunter, J.~D.}
-\newblock \bibinfo{title}{Matplotlib: A 2d graphics environment}.
-\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
-  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{90--95} (\bibinfo{year}{2007}).
-
-\bibitem{IPython}
-\bibinfo{author}{Pérez, F.} \& \bibinfo{author}{Granger, B.~E.}
-\newblock \bibinfo{title}{Ipython: A system for interactive scientific
-  computing}.
-\newblock \emph{\bibinfo{journal}{Computing in Science \& Engineering}}
-  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{21--29} (\bibinfo{year}{2007}).
-\newblock
-  \urlprefix\url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2007.53}.
-
-\bibitem{2016PhRvL.116x1102A}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{Properties of the Binary Black Hole Merger
-  GW150914}}.
-\newblock \emph{\bibinfo{journal}{Physical Review Letters}}
-  \textbf{\bibinfo{volume}{116}}, \bibinfo{pages}{241102}
-  (\bibinfo{year}{2016}).
-
-\bibitem{Stevenson:2017spin}
-\bibinfo{author}{{Stevenson}, S.}, \bibinfo{author}{{Berry}, C.~P.~L.} \&
-  \bibinfo{author}{{Mandel}, I.}
-\newblock \bibinfo{title}{{Hierarchical analysis of gravitational-wave
-  measurements of binary black hole spin-orbit misalignments}}.
-\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
-\newblock \eprint{1703.06873}.
-
-\bibitem{2015PhRvD..91d2003V}
-\bibinfo{author}{{Veitch}, J.} \emph{et~al.}
-\newblock \bibinfo{title}{{Parameter estimation for compact binaries with
-  ground-based gravitational-wave observations using the LALInference software
-  library}}.
-\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{91}},
-  \bibinfo{pages}{042003} (\bibinfo{year}{2015}).
-
-\bibitem{2010ApJ...725.2166H}
-\bibinfo{author}{{Hogg}, D.~W.}, \bibinfo{author}{{Myers}, A.~D.} \&
-  \bibinfo{author}{{Bovy}, J.}
-\newblock \bibinfo{title}{{Inferring the Eccentricity Distribution}}.
-\newblock \emph{\bibinfo{journal}{\apj}} \textbf{\bibinfo{volume}{725}},
-  \bibinfo{pages}{2166--2175} (\bibinfo{year}{2010}).
-
-\bibitem{2010PhRvD..81h4029M}
-\bibinfo{author}{{Mandel}, I.}
-\newblock \bibinfo{title}{{Parameter estimation on gravitational waves from
-  multiple coalescing binaries}}.
-\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{81}},
-  \bibinfo{pages}{084029} (\bibinfo{year}{2010}).
-
-\bibitem{2017arXiv170306869F}
-\bibinfo{author}{{Fishbach}, M.}, \bibinfo{author}{{Holz}, D.} \&
-  \bibinfo{author}{{Farr}, B.}
-\newblock \bibinfo{title}{{Are LIGO's Black Holes Made From Smaller Black
-  Holes?}}
-\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
-\newblock \eprint{arXiv:1703.06869}.
-
-\bibitem{2017arXiv170306223G}
-\bibinfo{author}{{Gerosa}, D.} \& \bibinfo{author}{{Berti}, E.}
-\newblock \bibinfo{title}{{First or second generation? Black hole census with
-  gravitational wave observations}}.
-\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2017}).
-\newblock \eprint{arXiv:1703.06223}.
-
-\bibitem{2016PhRvD..93l2003A}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{GW150914: First results from the search for binary
-  black hole coalescence with Advanced LIGO}}.
-\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{93}},
-  \bibinfo{pages}{122003} (\bibinfo{year}{2016}).
-
-\bibitem{2016arXiv161107531T}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{Effects of waveform model systematics on the
-  interpretation of GW150914}}.
-\newblock \emph{\bibinfo{journal}{ArXiv e-prints}}  (\bibinfo{year}{2016}).
-\newblock \eprint{arXiv:1611.07531}.
-
-\bibitem{2016PhRvD..94f4035A}
-\bibinfo{author}{{Abbott}, B.~P.} \emph{et~al.}
-\newblock \bibinfo{title}{{Directly comparing GW150914 with numerical solutions
-  of Einstein's equations for binary black hole coalescence}}.
-\newblock \emph{\bibinfo{journal}{\prd}} \textbf{\bibinfo{volume}{94}},
-  \bibinfo{pages}{064035} (\bibinfo{year}{2016}).
-
-\end{thebibliography}
-
 %\bibliographystyle{naturemag-arxiv}

### 2017-06-29 paper/CoverLetter2.tex
Updated cover letter.

% !TEX TS-program = pdflatexmk

\documentclass{letter}

\usepackage{hyperref}

\signature{Will M. Farr}

\address{Birmingham Institute for Gravitational Wave Astronomy\\School
  of Physics and Astronomy\\University of Birmingham\\Birmingham\\B15
  2TT\\United
  Kingdom\\\href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}}

\begin{document}

\begin{letter}{Dr.\ Leslie Sage\\968 National Press Building\\529 14th Street NW\\Washington DC 20045-1938\\United States}

\opening{Dear Dr. Sage:}

Thank you very much for notifying us of the provisional acceptance of
our manuscript for Nature Letters, ``Distinguishing Spin-Aligned and
Isotropic Black Hole Populations With Gravitational Waves.''  Since
you have received the last draft, our only edits have been to
re-format the paper appropriately for final submission.  We have 
\begin{itemize}
\item Removed the first footnote, and incorporated the last two
  footnotes into the main text.
\item Removed the URLs from references 32 and 34 (these references are
  complete without them).
\item Removed arXiv numbers from references to published papers.
\item Split the bibliography into two parts, one following the main
  text and one following the ``Methods'' section, with continuous
  numbering.
\item Moved the existing code availability statement to the end of the
  methods section and added a standard ``Data Availability'' statement
  immediately following (both code and data are stored in the same
  public repository on GitHub).
\item Moved figure captions to the end of the main section and methods
  section, and removed figures from the PDF produced by \LaTeX{}.
\item Corrected the references and title for the single table item to
  ``Extended Data Table 1.''
\end{itemize}

I am afraid that on my original submission I must have clicked ``Save
and Exit'' instead of ``Submit,'' so I apologize that this is reaching
you a week late.

\closing{Sincerely,}

\end{letter}

\end{document}

### 2017-06-30 paper/AlignedIsotropicGW.tex
Extended Data stuff in TIFF or EPS (ugh).

@@
 \begin{centering}
-\begin{tabular}{ c |  c  |  c  | c  }
-  \hline 
-  \hline
-Events & $\sigma_{I/A}$ &  $\sigma_{I/A}$ & $\sigma_{I/A}$ \\ 
- & ``Low'' & ``Flat'' & ``High'' \\ \hline
-GW150914 and GW151226 &  1.3 & 2.2 & 3.7 \\ %\hline
-All O1 events & 1.7 & 2.7 & 4.4 \\ %\hline
-All O1 events and GW170104 & \textbf{2.4} & 3.5 & 5.4 \\ \hline
-\end{tabular}
-\caption[]{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
+% \begin{tabular}{ c |  c  |  c  | c  }
+%   \hline 
+%   \hline
+% Events & $\sigma_{I/A}$ &  $\sigma_{I/A}$ & $\sigma_{I/A}$ \\ 
+%  & ``Low'' & ``Flat'' & ``High'' \\ \hline
+% GW150914 and GW151226 &  1.3 & 2.2 & 3.7 \\ %\hline
+% All O1 events & 1.7 & 2.7 & 4.4 \\ %\hline
+% All O1 events and GW170104 & \textbf{2.4} & 3.5 & 5.4 \\ \hline
+% \end{tabular}
+\caption[]{\textbf{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models.} The odds ratio is computed using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
 The number in bold is the result we quote in the main text.}

### 2017-07-25 paper/AlignedIsotropicGW.tex
Tweaking author contributions statement.
Put the figures and table back in (still at end of document sections).

--- 2017-06-29 paper/AlignedIsotropicGW.tex
@@
   No. NSF PHY11-25915.
-\item[Author Contributions] All authors contributed to the work
-  presented in this paper.
+\item[Author Contributions] All authors contributed at all stages to
+  the work presented in this paper.
 \item[Author Information] Reprints and permissions information is
@@
 \end{tabular}
-\caption[]{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
+\caption[]{\textbf{Significance $\sigma_{I/A}$ of the odds ratio between the isotropic and aligned models.} The odds ratio is computed using just GW150914 and GW151226, all 3 O1 events, and all 4 currently observed events (including GW170104).
 The number in bold is the result we quote in the main text.}
